### 2018-07-22 compact-encoding/doc/alt-encoding.tex
Add docs for alt-encoding

\documentclass[10pt]{article}

\usepackage{amsmath,amssymb,amsthm}
\usepackage[margin=1in]{geometry}
\usepackage{enumerate}
\usepackage{booktabs}
\usepackage{xspace}

\newcommand{\Omit}[1]{}
\newcommand{\tup}[1]{\langle #1 \rangle}
\newcommand{\denselist}{\topsep -1pt \itemsep -1pt}

\newcommand{\A}{\mathcal{A}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\M}{\mathcal{M}}
\renewcommand{\S}{\mathcal{S}}

\newcommand{\pre}[1]{\ensuremath{{\scriptsize\bullet}#1}}
\newcommand{\post}[1]{\ensuremath{#1{\scriptsize\bullet}}}
\newcommand{\act}[1]{\ensuremath{a(#1)}}

\newtheorem{definition}{Definition}

\begin{document}
\allowdisplaybreaks

\begin{center}
\Large\bf
Alternative encoding for learning feature-based reformulations \\
  {\normalsize Version 22/07/2018}
\end{center}

\bigskip\noindent
We aim at a reduction to SAT with parameters $K$ and $N$: $K$ bounds the size of
the abstract space by selecting $K$ features from the set of available features,
and $N$ bounds the number of abstract actions that can be used in the abstract
space.

The general idea is to use ``meta-variables'' for selecting the features that
define the abstract state model and then using ``meta-variables'' that define
the abstact actions over the abstract states.
The SAT problem ``constructs'' the abstract model and then verifies soundness
and completeness with respect to the transitions and states in the sample.

The set of available features is denoted with $\F$ and the set of sampled
transitions as $\M$.  The value of features in $\F$ at the states in the sample
$\M$ are given as input to the reduction in the form of a matrix $M$ that has
as many rows as the number of features in $\F$ and as many columns as the
number of states in $\M$.
We denote with $S$ and $T$ the set of states and transitions in $\M$.

If the unique states in $\M$ are $\{s_0,s_1,\ldots\}$, we use integer
$i$ to denote state $s_i$.
For boolean features $f$ and state $s_i$, the entry $m_{fi}$ in the matrix
$M$ is 1 iff the feature $f$ holds at state $s_i$.
For numerical features $f$ and state $s_i$, the entry $m_{fi}$ is the
value of $f$ at state $i$.

\bigskip\noindent
The size of the encoding is summarized in the table below where $S$ is number
of states, $T$ is number of transitions, $F$ is number of features, $L=\log_2(F)$,
$\Delta\leq L$, $D$ and $U$ are numbers for dummy and inexistent features
respectively, and $Q(L,K,N)=N(1+4K) + K(D+U) + (K-1)(L^2 + L) - 1$.

\begin{center}
  \begin{tabular}{lc}
    \toprule
      Parameters:   & $K=\text{\# selected features}$, $N=\text{\# abstract actions}$ \\
      \# variables: & $2SK(1 + N) + T(2K + 3N) + 2K(2N + L) - L$ \\
      \# clauses:   & $S(2KF + 7KN + N) + T(2KF + NK(10+\Delta) + 2N + 1) + Q(L,K,N)$ \\
    \bottomrule
  \end{tabular}
\end{center}

\medskip\noindent
The next table shows some experimental data over Blocksworld and Gripper:
\begin{center}
  \begin{tabular}{lcrrrrrr@{}rrr}
    \toprule
              &        &     &     & \multicolumn{3}{c}{Input}        && \multicolumn{3}{c}{SAT Theory} \\
    \cmidrule{5-7}
    \cmidrule{9-11}
      problem & status & $K$ & $N$ & \#states & \#trans. & \#features && \#vars & \#clauses & Minisat \\
    \midrule
     Blocks04 &    SAT &   5 &   8 &      125 &      272 &         73 &&  20,721 &   472,560 &   1.63 \\
     Blocks04 &  UNSAT &   5 &   7 &      125 &      272 &         73 &&  18,635 &   449,815 & $>$250 \\

    \midrule
     Blocks05 &    SAT &   5 &   8 &      866 &    2,090 &        102 && 149,223 & 4,387,384 &  16.31 \\
     Blocks05 &  UNSAT &   5 &   7 &      866 &    2,090 &        102 && 134,273 & 4,216,157 & $>$250 \\

    \midrule
    Gripper03 &    SAT &   5 &  10 &       88 &      280 &         12 &&  21,116 &   264,029 &   0.28 \\
    Gripper03 &  UNSAT &   5 &   9 &       88 &      280 &         12 &&  19,376 &   242,080 &   1.43 \\

    \midrule
    Gripper04 &    SAT &   5 &  10 &      256 &      896 &         14 &&  64,236 &   854,955 &   5.00 \\
    Gripper04 &  UNSAT &   5 &   9 &      256 &      896 &         14 &&  58,968 &   785,686 &   6.86 \\

    \midrule
    Gripper05 &    SAT &   5 &  10 &      704 &    2,624 &         16 && 182,636 & 2,546,913 &   2.47 \\
    Gripper05 &  UNSAT &   5 &   9 &      704 &    2,624 &         16 && 167,704 & 2,345,740 &  39.16 \\

    \midrule
    Gripper06 &    SAT &   5 &  10 &    1,856 &    7,232 &         18 && 493,685 & 7,518,671 &  39.25 \\
    Gripper06 &  UNSAT &   5 &   9 &    1,856 &    7,232 &         18 && 453,409 & 6,931,130 & 110.45 \\

    \bottomrule
  \end{tabular}
\end{center}

\section{Preliminaries}

The sample consists of transitions of the form $t=\tup{s,a,s'}$ that
are assumed to come from one (**** CHECK: or more than one? ****) instances
of a generalized problem.

The $K$ features selected from $\F$ define an abstract space given by the
$2^K$ boolean valuations of the $K$ features. For boolean features, such a
valuation contains the value of the feature. For numerical features, the
boolean value denotes whether the numerical feature is equal to or bigger
than zero.
An abstract state represents a set of concrete states in the sample.
Abstract actions look like a regular STRIPS action on the abstract space.
An abstract action $\tilde a$ is made of a precondition and an effect, both
being partial valuations of the $K$ features. For boolean features $f$, $f$
(resp.\ $\neg f$) appears in the precondition iff $\tilde a$ requires $f$
to hold (resp.\ not hold) in the abstract state, and $f$ (resp.\ $\neg f$)
appears in the effect if $\tilde a$ makes $f$ true (resp.\ false).
For numerical effects, positive (resp.\ negative) $f$-literals in the
precondition stand for the condition $f>0$ (resp.\ $f=0$), and positive
(resp.\ negative) $f$-literals in the effect stand for an increment
(resp.\ decrement) of $f$.
Hence, whether the feature $f$ is boolean or numeric, we can specify
the precondition or effect of an abstract action on $f$ with a boolean
$f$-literal.

Once abstract actions are defined over the abstract space, verifying
soundness is just a matter of checking whether for each abstract
transition, there is a matching concrete transition.
That is, let us suppose that the abstract state $\tilde s$ is mapped
by the abstract action $\tilde a$ into the abstract state $\tilde s'$.
The abstract action $\tilde a$ is made of a precondition and effect.
The precondition must hold at $\tilde s$ while the effect maps $\tilde s$
into $\tilde s'$.
We say that $\tilde a$ is sound (wrt $\M$) iff for each concrete state
$s$ on which the precondition of $\tilde a$ holds, there is transition
$t=\tup{s,a,s'}$ in $\M$ that ``matches'' the abstract transition
$\tilde t=\tup{\tilde s,\tilde a,\tilde s'}$.
That is, $s$ is represented by $\tilde s$ (i.e.\ for each of the $K$
features $f$, the boolean value of $f$ in $s'$ and $\tilde s'$ coincide),
$s'$ is represented by $\tilde s'$, and the selected numerical features
change in the transition $t$ as indicated by the abstact action
$\tilde a$: if $f$ is increased (resp.\ decreased) by $\tilde a$,
then $f$ increases (resp.\ decreases) along the transition $t$.

Completeness is the opposite. The abstract model is complete if for each
transition $t=\tup{s,a,s'}$ there is an abstract transition
$\tilde t=\tup{\tilde s,\tilde a,\tilde s'}$ that matches $t$.

Let $\sim$ be the equivalance relation over concrete states defined by
the $K$ selected features: $s\sim s'$ iff for each selected feature $f$,
the boolean value of $f$ at $s$ and $s'$ are the same. For state $s$
and abstract state $\tilde s$, we write $s\sim\tilde s$ when the boolean
value of the $K$ selected features coincide in $s$ and $\tilde s$.
Likewise, we denote with $[s]$ and $[\tilde s]$ the equivalence class
for state $s$ and the set $\{s:s\sim\tilde s\}$ respectively, and
we write $t\sim\tilde t$, for transition $t$ and abstract transition $\tilde t$,
when $t$ matches $\tilde t$.

\section{Encoding}

The input to the reduction consists of the matrix $M$ and the parameters
$K$ and $N$. We assume that the set of feature $\F$ is enumerated as
$f_0, f_1,\ldots$ in such a way that the boolean features appear only
after the numerical features, and that the first boolean feature $f_i$
has index $i$ that is a power of 2. The gap between the last numerical
feature and the first boolean feature is assumed to be filled with dummy
features.
With this encoding, checking whether a selected feature $f_i$ is numerical
just amounts to check whether its index is less than the index of the first
boolean feature.

For each $j\in K$, let $F^j$ be a multi-valued (meta-)variable with domain
$F$ that denotes the $j$-th feature selected from $\F$.
Each $F^j$ variable is encoded with $\lceil \log_2 F \rceil$ propositional
variables $F^j_k$ where $F$ is the number of features in $\F$.
To ensure different features are selected, it is enough that $F^i<F^j$ for
$i<j$.

In the construction of the abstract model, we only consider abstract actions
that put a precondition on the features affected by the action; i.e., an
abstract action cannot affect a feature for which there is no precondition,
but it may put a precondition on an unaffected feature.

Given $K$ selected features, there are at most $9^K$ different abstract actions.
Observe that each abstract action is the result of selecting a subset of affected
features from the $K$ features and considering the ways in which they can
be affected (2 ways for each feature), and selecting a subset of features
for preconditions and considering the ways the can appear as such (2 ways
for each feature).\footnote{The bound is not tight because we will require
  $f>0$ in precondition of actions that decrement $f$.}
Hence, a bound on the total number of abstract actions is
\begin{alignat}{1}
  \sum_{k=1}^K \binom{K}{k} 2^k \times \sum_{k=0}^{K} \binom{K}{k} 2^k\ =\ 3^K \times (3^K - 1)\ \leq\ 9^K \,.
\end{alignat}
Since $K$ is constant, this number is also constant.

Let $A^j$, for $j\in\{0,\ldots,N-1\}$, denote the $j$-th abstract action.
Instead of using an exact enumeration on the different abstract actions,
we consider the following loose enumeration.
As mentioned above, each abstract action is the result of selecting a subset
of the selected features as affected, putting an effect over such features,
and then adding preconditions.
Thus, each abstract action can be specified with $K$ bits to selected the
subset of affected features, $K$ bits for speciyfing the effect on each affected
feature, $K$ additional bits that select preconditions, and a final
block of $K$ bits that specify the value of the preconditions.
In total, we need $4K$ bits to specify an abstract action.
We use propositional variables $A^j_{i,k}$ for $j\in N$, $i\in\{1,\ldots,4\}$,
and $k\in\{0,\ldots,K-1\}$ to specify the abstract action $A^j$ where
\begin{enumerate}[--]\denselist
  \item the group $A^j_{1,k}$ selects the affected features,
  \item the group $A^j_{2,k}$ sets the effect for the selected features,
  \item the group $A^j_{3,k}$ selects the features for preconditions, and
  \item the group $A^j_{4,k}$ sets the precondition for the additional features.
\end{enumerate}
We assume that the precondition $f>0$ must be present when $f$ is decremented,
and we require that each abstract action affects at least one feature.
These restrictions are enforced with:
\begin{alignat}{1}
  &A^j_{2,k} \rightarrow A^j_{1,k} \,, \\
  &A^j_{4,k} \rightarrow A^j_{3,k} \,, \\
  &A^j_{1,k} \land \neg A^j_{2,k} \land [\text{$F^k$ is numeric}] \rightarrow A^j_{4,k} \,, \\
  &A^j_{1,0} \lor A^j_{1,1} \lor \cdots \lor A^j_{1,K-1} \,.
\end{alignat}
The total number of such clauses is $3NK + N$. On the other hand, symmetries
are broken among abstract actions with a weak ordering on them and by
requiring $A^0\leq A^1 \leq \cdots \leq A^{N-1}$. Abstract actions are ordered
using the $K$ bits for selected the affected features, for each $j\in\{1,\ldots,N-1\}$
and $k\in\{0,\ldots,K-1\}$:
\begin{alignat}{1}
  &\neg A^j_{1,0} \land \cdots \land \neg A^j_{1,k-1} \land A^j_{1,k} \rightarrow A^{j-1}_{1,0} \lor \cdots \lor A^{j-1}_{1,k} \,.
\end{alignat}
The number of clauses in (2)--(8) is $3NK + N + (N-1)K = N(1 + 4K) - K$.

\medskip\noindent
The clauses needed for the $F^k_t$ variables establish that no dummy
or inexistent feature is selected, and enforce the strict ordering
among features. The first group corresponds to the following clauses
where $d$ is an index of a dummy or inexistent feature:
\begin{alignat}{1}
  &\bigvee_{t:bit(t,d)} \neg F^k_t \ \lor \bigvee_{t:\neg bit(t,d)} F^k_t
\end{alignat}
where $bit(t,d)$ is a boolean constant for the $t$-th bit in the binary
expansion of $d$.  The ordering among selected featues is enforced with
\begin{alignat}{1}
  &df(k,0) \lor df(k,1) \lor \cdots \lor df(k,L)
\end{alignat}
for $k\in\{0,\ldots,K-2\}$ and $L=\log_2(F)$, and the boolean
variables $df(k,t)$ satisfy
\begin{alignat}{1}
  &df(k,t) \rightarrow \neg F^k_t \,, \\
  &df(k,t) \rightarrow \neg F^{k+1}_t \,, \\
  &df(k,t) \land F^k_i \rightarrow F^{k+1}_i \qquad \text{for $i\in\{t+1,\ldots,L-1\}$} \,, \\
  &df(k,t) \land \neg F^{k+1}_i \rightarrow F^k_i \qquad \text{for $i\in\{t+1,\ldots,L-1\}$} \,.
\end{alignat}
The total number of clauses in (7)--(12) is $K(D+U) + (K-1)(1 + 2L + L(L-1))$
where $D$ and $U$ are the number of dummy and inexistent features respectively.
The total number of clauses in (2)--(12) is
$Q(L,K,N)=N(1+4K) + K(D+U) + (K-1)(L^2 + L) - 1$.

\subsection{Value of selected features at states}

We define the variable $\phi(k,i)$ that tells whether the selected feature $F^k$
holds at state $s_i$ in the sample; $\phi(k,i)$ can be defined in two equivalent
ways:
\begin{alignat}{1}
  \label{eq:phi:1}
  \phi(k,i)\
    &\equiv\ \bigwedge_{\ell \in \F} [F^k = f_\ell] \rightarrow (m_{\ell i} > 0) \
     \equiv\ \bigwedge_{\ell \in \F: m_{\ell i}=0} \neg [F^k = f_\ell] \
     \equiv\ \bigwedge_{\ell \in \F: m_{\ell i}=0} \neg \biggl( \bigwedge_t F^k_t \leftrightarrow bit(t,\ell) \biggr) \\
  \label{eq:phi:2}
  \phi(k,i)\
    &\equiv\ \bigvee_{\ell \in \F} [F^k = f_\ell] \land (m_{\ell i} > 0) \
     \equiv\ \bigvee_{\ell \in \F: m_{\ell i}>0} [F^k = f_\ell] \
     \equiv\ \bigvee_{\ell \in \F: m_{\ell i}>0} \biggl( \bigwedge_t F^k_t \leftrightarrow bit(t,\ell) \biggr)
\end{alignat}
where $[F^k=f_\ell]$ denotes the formula that says the $k$-th selected feature is
$f_\ell$.
Therefore, for each $k\in K$ and $i\in\M$, the variable $\phi(k,i)$ is defined
with the following implications:
\begin{alignat}{1}
  \phi(k,i)\ &\rightarrow\ \biggl(\bigvee_{t:bit(t,\ell)} \neg F^k_t\biggr) \lor \biggl(\bigvee_{t:\neg bit(t,\ell)} F^k_t\biggr) \,,
               \quad \text{for each $\ell$ such that $m_{\ell i}=0$,} \\
  \phi(k,i)\ &\leftarrow\  \biggl(\bigwedge_{t:bit(t,\ell)} F^k_t\biggr) \land \biggl(\bigwedge_{t:\neg bit(t,\ell)} \neg F^k_t\biggr) \,,
               \quad \text{for each $\ell$ such that $m_{\ell i}>0$.}
\end{alignat}
These account for exactly $SKF$ clauses where $S$ is the number of states in $\M$
and $F$ is the number of features in $\F$.
Likewise, we define the variable $\phi^*(k,i)$ that tells whether the selected
feature $F^j$ \emph{does not} hold at $s_i$:
\begin{alignat}{1}
  \label{eq:phi*:1}
  \phi^*(k,i)\
    &\equiv\ \bigwedge_{\ell \in \F} [F^k = f_\ell] \rightarrow (m_{\ell i} = 0) \
     \equiv\ \bigwedge_{\ell \in \F: m_{\ell i}>0} \neg [F^k = f_\ell] \
     \equiv\ \bigwedge_{\ell \in \F: m_{\ell i}>0} \neg \biggl( \bigwedge_t F^k_t \leftrightarrow bit(t,\ell) \biggr) \\
  \label{eq:phi*:2}
  \phi^*(k,i)\
    &\equiv\ \bigvee_{\ell \in \F} [F^k = f_\ell] \land (m_{\ell i} = 0) \
     \equiv\ \bigvee_{\ell \in \F: m_{\ell i}=0} [F^k = f_\ell] \
     \equiv\ \bigvee_{\ell \in \F: m_{\ell i}=0} \biggl( \bigwedge_t F^k_t \leftrightarrow bit(t,\ell) \biggr)
\end{alignat}
with the following $SKF$ clauses:
\begin{alignat}{1}
  \phi^*(k,i)\ &\rightarrow\ \biggl(\bigvee_{t:bit(t,\ell)} \neg F^k_t\biggr) \lor \biggl(\bigvee_{t:\neg bit(t,\ell)} F^k_t\biggr) \,,
                \quad \text{for each $\ell$ such that $m_{\ell i}>0$,} \\
  \phi^*(k,i)\ &\leftarrow\  \biggl(\bigwedge_{t:bit(t,\ell)} F^k_t\biggr) \land \biggl(\bigwedge_{t:\neg bit(t,\ell)} \neg F^k_t\biggr) \,,
                \quad \text{for each $\ell$ such that $m_{\ell i}=0$.}
\end{alignat}

\subsection{Soundness}

We express soundness with the following formulas, one for each state $s$ in the
sample $\M$, and for each $j\in\{0,\ldots,N-1\}$:
\begin{alignat}{1}
  \label{eq:soundness}
  app(A^j,[s]) \rightarrow \textstyle\bigvee_{t=\tup{s,a,s'}\in\M} t \sim \tup{[s],A^j,res(A^j,[s])}
\end{alignat}
where $app(A^j,[s])$ means that the abstract action denoted by $A^j$ is applicable
at the abstract state $\tilde s$ such that $s\sim\tilde s$, $res(A^j,[s])$ denotes
the abstract state that results of applying $A^j$ in $\tilde s$, and
$\tilde t = \tup{[s],A^j,res(A^j,[s]}$ is the abstract transition from $\tilde s$
to $\tilde s'=res(A^j,[s])$ induced by $A^j$.

For $s_i=s$, the formula $app(A^j,[s])$ is captured with:
\begin{alignat}{1}
  app(A^j,[s])\
    &\equiv\ \bigwedge_{k} \biggl(A^j_{4,k} \rightarrow \phi(k,i)\biggr)\ \land\
             \bigwedge_{k} \biggl(A^j_{3,k} \land \neg A^j_{4,k} \rightarrow \phi^*(k,i)\biggr)\
     \equiv\ \bigwedge_{k} B_1(j,i,k) \land B_2(j,i,k)
\end{alignat}
where the auxiliary variables $B_1(j,i,k)$ and $B_2(j,i,k)$ are given by:
\begin{alignat}{1}
  &B_1(j,i,k) \land A^j_{4,k} \rightarrow \phi(k,i)\,, \\
  &\neg A^j_{4,k} \rightarrow B_1(j,i,k)\,, \\
  &\phi(k,i) \rightarrow B_1(j,i,k)\,, \\[1em]
  &B_2(j,i,k) \land A^j_{3,k} \land \neg A^j_{4,k} \rightarrow \phi^*(k,i)\,, \\
  &\neg A^j_{3,k} \rightarrow B_2(j,i,k)\,, \\
  &A^j_{4,k} \rightarrow B_2(j,i,k)\,, \\
  &\phi^*(k,i) \rightarrow B_2(j,i,k)\,.
\end{alignat}
In total, we need $7NSK$ clauses to define the auxiliary $B$ variables.

Let us focus on the consequent of the implication that defines soundness. We need
an expression for $t \sim \tup{[s],A^j,res(A^j,[s])}$ where $t=\tup{s,a,s'}$ is
a transition in $\M$, $A^j$ is an abstract action, and $res(A^j,[s])$ denotes
the abstract state that results of applying $A^j$ on $[s]$.
By definition, $\tup{s,a,s'}\sim\tup{\tilde s,\tilde a,\tilde s'}$ iff
1)~$s\sim\tilde s$, 2)~$s'\sim\tilde s'$, and 3)~the numeric features change
in both transitions in a compatible way.
In our case, $t=\tup{s_i,a,s_\ell}$ and $\tilde t=\tup{[s_i],A^j,res(A^j,[s_i])}$
and (1) is always satisfied.
For (2), we use the auxiliary variable $RES(i,\ell,j)$ to encode $s_\ell\sim res(A^j,[s_i])$:
\begin{alignat}{1}
  RES(i,\ell,j)\
    &\equiv\  \bigwedge_{k} \biggl(A^j_{2,k} \rightarrow \phi(k,\ell)\biggr)\ \land\
              \bigwedge_{k} \biggl(A^j_{1,k} \land \neg A^j_{2,k} \land \neg [\text{$F^k$ is numeric}] \rightarrow \phi^*(k,\ell)\biggr)\ \land \\
    &\quad\ \ \bigwedge_{k} \biggl(\neg A^j_{1,k} \land \phi(k,i) \rightarrow \phi(k,\ell)\biggr)\ \land\
              \bigwedge_{k} \biggl(\neg A^j_{1,k} \land \phi^*(k,i) \rightarrow \phi^*(k,\ell)\biggr)\,.
\end{alignat}
Since $RES(i,\ell,j)$ only appears in the consequent of the implications defining soundness,
it is enough to enforce the following implications:
\begin{alignat}{1}
  &RES(i,\ell,j) \land A^j_{2,k} \rightarrow \phi(k,\ell)\,, \\
  &RES(i,\ell,j) \land A^j_{1,k} \land \neg A^j_{2,k} \land \neg [\text{$F^k$ is numeric}] \rightarrow \phi^*(k,\ell)\,, \\
  &RES(i,\ell,j) \land \neg A^j_{1,k} \land \phi(k,i) \rightarrow \phi(k,\ell)\,, \\
  &RES(i,\ell,j) \land \neg A^j_{1,k} \land \phi^*(k,i) \rightarrow \phi^*(k,\ell)\,.
\end{alignat}
There are 4 implications for each transition $t=\tup{s_i,a,s_\ell}$ in $\M$,
each $j\in\{0,\ldots,N-1\}$, and each $k\in\{0,\ldots,K-1\}$.

The final requirement to enforce $t\sim\tilde t$ is that the numerical features
change in both transitions in a compatible manner. For each numerical feature $f$,
if $f$ is increased (resp.\ decreased) by $A^j$, then $f$ increases (resp.\ decreases)
in the transition from $s$ to $s'$. We use auxiliary variables $INC(k,i,\ell)$
(resp.\ $DEC(k,i,\ell)$) to denote that feature $k$ increases (resp.\ decreases)
in the transition $\tup{s_i,a,s_\ell}$ in $\M$.\footnote{**** CHECK: Can we have
  two transitions with equal source and destination but with different actions?
  Does it matter? It seems that it doesn't matter since whether $f$ increases
  or decreases in a transition only depent of the values of $f$ at the states.
  ****}
The $INC(k,i,\ell)$ variable can be expressed in two different ways:
\begin{alignat}{1}
  INC(k,i,\ell)\ &\equiv\ \bigwedge_{j\in\F} [F^k = f_j] \rightarrow (m_{ji} < m_{j\ell})\
                  \equiv\ \bigwedge_{j\in\F:m_{ji}\geq m_{j\ell}} \neg [F^k = f_j] \\
                 &\equiv\ \bigwedge_{j\in\F:m_{ji}\geq m_{j\ell}} \neg \biggl(\bigwedge_t F^k_t \leftrightarrow bit(t,j)\biggr) \,, \\
  INC(k,i,\ell)\ &\equiv\ \bigvee_{j\in\F} [F^k = f_j] \land (m_{ji} < m_{j\ell})\
                  \equiv\ \bigvee_{j\in\F:m_{ji}<m_{j\ell}} [F^k = f_j] \\
                 &\equiv\ \bigvee_{j\in\F:m_{ji}<m_{j\ell}} \biggl(\bigwedge_t F^k_t \leftrightarrow bit(t,j)\biggr) \,.
\end{alignat}
We therefore use the following implications to define $INC(k,i,\ell)$:
\begin{alignat}{1}
  INC(k,i,\ell)\ &\rightarrow\ \biggl(\bigvee_{t:bit(t,j)} \neg F^k_t\biggr) \lor \biggl(\bigvee_{t:\neg bit(t,j)} F^k_t\biggr) \,,
                 \quad \text{for each $j$ such that $m_{ji} \geq m_{j\ell}$,} \\
  INC(k,i,\ell)\ &\leftarrow\  \biggl(\bigwedge_{t:bit(t,j)} F^k_t\biggr) \land \biggl(\bigwedge_{t:\neg bit(t,j)} \neg F^k_t\biggr) \,,
                 \quad \text{for each $j$ such that $m_{ji} < m_{j\ell}$.}
\end{alignat}
The $DEC(k,i,\ell)$ variables are defined analogously with the implications:
\begin{alignat}{1}
  DEC(k,i,\ell)\ &\rightarrow\ \biggl(\bigvee_{t:bit(t,j)} \neg F^k_t\biggr) \lor \biggl(\bigvee_{t:\neg bit(t,j)} F^k_t\biggr) \,,
                 \quad \text{for each $j$ such that $m_{ji} \leq m_{j\ell}$,} \\
  DEC(k,i,\ell)\ &\leftarrow\  \biggl(\bigwedge_{t:bit(t,j)} F^k_t\biggr) \land \biggl(\bigwedge_{t:\neg bit(t,j)} \neg F^k_t\biggr) \,,
                 \quad \text{for each $j$ such that $m_{ji} > m_{j\ell}$.}
\end{alignat}
Therefore, there are $2TK$ such auxiliaries variables that are defined with $2TFK$ implications.
The following auxiliary variables capture $\tup{s_i,a,s_\ell}\sim\tup{[s_i],A^j,res(A^j,[s_i)}$:
\begin{alignat}{1}
  MATCH(i,\ell,j)\ &\equiv\ RES(i,\ell,j)\ \land \\
                   &\quad\ \, \bigwedge_k \biggl([\text{$F^k$ is numeric}] \rightarrow (A^j_{2,k} \leftrightarrow INC(k,i,\ell))\biggr)\ \land \\
                   &\quad\ \, \bigwedge_k \biggl(A^j_{1,k} \land [\text{$F^k$ is numeric}] \rightarrow (\neg A^2_{2,k} \leftrightarrow DEC(k,i,\ell))\biggr) \,.
\end{alignat}
As these variables only appear in the consequent of implications, it is sufficient to only
enforce:
\begin{alignat}{1}
  &MATCH(i,\ell,j) \rightarrow RES(i,\ell,j) \,, \\
  &MATCH(i,\ell,j) \land [\text{$F^k$ is numeric}] \land A^j_{2,k} \rightarrow INC(k,i,\ell) \,, \\
  &MATCH(i,\ell,j) \land [\text{$F^k$ is numeric}] \land INC(k,i,\ell) \rightarrow A^j_{2,k} \,, \\
  &MATCH(i,\ell,j) \land [\text{$F^k$ is numeric}] \land A^j_{1,k} \land \neg A^j_{2,k} \rightarrow DEC(k,i,\ell) \,, \\
  &MATCH(i,\ell,j) \land [\text{$F^k$ is numeric}] \land DEC(k,i,\ell) \rightarrow A^j_{1,k} \,, \\
  &MATCH(i,\ell,j) \land [\text{$F^k$ is numeric}] \land DEC(k,i,\ell) \rightarrow \neg A^j_{2,k} \,.
\end{alignat}
There are $TN + 5TNK$ such implications.

Finally, soundness is enforced with $SN$ implications, one for each state $s_i$ and $j\in\{0,\ldots,N-1\}$:
\begin{alignat}{1}
  \biggl(\bigwedge_k B_1(j,i,k) \land B_2(j,i,k)\biggr)\ \rightarrow
  \bigvee_{t=\tup{s_i,a,s_\ell}} MATCH(i,\ell,j) \,.
\end{alignat}

\subsection{Completeness}

For each transition $t=\tup{s,a,s'}$ in $\M$, there is an abstract transition $\tilde t$ such that $t\sim\tilde t$;
that is, for each transition $t=\tup{s_i,a,s_\ell}$ in $\M$:
\begin{alignat}{1}
  \label{eq:completeness}
  \textstyle\bigvee_{j} app(A^j,[s_i]) \land t \sim \tup{[s_i],A^j,res(A^j,[s_i])} \,.
\end{alignat}
Since $MATCH(i,\ell,j)$ implies $\tup{s_i,a,s_\ell}\sim\tup{[s_i],A^j,res(A^j,[s_i])}$,
completeness can be capture with
\begin{alignat}{1}
  \label{eq:completeness}
  \textstyle\bigvee_{j} APPMATCH(i,\ell,j) \,,
\end{alignat}
one for each transition $\tup{s_i,a,s_\ell}$ in $\M$, where the auxiliary variables
$APPMATCH(i,\ell,j)$ satisfy
\begin{alignat}{1}
  &APPMATCH(i,\ell,j) \rightarrow MATCH(i,\ell,j) \,, \\
  &APPMATCH(i,\ell,j) \rightarrow B_1(j,i,k) \,, \\
  &APPMATCH(i,\ell,j) \rightarrow B_2(j,i,k)
\end{alignat}
for $k=0,\ldots,K-1$.

\end{document}

### 2018-07-22 compact-encoding/psvn/blocksworld/features.cpp
Add psvn

@@ -1,0 +1,534 @@
+/*
+This program computes the distance to goal (i.e. the cost of the least-cost path to the goal)
+of every state from which the goal can be reached.
+It does this by executing Dijkstra's algorithm backwards from the goal.
+It prints on stdout each state and its distance (distance first, then the state) and, if a filename is
+provided as a command line argument, it prints the state_map it builds to that file.
+
+Copyright (C) 2013 by the PSVN Research Group, University of Alberta
+*/
+
+#include <iostream>
+#include <fstream>
+#include <string>
+#include <map>
+#include <set>
+#include <vector>
+#include "priority_queue.hpp"
+
+using namespace std;
+
+class Feature {
+  protected:
+    mutable int value_;
+  public:
+    Feature() : value_(0) { }
+    virtual ~Feature() { }
+    virtual string name() const = 0;
+    virtual void calculate(int num_blocks, const state_t &state) const = 0;
+    int value() const { return value_; }
+};
+
+class BooleanFeature : public Feature {
+  public:
+    BooleanFeature() { }
+    virtual ~BooleanFeature() { }
+};
+
+class NumericalFeature : public Feature {
+  public:
+    NumericalFeature() { }
+    virtual ~NumericalFeature() { }
+    int num_above(int block, int num_blocks, const state_t &state) const {
+        int count = 0;
+        if( state.vars[block] == 0 ) { // not clear(block)
+            for( int b = 1; b <= num_blocks; ++b ) {
+                if( state.vars[b] == 1 ) {
+                    int t = b;
+                    while( (t != 0) && (t != block) ) {
+                        ++count;
+                        t = state.vars[num_blocks + t];
+                    }
+                    if( t == block ) break;
+                }
+                count = 0;
+            }
+            assert(count > 0);
+        }
+        return count;
+    }
+    int num_below(int block, int num_blocks, const state_t &state) const {
+        int count = 0;
+        int b = state.vars[num_blocks + block];
+        while( b != 0 ) {
+            ++count;
+            b = state.vars[num_blocks + b];
+        }
+        return count;
+    }
+    int num_other(int block, int num_blocks, const state_t &state) const {
+        int nabove = num_above(block, num_blocks, state);
+        int nbelow = num_below(block, num_blocks, state);
+        assert(1 + nbelow + nabove <= num_blocks);
[…]

### 2018-07-22 compact-encoding/psvn/blocksworld/gen_blocks.cc

@@ -1,0 +1,281 @@
+#include <cassert>
+#include <iostream>
+
+using namespace std;
+
+class Blocks {
+  protected:
+    int num_blocks_;
+
+  public:
+    Blocks(int num_blocks) : num_blocks_(num_blocks) { }
+    ~Blocks() { }
+
+    string block(int i) const {
+        assert(i >= 0);
+        return i == 0 ? string("0") : string("") + char(int('A' + i - 1));
+    }
+
+    void print_stack(ostream &os, int x, int y) const {
+        // PRE: holding
+        os << block(1 + x) << "  ";
+
+        // PRE: clear
+        for( int i = 0; i < num_blocks_; ++i ) {
+            if( (i == x) || (i == y) )
+                os << " 1";
+            else
+                os << " -";
+        }
+        os << "  ";
+
+        // PRE: on
+        for( int i = 0; i < num_blocks_; ++i ) {
+            if( i == x )
+                os << " 0";
+            else
+                os << " -";
+        }
+
+        // arrow + POST: holding
+        os << "  =>  0  ";
+
+        // POST: clear
+        for( int i = 0; i < num_blocks_; ++i ) {
+            if( i == x )
+                os << " 1";
+            else if( i == y )
+                os << " 0";
+            else
+                os << " -";
+        }
+        os << "  ";
+
+        // POST: on
+        for( int i = 0; i < num_blocks_; ++i ) {
+            if( i == x )
+                os << " " << block(1 + y);
+            else
+                os << " -";
+        }
+        os << "  ";
+
+        // label
+        os << "LABEL stack_" << block(1 + x) << "_" << block(1 + y) << endl;
+    }
+
+    void print_unstack(ostream &os, int x, int y) const {
+        // PRE: holding
+        os << "0  ";
+
+        // PRE: clear
+        for( int i = 0; i < num_blocks_; ++i ) {
+            if( i == x )
+                os << " 1";
+            else if( i == y )
+                os << " 0";
+            else
+                os << " -";
+        }
+        os << "  ";
+
+        // PRE: on
+        for( int i = 0; i < num_blocks_; ++i ) {
+            if( i == x )
+                os << " " << block(1 + y);
+            else
+                os << " -";
+        }
+
+        // arrow + POST: holding
+        os << "  =>  " << block(1 + x) << "  ";
+
+        // POST: clear
+        for( int i = 0; i < num_blocks_; ++i ) {
[…]

### 2018-07-22 compact-encoding/psvn/global/dist.cpp

@@ -1,0 +1,80 @@
+/*
+This program computes the distance to goal (i.e. the cost of the least-cost path to the goal)
+of every state from which the goal can be reached.
+It does this by executing Dijkstra's algorithm backwards from the goal.
+It prints on stdout each state and its distance (distance first, then the state) and, if a filename is
+provided as a command line argument, it prints the state_map it builds to that file.
+
+Copyright (C) 2013 by the PSVN Research Group, University of Alberta
+*/
+
+#include <vector>
+#include "priority_queue.hpp"
+
+int main( int argc, char **argv )
+{
+    state_t state, child;   // NOTE: "child" will be a predecessor of state, not a successor
+    int d, ruleid;
+    ruleid_iterator_t iter;
+
+    PriorityQueue<state_t> open; // used for the states we have generated but not yet expanded (the OPEN list)
+    state_map_t *map = new_state_map(); // contains the cost-to-goal for all states that have been generated
+    FILE *file; // the final state_map is written to this file if it is provided (command line argument)
+
+    /* add goal states */
+    first_goal_state( &state, &d ); do {
+        state_map_add( map, &state, 0 );
+        open.Add(0, 0, state );
+    } while( next_goal_state( &state, &d ) );
+
+    d = 0;
+    while( !open.Empty() ) {
+
+        /* get current distance from goal */
+        d = open.CurrentPriority();
+
+        /* get state */
+        state = open.Top();
+        open.Pop();
+        
+        /* check if we already expanded this state.
+	   (entries on the open list are not deleted if a cheaper path to a state is found) */
+        const int *best_dist = state_map_get( map, &state );
+        assert(best_dist != NULL);
+        if (*best_dist < d)
+            continue;
+        
+/* print the distance then the state */
+        printf("%d  ",d);
+        print_state(stdout,&state);
+        printf(" \n");
+
+        /* look at all predecessors of the state */
+        init_bwd_iter( &iter, &state );
+        while( ( ruleid = next_ruleid( &iter ) ) >= 0 ) {
+            apply_bwd_rule( ruleid, &state, &child );
+            const int child_d = d + get_bwd_rule_cost( ruleid );
+
+            /* check if either this child has not been seen yet or if
+               there is a new cheaper way to get to this child. */
+            const int *old_child_d = state_map_get( map, &child );
+            if ( old_child_d == NULL || *old_child_d > child_d ) {
+                /* add to open with the new distance */
[…]

### 2018-07-22 compact-encoding/psvn/global/distSummary.cpp

@@ -1,0 +1,93 @@
+/*
+This program computes the distance to goal (i.e. the cost of the least-cost path to the goal)
+of every state from which the goal can be reached.
+It does this by executing Dijkstra's algorithm backwards from the goal.
+It prints on stdout the number of states at each distance from goal and, if a filename is
+provided as a command line argument, it prints the state_map it builds to that file.
+
+Copyright (C) 2013 by the PSVN Research Group, University of Alberta
+*/
+
+#include <vector>
+#include "priority_queue.hpp"
+
+int main( int argc, char **argv )
+{
+    int64_t totalNodes, numAtD;  // counters
+    state_t state, child;   // NOTE: "child" will be a predecessor of state, not a successor
+    int d, ruleid;
+    ruleid_iterator_t iter;
+
+    PriorityQueue<state_t> open; // used for the states we have generated but not yet expanded (the OPEN list)
+    state_map_t *map = new_state_map(); // contains the cost-to-goal for all states that have been generated
+    FILE *file; // the final state_map is written to this file if it is provided (command line argument)
+
+    /* add goal states */
+    first_goal_state( &state, &d ); do {
+        state_map_add( map, &state, 0 );
+        open.Add(0, 0, state );
+    } while( next_goal_state( &state, &d ) );
+
+    totalNodes = 0;
+    d = 0;
+    numAtD = 0;
+    while( !open.Empty() ) {
+
+        /* get current distance from goal; since operator costs are
+           non-negative this distance is monotonically increasing */
+        if (open.CurrentPriority() > d) {
+            printf( "%"PRId64" states at distance %d\n", numAtD, d );
+            d = open.CurrentPriority();
+            numAtD = 0;
+        }
+
+        /* get state */
+        state = open.Top();
+        open.Pop();
+        
+        /* check if we already expanded this state.
+	   (entries on the open list are not deleted if a cheaper path to a state is found) */
+        const int *best_dist = state_map_get( map, &state );
+        assert(best_dist != NULL);
+        if (*best_dist < d)
+            continue;
+        
+        numAtD++;
+        totalNodes++;
+
+        /* look at all predecessors of the state */
+        init_bwd_iter( &iter, &state );
+        while( ( ruleid = next_ruleid( &iter ) ) >= 0 ) {
+            apply_bwd_rule( ruleid, &state, &child );
+            const int child_d = d + get_bwd_rule_cost( rule );
+
[…]

### 2018-07-22 compact-encoding/psvn/global/succ.c

@@ -1,0 +1,62 @@
+/* This program reads a state from stdin and prints out its successors.
+
+Copyright (C) 2013 by the PSVN Research Group, University of Alberta
+*/
+
+#include <stdlib.h>
+#include <stdio.h>
+#include <inttypes.h>
+#include <assert.h>
+#include <sys/time.h>
+
+#define  MAX_LINE_LENGTH 999 
+
+int main( int argc, char **argv )
+{
+// VARIABLES FOR INPUT
+    char str[ MAX_LINE_LENGTH +1 ] ;
+    ssize_t nchars; 
+    state_t state; // state_t is defined by the PSVN API. It is the type used for individual states.
+
+// VARIABLES FOR ITERATING THROUGH state's SUCCESSORS
+    state_t child;
+    ruleid_iterator_t iter; // ruleid_terator_t is the type defined by the PSVN API successor/predecessor iterators.
+    int ruleid ; // an iterator returns a number identifying a rule
+    int childCount;
+
+// READ A LINE OF INPUT FROM stdin
+    printf("Please enter a state followed by ENTER: ");
+    if ( fgets(str, sizeof str, stdin) == NULL ) {
+	printf("Error: empty input line.\n");
+	return 0; 
+    }
+
+// CONVERT THE STRING TO A STATE
+    nchars = read_state( str, &state );
+    if (nchars <= 0) {
+	printf("Error: invalid state entered.\n");
+	return 0; 
+    }
+
+    printf("The state you entered is: ");
+    print_state( stdout, &state );
+    printf("\n");
+
+// LOOP THOUGH THE CHILDREN ONE BY ONE
+    childCount = 0;
+    init_fwd_iter( &iter, &state );  // initialize the child iterator 
+      while( ( ruleid = next_ruleid( &iter ) ) >= 0 ) {
+//    while( -1 >= 0 ) {
+	apply_fwd_rule( ruleid, &state, &child );
+	++childCount;
+    	printf("child %d. ",childCount);
+	print_state( stdout, &child );
+    	//printf("  %s (cost %d)\n",get_fwd_rule_label(ruleid),get_fwd_rule_cost(ruleid));
+    	printf("  %s (cost %d), goal=%d\n",get_fwd_rule_label(ruleid),get_fwd_rule_cost(ruleid),is_goal(&child));
+    } // end while... no more children
+    if (childCount == 0) {
+	printf("Your state has no children.\n");
+    }
+
+    return 0;
+} // end main

### 2018-07-22 compact-encoding/psvn/gripper/features.cpp

@@ -1,0 +1,384 @@
+/*
+This program computes the distance to goal (i.e. the cost of the least-cost path to the goal)
+of every state from which the goal can be reached.
+It does this by executing Dijkstra's algorithm backwards from the goal.
+It prints on stdout each state and its distance (distance first, then the state) and, if a filename is
+provided as a command line argument, it prints the state_map it builds to that file.
+
+Copyright (C) 2013 by the PSVN Research Group, University of Alberta
+*/
+
+#include <iostream>
+#include <fstream>
+#include <string>
+#include <map>
+#include <set>
+#include <vector>
+#include "priority_queue.hpp"
+
+using namespace std;
+
+class Feature {
+  protected:
+    mutable int value_;
+  public:
+    Feature() : value_(0) { }
+    virtual ~Feature() { }
+    virtual string name() const = 0;
+    virtual void calculate(int num_balls, const state_t &state) const = 0;
+    int value() const { return value_; }
+};
+
+class BooleanFeature : public Feature {
+  public:
+    BooleanFeature() { }
+    virtual ~BooleanFeature() { }
+};
+
+class NumericalFeature : public Feature {
+  public:
+    NumericalFeature() { }
+    virtual ~NumericalFeature() { }
+    int num_ball_in_room(int room, int num_balls, const state_t &state) const {
+        int count = 0;
+        for( int b = 0; b < num_balls; ++b )
+            count += state.vars[3 + b] == room ? 1 : 0;
+        return count;
+    }
+};
+
+
+class DummyFeature : public BooleanFeature {
+  public:
+    DummyFeature() { }
+    virtual ~DummyFeature() { }
+    virtual string name() const {
+        return string("dummy()");
+    }
+    virtual void calculate(int num_balls, const state_t &state) const {
+        value_ = 0;
+    }
+};
+
+class HoldingBall : public BooleanFeature {
+  protected:
+    int ball_;
+    int gripper_;
+  public:
+    HoldingBall(int ball, int gripper) : ball_(ball), gripper_(gripper) { }
+    virtual ~HoldingBall() { }
+    virtual string name() const {
+        return string("holding(") + to_string(ball_) + "," + to_string(gripper_) + ")";
+    }
+    virtual void calculate(int num_balls, const state_t &state) const {
+        value_ = state.vars[1 + gripper_] == 1 + ball_;
+    }
+};
+
+class Holding : public BooleanFeature {
+  protected:
+    int gripper_;
[…]

### 2018-07-22 compact-encoding/psvn/gripper/gen_gripper.cc

@@ -1,0 +1,185 @@
+#include <cassert>
+#include <iostream>
+
+using namespace std;
+
+class Gripper {
+  protected:
+    int num_balls_;
+
+  public:
+    Gripper(int num_balls) : num_balls_(num_balls) { }
+    ~Gripper() { }
+
+    string ball(int i) const {
+        assert(i > 0);
+        return string("b") + to_string(i);
+    }
+
+    void print_move(ostream &os, int from, int to) const {
+        // PRE
+        os << from << "  - - ";
+        for( int i = 0; i < num_balls_; ++i )
+            os << " -";
+
+        // arrow
+        os << "  =>  ";
+
+        // POST
+        os << to << "  - - ";
+        for( int i = 0; i < num_balls_; ++i )
+            os << " -";
+        os << "  ";
+
+        // label
+        os << "LABEL move_" << from << "_" << to << endl;
+    }
+
+    void print_pick(ostream &os, int r, int b, char gripper) const {
+        // PRE: robby
+        os << r << "  ";
+
+        // PRE: grippers
+        os << (gripper == 'L' ? "0 -" : "- 0");
+        os << " ";
+
+        // PRE: balls
+        for( int i = 0; i < num_balls_; ++i ) {
+            if( i == b )
+                os << " " << r;
+            else
+                os << " -";
+        }
+                
+        // arrow
+        os << "  =>  ";
+
+        // POST: robby
+        os << r << "  ";
+
+        // POST: grippers
+        if( gripper == 'L' )
+            os << ball(1 + b) << " -";
+        else
+            os << "- " << ball(1 + b);
+        os << " ";
+
+        // POST: balls
+        for( int i = 0; i < num_balls_; ++i ) {
+            if( i == b )
+                os << " " << gripper;
+            else
+                os << " -";
+        }
+        os << "  ";
+
+        // label
+        os << "LABEL pick_" << r << "_" << ball(1 + b) << "_" << gripper << endl;
+    }
+
+    void print_drop(ostream &os, int r, int b, char gripper) const {
+        // PRE: robby
+        os << r << "  ";
+
+        // PRE: grippers
+        if( gripper == 'L' )
+            os << ball(1 + b) << " -";
+        else
+            os << "- " << ball(1 + b);
+        os << " ";
+
+        // PRE: balls
+        for( int i = 0; i < num_balls_; ++i ) {
+            if( i == b )
[…]

### 2018-07-22 compact-encoding/psvn/priority_queue.hpp

@@ -1,0 +1,162 @@
+/*
+  Copyright (C) 2011 by the PSVN Research Group, University of Alberta
+*/
+
+#ifndef _PRIORITY_QUEUE_HPP
+#define _PRIORITY_QUEUE_HPP
+
+#include <queue>
+#include <cstdio>
+#include <cassert>
+#include <cstdlib>
+
+template <typename T>
+class PriorityQueue
+{
+public:
+    PriorityQueue();
+    int Add(int f, int g, const T& data);
+    const T& Top() const;
+    int CurrentPriority() const;
+    bool Empty() const;
+    void Pop();
+    void Clear();
+
+    void Modify(int f, int g, int index, const T& data);
+
+private:
+    /* Min priority queue on f. 
+       If same f: prioritize on max g. */
+    class Bucket
+    {
+    public:
+        void Pop();
+        bool Empty() const;
+        int Add(int g, const T& data);
+        const T& Top() const;
+        void Clear();
+        void Modify(int g, int index, const T& data);
+    private:
+        std::vector<std::vector<T> > m_stacks;
+    };
+    int m_max;
+    int m_at;
+    std::vector<Bucket> m_buckets;
+};
+
+template<typename T>
+const T& PriorityQueue<T>::Bucket::Top() const
+{
+    assert(!m_stacks.empty());
+    return m_stacks.back().back();
+}
+
+template<typename T>
+void PriorityQueue<T>::Bucket::Pop()
+{
+    m_stacks.back().pop_back();
+    while (!m_stacks.empty() && m_stacks.back().empty())
+        m_stacks.pop_back();
+}
+
+template<typename T>
+bool PriorityQueue<T>::Bucket::Empty() const
+{
+    return m_stacks.empty();
+}
+
+template<typename T>
+int PriorityQueue<T>::Bucket::Add(int g, const T& data)
+{
+    if (m_stacks.size() <= (size_t)g)
+        m_stacks.resize(g + 1);
+    m_stacks[g].push_back(data);
+    return m_stacks[g].size() - 1;
+}
+
+template<typename T>
+void PriorityQueue<T>::Bucket::Modify(int g, int index, const T& data)
+{
+    // if (g >= (int)m_stacks.size() || index >= m_stacks[g].size()) {
+    //     printf("%d %d | %d %d\n", g, index, m_stacks.size(), m_stacks[g].size());
+    // }
[…]

### 2018-07-22 compact-encoding/psvn/psvn.cpp

@@ -1,0 +1,438 @@
+/*
+Copyright (C) 2011 by the PSVN Research Group, University of Alberta
+*/
+
+#include "psvn.hpp"
+#include <assert.h>
+#include <sstream>
+
+ostream &operator<<( ostream &os, const PSVN &psvn )
+{
+  size_t i, j;
+
+  /* find un-used domains */
+  vector<bool> domainUsed( psvn.numVariables(), false );
+  for( i = 0; i < psvn.numVariables(); ++i ) {
+    domainUsed[ psvn.variableDomain( i ) ] = true;
+  }
+
+  bool explicitDomains = false;
+  for( i = 0; i < psvn.numDomains(); ++i ) {
+    if( !domainUsed[ i ] ) {
+      continue;
+    }
+    if( sscanf( psvn.domainName( i ).c_str(), "%zu", &j ) ) {
+      /* don't print out integer domains */
+      continue;
+    }
+
+    explicitDomains = true;
+    os << "DOMAIN " << psvn.domainName( i )
+	 << " " << psvn.domainValues( i ).size() << endl;
+    os << " ";
+
+    for( j = 0; j < psvn.domainValues( i ).size(); ++j ) {
+      os << " " << psvn.domainValues( i )[ j ];
+    }
+    os << endl;
+  }
+  if( explicitDomains ) {
+    os << endl;
+  }
+
+  os << psvn.numVariables() << endl;
+  for( i = 0; i < psvn.numVariables(); ++i ) {
+    if( i ) {
+      os << " ";
+    }
+    os << psvn.domainName( psvn.variableDomain( i ) );
+  }
+  os << endl << endl;
+
+  for( i = 0; i < psvn.numRules(); ++i ) {
+
+    for( j = 0; j < psvn.numVariables(); ++j ) {
+      if( j ) {
+	os << " ";
+      }
+      os << psvn.ruleTests( i )[ j ];
+    }
+
+    os << " =>";
+
+    for( j = 0; j < psvn.numVariables(); ++j ) {
+      os << " " << psvn.ruleActions( i )[ j ];
+    }
+
+    os << " LABEL " << psvn.ruleLabel( i );
+    os << " COST " << psvn.ruleCost( i ) << endl;
+  }
+  os << endl;
+
+  for( i = 0; i < psvn.numGoalRules(); ++i ) {
+
+    os << "GOAL";
+    for( j = 0; j < psvn.numVariables(); ++j ) {
[…]

### 2018-07-22 compact-encoding/psvn/psvn.hpp

@@ -1,0 +1,160 @@
+/*
+Copyright (C) 2011 by the PSVN Research Group, University of Alberta
+*/
+
+#ifndef _GAMEFILE_HPP
+#define _GAMEFILE_HPP
+
+#include <vector>
+#include <string>
+#include <iostream>
+#include <fstream>
+#include <stdlib.h>
+
+using namespace std;
+
+/** Marks a variable as unused. 
+    More portable than a #pragma directive. */
+template <class T>
+inline void UNUSED(const T&)
+{
+}
+
+
+class PSVN {
+private:
+  friend ostream &operator<<( ostream &os, const PSVN &psvn );
+
+  PSVN() {}
+
+protected:
+  vector<string> m_domainNames;
+  vector<vector<string> > m_domainValues;
+
+  vector<size_t> m_variableDomains;
+
+  vector<vector<string> > m_ruleTests;
+  vector<vector<string> > m_ruleActions;
+  vector<string> m_ruleLabels;
+  vector<size_t> m_ruleCosts;
+
+  vector<vector<string> > m_goalRules;
+
+  /* parse a file into whitespace separated uppercase tokens */
+  static vector<string> getTokens( istream &os );
+
+public:
+
+  /* read a PSVN description from a file
+     returns a newly allocated PSVN description on success
+     returns NULL on failure
+     NOTE: THIS WILL READ ALL INPUT UP TO END OF FILE! */
+  static PSVN *fromInput( istream &is );
+
+  size_t numDomains() const { return m_domainNames.size(); }
+  const vector<string> &domainNames() const { return m_domainNames; }
+  const string &domainName( const size_t d ) const {
+    return m_domainNames[ d ];
+  }
+
+  /* returns the index of a named domain 'name' in list 'domainNames',
+     or numDomains() if it does not exist */
+  size_t getDomain( const string &name ) {
+    size_t i;
+    for( i = 0; i < m_domainNames.size(); ++i ) {
+      if( m_domainNames[ i ].compare( name ) == 0 ) {
+	break;
+      }
+    }
+    return i;
+  }
+
+  bool domainIsUsed( const size_t domain ) const {
+    for( size_t v = 0; v < m_variableDomains.size(); ++v ) {
+      if( m_variableDomains[ v ] == domain ) { return true; }
+    }
+    return false;
+  }
+
+  const vector<vector<string> > &domainValues() const {
+    return m_domainValues;
+  }
+  const vector<string> &domainValues( const size_t domain ) const {
+    return m_domainValues[ domain ];
+  }
[…]

### 2018-07-22 compact-encoding/psvn/psvn2c.cpp

@@ -1,0 +1,3408 @@
+/*
+Copyright (C) 2011, 2012, 2014 by the PSVN Research Group, University of Alberta
+*/
+
+#include <fstream>
+#include <stdlib.h>
+#include <string.h>
+#include <assert.h>
+#include <limits.h>
+#include "psvn2c.hpp"
+
+/*
+The original (2011) version of the move pruning code tested two things:
+(a) is the newly generated operator sequence redundant with a previously generated sequence?, and
+(b) is the some previously generated sequence redundant with the newly generated operator sequence?
+In 2012 we discovered that it was unsafe to do both of these, so the current version only does (a).
+This is thoroughly explained in Burch and Holte's SoCS 2012 paper, and the correctness of (a) is formally proven.
+
+I have left the original 2011 code here in case anyone wants to reproduce those results and/or
+test other ways of making move pruning safe.  It has been disabled by putting it inside #ifdef's.
+To reinstate it, just move the following #define outside this comment block.
+#define UnsafeVersion
+*/
+
+/* Verbosity Levels */
+#define V_QUIET 0
+#define V_USER  1
+#define V_DEBUG0 2
+#define V_DEBUG1 3
+#define V_DEBUG2 4
+#define V_DEBUG_MP 5
+
+
+/* DEFINE THE DEFAULT SETTINGS AND MAXIMUM VALUES FOR THE COMMAND LINE PARAMETERS */
+#define  historyLenMAX 99
+#define  fwdHistoryLenDEFAULT  1
+#define  bwdHistoryLenDEFAULT  1
+#define  bwdMovesDEFAULT       true
+#define  ruleGroupSizeDEFAULT  100
+#define  ruleGroupSizeMAX  1000
+#define  removeDuplicateRulesDEFAULT  false
+#define  dynamicAbstractionsDEFAULT  false
+#define  AbstractionDEFAULT  true
+#define  state_mapDEFAULT  true
+#define  verbosityDEFAULT V_QUIET
+#define  verbosityMAX 9
+
+
+#define ASSERT( X )
+
+static bool valueInSet( const Value &val, const vector<Value> &values )
+{
+  size_t i;
+
+  for( i = 0; i < values.size(); ++i ) {
+    if( val == values[ i ] ) {
+      return true;
+    }
+  }
+
+  return false;
+}
+
[…]

### 2018-07-22 compact-encoding/psvn/psvn2c.hpp

@@ -1,0 +1,606 @@
+/*
+Copyright (C) 2011 by the PSVN Research Group, University of Alberta
+*/
+
+#ifndef _GAME_H
+#define _GAME_H
+
+#include <vector>
+#include <string>
+#include <iostream>
+#include <fstream>
+#include "psvn.hpp"
+
+using namespace std;
+
+
+class Game;
+class Context;
+class VarTest;
+
+enum ValueType { Constant, Label };
+
+class Value {
+protected:
+  ValueType m_type;
+  size_t m_value;
+
+public:
+  Value( ValueType type, size_t value )
+    : m_type( type ), m_value( value ) {}
+
+  bool operator==( const Value &v ) const {
+    if( m_type == v.m_type && m_value == v.m_value ) { return true; }
+    return false;
+  }
+
+  const ValueType type() const { return m_type; }
+  const size_t value() const { return m_value; }
+
+  string toString() const {
+    string s; char t[ 32 ];
+    if( m_type == Label ) { s.append( "[" ); }
+    sprintf( t, "%zu", m_value ); s.append( t );
+    if( m_type == Label ) { s.append( "]" ); }
+    return s;
+  }
+};
+
+
+class Rule {
+protected:
+  size_t m_id;
+  string m_name;
+  size_t m_cost;
+  vector<vector<Value> > m_tests;
+  vector<vector<Value> > m_actions;
+  bool m_modified;	// used in RuleTrees.  A test taken resulted
+			// in the rule being modified (e.g. test removed)
+
+public:
+  Rule( const size_t id, const string name, const size_t cost,
+	const vector<vector<Value> > &tests,
+	const vector<vector<Value> > &actions )
+    : m_id( id ),
+      m_name( name ),
+      m_cost ( cost ), 
+      m_tests( tests ),
+      m_actions( actions ),
+      m_modified( false ) {
+    assert( tests.size() == actions.size() );
+  }
+
+  bool operator==( const Rule &rule ) const;
+
+  bool isEmpty() const;
+
+  const size_t id() const { return m_id; }
+  const string name() const { return m_name; }
+  const size_t cost() const { return m_cost; }
+  const vector<vector<Value> > &tests() const { return m_tests; }
+  const vector<vector<Value> > &actions() const { return m_actions; }
+
+  /* only used in RuleTrees */
+  const bool wasModified() const { return m_modified; }
+
+  void setId( const size_t id ) { m_id = id; }
+
[…]

### 2018-07-22 compact-encoding/psvn/psvn2c_abstraction.c

@@ -1,0 +1,325 @@
+/*
+Copyright (C) 2011, 2014 by the PSVN Research Group, University of Alberta
+*/
+
+typedef struct {
+  int size;
+  var_t *v;
+} abst_array_t;
+
+typedef struct {
+  var_t *value_map[ NUMDOMAINS ];
+  uint8_t project_away_var[ NUMVARS ];
+  abst_array_t* mapped_in[ NUMDOMAINS ];
+  int* fwd_rule_label_sets;
+  int* bwd_rule_label_sets;
+} abstraction_t;
+
+
+static abstraction_t* allocate_abstraction()
+{
+    int i;
+    int64_t s;
+    abstraction_t* abst = (abstraction_t *)malloc( sizeof( *abst ) );
+    if( abst == NULL )
+        return NULL;
+    
+    for( s = 0, i = 0; i < NUMDOMAINS; ++i ) {
+        s += domain_sizes[ i ];
+    }
+    abst->value_map[ 0 ]
+        = (var_t * )malloc( sizeof( abst->value_map[ 0 ][ 0 ] ) * s );
+    if( abst->value_map[ 0 ] == NULL ) {
+        free( abst );
+        return NULL;
+    }
+    abst->mapped_in[ 0 ] 
+        = (abst_array_t*)malloc(sizeof(abst->mapped_in[0][0]) * s);
+    if (abst->mapped_in[ 0 ] == NULL) {
+        free(abst->value_map[0]);
+        free(abst);
+        return NULL;
+    }
+    for( s = domain_sizes[ 0 ], i = 1;
+         i < NUMDOMAINS;
+         s += domain_sizes[ i ], ++i ) 
+    {
+        abst->value_map[ i ] = &abst->value_map[ 0 ][ s ];
+        abst->mapped_in[ i ] = &abst->mapped_in[ 0 ][ s ];
+    }
+
+    for (i = 0; i < NUMDOMAINS; ++i){
+        abst->mapped_in[ i ][ 0 ].v 
+            = (var_t*) malloc (sizeof(var_t) * domain_sizes[i]);
+    }
+
+    return abst;
+}
+
+
+static void destroy_abstraction( abstraction_t *abst )
+{
+    int i;
+    for (i = 0; i < NUMDOMAINS; ++i)
+        free ( abst->mapped_in[i][0].v );
+    free( abst->mapped_in[ 0 ] );
+    free( abst->value_map[ 0 ] );
+    free( abst );
+}
+
+
+
[…]

### 2018-07-22 compact-encoding/psvn/psvn2c_core.c

@@ -1,0 +1,436 @@
+/*
+Copyright (C) 2011, 2014 by the PSVN Research Group, University of Alberta
+*/
+
+#include <string.h>
+
+/* copy a state */
+#define copy_state(dest_ptr,src_ptr) memcpy(dest_ptr,src_ptr,sizeof(var_t)*NUMVARS)
+
+/* compare two state pointers
+   returns 0 if equal, non-zero otherwise */
+#define compare_states(a,b) memcmp(a,b,sizeof(var_t)*NUMVARS)
+
+
+/* print a state to a file
+   returns number of characters on success, -1 on failure */
+static ssize_t print_state( FILE *file, const state_t *state )
+{
+  size_t len, t; int i;
+  for( len = 0, i = 0; i < NUMVARS; ++i ) {
+    t = fprintf( file, "%s ", var_domain_names[ i ][ state->vars[ i ] ] );
+    if( t < 0 ) { return -1; }
+    len += t;
+  }
+  return len;
+}
+
+/* print a state to a string
+   returns number of characters on success, -1 on failure */
+static ssize_t sprint_state( char *string,const size_t max_len,const state_t *state )
+{
+  size_t len, t; int i;
+  for( len = 0, i = 0; i < NUMVARS; ++i ) {
+    t = snprintf( &string[ len ], max_len - len, "%s ",
+		  var_domain_names[ i ][ state->vars[ i ] ] );
+    if( t < 0 ) { return -1; }
+    len += t;
+  }
+  if( len >= max_len ) { return -1; } else { string[ len ] = 0; }
+  return len;
+}
+
+/* read a state from a string
+   returns number of characters consumed on success, -1 on failure
+   NOTE: the part of the string representing the state must be
+   followed either by whitespace (which will all be consumed) or
+   the end of string marker */
+static ssize_t read_state( const char *string, state_t *state )
+{
+  size_t len, t; int i; var_t j;
+  len = 0;
+  while( (string[len] !=0) && isspace(string[len]) ) { len++; } // skip leading white space
+  for( i = 0; i < NUMVARS; ++i ) {
+    for( j = 0; j < domain_sizes[ var_domains[ i ] ]; ++j ) {
+      t = strlen( var_domain_names[ i ][ j ] );
+      if( !strncasecmp( var_domain_names[ i ][ j ], &string[ len ], t ) ) {
+          if( (string[len+t] ==0) || isspace(string[len+t]) ) { // require a terminator after the match
+	      state->vars[ i ] = j;
+	      len += t;
[…]

### 2018-09-03 domains/grid-circles/generator.py
Playing with grid circles

@@ -127,7 +127,7 @@ def generate_propositional_domain(gridsize, num_rewards, num_blocks, add_noop=Fa
 
 def main():
 
-    add_noop = True
+    add_noop = False
     for gridsize in [3, 5, 7, 10]:
         num_blocks_and_rewards = gridsize
         problem = generate_propositional_domain(gridsize, num_blocks_and_rewards-2, num_blocks_and_rewards-2, add_noop)

### 2018-09-03 learn_actions.py

@@ -243,7 +243,7 @@ class ModelTranslator(object):
         assert s in self.optimal_states
 
         for s_prime in self.transitions[s]:
-            if s_prime < s or s_prime not in self.optimal_states:  # We want only optimal transitions
+            if not self.is_optimal_transition(s, s_prime):
                 continue
 
             forward_clause_literals = [d1_literal]
@@ -260,13 +260,20 @@ class ModelTranslator(object):
             self.np_qchanges[(s0, s1)] = val = self.compute_qchanges(s0, s1)
             return val
 
+    def is_optimal_transition(self, s, sprime):
+        """ Return whether the give transition is optimal.
+        To compute so, we make use of the fact that states are generated breadth-first, and thus given two
+        "optimal" states, only one of the transitions is truly optimal. We might want to improve this in the sampling
+        so thas this looks less hacky and is more robust to other generation strategies.
+        """
+        return sprime > s and sprime in self.optimal_states
+
     def iterate_over_optimal_transitions(self):
         """ """
         for s in self.optimal_states:
             for sprime in self.transitions[s]:
-                if sprime < s or sprime not in self.optimal_states:
-                    continue
-                yield (s, sprime)
+                if self.is_optimal_transition(s, sprime):
+                    yield (s, sprime)
 
     def iterate_over_all_transitions(self):
         for s in self.transitions:

### 2018-09-03 doc/experiments.tex
Add some latex notes on experiments

\documentclass[12pt]{article}

\topmargin 0in
\headheight 0in
\headsep 0in
\oddsidemargin 0in
\evensidemargin 0in
\textheight 9in
\textwidth 6.5in
\parskip 2ex

\usepackage{color,soul}
\newcommand\hg[1]{\sethlcolor{green}\hl{HG: #1}\sethlcolor{cyan}}
\newcommand\bb[1]{\sethlcolor{cyan}\hl{BB: #1}}
\newcommand\gf[1]{\sethlcolor{yellow}\hl{GF: #1}}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{enumerate}
\usepackage{upgreek}
\usepackage[normalem]{ulem}
\usepackage{natbib}

\newcommand{\Omit}[1]{}
\newcommand{\denselist}{\itemsep -1pt\partopsep 0pt}
\newcommand{\tup}[1]{\langle #1 \rangle}
\newcommand{\pair}[1]{\langle #1 \rangle}
\newcommand{\tuple}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\set}[1]{\ensuremath{\left\{#1 \right\}}}
\newcommand{\setst}[2]{\ensuremath{\left\{#1 \mid #2 \right\}}}
\newcommand{\abs}[1]{\ensuremath{\left\vert{#1}\right\vert}}
\newcommand{\citeay}[1]{\citeauthor{#1} [\citeyear{#1}]}
\newcommand{\citeaywithnote}[2]{\citeauthor{#1} (\citeyear[#2]{#1})}
\newcommand{\mli}[1]{\mathit{#1}} % Multi-letter identifier, see https://tex.stackexchange.com/a/129434

\newtheorem{definition}{Definition}
\newtheorem{definitionandtheorem}[definition]{Definition and Theorem}
\newtheorem{theorem}[definition]{Theorem}
\newtheorem{lemma}[definition]{Lemma}
\newtheorem{proposition}[definition]{Proposition}
\newtheorem{corollary}[definition]{Corollary}
\newtheorem{example}[definition]{Example}

\def\withproofs{0} % to include proofs in the draft: 0=don't include, 1=include
\newcommand{\nosuchproof}{\textcolor{red}{\bf No such proof yet...}}
\newcommand{\problem}{\textcolor{red}{\bf **** PROBLEM ****}}
\newcommand{\alert}[1]{\textcolor{red}{#1}}
\newcommand{\CHECK}[1]{\textcolor{red}{\bf Note: #1}}

\newcommand{\C}{\mathcal{C}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}

\newcommand{\true}{\mathsf{true}}
\newcommand{\false}{\mathsf{false}}
\newcommand{\Eff}{{\mathit{Eff}}}

\newcommand{\abst}[2]{\tup{#1;#2}}
\newcommand{\Rule}[2]{\ensuremath{#1 \Rightarrow #2}}

% decrements and increments
\newcommand{\pplus}{\hspace{-.05em}\raisebox{.15ex}{\footnotesize$\uparrow$}}
\newcommand{\mminus}{\hspace{-.05em}\raisebox{.15ex}{\footnotesize$\downarrow$}}

% numerical effects
\newcommand{\dec}[1]{#1:=#1-1}
\newcommand{\inc}[1]{#1:=#1+1}

% qualitative effects
\newcommand{\Dec}{Dec}
\newcommand{\Inc}{Inc}
\newcommand{\RInc}{\ensuremath{\text{Inc}^*}}  % relaxed increase

% STRIPS actions
\newcommand{\Stack}{\text{Stack}}
\newcommand{\Unstack}{\text{Unstack}}
\newcommand{\Pickup}{\text{Pickup}}
\newcommand{\Putdown}{\text{Putdown}}
\newcommand{\Move}{\text{Move}}

% Examples
\newcommand{\Example}{\medskip\noindent\textbf{Example.}\xspace}

\begin{document}

\begin{center}
  \textbf{\large Learning Relational Features for Generalized Planning} \\
  BB, GF, HG; Revised \today
\end{center}

\section{Examples}

\subsection{Clearing a Block in Blocksworld}
We want to achieve a goal $clear(x)$ for a certain block $x$, starting from
any block configuration, where $x$ is a domain parameter.
When in validation mode, our approach validates
the feature set described in \citep{bonet2018features} as being sound and complete.
That feature set has a total aggregated complexity of $33$, and its most complex feature,
labelled as $m(x)$ (which denotes the set of blocks in any state which
is not in the same tower as block $x$ and is not being held either), has complexity $18$.

\paragraph{Learning on $3$ blocks.}
If we feed our approach with the full state space of a 3-block blocksworld, we learn an abstract space based on 5 features (total complexity: 12):

\begin{itemize}
 \item $F_1 \equiv holding$: whether some block is being held.
 \item $F_2 \equiv ontable \sqcap \set{x}$: whether block $x$ is on the table.
 \item $F_3 \equiv clear \sqcap \set{x}$: whether block $x$ is clear.

 \item $F_4 \equiv \abs{\neg ontable}$: Number of blocks not in the table.
 \item $F_5 \equiv \abs{ontable \sqcap clear}$: Number of blocks that are clear and on the table.

\end{itemize}

\noindent
The abstract action model that derives from these features is overfit to the fact that there are only three blocks.
% An example is the use of the condition $\neg F_1 \land (F_5 = 0)$ as a simple surrogate for the more complex concept
% that would describe that ``all blocks are on a single tower'' --- only if we have a number of blocks $n \leq 3$ both 
% concepts describe the same thing.
An example is the use of the condition $\neg F_1 \land \neg F_2 \land \neg F_3$ as a surrogate for the 
condition ``there is one single block on top of block $x$''. The latter can also be expressed with concept-based
features, but they have higher complexity, hence our approach prefers the former (this is because 
with only 3 blocks, if $x$ is not being held, not on the table and not clear, then it must be in the middle
of a 3-block tower). We hence learn an action model that is sound with respect to the samples, but not sound \emph{in general}.

\paragraph{Learning on $4$ blocks.}
If we expand the full 4-block state space, 
 we learn an abstract space based on 7 features (total complexity: 20):

\begin{itemize}
 \item $F_1 \equiv holding$: whether some block is being held.
 \item $F_2 \equiv ontable \sqcap \set{x}$: whether block $x$ is on the table.
 \item $F_3 \equiv holding \sqcap \set{x}$: whether block $x$ is being held.
 \item $F_4 \equiv \abs{\exists on . \top}$: Number of blocks that are on top of some other block.
 \item $F_5 \equiv \abs{\exists on^* . \set{x}}$: Number of blocks on top of block $x$.
 \item $F_6 \equiv \abs{\exists on . \neg ontable}$: Number of blocks that have at least two blocks below.
 \item $F_7 \equiv \abs{ontable \sqcap clear}$: Number of blocks that are clear and on the table.

\end{itemize}

\noindent
$F_5$ is the concept $n(x)$ from \citep{bonet2018features}, but the set of features does not generalize well.
\gf{(Haven't figured out the details, but checked it with our approach on larger BW instances and doesn't work)}

\paragraph{Learning on $5$ blocks with random sampling.}

Only with $5$ blocks, a similar set of features that that of \citep{bonet2018features}, with 5 features, turns out to be the simplest
which is sound and complete. The total complexity of the set is 26:

\begin{itemize}
 \item $F_1 \equiv handempty$: whether the hand is empty.
 \item $F_2 \equiv ontable \sqcap \set{x}$: whether block $x$ is on the table.
 \item $F_3 \equiv holding \sqcap \set{x}$: whether block $x$ is being held.
 \item $F_4 \equiv \abs{\exists on^* . \set{x}}$: Number of blocks on top of block $x$.
 \item $F_5 \equiv \abs{C_1 \sqcap C_2 \sqcap \neg \set{x}}$,
\end{itemize}

\noindent where 
$C_1 \equiv \forall on^* . \neg \set{x}$ denotes the set of blocks which are not above $x$,
$C_2 \equiv \forall (on^{-1})^* . \neg \set{x}$ the set of blocks which are not below $x$,
and the whole conjunction $F_5$ thus denotes the set of blocks not in the same tower as $x$.

This abstraction we computed by using a slightly different approach, in which the whole state space is expanded, but then
a certain number of states are \textbf{randomly sampled}, with the idea of providing a more diverse set of situations, for a fixed number
of states and transitions. In this case, with as few as $40$ state expansions, we obtain the reported abstraction.
The space of all concept-based features of size $k \leq 15$, which is the size of the feature $F_5$ above,
contains a total of $6858$ non-redundant features.
The resulting max-sat problem features $20006$ variables and $25.99$ million clauses.

\paragraph{Learning on $5$ blocks with optimal-state marking.}
An alternative to scale up is to pursue completeness only with respect to states lying on some optimal path.

\subsection{Blocksworld: $on(x, y)$ from 2 Towers}

Let's assume we want to achieve a goal $on(x, y)$, where $x$ and $y$ are domain parameters which denote
blocks that start in two different towers.

We here use directly the optimal-state marking strategy.

We train the approach with two small instances:
\begin{enumerate}
 \item One where we are holding block block $x$, and we have a single tower $D C B$ (i.e. D is on C, which is on B, and B is on the table).
 \item Another instance where we have two towers. Tower \#1 is B D C, tower \#2 is F E A.
\end{enumerate}

The abstract space learnt has 6 features (total aggregated complexity: 20)

\begin{itemize}
 \item $handempty$.
 \item $n(x)$: Number of blocks above $x$.
 \item $n(y)$: Number of blocks above $x$.
 \item $holding(x)$.
 \item $ontable(x)$.
 \item $F_6 \equiv \set{x} \sqcap (\forall on . \set{y})$: Whether block $x$ is either on $y$, on the table, or held.
\end{itemize}

The feature $ontable(x)$ is necessary to properly identify the goal, which is $F_6 \land handempty \land \neg ontable(x)$.
\gf{Not totally what we wanted, but looks good}

(\gf{Experiment ID: bw\_on\_x\_y\_completeness\_opt})

\Omit{
\subsection{Creating a Tower in Blocksworld}

With a 5-block state space, and using concepts up to size 15, the abstraction we find overfits
to the fact that there are only that many blocks. To illustrate, the existence of a tower
of size $3$ can be expressed with the feature
$\abs{\exists on . (\exists on . ontable)} > 0$,
that of a tower of size $4$, with the concept
$\abs{\exists on .\left[ \exists on . (\exists on . ontable)\right]} > 0$, and so on.

Interestingly, expressing the notion of ``a single tower'' using our concept-based language turns out to be harder
that expression the notion of ``a single tower with block $x$ at the bottom''. Whereas no truly generalized expression
for the first is found through our approach with a maximum feature size of 15, the feature 
$$\abs{\neg \set{x} \sqcap (\forall on^* . \neg \set{x})},$$
%     not_x_and_all_below_not_x = AndConcept(not_table, AndConcept(all_below_not_x, not_x, "object"), "object")
\noindent denotes how many blocks $y \neq x$ are not above $x$ (when this is 0, we must necessarily
have a tower based on $x$), and has only size $8$.
}

\subsection{Gripper}

In Gripper a robot with two grippers must move a number of balls which are in a certain room $A$ to a certain target room $B$,
every gripper being only able to hold one single ball. We use the standard encoding of Gripper with 
\emph{move}, \emph{pick} and \emph{drop} actions, and predicates that denote the location of balls and robot,
and whether each gripper is holding some ball or is empty. We consider the target room $x$ as the only domain parameter.
Distance features play no role in this problem, as all rooms are connected with the target room.

We learn on a small instance with only 3 balls, and expanding the full state space with 88 states and 280 transitions.
With a maximum feature size of 10 and a max. of
3 iterations of the concept-generating grammar, our approach learns an abstract space based on the following features:\footnote{
Interestingly, the max-sat compilation is quite large, in the order of roughly 400K variables and 40 million clauses, but the characteristics
of the state space allow the max-sat solver to solve it in around 20 seconds on a standard laptop.}

\begin{itemize}
 \item $\abs{free}$: Number of empty grippers.
 \item $at\_robby \sqcap \set{x}$: Whether the robot is in the target room.
 \item $\exists at . \set{x}$: Number of balls in the target room.
 \item $\exists at . \neg \set{x}$: Number of balls not in the target room.
 \item $\exists carry . \top$: Number of balls being held.
\end{itemize}

The abstract state space results in an abstract action model which is sound and complete in general.
\gf{double-check}

\gf{I also have a generalization of gripper with $n$ rooms and $m$ robots, but haven't tested it thoroughly. Might make for a more interesting discussion.}

\subsection{Moving in a Grid}

An agent must move between two cells in a given bidimensional grid, with
the position of the represented as a pair $(x, y)$ of Cartesian coordinates in the range $[1, M]$,
and we use the standard natural successor relation $succ(a, b)$ to denote that $b = a + 1$.
From a simple $5 \times 5$ grid where the agent needs to go from one corner to the opposite one, 
we learn an abstract state space 
with two features
$\abs{\exists succ^*. x}$ and
$\abs{\exists (succ^{-1})^*. x}$
(and likewise for coordinate $y$, for a total of four features),
which denote, respectively, the distance between the current $x$ coordinate and the left and right limits of the grid.

\subsection{Graph Traversal}

An agent must traverse an arbirtary graph from a source to a target node. The problem is encoded 
using a set of nodes plus an adjacency relation between nodes. When allowing distance features, our approach learns an abstract
state space made up of a single feature 
\[D = dist(at, adjacent, at_g)\]
that represents the minimum distance between the current and the target node. The abstract action model is:
\begin{enumerate}
 \item $\tuple{D>0; D\downarrow}$
 \item $\tuple{D>0; \RInc(D)}$
 \item $\tuple{D=0; \RInc(D)}$
\end{enumerate}

Using \emph{goal concepts} is of course required to generate this feature; without them, no abstract model is found.
Using a the relaxed \emph{increase} semantics is also required.
Depending on the topology of the graph, a \emph{no-op} concrete action might be necessary for the above abstract model to be sound.\footnote{
This happens when there is some node $v$ (e.g. in a line graph) from which the distance $D>0$ to the target node can not be increased, but this
node $v$ is abstracted into the same abstract state as other nodes $D>0$ for which it can be increased. In this case, the action above with effect
$\RInc(D)$ requires a concrete no-op in order to be sound.
}

\subsection{Collecting Rewards in a Grid}

An agent moves in a bidimensional grid collecting rewards that are on random cells of the grid. The goal is to collect all available reward.
The concrete actions allow the agent to move between any two adjacent cells, and to pick the reward on the current cell, if any (only one reward item per cell is allowed).
The encoding uses an adjacency relation, not a Cartesian coordinate system (the distinction is relevant).

The approach learns a simple abstraction with two features $R = \abs{reward}$ and $D = dist(at, adjacent, reward)$, the first denoting the number of uncollected rewards,
and the second the minimum distance between the current cell and any cell with reward. The resulting abstract action model has the following 5 actions (names are only a posteriori manual labels):

\begin{enumerate}
 \item $\text{pick-reward}: \tuple{R>0, D=0; \Dec(R), \RInc(D)}$

 \item $\text{move-to-reward}: \tuple{D>0; \Dec(D)}$
 \item $\text{move-away-from-reward}: \tuple{D=0, R>0; \RInc(D)}$
 \item $\text{move-otherwise}: \tuple{D>0; \RInc(D)}$
\end{enumerate}

This example however illustrates the shortcomings of the relaxed increase semantics. 
If we discard the previous solution, another more complex state space is learnt, based only on the feature
\[D = dist(at, adjacent, T), \]

\noindent where the concept $T \equiv \exists adjacent^*.reward$ denotes the set of all cells from which the reward is reachable.
The distance $D$ is thus always $0$, except when no reward is left on the grid, in which case it is $\infty$. This allows to distinguish between goal
and non-goal states, but the resulting abstract action model is:
\begin{enumerate}
 \item $\tuple{D>0; \RInc(D)}$
 \item $\tuple{D=0; \RInc(D)}$
\end{enumerate}

which is not too useful. The problem here is that $D$ can only increase from $0$ to $\infty$, and the relaxed INC semantics
allow this set of abstract actions to be sound.

{
\small
\bibliographystyle{named}
\bibliography{control}
}

\end{document}

### 2018-09-04 doc/paper.tex
Hector's draft

\def\year{2019}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
\usepackage{aaai19}  %Required
\usepackage{times}  %Required
\usepackage{helvet}  %Required
\usepackage{courier}  %Required
\usepackage{url}  %Required
\usepackage{graphicx}  %Required
\frenchspacing  %Required
\setlength{\pdfpagewidth}{8.5in}  %Required
\setlength{\pdfpageheight}{11in}  %Required
%PDF Info Is Required:
  \pdfinfo{
/Title ()
/Author ()}
\setcounter{secnumdepth}{0}

% \usepackage{soul}
% \usepackage[hidelinks]{hyperref}
% \usepackage[utf8]{inputenc}
% \usepackage[small]{caption}
%\usepackage{helvet}
%\usepackage{graphicx}
%\usepackage{url}

%\usepackage{color,soul}
%\newcommand\sr[1]{\sethlcolor{yellow}\hl{SR: #1}\sethlcolor{cyan}}
%\newcommand\hg[1]{\sethlcolor{green}\hl{HG: #1}\sethlcolor{cyan}}
%\newcommand\gdg[1]{\sethlcolor{yellow}\hl{GDG: #1}\sethlcolor{cyan}}
%\newcommand\bb[1]{\sethlcolor{cyan}\hl{BB: #1}}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{enumerate}
%\usepackage{upgreek}
%\usepackage[normalem]{ulem}

%\usepackage[small]{caption}

\usepackage{xspace}
% \usepackage{ifthen}
\usepackage{tikz}
% \usetikzlibrary{shapes.geometric}
% \pgfdeclareimage[width=1cm]{dirt}{dirt2}

% \usetikzlibrary{shapes.geometric}
% \pgfdeclareimage[width=1cm]{dirt}{dirt2}

\newcommand{\Omit}[1]{}
\newcommand{\denselist}{\itemsep -1pt\partopsep 0pt}
\newcommand{\tup}[1]{\langle #1 \rangle}
\newcommand{\pair}[1]{\langle #1 \rangle}
% \newcommand{\citeay}[1]{\citeauthor{#1} [\citeyear{#1}]}
\newcommand{\citeay}[1]{\citeauthor{#1} (\citeyear{#1})}
\newcommand{\citeaywithnote}[2]{\citeauthor{#1} (\citeyear[#2]{#1})}

\newtheorem{definition}{Definition}
\newtheorem{definitionandtheorem}[definition]{Definition and Theorem}
\newtheorem{theorem}[definition]{Theorem}
\newtheorem{lemma}[definition]{Lemma}
\newtheorem{proposition}[definition]{Proposition}
\newtheorem{corollary}[definition]{Corollary}
\newtheorem{example}[definition]{Example}

\def\withproofs{0} % to include proofs in the draft: 0=don't include, 1=include
\newcommand{\nosuchproof}{\textcolor{red}{\bf No such proof yet...}}
\newcommand{\problem}{\textcolor{red}{\bf **** PROBLEM ****}}
\newcommand{\alert}[1]{\textcolor{red}{\bf #1}}
\newcommand{\CHECK}[1]{\textcolor{red}{\bf Note: #1}}

\newcommand{\hobs}{\ensuremath{h_{obs}}\xspace}
\newcommand{\indicator}[1]{[\![#1]\!]}
\newcommand{\C}{\mathcal{C}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\dexp}{\text{2-}\!\exp}

\newcommand{\Pn}{\ensuremath{P_n}\xspace}
\newcommand{\Pnm}{\ensuremath{P_{n,m}}\xspace}
\newcommand{\muwf}{\ensuremath{\mu_{w\!f}}\xspace}

\newcommand{\true}{\mathsf{true}}
\newcommand{\false}{\mathsf{false}}
\newcommand{\Eff}{{\mathit{Eff}}}

%% C++
\newcommand{\CC}{C\nolinebreak\hspace{-.05em}\raisebox{.4ex}{\tiny\bf +}\nolinebreak\hspace{-.10em}\raisebox{.4ex}{\tiny\bf +}}
%%% \def\CC{{C\nolinebreak[4]\hspace{-.05em}\raisebox{.4ex}{\tiny\bf ++}}}
%%% \newcommand{\pplus}{\!\hspace{-.05em}\raisebox{.4ex}{\tiny\bf +}\nolinebreak\hspace{-.10em}\raisebox{.4ex}{\tiny\bf +}}
%%% \newcommand{\mminus}{\!\hspace{-.05em}\raisebox{.4ex}{\tiny\bf -}\nolinebreak\hspace{-.10em}\raisebox{.4ex}{\tiny\bf - }}
%%%\newcommand{\pplus}{\hspace{-.05em}\raisebox{.2ex}{\footnotesize\mit +}\nolinebreak\hspace{-.10em}\raisebox{.2ex}{\footnotesize\mit +}}
%%%\newcommand{\mminus}{\hspace{-.05em}\raisebox{.2ex}{\footnotesize\mit -}\nolinebreak\hspace{-.07em}\raisebox{.2ex}{\footnotesize\mit -}}

% abstract actions and plans
\newcommand{\abst}[2]{\tup{#1;#2}}
\newcommand{\Rule}[2]{\ensuremath{#1 \Rightarrow #2}}

% decrements and increments
\newcommand{\pplus}{\hspace{-.05em}\raisebox{.15ex}{\footnotesize$\uparrow$}}
\newcommand{\mminus}{\hspace{-.05em}\raisebox{.15ex}{\footnotesize$\downarrow$}}

% numerical effects
\newcommand{\dec}[1]{#1:=#1-1}
\newcommand{\inc}[1]{#1:=#1+1}

% qualitative effects
\newcommand{\Dec}{Dec}
\newcommand{\Inc}{Inc}

% STRIPS actions
\newcommand{\Stack}{\text{Stack}}
\newcommand{\Unstack}{\text{Unstack}}
\newcommand{\Pickup}{\text{Pickup}}
\newcommand{\Putdown}{\text{Putdown}}
\newcommand{\Move}{\text{Move}}

% Examples
\newcommand{\Example}{\medskip\noindent\textbf{Example.}\xspace}

% \pdfinfo{ ..}

\title{Learning features and abstract actions for computing generalized plans}

% \author{B. Bonet, G. Frances, H. Geffner} % Alphabetical
% 
\author{\textbf{Submission \#5695}}

\begin{document}

\maketitle

\begin{abstract}
  Generalized planning is concerned with the   computation of  plans that solve multiple instances.
  Recently, it has been shown that such plans can be expressed as mappings of feature values into actions, and that they
  can often be  computed by means   of fully observable non-deterministic (FOND) planners after a sequence of
  transformations. The actions in such  plans, however,   are not the actions in the  instances themselves,  which  are
  not necessarily common to  other instances,   but  abstract actions that track the effect of the  actions
  on the given set of features. The formulation assumes  that both the features and the abstract actions are given.
  In this work, we address this limitation by  showing  how to learn them.  The resulting account of generalized
  planning  combines learning and planning in a novel way: a \emph{learner},  based on a min SAT formulation, yields
  the features and abstract actions  from    state transitions sampled  from some problem instances, 
  and a FOND \emph{planner} uses this information,   suitably   transformed as before,  to produce the general plans.
  Correctness guarantees are given and 
  experimental results on several  domains are reported.
\end{abstract}

\section{Introduction}

Generalized planning studies the computation of plans for  solving  multiple  instances at once
\cite{srivastava08learning,bonet09automatic,srivastava:generalized,hu:generalized,BelleL16,anders:generalized}.
For example, the plan  that iteratively  picks  a  clear block above $x$ 
and places it  on the table, achieves the goal $clear(x)$
in \emph{any} instance of the blocks world where the gripper is initially empty.
Once this  general plan or  policy is derived it can be  applied to solve an infinite collection
of instances that involve different initial states, different objects, and hence  different (ground)  actions.

In the basic formulation due  to \citeay{hu:generalized}, a generalized plan
is a mapping of observations into actions that are assumed to be common
among  all the instances. More recently, this formulation has been extended by
\citeay{bonet:ijcai2018} to account for relational domains like Blocks world  where the set of
objects and  actions change from instance to instance.
For this,   the observations  are replaced by  a given  set of boolean
and numerical features $F$ and a set of \emph{abstract actions} $A_F$.
The  abstract actions are  \emph{sound} and \emph{complete} if they track the
effects  of the actions on  the features in a suitable way.
The resulting generalized plans  map feature values into abstract actions, and soundness
ensures that the application  of an   abstract action  results in the application of
a concrete action with a similar  effect  on  the features. Moreover, the form of the abstract actions
ensures that such plans can be computed using fully observable non-deterministic (FOND) planners
after some transformations.

\Omit{
Abstract actions can increase or decrease numerical variables  $n$
associated with the numerical features (e.g., number of blocks above $x$), and while
increments  are transformed into deterministic propositional  effects $n > 0$ where $n > 0$ is
the negation of the proposition $n=0$, decrements  are  transformed  into
non-deterministic (disjunctive)  propositional effects $n > 0 \, | \, n=0$. 
}

Bonet's and Geffner's  formulation of generalized planning, however, assumes that the
features and the abstract actions are given.   In this work, we address this limitation.
We achieving this by  \emph{showing  how features and abstract actions  can be learned from
the primitive predicates used to define the instances and state transitions
sampled from them.} For example,  in \cite{bonet:ijcai2018}, 
the general policy for achieving the goal $clear(x)$  is obtained 
using a FOND planner by introducing the boolean feature $H$ for holding a block,
the numerical feature $n(x)$ for counting the  number of blocks above $x$, and  two   abstract  actions:
one  with  preconditions  $\neg H$ and $n(x) > 0$, and effects $H$ and $n(x)\mminus$
(decrement $n(x)$), the other with precondition $H$ and effect $\neg H$.
In this work, we show how  to obtain such  policies without
having to provide the  features and  abstract actions by hand.
The abstraction is learned instead from the domain predicates
and  state transitions sampled from some  instances.

The work has relation to a number of research threads in planning, knowledge representation
and machine learning.  We make use of SAT solvers and description logics for learning the features
and abstract actions \cite{sat-handbook,dl-handbook}. The abstract actions provide a model from which the plans are obtained via
transformations and FOND planners \cite{geffner:book,ghallab:book}.
The model is not the model of an instance but a generalized model for  obtaining  plans
that work for multiple instances. In this sense, the work is different than
action model learning \cite{yang:action-learning} and model-based reinforcement learning
and closer in aims to work on learning general policies from examples or experience
\cite{martin:concept,fern:bias,mazebase,general-drl}. Generalized planning has  also been formulated as a problem in first-order logic
\cite{srivastava:generalized} and general plans over finite horizons have also been derived using  first-order regression
\cite{boutilier2001symbolic,wang2008first,van2012solving}. Our approach differs from first-order approaches in the use
efficient propositional planners, and from purely learning approaches, in the \emph{formal guarantees} that are characteristic of planning:
if the learned abstract actions are actually sound,  the resulting general plan must be correct.

The paper is organized as follows. We provide the background first and  show then  how features and
abstract actions can be  learned by enforcing soundness and completeness over a set of samples  using 
a  pool of features obtained the   domain  predicates. The full  computational model  is then
summarized,  following by the experimental results and a discussion.

\section{Background}

We review generalized planning, abstract actions, and computation of solutions 
following  \citeay{bonet:ijcai2018}.

\subsection{Generalized Planning}

A \emph{generalize planning problem} $\Q$ is a collection of planning instances $P$.
An instance $P$ is a  classical planning problems expressed in some compact language as a
tuple $P=\tup{V,I,A,G}$ where $V$ is a set of state variables that can take a finite set of values (boolean or not), $I$
is a set of atoms over $V$ defining an initial state $s_0$, $G$ is a set of atoms or literals over $V$ describing the goal states,
and $A$ is a set of actions $a$ with their preconditions and effects that define the set $A(s)$ of actions applicable
in any state $s$, and the successor state function $f(a,s)$, $a \in A(s)$. A state is a valuation over $V$, and a solution to
$P$ is an applicable action sequence $\pi=a_0,\ldots,a_n$ that generates  a state sequence $s_0,s_1,\ldots,s_{n}$
where $s_n$ is a goal state (makes $G$ true). In this sequence, $a_i \in A(s_i)$ and $s_{i+1}=f(a_i,s_i)$ for $i=0, \ldots, n-1$.
A state $s$ is reachable in $P$ if $s=s_n$ for one such sequence. A solution to the \emph{generalized problem} $\Q$ is
a solution to all instances $P \in Q$. The form of such solutions is described  below.

A \emph{feature} $f$ for a class $\Q$ of problems represents a function $\phi_f$
that takes an instance $P$ from $\Q$ and a state $s$ reachable in $P$, 
and results in a value denoted as $\phi_f(s)$. 
A feature is \emph{boolean}  if it results in a boolean value, and  \emph{numeric} if it results
in a numerical value, that we assume to be integer and non-negative. The features $H$
and $n(x)$ represent two features in the Blocks world (BW): the first a boolean
feature encoding that   the gripper is not empty; the second, a numerical
feature encoding  the number of blocks above  $x$.

Symbols like $x$ to denote \emph{parameters} whose value depends on the instance  $P$ in  $\Q$.
For example, if $\Q_{clear}$  denotes  the BW instances with goals of the form $clear(x)$,
then the value of $x$ in an instance $P$ with goal $clear(a)$ is  $a$.

\subsection{Abstract Actions}

An \emph{abstract action}  for a generalized problem $\Q$ and a set of features $F$
is a pair $\bar{a}=\abst{Pre}{\Eff}$ where $Pre$ and $\Eff$ are the  preconditions
and effects expressed in terms of a  set $V_F$ of boolean and numerical \emph{state variables}
associated with the features in $F$. Preconditions and effects over boolean state variables $p$ are literals of the form $p$
and $\neg p$, that abbreviate the  atoms $p=true$ and $p=false$, while  over  numerical state variables $n$,
preconditions are of  the form $n=0$ and $n > 0$, and effects are  of the form $n\mminus$ (decrements) and
$n\pplus$ (increments).

Features $f$ refer to \emph{state functions} $\phi_f(s)$  in the instances of the generalized problem $\Q$,
but to \emph{state variables} in the abstraction.  Likewise, an \emph{abstract state} is  a truth valuation over the
state variables  $p$ associated with boolean features and the atoms $n=0$ over  state variables  $n$ associated with numerical features.
The abstract state $\bar{s}$  that \emph{corresponds} to a concrete state $s$ is the truth valuation that makes
$p$ true iff $\phi_p(s)=true$, and $n=0$ true iff $\phi_n(s) = 0$, and an abstract action $\bar{a}$
is applicable in a state $s$ if its preconditions are satisfied in $\bar{s}$.

A set of abstract actions $A_F$ is \emph{sound}  relative to a generalized problem $\Q$ 
iff for any reachable state $s$ over an instance $P$ in $\Q$, if an abstract action $\bar{a}$ in $A_F$
is applicable in $\bar{s}$, then there is a concrete action $a$ in $P$ that is applicable in $s$
that has the \emph{same qualitative  effects as}  $\bar{a}$ over the features. The action $a$
\emph{represents} or \emph{instantiates}  the abstract action $\bar{a}$ in $s$.

Similarly, a set of abstract actions $A_F$ is \emph{complete}
iff for any reachable state $s$ over an instance $P$ in $\Q$, if a concrete action $a$ in $P$
that is applicable in $s$, then there is an abstract action $\bar{a}$ in $A_F$
that is  applicable in $\bar{s}$ that has the \emph{same qualitative effects over the features as $a$.}

The actions $a$ and $\bar{a}$ \emph{have the same qualitative  effects} in a state $s$ 
when  they have the same   effects on the boolean features $p$,  and the same \emph{qualitative effects} on the numerical
features $n$; i.e., $a$ decreases (resp. increases) the value of the function $\phi_n(s)$ in $s$ iff
$\bar{a}$ decreases (resp. increases) the value of the state variable  $n$.

\begin{example}
  For the features $F=\{H,n(x)\}$, the abstract action  $\bar{a}=\abst{\neg H,n(x)>0}{H,n(x)\mminus}$ 
  is sound in  problem $\Q_{clear}$ that includes the BW instances with stack and unstack actions.
  Indeed, if $s$ is a reachable state in an instance $P$ and $\bar{a}$ is applicable in $\bar{s}$, 
  the gripper must be empty in $s$ and some block  must be above $x$. If $A$ is the clear block above $x$ and $on(A,B)$ is true in $s$,
  then the  action $a=unstack(A,B)$  will be  applicable in $s$ and have the same qualitative effects of  $\bar{a}$
  over $F$: they both make $H$ true and decrease the value of $n(x)$. The abstract action $\bar{a'}=\abst{H}{\neg H}$ is also sound.
\end{example}

For a set of abstract actions to be complete in this example, it should 
should contain  abstracts that track the  effects of the other  actions
on the features such as stacking a block above $x$, picking up a block that is not above $x$, and so on.
It may well be possible that a particular feature set $F$ can't support a set of abstract actions
$A_F$ that is both sound and complete relative to  $\Q$. Soundness, however,  is crucial for deriving general plans as it ensures
that in any state of any instance of $\Q$,  an applicable  abstract action  can be ``instantiated'' 
by   one or more concrete actions that have the same qualitative effects on the features.

\subsection{Solutions}

A solution to a  generalized problem $\Q$ given the  features $F$ and  abstract actions $A_F$
is a partial  function $\pi$ that maps \emph{abstract states} into \emph{abstract actions} such that
$\pi$ solves all instances $P$ in $\Q$. The plan or policy $\pi$ induces a  trajectory $s_0,a_0,s_1, \ldots, s_n$ in an instance $P$ of $\Q$
iff a)~$s_0$ is the initial state in $P$, b)~$a_i$ is one of the actions that instantiate 
$\pi(\bar{s_i})$ in $s_i$, c)~$s_{i+1}=f(a_i,s_i)$ if $a_i$ is applicable in $s_i$,
and d)~$s_n$ is a goal  state of $P$,  $n=\infty$, or $\pi(\bar{s_n})$ is undefined, denotes an abstract action
that is not applicable in $\bar{s_n}$, or no applicable action $a_n$ instantiates  $\pi(\bar{s_n})$.
The policy $\pi$ solves $P$ iff all the induced trajectories reach the goal of $P$.

\begin{example}
A solution for the generalized problem $\Q_{clear}$  with feature set $F=\{H,n(x)\}$
is the policy $\pi$  given by the rules $\Rule{\neg H, n(x)>0}{\bar{a}}$ and $\Rule{H, n(x)>0}{\bar{a}'}$
for the abstract actions $\bar{a}$ and $\bar{a'}$ above. The policy picks blocks above $x$ and puts them aside
(not above $x$) until $n(x)$ becomes zero.
\end{example}

\subsection{Computation}

The steps for obtaining general policies $\pi$ for a generalized problem $\Q$
using the features $F$ and the  abstract actions $A_F$ are as follows \cite{bonet:ijcai2018}:

\medskip

\noindent 1. The state variables $V_F$ and the abstract actions $A_F$ are extended
  with initial and goal  formulas $I_F$ and $G_F$  over $V_F$ to yield a   \emph{numerical, non-deterministic planning problem}
  $Q_F=\tup{V_F,I_F,G_F,A_F}$. For soundness,  $I_F$ must be  such that the initial states of instances $P$ in $\Q$ all satisfy $I_F$,
  while all states that satisfy $G_F$ must be goal states of $P$.\footnote{
   A state $s$ is said to satisfy $I_F$ or $G_F$ if  $\bar{s}$ does.}

\medskip

\noindent 2. The problem $Q_F=\tup{V_F,I_F,G_F,A_F}$ is converted into a \emph{boolean FOND}
  problem $Q'_F=\tup{V'_F,I'_F,G'_F,A'_F}$ by replacing 
  the numerical variables $n\in N$ by the symbols $n=0$, the first-order literals $n=0$ by propositional literals  $n=0$, 
  the effects $n\pplus$ by deterministic effects $n > 0$, and the effects $n\mminus$ by non-deterministic effects $n>0\,|\,n=0$.

  \medskip

\noindent 3. The strong cyclic solutions of  $Q'_F$ assume that the
  non-deterministic effects $n>0\,|\,n=0$ are \emph{fair} \cite{strong-cyclic},  
  yet  they are  \emph{conditionally fair} only: infinite decrements of $n$ ensure $n=0$  eventually but just without
  infinite  increments. The problem of obtanining  solutions of $Q'_F$ that assume conditional fairness only, the  terminating
  solutions in   \cite{srivastava:aaai2011}, is mapped into the problem of solving an  amended FOND $Q^+_F$
  \cite{bonet:ijcai2017}.\footnote{For this, we use a translation made available by the authors that fixes bugs in the  description found in the paper
     at https://github.com/bonetblai/qnp2fond.}

\medskip

The strong cyclic solutions of  $Q^+_F$ computed by FOND planners off-the-shelf   are solutions to the generalized problem $\Q_F$.

\begin{example}
  Let restrict  $\Q_{clear}$  to those  instances where  the gripper is initially empty
  and there are blocks on top of $x$. For $F=\{H,n(x)\}$  and $A_F=\{\bar{a},\bar{a'}\}$ as above, 
   $Q_F=\tup{V_F,I_F,G_F,A_F}$ can be defined with $I_F=\{\neg H, n(x) > 0\}$ and $G_F=\{n(x)=0\}$.
   $Q'_F$ is like $Q_F$ but with $n(x)=0$ regarded as a propositional symbol, $n(x) > 0$
  as its negation, and the effect $n(x)\mminus$ of  $\bar{a}$ replaced by  $\neg n(x)=0 \, | \, n(x)=0$.
  Since  no action in $A_F$  increases $n(x)$,   $Q^+_F=Q'_F$, and the strong cyclic solution  of $Q'_F$,
  that corresponds to the policy above, solves $\Q_{clear}$.
\end{example}

\section{Approximate Soundness and Completeness}

Bonet's and Geffner's formulation and computational model rest on 
features $F$ and  sound abstract actions $A_F$ provided by hand.
The contribution of this work is to learn them both from samples:

\begin{definition}
 For a generalized problem $\Q$, a \emph{sample set}  $\cal S$ is a non-empty set of tuples  $(s,s',P)$
 such that 1)~$P$ is an instance of $\Q$, 2)~$s$ is a reachable state in $P$, 3)~$s'$
 is a possible succesor of $s$ in $P$; i.e., $s'=f(a,s)$ for some action $a$ applicable in $s$,
 4)~$\cal S$ is closed in the sense  that if $(s,s',P)$ is in $\cal S$, then $(s,s'',P)$ is in $\cal S$  if $s''$ is another possible succesor of $s$ in $P$.
\end{definition}

By assuming that sampled states $s$ are tagged with the instance $P$, 
we can abbreviate the tuples $(s,s',P)$ as $(s,s')$, and refer to
the states $s'$ in  the pairs $(s,s') \in {\cal S}$ as the successors
of $s$ in ${\cal S}$. Notice that the closure condition 4) ensures that
states $s$ appearing \emph{first} in pairs $(s,s')$ must be fully expanded in ${\cal S}$;
namely, all possible transitions $(s,s'')$ must be in the sample as well. We call them
the \emph{expanded} states in ${\cal S}$.

The features and abstract actions for $\Q$ are  learned by
enforcing an approximate form of soundness and completeness
over the samples. Like before,  we take a transition $(s,s')$ and an abstract action $\bar{a}$
to have  the same qualitative effects over a boolean  feature  iff they both make it true , they both make it false,  or
they both leave it unchanged,  and over a numerical feature  iff they both decrease it,  they both increase it,  or they both leave it unchanged.

\begin{definition}
  A set of abstract actions $A_F$ is sound \emph{relative to a  sample set} $\cal S$ for $\Q$
  iff for  any  abstract action $\bar{a}$ in $A_F$ that is applicable in  a expanded state $s$ of ${\cal S}$,
  there is a transition $(s,s')$ in $\cal S$ that has the same qualitative  effects over $F$ as $\bar{a}$.
\end{definition}

\begin{definition}
A set of abstract actions $A_F$ is complete   \emph{relative to a sample set} $\cal S$ for $\Q$
iff for each transition  $(s,s')$ in  $\cal S$, there is an abstract action $\bar{a}$ in $A_F$
with the same qualitative effects over in $F$ that is applicable in $s$.
\end{definition}

For a sufficiently large sample set, these approximate notions of soundness and completeness
converge to the exact notions.

\section{Learning  features and  abstract actions}

For computing the features and abstract actions
we define  a  formula  $T({\cal S,F})$ that  is \emph{satisfiable} 
iff there is a set of features $F  \subseteq {\cal F}$ and a set of abstract actions $A_F$ over $F$
that is sound and complete relative to the sample set ${\cal S}$.

\Omit{The  sets  $F$ and $A_F$ can
     then be read off  from any  satisfying assignment. Later we will add weights
     to so
     me of the clauses to establish a preference over these assignments
     (e.g., for  minimizing $|F|$).}

\subsection{SAT Encoding}

For each transition  $(s, s') \in \mathcal{S}$ and each feature $f \in \mathcal{F}$,
$\Delta_f(s, s') \in \{+, -, \uparrow, \downarrow, \bot\}$ will denote the \emph{qualitative change of value}
of feature $f$ along the transition $(s, s')$, which can go from false to true ($+$), from true to false ($-$) or remain unchanged ($\bot$),
for boolean features, and can increase ($\uparrow$), decrease ($\downarrow$), or remain unchanged ($\bot$), for numerical features.
The \emph{propositional variables}  in $T({\cal S,F})$ are then:

  \begin{itemize}
  \item  $selected(f)$ for each $f \in {\cal F}$,  true iff  $f$  selected  (in  $F$).
  \item $D_1(s,t)$ for  states   $s$ and $t$ expanded in ${\cal S}$,  true iff selected features distinguish  $s$ from $t$;
      i.e.  if $s$ and $t$ disagree on the truth value of a selected boolean feature $p$ or atom $n=0$ for a selected numerical feature $n$.
      % Since $D_1$ is symmetric, only symbols for  $s < t$ are created   for  some  arbitrary, fixed order over the states in ${\cal S}_1$.
  \item $D_2(s, s', t, t')$ for  each $(s, s')$ and $(t, t')$ in $\mathcal{S}$,
    true iff selected features $f$ distinguish the two transitions; i.e., if $\Delta_f(s, s')\not=\Delta_f(t,t')$. 
%      As for $D_1$, the number of $D_2$ variables is cut in half due to symmetries.
  \end{itemize}

\medskip

\noindent The  first  formulas in $T({\cal S,\cal F})$ capture the meaning of $D_1$:

\begin{equation}
  D_1(s, t) \Leftrightarrow \bigvee_{f} \, selected(f)
  \label{d1}
\end{equation}

\noindent where $f$ ranges over features in $\mathcal{F}$ with  different qualitative values in $s$ and $t$; i.e.
boolean features $p$ with different truth value in $s$ and $t$, and numerical features $n$ with different truth values
for atom $n=0$ in $s$ and $t$.

\noindent The second formulas encode  the meaning of $D_2$:

\begin{equation}
  D_2(s, s', t, t') \Leftrightarrow \bigvee_f \,  selected(f)
\end{equation}

\noindent where $f$  ranges over features $f \in \mathcal{F}$ that have the same qualitative values in $s$ and $t$
but which change differently in the two transitions; i.e.,  $\Delta_f(s, s') \neq \Delta_f(t, t')$.

The third  class of formulas relate $D_1$ and $D_2$ by  enforcing  \emph{soundness} and \emph{completeness} over  the samples in ${\cal S}$.
Yet, since the abstract actions  are not given but are derived, the qualitative changes in each  transition $(s,s')$ in ${\cal S}$
are  taken as templates  of abstract actions. Thus, if $s$ and $t$ are not distinguished by the selected features, 
for each transition $(s, s') \in \mathcal{S}$, there must be a transition $(t, t') \in \mathcal{S}$
such that that the two transitions are not distinguished by the selected features either.
This is expressed as

\begin{equation}
  \neg D_1(s, t) \, \Rightarrow \, \bigvee_{t'} \neg D_2(s, s', t, t')
  \label{bridge1}
\end{equation}

\noindent where $s$ and $t$ are expanded states in  ${\cal S}$, $s'$ is a successor of $s$ in ${\cal S}$,
and $t'$ ranges over the successors of $t$ in ${\cal S}$.

%%% --
\Omit{ % left out: this is for efficiency only, and is not strictly needed. Also, doesn't fit as such with goal-marking
  Refer to this as an ``efficiency optimization'' ..

  such that $s < t$, and transition $(s,s')$ in ${\cal S}$. Due to the symmetry breaking caused by the constraint $s < t$,
  the following formulas are also needed

\begin{equation}
  \neg D_1(s, t) \, \Rightarrow \, \bigvee_{s'} \neg D_2(s, s', t, t')
    \label{bridge2}
\end{equation}

\noindent where the  iteration in the right hand disjunction goes over the different successor states $s'$ of the first argument
$s$ of $D_1$ that are in the sample set ${\cal S}$, $s$ and $t$ are states in ${\cal S}_1$ as before with $s < t$, and $t'$ is a
successor state of $t$ in ${\cal S}$; i.e. $(t,t') \in {\cal S}$.
}
%% --

The last  set of clauses force the selected features to distinguish goal from non-goal states as:

\begin{equation}
  D_1(s,t) 
\label{goal}
\end{equation}

\noindent where  $s$ and $t$  are expanded states  ${\cal S}$ such that exactly one of them is a goal state. 
For this, it is assumed that  states $s$ in   tuples $(s,s',P)$ are labeled as   goal states  of $P$ or non-goals.

The theory $T({\cal S},{\cal F})$ given by the formulas  \ref{d1})--(\ref{goal}
has a number of atoms that is linear in the size of the feature pool $\cal F$,
and quadratic in the number of expanded  states $s$  and transitions $(s,s')$ in ${\cal S}$.
The resulting number of clauses is ....  , which can be roughly cut in half by  symmetry considerations.

In the experiments we will   consider also  a smaller theory  that avoids the consideration of all pairs
of expanded states or all pairs of transitions in $\cal S$]. For this, some expanded states $s$ and transitions $(s,s')$ in
tuples $(s,s',P)$ are marked as \emph{goal relevant}. This is achieved by computing some plans for some instances $P$ in $\Q$, 
and by adding the resulting state transitions to ${\cal S}$ at the same time that the first
arguments $s$ and $(s,s')$ in the  $D_1$ and $D_2$ atoms are restricted to be marked.
If the set of marked states and transitions is much smaller than the set of unmarked ones, 
then the resulting formula,  $T_G({\cal S},{\cal F})$ will be much smaller than $T({\cal S},{\cal F})$.
The goal marking scheme that involves the theory $T_G({\cal S},{\cal F})$ in place of $T({\cal S},{\cal F})$
amounts to enforcing soundness on  the sample set ${\cal S}$ but not completeness, yet this does
not affect the resulting  formal guarantees that depend on soundness only.

\subsection{Extracting $F$ and  $A_F$}

If  $\sigma$ is  a satisfying assignment of the theory $T({\cal S},{\cal F})$, 
we define $F_{\sigma}$ as the set of  features $f$ in $\cal F$ such that $selected(f)$ is true in $\sigma$,
and $A_{\sigma}$ as the set of abstract actions that capture  the  transitions $(s,s')$ in ${\cal S}$
with duplicates and redundancies removed. The abstract action $\bar{a}=\abst{Pre}{\Eff}$
that captures the transition $(s,s')$ has the \emph{precondition} $p$ (resp. $\neg p$)
if $p$ is a boolean feature in $F_{\sigma}$ that is true (resp. false) in $s$, and has
the precondition $n=0$ (resp. $n > 0$) if $n$ is a numerical feature in $F_{\sigma}$
such that $n=0$ is true (resp. false) in $s$. Similarly, $\bar{a}$ has the \emph{effect}
$p$ (resp. $\neg p$) if $p$ is a boolean feature in $F_{\sigma}$ that is true (resp. false)
in $s'$ but false (resp. true) in $s$, and the effect $n\mminus$ (resp. $n\pplus$) if $n$ is a numerical feature in $F_{\sigma}$
whose values decreases (resp. increases) in the transition from $s$ to $s'$. 
If two abstract actions $\bar{a}$ and $\bar{a}$ differ only in the sign of a precondition,
the precondition can be dropped, removing the duplicates and iterating until  no more simplifications
can be done. They key properties are:

\begin{theorem}
  The theory $T({\cal S},{\cal F})$ is satisfiable iff there is a set of features $F \subseteq {\cal F}$
  and a set of abstract actions $A_F$ over $F$ such that $A_F$ is sound and complete relative to  ${\cal S}$.
\end{theorem}

\begin{theorem}
  If $\sigma$ is a satisfying assigment of $T({\cal S},{\cal F})$, then the abstraction actions in $A_{\sigma}$ 
  are  sound and complete  relative to  ${\cal S}$.
\end{theorem}

\begin{theorem}
  If $\sigma$ satisfies  $T_G({\cal S},{\cal F})$, then $A_{\sigma}$   are  sound  relative to  ${\cal S}$.
\end{theorem}

\Omit{
%% commented out
The goal expression $G_F$ can be provided explicitly  in terms of the primitive predicates,
but  more generally can be extracted  from the satisfying assigment $\sigma$ as well.
For this, $G_F = G_{\sigma}$ is defined as the DNF formula whose terms 
correspond to the abstract states over the selected features in $F_{\sigma}$ 
that are false in all expanded states in ${\cal S}$ marked as non-goals
and true in some expanded goal state. This DNF formula can   be simplified in standard ways. 
Due to  (\ref{d1}) and (\ref{goal}) that force the selected features in $F_{\sigma}$
to   distinguish goal  from non-goal states, we have that:

\begin{theorem}
For a satisfying assigment $\sigma$  of $T({\cal S},{\cal F})$, 
$s$ is expanded goal   state in ${\cal S}$  iff  $s$ satisfies  $G_F=G_{\sigma}$.
\end{theorem}
}

Different  satisfying  assignments $\sigma$,  give rise to different  features $F_{\sigma}$
and abstract actions $A_{\sigma}$. More meaningful features are found by looking for
satisfying assignments $\sigma$  that minimizes a cost measure such as $|F_{\sigma}|$
or  more generally $\sum_{f \in F_{\sigma}} cost(f)$. We will  achieve this  constrained
optimization through the use  Min SAT or Max-Weighted SAT solvers,  where it suffices to give
infinite weights  to the clauses resulting from $T({\cal S},{\cal F})$ and weights given by $cost(f)$ 
to the ``soft''  unit clauses  $\neg selected(f)$.

\section{Feature Pool}

The computation of the feature pool $\cal F$ used in $T({\cal S},{\cal F})$
relies on the predicates used to encode the instances in $\Q$.
A simple grammar   allow  us   to construct new predicates from them, 
and the boolean and numerical features in the pool
are defined from the resulting \emph{unary predicates}. 
The key issue is how to define new predicates from old  ones. For this 
we take a solution off-the-shelf: \emph{description logics (DL)},  called  also
\emph{concept languages},  are grammars for doing exactly this.
Concepts $C$  refer  to predicates of arity 1 and roles $R$  to predicates of arity 2.
Concept languages have been used before for  generating general policies using purely learning methods  \cite{martin:concept,fern:bias}.
For details on the syntax and semantics of description logics; see \cite{dl-handbook}.

\subsection{Grammar}

The DL grammar is given by the rules:

  \begin{itemize}
  \item $C \ \leftarrow\ \ C_p, C_u, C_x$, primitive, universal,  nominal c's,\footnote{Denotation of $C_x$ for parameter $x$ is $\{x\}$.}
  \item $C \ \leftarrow\  \neg C, C \sqcap C'$, negation and conjunction, 
  \item $C \ \leftarrow\  \exists R.C , \forall R.C$, concepts that use of  roles: first denotes $x$ s.t. for some  $y$,  $R(x,y) \land C(y)$.
  Second,  $x$ s.t. for all $y$, $R(x,y) \rightarrow C(y)$,
  \item $C \ \leftarrow\  R=R'$, stands for $x$ s.t. set of $y$'s  s.t. $R(x,y)$ equal to set of $y$'s s.t. $R'(x,y)$,
    \item $R \leftarrow R_p, R^{-1}, R^*$ ; primitive, inverses, transitive.
%     \item $R \leftarrow R \circ R'$, role composition: $(x,y)$ in $R \circ R'$ if $(x,z)$ in $R$ and $(z,y)$ in $R'$.
%     \item $R \leftarrow R:C$, denotes pairs $(x,y)$ in $R$ s.t. $C(y)$.
 \end{itemize}

The  compositional  semantics   of these rules allows us to determine the denotations (extensions)   $C^s$ and $R^s$  of
the resulting unary and binary predicates from the denotation of the primitive predicates in any  state $s$.

** Check what is used here and not; see what needs to be explained; keep concise, improve lay out.

\subsection{Pool}

The set of  concepts and roles  with complexity bounded by $k$ is denoted as ${\cal G}(k)$
with their  complexity given by the    minimum number of  rules for generating them.
The size  of this set is reduced by incrementally pruning ``duplicate'' concepts,
those whose denotation coincides with the denotation of a another concept over all
the sampled states in ${\cal S}$.

The pool of features ${\cal F}={\cal F}^k$  is defined from   ${\cal G}^k$
and contains the  features $f$ that represent the functions $\phi_f(s)$
as follows:

\begin{itemize} 
\item Numerical   $n_C$ for each $C$  that  stand for cardinality $|C^s|$,
\item Boolean  $b_C$ for each   $C$ that are  true in $s$ iff $|C^s| > 0$,
\item Boolean   $b_p$ for each primitive predicate $p$ of arity $0$,
\item Numerical $\textit{dist}(C_1,R:C,C_2)$,  that represent that smallest $n$ such that for $x_1, \ldots, x_n$, 
$C_1^s(x_1)$, $C_2^s(x_{n})$, and $R:C(x_i,x_{i+1})$ for $i=1, \ldots, n$. 
\end{itemize}

The role $R:C$ is such that its denotation contains the pairs $(x,y)$ in $R$ such that $y$ is in $C$.
Some pruning is done in this space of features as well. E.g., numerical features $n_C$ are pruned from the pool
when there are  no states $s$ and $s'$ in the sample set for which $n_c(s)=0$ and $n_c(s') > 1$,
while boolean features $n_b$  are pruned if uniformly true or false for all $s$, or if $n_C$ not pruned. 
The measure  $cost(f)$ is set to the complexity of $C$ for $n_C$ and $b_C$, to $0$ for $b_p$
and to the complexities of $C_1$, $R$, $C$, and $C_2$ added up for $\textit{dist}(C_1,R:C,C_2)$.
Only features with cost bounded by $k$ are allowed in $\cal F$.

\section{Computational Model: Summary}

In summary, the steps for computing general plans are:

\begin{enumerate}
\item a sample set ${\cal S}$ is computed from a few  instances $P$ of  a given generalized problem $\Q$ 
\item a pool of features $\cal F$ is obtained from the predicates used in the instances $P$, the  grammar, and bound $k$,
\item a satisfying assignment $\sigma$ of the theory $T({\cal S},{\cal F})$ or $T_G({\cal S},{\cal F})$
that minimizes  $\sum_{f \in F_{\sigma}} cost(f)$ is obtained using a Weighted-MAX SAT solver,
\item the features $F$ and abstract actions $A_F$  are extracted from $\sigma$ as $F_{\sigma}$ and $A_{\sigma}$,
\item the problem $Q_F=\tup{V_F,I_F,G_F,A_F}$ is  defined with initial and goal conditions $I_F$ and $G_F$ provided by hand
to match $\Q$, 
\item the FONDP  $Q^+_F$, constructed from $Q_F$,  is solved by a FOND planner off-the-shelf
\end{enumerate}

Note that the  policy $\pi$ that results from the last step  deals with propositional symbols
that correspond to atoms $p$ and $n=0$ for boolean and numerical features $p$ and $n$ in $F$. 
When this policy is applied to an instance $P$ of $\Q$, however, such features denote state functions
which are determined by the  semantics of the concepts and roles used in  the features.
(e.g., $n_C$ stands for feature function $\phi_{n_c}(s)=|C^s|$).
The exception are the boolean features $b_p$ associated with
predicates $p$ of zero arity which are treated as fluents.

If we write $A \models B$ to express that all the states that satisfy $A$, satisfy $B$, 
then the formal guarantees can be expressed as:

\begin{theorem}
If the abstract actions $A_F$ that are sound relative to the sample set $\cal S$  are actually sound,
then a   policy $\pi$ that solves  $Q^+_F$  is guaranteed to solve  all instances $P$ of $\Q$
with initial and goal situations $I$ and $G$ s.t. $I \models  I_F$ and $G_F \models G$.
\end{theorem}

This means that a  policy $\pi$ obtained with a FOND planner from $Q^+_F$ 
solves solves the  generalized problem $\Q$, provided that the  abstract
actions learned  are  sound and  that $\Q$ is restrictedto the instances
$P$ whose initial and goal conditions comply with the conditions on $I_F$ and $G_F$.
Since we don't have a language for specifying $\Q$ formally, we can take the formulas
$I_F$ and $G_F$ indeed as a partial formal specification of $\Q$. The theorem then says that
if soundness over the samples, ensure formal soundness over $\Q$, the policy that solves the
FONDP $Q^+_F$, solves $\Q$. The result holds whether $T$ or $T_G$ (goal marking) is used in step 3.
In principle, we could learn the formulas $I_F$ and $G_F$ from the sample as well,  but with different guarantees.

\section{Experimental Results}

We evaluate the computational model given by steps 1--6 above over several domains. *Fill up general  details in 1--6
that don't change from problem to problem: how ${\cal S}$ obtained, value of $k$, SAT solver used, FOND planners, translator, etc.
The goal marking scheme that  yields the smaller  theory $T_G({\cal S},{\cal F})$ is obtained an optimal plan over
a few (always less than xxxx) instances $P$.

Experiments setting: cpu, cutoffs, etc.

For each problem: 1--6 details that vary from problem to problem. Follow same structure:
$\Q$, $P$s, primitive predicates, number of atoms and clauses  of $T$ and/or $T^G$,
SAT times,  $F$, $|A_F|$;  number of FOND atoms and actions in $Q^+_F$ (include those in $F$, $A_F$ plus translation), $I_F$ and $G_F$,
FOND time, resulting policy.  Numbers can go all in a table.

** Mention: symbold DRL paper by Shanahan in

\section{Discussion}

We have introduced a scheme for computing general plans that mixes learning and planning: a learner infers
a general abstraction made up of features and abstract actions by enforcing  a notion of soundness and completeness
over a suitable set of samples, which is transformed and fed into a FOND planner.
The number of samples required for obtaining correct general plans is very small,
as the learner doesn't have to learn how to plan,  but just to identify the
general features that are relevant for the planner. Moreover, unlike, purely
learning approaches, the scope and correctness of the resulting general plans
can be assessed.  We have carried out a number of experiments over several
domains illustrating the versatility of the approach. In the future, we
want to explore the use of the learned  features and actions
in classical planning, the study of  formal characterizations
of the generalized problem $\Q$ and  the possibilities
of testing soudness over $\Q$ automatically, and ways
for improving the scalability of the approach, for example,
by collecting samples intelligently and incrementally,
from the states where the computed  policy fails,
and by reusing the features and actions obtained
for some goal for more complex goals.

\Omit{

** discussion  of experiments; things we do well and not, scope, lessons. limitations (expressive? scalability?)

** briefly related work, expanding paragraph in intro.

** summary and cconlusions
}

\bibliographystyle{aaai}
\bibliography{control}

\end{document}

### 2018-09-04 experiments/blocks.py
prepare some cluster experiments

@@ -39,8 +39,11 @@ def experiment(experiment_name=None):
     # card[And(And(Forall(Star(on),Not({a})), Forall(Star(Inverse(on)),Not({a}))), And(Not(holding), Not({a})))] 18
     # And indeed learnt the correct state space!!!
     aaai_ijcai_features_on_clear_5_rnd = dict(
-        instances="instance_5_clear_x_1.pddl",
-        num_states=2000, num_sampled_states=40, random_seed=12,
+        instances=[
+            "inst_clear_x_1.pddl",
+            # "inst_clear_x_2.pddl",
+        ],
+        num_states=1000, num_sampled_states=80, random_seed=10,
         max_concept_size=18, max_concept_grammar_iterations=3,
         concept_generator=None, parameter_generator=add_bw_domain_parameters,
         feature_namer=ijcai_paper_bw_feature_namer,)
@@ -73,7 +76,6 @@ def experiment(experiment_name=None):
                                             num_states=2000, max_width=[2],
                                             num_sampled_states=100,
                                             )
-
 
     #
     ijcai_features_on_clear_5 = update_dict(aaai_ijcai_features_on_clear_5_rnd, num_sampled_states=None)

### 2018-09-05 experiments/blocks.py
Make pipeline more uniform accross different experiments and print some extra info

@@ -60,8 +60,8 @@ def experiment(experiment_name=None):
         instances="instance_5_clear_x_1.pddl",
         num_states=2000, max_width=[-1],
         num_sampled_states=300,
-        complete_only_wrt_optimal=True,
-        max_concept_size=18, max_concept_grammar_iterations=3,
+        # complete_only_wrt_optimal=True,
+        max_concept_size=8, max_concept_grammar_iterations=3,
         concept_generator=None, parameter_generator=add_bw_domain_parameters,
         feature_namer=ijcai_paper_bw_feature_namer,
     )
@@ -125,7 +125,7 @@ def experiment(experiment_name=None):
                                                   sampling="all",
                                                   # enforce_features=get_on_x_y_feature,
                                                   # feature_generator=features_clear_x,
-                                                  max_concept_size=10, max_concept_grammar_iterations=3,
+                                                  max_concept_size=8, max_concept_grammar_iterations=3,
                                                   concept_generator=None, parameter_generator=add_bw_domain_parameters_2,
                                                   feature_namer=ijcai_paper_bw_feature_namer,
                                                   )

### 2018-09-05 experiments/gripper.py

@@ -38,7 +38,7 @@ def experiment(experiment_name=None):
         num_states=2000, max_width=[-1],
         num_sampled_states=100,
         complete_only_wrt_optimal=True,
-        max_concept_size=10, max_concept_grammar_iterations=3,
+        max_concept_size=8, max_concept_grammar_iterations=3,
         concept_generator=None, parameter_generator=add_domain_parameters,
         feature_namer=feature_namer,)
 

### 2018-09-05 sampling.py

@@ -102,7 +102,8 @@ def sample_generated_states(config, rng):
         remap_sample_expanded_states(set(selected), states, goal_states, transitions, optimal_transitions, set(root_states))
 
     expanded = lambda s: len(transitions[s]) > 0
-    logging.info("Selected (states / goals / optimal tx): {} / {} / {}".format(len(states), len(goals), len(optimal_transitions)))
+    total_tx = sum(len(transitions[sid]) for sid, _ in states.values())
+    logging.info("Resampling (states / goals / tx / optimal tx): {} / {} / {} / {}".format(len(states), len(goals), total_tx, len(optimal_transitions)))
     log_sampled_states(states, goals, transitions, expanded, optimal_transitions, root_states, config.resampled_states_filename)
     return states, goals, transitions, optimal_transitions, root_states
 

### 2018-09-05 doc/paper.tex
Latex table draft

@@
 \setlength{\pdfpageheight}{11in}  %Required
-%PDF Info Is Required:
-  \pdfinfo{
-/Title ()
-/Author ()}
+
 \setcounter{secnumdepth}{0}
@@
 %\usepackage[normalem]{ulem}
-
-%\usepackage[small]{caption}
+\usepackage{multirow}
+% \usepackage[small]{caption}
 
@@
 \newcommand{\pair}[1]{\langle #1 \rangle}
+\newcommand{\tuple}[1]{\ensuremath{\langle #1 \rangle}}
+\newcommand{\set}[1]{\ensuremath{\left\{#1 \right\}}}
+\newcommand{\setst}[2]{\ensuremath{\left\{#1 \mid #2 \right\}}}
+\newcommand{\abs}[1]{\ensuremath{\left\vert{#1}\right\vert}}
 % \newcommand{\citeay}[1]{\citeauthor{#1} [\citeyear{#1}]}
@@
 
-% \pdfinfo{ ..}
+%PDF Info Is Required (no accents, etc.!):
+\pdfinfo{
+/Title ()
+/Author ()}
 
@@
 
-Bonet's and Geffner's  formulation of generalized planning, however, assumes that the
+Bonet and Geffner's  formulation of generalized planning, however, assumes that the
 features and the abstract actions are given.   In this work, we address this limitation.
@@
 
-Bonet's and Geffner's formulation and computational model rest on 
+Bonet and Geffner's formulation and computational model rest on 
 features $F$ and  sound abstract actions $A_F$ provided by hand.
@@
 
-The theory $T({\cal S},{\cal F})$ given by the formulas  \ref{d1})--(\ref{goal}
+The theory $T({\cal S},{\cal F})$ given by the formulas  (\ref{d1})--(\ref{goal})
 has a number of atoms that is linear in the size of the feature pool $\cal F$,
@@
 relies on the predicates used to encode the instances in $\Q$.
-A simple grammar   allow  us   to construct new predicates from them, 
+A simple grammar allows  us   to construct new predicates from them, 
 and the boolean and numerical features in the pool
@@
   \begin{itemize}
-  \item $C \ \leftarrow\ \ C_p, C_u, C_x$, primitive, universal,  nominal c's,\footnote{Denotation of $C_x$ for parameter $x$ is $\{x\}$.}
-  \item $C \ \leftarrow\  \neg C, C \sqcap C'$, negation and conjunction, 
-  \item $C \ \leftarrow\  \exists R.C , \forall R.C$, concepts that use of  roles: first denotes $x$ s.t. for some  $y$,  $R(x,y) \land C(y)$.
+  \item $C \ \leftarrow\ \ C_p, C_u, C_x$: primitive, universal,  nominal c's,\footnote{Denotation of $C_x$ for parameter $x$ is $\{x\}$.}
+  \item $C \ \leftarrow\  \neg C, C \sqcap C'$: negation and conjunction, 
+  \item $C \ \leftarrow\  \exists R.C , \forall R.C$: concepts that use of  roles: first denotes $x$ s.t. for some  $y$,  $R(x,y) \land C(y)$.
   Second,  $x$ s.t. for all $y$, $R(x,y) \rightarrow C(y)$,
-  \item $C \ \leftarrow\  R=R'$, stands for $x$ s.t. set of $y$'s  s.t. $R(x,y)$ equal to set of $y$'s s.t. $R'(x,y)$,
-    \item $R \leftarrow R_p, R^{-1}, R^*$ ; primitive, inverses, transitive.
+  \item $C \ \leftarrow\  R=R'$: stands for $x$ s.t. set of $y$'s  s.t. $R(x,y)$ equal to set of $y$'s s.t. $R'(x,y)$,
+    \item $R \leftarrow R_p, R^{-1}, R^*$: primitive, inverse, transitive closure.
 %     \item $R \leftarrow R \circ R'$, role composition: $(x,y)$ in $R \circ R'$ if $(x,z)$ in $R$ and $(z,y)$ in $R'$.
@@
 
-The set of  concepts and roles  with complexity bounded by $k$ is denoted as ${\cal G}(k)$
+The set of all concepts and roles with complexity no larger than $k$ is denoted as ${\cal G}^k$,
 with their complexity given by the minimum number of rules for generating them.
@@
 
-We evaluate the computational model given by steps 1--6 above over several domains. *Fill up general  details in 1--6
-that don't change from problem to problem: how ${\cal S}$ obtained, value of $k$, SAT solver used, FOND planners, translator, etc.
+We evaluate the computational model described in steps 1--6 above over several domains.
+%
+
+% *Fill up general details in 1--6 that don't change from problem to problem
+%how ${\cal S}$ obtained
+The sample set $\mathcal{S}$ is obtained following one of two possible strategies.
+In the \emph{exhaustive} sampling strategy,
+$\mathcal{S}$ contains, for every one of the training instances $P$,
+all the states generated on a breadth-first search from the initial state in $P$
+that is run until $m$ expansions have been performed. If that set includes no goal state,
+an optimal plan for $P$ is computed and all of the states along the plan are included in $\mathcal{S}$ as well,
+as we need at least one goal state in the sample set.
+%
+In the \emph{random} sampling strategy, a large number $M$ of states are generated, potentially exhausting the full state space,
+and then a certain number $m <\!< M$ of them is sampled uniformly at random. If no goal state is found among those $m$,
+one random goal state from the larger set is drawn at random and added into the sample.
+
+The feature complexity bound $k$ in the feature pool $\mathcal{F}$ is set to a value of $k=10$, unless no solution is found
+to the theory $T({\cal S},{\cal F})$, in which case it is increased to $k=20$. 
+% We have considered strategies to increase the bound adaptively
+The resulting theory $T({\cal S},{\cal F})$ is solved with the Open-WBO MAXSAT solver \cite{martins2014open}.
+
+** Details on the FOND planner, QNP translation, etc. **
+
+The whole computational pipeline made up of steps 1--6 is processed on Intel Xeon E5-2660 CPUs,
+with time and memory cutoffs of 1h and 32GB.
+
+The number of instances $P$ is never larger than $3$, and they are selected XXX THIS NEEDS TO BE WELL JUSTIFIED...
+
 The goal marking scheme that  yields the smaller  theory $T_G({\cal S},{\cal F})$ is obtained an optimal plan over
-a few (always less than xxxx) instances $P$.
+a few (always less than xxxx) instances $P$. ??
 
-Experiments setting: cpu, cutoffs, etc.
+\begin{table*}[t]
+\centering
+% \addtolength{\tabcolsep}{-1pt}
+{
+% \scriptsize
+% \footnotesize
+\small
+\def\arraystretch{1}%  1 is the default
 
-For each problem: 1--6 details that vary from problem to problem. Follow same structure:
-$\Q$, $P$s, primitive predicates, number of atoms and clauses  of $T$ and/or $T^G$,
-SAT times,  $F$, $|A_F|$;  number of FOND atoms and actions in $Q^+_F$ (include those in $F$, $A_F$ plus translation), $I_F$ and $G_F$,
-FOND time, resulting policy.  Numbers can go all in a table.
+\newcommand{\rowspacing[1]}{&&&&&&&&&&&&&&&\\[\dimexpr-\normalbaselineskip+2pt]}
+
+\begin{tabular}{%
+@{}l%
+@{\extracolsep{1.5em}}r%@{}l%
+@{\extracolsep{1.5em}}r%
+@{\extracolsep{8pt}}r%
+%
+@{\extracolsep{1.5em}}r%
+@{\extracolsep{8pt}}r%
+@{\extracolsep{8pt}}r%
+@{\extracolsep{8pt}}r%
+%
+@{\extracolsep{1.5em}}r%
+@{\extracolsep{8pt}}r%
+@{\extracolsep{8pt}}r%
+@{\extracolsep{8pt}}r%
+%
+@{\extracolsep{1.5em}}r%
+@{\extracolsep{8pt}}r%
+@{\extracolsep{8pt}}r%
+@{\extracolsep{8pt}}r%
+%
+@{}}
+
+% @{\extracolsep{3pt}}
+
+ & $I$ & $\abs{\mathcal{S}}$ & $k$ & $\abs{\mathcal{F}^k}$ & \multicolumn{2}{c}{$T$} & \multicolumn{2}{c}{$T_G$} &  $C_{\text{SAT}}$ & $\abs{F}$ & $\abs{A_F}$ & $\abs{I_F}$  & $\abs{G_F}$ & $\abs{Q^+_F}$ & $C_{\text{FOND}}$ \\ %
+
+\rowspacing 
+\cline{6-7} \cline{8-9}%
+\rowspacing
+
+ & &     &      &         & $\#V$ & $\#C$           & $\#V$ & $\#C$             &     &&&&&& \\%
+
+\rowspacing \hline \rowspacing
+\multirow{1}{*}{$\Q_{clear}$}
+& 1     &     927 & 8  & 322 & 535K & 59.6M & 7.7K & 767K   &  0.01 & 3 & 2   &  - & - & - & - \\% Total complexity of F: 8
+
+\rowspacing \hline \rowspacing
+\multirow{1}{*}{$\Q_{on\text{-}2\text{-}tow}$}   
+& 3 &        420  & 8 & 657 & -K & -K & 18.3K &     3.3M    &  0.02&  5 & 7   &  - & - & - & - \\% Total complexity of F: 18
+
+\rowspacing \hline \rowspacing
+\multirow{1}{*}{$\Q_{gripper}$}    
+& 2 &         403 & 8 & 130 & 93K & 4.7M & 8.1K &   358K    &  0.01 & 4 & 5&  - & - & - & - \\% Total complexity of F: 10
+
+\rowspacing \hline \rowspacing
+\multirow{1}{*}{$\Q_{reward}$}
+& 2 &         568 & 8 & 280 & 184K & 11.9M & 15.9K & 1.2M   &  0.01 & 2 & 2 &  - & - & - & - \\% Total complexity of F: ?
+\end{tabular}
+
+}
+\caption{{
+  Experimental Results on 4 different generalized planning problems. 
+  $I$: number of training instances; 
+  $\abs{\mathcal{S}}$: sample set size, in total number of transitions; 
+  $k$: max. feature comlexity; $\abs{\mathcal{F}^k}$: size of the pool of features;
+  $\#V$, $\#C$: Number of variables and clauses of the theories $T({\cal S},{\cal F})$ and $T_G({\cal S},{\cal F})$;
+  $C_{\text{SAT}}$, $C_{\text{FOND}}$: computation time (sec.) of the SAT and FOND solvers, resp.;
+  $\abs{F}$: number of selected features;
+  $\abs{A_F}$: number of abstract actions;
+  $\abs{I_F}$: ???;
+  $\abs{G_F}$: ??? ;
+  $\abs{Q^+_F}$: ??? .
+  All figures that depend on the SAT theory are given for $T_G({\cal S},{\cal F})$.
+}}
+\label{tab:exp-result-data}
+\end{table*}
+% 2018-09-05 00:54:08 INFO     Selected (states / goals / optimal tx): 220 / 34 / 20
+
+Table~\ref{tab:exp-result-data} summarizes the relevant data for each of the test benchmarks.
+
+% For each problem: 1--6 details that vary from problem to problem. Follow same structure:
+% $\Q$, $P$s, primitive predicates, number of atoms and clauses  of $T$ and/or $T^G$,
+% SAT times,  $F$, $|A_F|$;  number of FOND atoms and actions in $Q^+_F$ (include those in $F$, $A_F$ plus translation), $I_F$ and $G_F$,
+% FOND time, resulting policy.  Numbers can go all in a table.
+
+\paragraph{Clearing a block.}
+The generalized problem $\Q_{clear}$ aims at achieving the goal $clear(x)$ for any block $x$
+in the Blocks world. 
+We take the standard encoding with primitive predicates $on(a, b)$, $clear(b)$, $ontable(b)$, $holding(b)$ and $handempty$.
+
+The abstract action model features one action \emph{pick-above-$x$} to pick the topmost block above $x$,
+and one action \emph{place-aside} to put it away, both described in terms of features $H$, $X$ and $n(x)$, 
+that denote, respectively, whether any block is being held, whether block $x$ is being held,
+and the number of blocks above $x$.\footnote{
+Action and features names are assigned manually by us to increase readability, but correspond faithfully
+to the learnt entities.
+}
+%
+The model is provably correct, but not complete in general, as it lacks e.g.
+any action to pick block $x$.
+However, it is sufficient to obtain a policy with a single loop that will 
+alternatively \emph{pick-above-$x$} and \emph{place-aside} until block $x$ is clear,
+which provably \emph{achieves the goal from any configuration of blocks, for any number of blocks}.
+
+% Action 1:
+%         PRE: NOT holding(a), bool[holding]
+%         EFFS: NOT bool[holding]
+% 
+% Action 2:
+%         PRE: NOT bool[holding], NOT holding(a), n(a) > 0
+%         EFFS: DEC n(a), bool[holding]
+
+% Loop:
+%   Action 2 [Pick block above x]
+%   If n(a) = 0 Break
+%   Action 1 [Put block aside]
+%
+
+\paragraph{Stacking two blocks from different towers.}
+The problem $\Q_{on\text{-}2\text{-}tow}$ is aims at achieving a goal of the form $on(x,y)$
+in the Blocks world for any two blocks $x$ that are initially not in the same tower.
+We take the same encoding as above.
+
+\paragraph{Gripper.}
+The problem $\Q_{gripper}$ is based on the classical Gripper problem, where a robot with two arms needs to 
+move a number of blocks between two different rooms, each arm being able to carry only one ball at a time.
+We use the standard encoding with primitive predicates $at\text{-}robby(l)$, $at\text{-}ball(b, l)$, $free(g)$, $carry(b, g)$ that denote,
+respectively, whether the robot (the ball) is at location $l$, whether gripper $g$ is free, and whether gripper $g$ carries ball $b$,
+plus type unary predicates $room$, $ball$, and $gripper$.
+
+\paragraph{Collecting rewards in a maze.}
+The objective in the problem $\Q_{reward}$ is to move in an $n \times m$ grid to pick up certain
+rewards that are spread out in certain grid cells, while avoiding other cells which are considered as \emph{blocked}.
+The encoding has primitive predicates $reward(c)$, $at(c)$, $blocked(c)$ and $adjacent(c_1, c_2)$ that denote
+respectively the position of the rewards, of the agent, of the blocked cells, and the grid topology.
+We use two single instances of the problem with size $5 \times 5$ and $4 \times 4$ and different distributions of blocks and rewards.
 

### 2018-09-05 doc/paper.tex
draft progress

@@
 In the experiments we will consider also a smaller theory  that avoids the consideration of all pairs
-of expanded states or all pairs of transitions in $\cal S$]. For this, some expanded states $s$ and transitions $(s,s')$ in
+of expanded states or all pairs of transitions in $\cal S$. For this, some expanded states $s$ and transitions $(s,s')$ in
 tuples $(s,s',P)$ are marked as \emph{goal relevant}. This is achieved by computing some plans for some instances $P$ in $\Q$, 
@@
 
-The feature complexity bound $k$ in the feature pool $\mathcal{F}$ is set to a value of $k=10$, unless no solution is found
-to the theory $T({\cal S},{\cal F})$, in which case it is increased to $k=20$. 
+The feature complexity bound $k$ in the feature pool $\mathcal{F}$ is set to a value of $k=8$.
+In our experiments, we increased this value to $k=16$ if no model was found
+for the logical theories $T$ and $T_G$, and although it would be more effective
+to devise an adaptive strategy to do so incrementally or through binary search,
+that is beyond the scope of this work.
+% theory $T({\cal S},{\cal F})$, in which case it is increased to $k=20$. 
 % We have considered strategies to increase the bound adaptively
-The resulting theory $T({\cal S},{\cal F})$ is solved with the Open-WBO MAXSAT solver \cite{martins2014open}.
+The resulting theories are then solved with the Open-WBO MAXSAT solver \cite{martins2014open}.
 
@@
 {
-% \scriptsize
-% \footnotesize
+% \def\arraystretch{1}%  1 is the default
 \small
-\def\arraystretch{1}%  1 is the default
 
@@
 \multirow{1}{*}{$\Q_{on\text{-}2\text{-}tow}$}   
-& 3 &        420  & 8 & 657 & -K & -K & 18.3K &     3.3M    &  0.02&  5 & 7   &  - & - & - & - \\% Total complexity of F: 18
+& 3 &        420  & 8 & 657 & 128K & 25.8M & 18.3K &     3.3M    &  0.02&  5 & 7   &  - & - & - & - \\% Total complexity of F: 18
+% Max-sat problem has 128217 variables and 25825428 unique clauses (with repetitions: 25825428). Top weight is 4577
 
@@
 We take the standard encoding with primitive predicates $on(a, b)$, $clear(b)$, $ontable(b)$, $holding(b)$ and $handempty$.
+%
+With one single instance $\Q_{clear}$ with 5 blocks, we are able to learn
+an abstract action model which features two actions:
 
-The abstract action model features one action \emph{pick-above-$x$} to pick the topmost block above $x$,
-and one action \emph{place-aside} to put it away, both described in terms of features $H$, $X$ and $n(x)$, 
+- \emph{pick-above-$x$}: \tuple{\neg X, \neg H, n(x) > 0; n(x) \downarrow},
+
+- \emph{place-aside}: \tuple{\neg X, H; \neg H},
+
+\noindent both described in terms of features $H: \abs{holding}$, $X: \abs{holding \sqcap \set{x}}$ and $n(x)\equiv \abs{\exists on^* . \set{x}}$, 
 that denote, respectively, whether any block is being held, whether block $x$ is being held,
@@
 The problem $\Q_{on\text{-}2\text{-}tow}$ is aims at achieving a goal of the form $on(x,y)$
-in the Blocks world for any two blocks $x$ that are initially not in the same tower.
-We take the same encoding as above.
+in the Blocks world for any two blocks $x$ and $y$ that are initially not in the same tower.
+We take the same encoding as above. Based on 3 different instances with up to 14
+blocks, we are able to learn an abstract action model with 5 features and 7 actions.
+The set of features is made up of  $X$, $n(x)$ and $n(y)$, with the same syntax and semantics as given before,
+as well as $E: handempty$
+and $XY: [(\exists on^{-1} . \set{x})\sqcap \set{b}]$,
+the second of which precisely denotes that block $x$ is on block $y$.
+The set of actions includes actions to pick up the topmost block above $x$, to pick up the topmost block above $y$,
+to put a block being held \emph{aside}, i.e. not above $x$ nor above $y$, to pick $x$, 
+to place $x$ on $y$ and to place $x$ on the table. This last action will indeed be necessary in case block $x$ is initially being held
+but block $y$ is not clear.
+
+DISCUSS POLICY WHEN AVAILABLE. / 
+DISCUSS WHY WOULDN'T WORK WHEN IN SAME TOWER.
 
 \paragraph{Gripper.}
-The problem $\Q_{gripper}$ is based on the classical Gripper problem, where a robot with two arms needs to 
-move a number of blocks between two different rooms, each arm being able to carry only one ball at a time.
+The problem $\Q_{gripper}$ is based on the classical Gripper problem, where a robot with two grippers needs to 
+move a number of blocks from one room into another \emph{target} room $x$, which we take to be a parameter of the problem,
+each gripper of the robot being able to carry only one ball at a time.
 We use the standard encoding with primitive predicates $at\text{-}robby(l)$, $at\text{-}ball(b, l)$, $free(g)$, $carry(b, g)$ that denote,
 respectively, whether the robot (the ball) is at location $l$, whether gripper $g$ is free, and whether gripper $g$ carries ball $b$,
-plus type unary predicates $room$, $ball$, and $gripper$.
+plus unary type predicates $room$, $ball$, and $gripper$.
+%
+We feed our pipeline with 2 training instances of the problem with only 2 rooms and at most 4 balls.
+The resulting abstract action model is based on the following features:
+
+- $C: \exists carry . \top$ (number of balls carried by the robot),
+
+- $X: at\_robby \sqcap \set{x}$ (whether robot in target room),
+
+- $G: \abs{free}$ (number of empty grippers), and
+
+- $B: \exists at . \neg \set{x}$ (number of balls not in target room),
+
+\noindent and contains the following actions :
+
+- \emph{leave-ball(s)-at-$x$}: \tuple{C>0, X;  C\downarrow, G \uparrow}
+
+- \emph{move-to-$x$-half-loaded}: \tuple{\neg X, B=0, C>0, G>0;   X}
+
+- \emph{move-to-$x$-fully-loaded}: \tuple{\neg X, C>0, G=0;   X}
+
+- \emph{pick-ball-not-in-$x$}: \tuple{\neg X, B > 0, G > 0;   B \downarrow, G \downarrow, C\uparrow}
+
+- \emph{leave-$x$}: \tuple{X, C=0, G > 0;   \neg X}
+
+The model is not complete, as it e.g. features no action to pick a ball from room $x$,
+but is goal-oriented as a result of using the SAT theory $T_G$, in that it features all of the actions
+necessary to achieve the goal.
+This can be seen e.g. in the fact that the action to move \emph{move-to-$x$-half-loaded}
+can only be performed when no more balls remain in the other room.
+Additionally, the model is \emph{provably correct for any gripper instance with an arbitrary number of balls, 
+rooms and grippers}. CHECK !!
+
+The policy that results ... (!!??)
+
+% Action 1:
+%         PRE: ncarried > 0, robot-at-B
+%         EFFS: DEC ncarried, INC nfree-grippers
+% 
+% Action 2:
+%         PRE: NOT robot-at-B, nballs-A = 0, ncarried > 0, nfree-grippers > 0
+%         EFFS: robot-at-B
+% 
+% Action 3:
+%         PRE: NOT robot-at-B, ncarried > 0, nfree-grippers = 0
+%         EFFS: robot-at-B
+% 
+% Action 4:
+%         PRE: NOT robot-at-B, nballs-A > 0, nfree-grippers > 0
+%         EFFS: DEC nballs-A, DEC nfree-grippers, INC ncarried
+% 
+% Action 5:
+%         PRE: ncarried = 0, nfree-grippers > 0, robot-at-B
+%         EFFS: NOT robot-at-B
 
@@
 We use two single instances of the problem with size $5 \times 5$ and $4 \times 4$ and different distributions of blocks and rewards.
+We learn a set with 2 features:
 
-** Mention: symbold DRL paper by Shanahan in
+- $R: \abs{reward}$ (number of remaining rewards), and
+
+- $D: dist(at, adjacent:(\neg blocked), reward)$ (distance from current cell to closest cell with reward, traversing adjacent, unblocked cells only).
+
+\noindent The action model has also 2 actions:
+
+- \emph{pick-reward}: \tuple{D=0, R > 0;  R \downarrow, D \uparrow}
+
+- \emph{move-towards-closest-reward}: \tuple{R>0, D>0; D \downarrow}
+
+The model is not complete ... 
+...
+
+The policy ... 
+... EXPAND A BIT - TOO TIRED NOW.
+Discuss suboptimality of the policy.
+Perhaps discuss D=$\infty$ when no reward?
+
+% Action 1 [PICK-REWARD]:
+% 	PRE: min-dist[at, Restrict(adjacent,Not(blocked)), reward] = 0, num-rewards > 0
+% 	EFFS: DEC num-rewards, INC min-dist[at, Restrict(adjacent,Not(blocked)), reward]
+% 
+% Action 2 [MOVE-TO-CLOSEST-REWARD]:
+% 	PRE: min-dist[at, Restrict(adjacent,Not(blocked)), reward] > 0, num-rewards > 0
+% 	EFFS: DEC min-dist[at, Restrict(adjacent,Not(blocked)), reward]
+
+\paragraph{Summary.}
+XXX Perhaps mention general aspects of the results here, if space allows, e.g.:
+\begin{itemize}
+ \item Running times of the SAT solver are negligible.
+ \item Minimization of precondition DNFs
+ \item Theory T
+ \item Sample sparsity - we use very few instances, rather few transitions.
+ \item ** Mention: symbold DRL paper by Shanahan in 
+\end{itemize}
 

### 2018-09-05 doc/paper.tex
Some progress incorporating H's comments

@@
 In the \emph{random} sampling strategy, a large number $M$ of states are generated, potentially exhausting the full state space,
-and then a certain number $m <\!< M$ of them is sampled uniformly at random. If no goal state is found among those $m$,
+and then a certain number $m \ll M$ of them is sampled uniformly at random. If no goal state is found among those $m$,
 one random goal state from the larger set is drawn at random and added into the sample.
 
-The feature complexity bound $k$ in the feature pool $\mathcal{F}$ is set to a value of $k=8$.
-In our experiments, we increased this value to $k=16$ if no model was found
-for the logical theories $T$ and $T_G$, and although it would be more effective
-to devise an adaptive strategy to do so incrementally or through binary search,
-that is beyond the scope of this work.
+The complexity bound $k$ in the feature pool $\mathcal{F} = \mathcal{F}^k$ is set to a value of $k=8$,
+with just one difference in the concept generation process: distance features are only generated for the last example,
+meant to illustrate their usefulness.
+% In our experiments, we increased this value to $k=16$ if no model was found
+% for the logical theories $T$ and $T_G$, and although it would be more effective
+% to devise an adaptive strategy to do so incrementally or through binary search,
+% that is beyond the scope of this work.
 % theory $T({\cal S},{\cal F})$, in which case it is increased to $k=20$. 
@@
 
-The number of instances $P$ is never larger than $3$, and they are selected XXX THIS NEEDS TO BE WELL JUSTIFIED...
+The number of instances $P$ used to generate the samples was kept small 
+by choosing the instances $P$ manually, like a teacher. 
+Very few instances turn out to be required for generating general plans.
+The learning scheme is not supervised, as the training set are state 
+transitions, not state action pairs to be generalized; but is not fully unsupervised either as currently 
+implemented, with the instances $P$ from $\Q$ selected by hand.
+We do not expect the results to be different if such instances
+were to be sampled at random, but the number of instances required for 
+producing sound abstract actions and valid general plans, as well as the size of the 
+theories, would likely be higher.
+
+All the results below are with goal marking using the theory $T_G({\cal S},{\cal F})$ that is
+smaller than $T({\cal S,F})$ and produces smaller abstract action sets  $A_F$, which are forced
+to be sound relative to the sample $\mathcal{S}$ but not complete, thus 
+avoiding abstract actions that may make  planning harder and but are not relevant to the goal.
+In general, the theory $T({\cal S},{\cal F})$ results in a significantly larger number of atoms and clauses,
+as seen in \ref{tab:exp-result-data}, which makes the MAXSAT solving process much harder.
 
@@
 \rowspacing \hline \rowspacing
-\multirow{1}{*}{$\Q_{on\text{-}2\text{-}tow}$}   
+\multirow{1}{*}{$\Q_{on}$}   
 & 3 &        420  & 8 & 657 & 128K & 25.8M & 18.3K &     3.3M    &  0.02&  5 & 7   &  - & - & - & - \\% Total complexity of F: 18
@@
 in the Blocks world. 
-We take the standard encoding with primitive predicates $on(a, b)$, $clear(b)$, $ontable(b)$, $holding(b)$ and $handempty$.
+We take the standard STRIPS representation with primitive predicates $on(a, b)$, $clear(b)$, $ontable(b)$, $holding(b)$ and $handempty$.
 %
@@
 \paragraph{Stacking two blocks from different towers.}
-The problem $\Q_{on\text{-}2\text{-}tow}$ is aims at achieving a goal of the form $on(x,y)$
+The problem $\Q_{on}$ is aims at achieving a goal of the form $on(x,y)$
 in the Blocks world for any two blocks $x$ and $y$ that are initially not in the same tower.
-We take the same encoding as above. Based on 3 different instances with up to 14
+We take the same STRIPS representation as above. Based on 3 different instances with up to 14
 blocks, we are able to learn an abstract action model with 5 features and 7 actions.
@@
 the second of which precisely denotes that block $x$ is on block $y$.
+
 The set of actions includes actions to pick up the topmost block above $x$, to pick up the topmost block above $y$,
 to put a block being held \emph{aside}, i.e. not above $x$ nor above $y$, to pick $x$, 
-to place $x$ on $y$ and to place $x$ on the table. This last action will indeed be necessary in case block $x$ is initially being held
+to place $x$ on $y$ and to place $x$ on the table.
+
+- \emph{pick-up-above-$x$}:
+
+- \emph{pick-up-above-$y$}:
+
+This last action will indeed be necessary in case block $x$ is initially being held
 but block $y$ is not clear.
@@
 each gripper of the robot being able to carry only one ball at a time.
-We use the standard encoding with primitive predicates $at\text{-}robby(l)$, $at\text{-}ball(b, l)$, $free(g)$, $carry(b, g)$ that denote,
+We use the standard STRIPS representation with primitive predicates $at\text{-}robby(l)$, $at\text{-}ball(b, l)$, $free(g)$, $carry(b, g)$ that denote,
 respectively, whether the robot (the ball) is at location $l$, whether gripper $g$ is free, and whether gripper $g$ carries ball $b$,
@@
 
-- \emph{leave-ball(s)-at-$x$}: \tuple{C>0, X;  C\downarrow, G \uparrow}
+- \emph{drop-ball(s)-at-$x$}: \tuple{C>0, X;  C\downarrow, G \uparrow}
 
@@
 
-- \emph{leave-$x$}: \tuple{X, C=0, G > 0;   \neg X}
+- \emph{drop-$x$}: \tuple{X, C=0, G > 0;   \neg X}
 
@@
 rewards that are spread out in certain grid cells, while avoiding other cells which are considered as \emph{blocked}.
-The encoding has primitive predicates $reward(c)$, $at(c)$, $blocked(c)$ and $adjacent(c_1, c_2)$ that denote
+This example is from recent work in symbolic deep
+reinforcement learning, where there is a penalty associated for 
+stepping in blocked cells instead \cite{garnelo2016towards}.
+%
+We represent the instances $P$ with primitive predicates $reward(c)$, $at(c)$, $blocked(c)$ and $adjacent(c_1, c_2)$ that denote
 respectively the position of the rewards, of the agent, of the blocked cells, and the grid topology.
@@
 
-\paragraph{Summary.}
-XXX Perhaps mention general aspects of the results here, if space allows, e.g.:
-\begin{itemize}
- \item Running times of the SAT solver are negligible.
- \item Minimization of precondition DNFs
- \item Theory T
- \item Sample sparsity - we use very few instances, rather few transitions.
- \item ** Mention: symbold DRL paper by Shanahan in 
-\end{itemize}
+\paragraph{Observations.}
+We conclude the section with a few general observations:
+
+\begin{enumerate}
+\item
+\emph{Soundness}: interestingly, abstract actions obtained from a small number of samples in
+$\mathcal{S}$ turn out to be sound not just for $\mathcal{S}$ but for the generalized problem $\Q$,
+as discussed above. Yet this does not need to be true for the resulting policies to be 
+correct for $\Q$.
+%
+Abstract actions $A_F$ do not actually have to be sound for a policy $\pi$
+obtained from them to be sound in $\Q$.
+Indeed, for the correctness of $\pi$ for solving $\Q$,
+it suffices for $A_F$ to be sound \emph{over the states reachable by $\pi$}
+in all the instances $P$ of $\Q$.
+
+For example, the abstract action \emph{pick-up-above-$y$} in $\Q_{on}$
+is sound relative to the sample set $\mathcal{S}$ used, but not relative to the general
+$\Q_{on}$. In any state $s$ where $x$ is a few blocks above $y$ and clear,
+there is no concrete action that instantiates \emph{pick-up-above-$y$},
+as picking up the topmost block above $y$ would necessarily make feature $X \equiv holding(x)$
+true, which the abstract action does not.
+%
+Keeping track of the information necessary to ensure general correctness would require indeed 
+a feature set $F$ with higher complexity.
+Yet if we start in states where $x$ and $y$ are in different towers, 
+states as $s$ will never be reached, and hence the policy is correct and solves $\Q_{on}$.
+
+\item
+\emph{Grammar and Feature Pool}: the general DL grammar we have described appears to be 
+adequate to generate a sufficiently expressive pool of features for the examples we use,
+but this does not have to be always so.
+To come up with a general strategy in the Towers of Hanoi, for example,
+the first action to be taken  (whether to move the smallest disk to target or buffer peg) 
+depends on the number of disks being odd or even,
+which cannot be easily expressed with the type of features generated by our grammar.
+
+\item
+\emph{Expressive and Computational Limitations}:
+The current approach
+
+*** XXX ***
+
+The use of primitive \emph{goal predicates} like $on_g$ with denotation
+given by the goal conjunctions in the concrete instances, as done in 
+\cite{martin-geffner:generalized}, could potentially alleviate this
+e.g. to be able to generate general policies for all BW problems.
+
+Why pipeline cannot be used to solve arbitrary problems. Eg, logistics.
+Que partes explotan? Sizes k and $\abs{\mathcal{S}}$?
+
+* Goal in $\Q$ needs to be expressible in generic terms (but this is not well formalized in our formulation)
+* Increasing number of domain parameters kills scalability.
+* How to select which are the good parameters if we want arbitrary goals? This is also related to the proper formalization
+  of the approach, which we havent done.
+
+%  \item Running times of the SAT solver are negligible.
+%  \item Minimization of precondition DNFs
+%  \item Theory T
+\end{enumerate}
 

### 2018-09-05 doc/paper.tex
Added Blai's changes

@@
 \newcommand{\Q}{\mathcal{Q}}
+\newcommand{\F}{\mathcal{F}}
+\renewcommand{\S}{\mathcal{S}}
+\newcommand{\G}{\mathcal{G}}
 \newcommand{\dexp}{\text{2-}\!\exp}
@@
   Generalized planning is concerned with the computation of plans that solve multiple instances.
-  Recently, it has been shown that such plans can be expressed as mappings of feature values into actions, and that they
-  can often be  computed by means   of fully observable non-deterministic (FOND) planners after a sequence of
-  transformations. The actions in such  plans, however,   are not the actions in the  instances themselves,  which  are
-  not necessarily common to  other instances,   but  abstract actions that track the effect of the  actions
-  on the given set of features. The formulation assumes  that both the features and the abstract actions are given.
-  In this work, we address this limitation by  showing  how to learn them.  The resulting account of generalized
-  planning  combines learning and planning in a novel way: a \emph{learner},  based on a min SAT formulation, yields
-  the features and abstract actions  from    state transitions sampled  from some problem instances, 
-  and a FOND \emph{planner} uses this information,   suitably   transformed as before,  to produce the general plans.
-  Correctness guarantees are given and 
-  experimental results on several  domains are reported.
+  Recently, it has been shown that generalized plans can be expressed as mappings of feature
+  values into actions, and that they can often be computed with fully observable non-deterministic
+  (FOND) planners on suitable FOND problems. %after a sequence of transformations.
+  The actions in such plans, however, are not the actions in the instances themselves, which are
+  not necessarily common to  other instances, but abstract actions that track the effect of the
+  actions on the set of features. The formulation assumes that both the features and the abstract
+  actions are given.
+  In this work, we address this limitation by showing how to automatically learn the features and
+  abstract actions. The resulting account of generalized planning combines learning and planning
+  in a novel way: a \emph{learner}, based on a min SAT formulation, yields the features and abstract
+  actions from state transitions sampled from some problem instances, and a FOND \emph{planner}
+  uses this information, suitably transformed, to produce the general plans.
+  Correctness guarantees are given and experimental results on several domains are reported.
 \end{abstract}
@@
 among  all the instances. More recently, this formulation has been extended by
-\citeay{bonet:ijcai2018} to account for relational domains like Blocks world  where the set of
+\citeay{bonet:ijcai2018} to account for relational domains like Blocksworld  where the set of
 objects and  actions change from instance to instance.
-For this,   the observations  are replaced by  a given  set of boolean
+In the new formulation, the observations  are replaced by  a set of boolean
 and numerical features $F$ and a set of \emph{abstract actions} $A_F$.
@@
 a concrete action with a similar  effect  on  the features. Moreover, the form of the abstract actions
-ensures that such plans can be computed using fully observable non-deterministic (FOND) planners
-after some transformations.
+ensures that generalized plans can be computed using fully observable non-deterministic (FOND) planners
+once the generalized planning problem is transformed into a FOND problem.
+%after some transformations.
 
@@
 
-Bonet and Geffner's  formulation of generalized planning, however, assumes that the
-features and the abstract actions are given.   In this work, we address this limitation.
-We achieving this by  \emph{showing  how features and abstract actions  can be learned from
-the primitive predicates used to define the instances and state transitions
-sampled from them.} For example,  in \cite{bonet:ijcai2018}, 
-the general policy for achieving the goal $clear(x)$  is obtained 
-using a FOND planner by introducing the boolean feature $H$ for holding a block,
-the numerical feature $n(x)$ for counting the  number of blocks above $x$, and  two   abstract  actions:
+Bonet and Geffner's  formulation of generalized planning assumes that the
+features and abstract actions are given. In this work we address this limitation
+by  \emph{showing  how the features and abstract actions  can be learned from
+the primitive predicates used to define the instances and a given sample of state transitions.}
+For example,  the general policy for achieving $clear(x)$ of \citeay{bonet:ijcai2018}
+is obtained using a FOND planner on an abstracton that consists of the boolean
+feature $H$ that tracks whether the gripper holds a block and a numerical feature
+$n(x)$ that counts the number of blocks above $x$, and two abstract actions:
 one with preconditions $\neg H$ and $n(x) > 0$, and effects $H$ and $n(x)\mminus$
-(decrement $n(x)$), the other with precondition $H$ and effect $\neg H$.
-In this work, we show how  to obtain such  policies without
-having to provide the  features and  abstract actions by hand.
-The abstraction is learned instead from the domain predicates
-and  state transitions sampled from some  instances.
-
-The work has relation to a number of research threads in planning, knowledge representation
-and machine learning.  We make use of SAT solvers and description logics for learning the features
-and abstract actions \cite{sat-handbook,dl-handbook}. The abstract actions provide a model from which the plans are obtained via
+(decrement $n(x)$), and the other with precondition $H$ and effect $\neg H$.
+In this work, we show how to obtain such policies without having to provide the
+features and abstract actions by hand.  The abstraction is learned instead from
+the domain predicates and state transitions sampled from some instances.
+
+The work has relation to a number of research threads in planning, knowledge
+representation, and machine learning. We make use of SAT solvers and description
+logics for learning features and abstract actions \cite{sat-handbook,dl-handbook}.
+The abstract actions provide a model from which the plans are obtained via
 transformations and FOND planners \cite{geffner:book,ghallab:book}.
-The model is not the model of an instance but a generalized model for  obtaining  plans
-that work for multiple instances. In this sense, the work is different than
-action model learning \cite{yang:action-learning} and model-based reinforcement learning
-and closer in aims to work on learning general policies from examples or experience
-\cite{martin:concept,fern:bias,mazebase,general-drl}. Generalized planning has  also been formulated as a problem in first-order logic
-\cite{srivastava:generalized} and general plans over finite horizons have also been derived using  first-order regression
-\cite{boutilier2001symbolic,wang2008first,van2012solving}. Our approach differs from first-order approaches in the use
-efficient propositional planners, and from purely learning approaches, in the \emph{formal guarantees} that are characteristic of planning:
-if the learned abstract actions are actually sound,  the resulting general plan must be correct.
-
-The paper is organized as follows. We provide the background first and  show then  how features and
-abstract actions can be  learned by enforcing soundness and completeness over a set of samples  using 
-a  pool of features obtained the   domain  predicates. The full  computational model  is then
-summarized,  following by the experimental results and a discussion.
+The model is not the model of an instance but a generalized model for obtaining
+plans that work for multiple instances. In this sense, the work is different than
+action model learning \cite{yang:action-learning} and model-based reinforcement
+learning and closer in aims to work on learning general policies from examples or
+experience \cite{martin:concept,fern:bias,mazebase,general-drl}.
+Generalized planning has also been formulated as a problem in first-order logic
+\cite{srivastava:generalized} and general plans over finite horizons have been
+derived using first-order regression \cite{boutilier2001symbolic,wang2008first,van2012solving}.
+Our approach differs from first-order approaches in the use of propositional
+planners, and from purely learning approaches, in the \emph{formal guarantees} that
+are characteristic of planning: if the learned abstract actions are actually sound,
+the resulting general plans are correct.
+
+The paper is organized as follows. We provide the background first and show
+how features and abstract actions can be learned by enforcing soundness and
+completeness over a set of samples using a pool of features obtained the
+domain predicates. The full computational model is then summarized, followed
+by experimental results and a discussion.
 
@@
 
-A \emph{generalize planning problem} $\Q$ is a collection of planning instances $P$.
-An instance $P$ is a  classical planning problems expressed in some compact language as a
-tuple $P=\tup{V,I,A,G}$ where $V$ is a set of state variables that can take a finite set of values (boolean or not), $I$
-is a set of atoms over $V$ defining an initial state $s_0$, $G$ is a set of atoms or literals over $V$ describing the goal states,
-and $A$ is a set of actions $a$ with their preconditions and effects that define the set $A(s)$ of actions applicable
-in any state $s$, and the successor state function $f(a,s)$, $a \in A(s)$. A state is a valuation over $V$, and a solution to
-$P$ is an applicable action sequence $\pi=a_0,\ldots,a_n$ that generates  a state sequence $s_0,s_1,\ldots,s_{n}$
-where $s_n$ is a goal state (makes $G$ true). In this sequence, $a_i \in A(s_i)$ and $s_{i+1}=f(a_i,s_i)$ for $i=0, \ldots, n-1$.
-A state $s$ is reachable in $P$ if $s=s_n$ for one such sequence. A solution to the \emph{generalized problem} $\Q$ is
-a solution to all instances $P \in Q$. The form of such solutions is described  below.
+A \emph{generalized planning problem} $\Q$ is a collection of planning instances $P$.
+An instance $P$ is a  classical planning problems expressed in some compact language
+as a tuple $P=\tup{V,I,G,A}$ where $V$ is a set of state variables that can take a
+finite set of values (boolean or not), $I$ is a set of atoms over $V$ defining an
+initial state $s_0$, $G$ is a set of atoms or literals over $V$ describing the goal
+states, and $A$ is a set of actions $a$ with their preconditions and effects that
+define the set $A(s)$ of actions applicable in any state $s$, and the successor
+state function $f(a,s)$, for any state $s$ and $a \in A(s)$.
+A state is a valuation over $V$, and a solution to $P$ is an applicable action
+sequence $\pi=a_0,\ldots,a_n$ that generates  a state sequence $s_0,s_1,\ldots,s_{n}$
+where $s_n$ is a goal state (makes $G$ true). In this sequence, $a_i \in A(s_i)$
+and $s_{i+1}=f(a_i,s_i)$ for $i=0, \ldots, n-1$. A state $s$ is reachable in $P$
+if $s=s_n$ for one such sequence. A solution to the \emph{generalized problem}
+$\Q$ is a solution to all instances $P \in Q$. The form of such solutions is
+described below.
 
 A \emph{feature} $f$ for a class $\Q$ of problems represents a function $\phi_f$
-that takes an instance $P$ from $\Q$ and a state $s$ reachable in $P$, 
-and results in a value denoted as $\phi_f(s)$. 
-A feature is \emph{boolean}  if it results in a boolean value, and  \emph{numeric} if it results
-in a numerical value, that we assume to be integer and non-negative. The features $H$
-and $n(x)$ represent two features in the Blocks world (BW): the first a boolean
-feature encoding that   the gripper is not empty; the second, a numerical
-feature encoding  the number of blocks above  $x$.
-
-Symbols like $x$ to denote \emph{parameters} whose value depends on the instance  $P$ in  $\Q$.
-For example, if $\Q_{clear}$  denotes  the BW instances with goals of the form $clear(x)$,
-then the value of $x$ in an instance $P$ with goal $clear(a)$ is  $a$.
+that takes an instance $P$ from $\Q$ and a state $s$ reachable in $P$, and results
+in a value denoted as $\phi_f(s)$. 
+A feature is \emph{boolean} if it results in a boolean value, and \emph{numeric}
+if it results in a numerical value, that we assume to be a non-negative integer.
+The features $H$ and $n(x)$ represent two features in the Blocksworld:
+the first is a boolean feature that tracks whether the gripper is empty, while
+the second is numerical and tracks the number of blocks above $x$.
+
+Symbols like $x$ denote \emph{parameters} whose value depends on the
+instance $P$ in $\Q$. For example, if $\Q_{clear}$ denotes the Blocksworld
+instances with goals of the form $clear(x)$, the value of $x$ in an instance
+$P$ with goal $clear(a)$ is $a$.
 
@@
 
-Features $f$ refer to \emph{state functions} $\phi_f(s)$  in the instances of the generalized problem $\Q$,
-but to \emph{state variables} in the abstraction.  Likewise, an \emph{abstract state} is  a truth valuation over the
-state variables  $p$ associated with boolean features and the atoms $n=0$ over  state variables  $n$ associated with numerical features.
-The abstract state $\bar{s}$  that \emph{corresponds} to a concrete state $s$ is the truth valuation that makes
-$p$ true iff $\phi_p(s)=true$, and $n=0$ true iff $\phi_n(s) = 0$, and an abstract action $\bar{a}$
-is applicable in a state $s$ if its preconditions are satisfied in $\bar{s}$.
-
-A set of abstract actions $A_F$ is \emph{sound}  relative to a generalized problem $\Q$ 
-iff for any reachable state $s$ over an instance $P$ in $\Q$, if an abstract action $\bar{a}$ in $A_F$
-is applicable in $\bar{s}$, then there is a concrete action $a$ in $P$ that is applicable in $s$
-that has the \emph{same qualitative  effects as}  $\bar{a}$ over the features. The action $a$
-\emph{represents} or \emph{instantiates}  the abstract action $\bar{a}$ in $s$.
-
-Similarly, a set of abstract actions $A_F$ is \emph{complete}
-iff for any reachable state $s$ over an instance $P$ in $\Q$, if a concrete action $a$ in $P$
-that is applicable in $s$, then there is an abstract action $\bar{a}$ in $A_F$
-that is  applicable in $\bar{s}$ that has the \emph{same qualitative effects over the features as $a$.}
-
-The actions $a$ and $\bar{a}$ \emph{have the same qualitative  effects} in a state $s$ 
-when  they have the same   effects on the boolean features $p$,  and the same \emph{qualitative effects} on the numerical
-features $n$; i.e., $a$ decreases (resp. increases) the value of the function $\phi_n(s)$ in $s$ iff
-$\bar{a}$ decreases (resp. increases) the value of the state variable  $n$.
+Features $f$ refer to \emph{state functions} $\phi_f(s)$ in the instances
+of the generalized problem $\Q$, but to \emph{state variables} in the
+abstraction. Likewise, an \emph{abstract state} is  a truth valuation over
+the state variables $p$ associated with the boolean features and the atoms
+$n=0$ over the state variables $n$ associated with the numerical features.
+The abstract state $\bar{s}$ that \emph{corresponds} to a concrete state $s$
+is the truth valuation that makes $p$ true iff $\phi_p(s)=true$, and $n=0$
+true iff $\phi_n(s) = 0$, and an abstract action $\bar{a}$ is applicable in
+a state $s$ if its preconditions are satisfied in $\bar{s}$.
+
+A set of abstract actions $A_F$ is \emph{sound} relative to a generalized
+problem $\Q$ iff for any reachable state $s$ over an instance $P$ in $\Q$
+and any abstract action $\bar{a}$ in $A_F$ that is applicable in $\bar{s}$,
+there is a concrete action $a$ in $P$ that is applicable in $s$ and that
+has the \emph{same qualitative effects as} $\bar{a}$ over the features.
+The action $a$ \emph{represents} or \emph{instantiates} the abstract action
+$\bar{a}$ on $s$. 
+%
+Similarly, a set of abstract actions $A_F$ is \emph{complete} iff for any
+reachable state $s$ over an instance $P$ in $\Q$ and any concrete action $a$
+in $P$ that is applicable in $s$, there is an abstract action $\bar{a}$ in
+$A_F$ that is applicable in $\bar{s}$ and that has the \emph{same qualitative
+effects over the features as $a$.}
+
+The actions $a$ and $\bar{a}$ \emph{have the same qualitative effects} in
+a state $s$ when they have the same effects on the boolean features $p$,
+and the same \emph{qualitative effects} on the numerical features $n$; i.e.,
+$a$ decreases (resp.\ increases) the value of the function $\phi_n(s)$ in
+$s$ iff $\bar{a}$ decreases (resp.\ increases) the value of the state
+variable  $n$.
 
@@
   For the features $F=\{H,n(x)\}$, the abstract action $\bar{a}=\abst{\neg H,n(x)>0}{H,n(x)\mminus}$ 
-  is sound in  problem $\Q_{clear}$ that includes the BW instances with stack and unstack actions.
-  Indeed, if $s$ is a reachable state in an instance $P$ and $\bar{a}$ is applicable in $\bar{s}$, 
-  the gripper must be empty in $s$ and some block  must be above $x$. If $A$ is the clear block above $x$ and $on(A,B)$ is true in $s$,
-  then the  action $a=unstack(A,B)$  will be  applicable in $s$ and have the same qualitative effects of  $\bar{a}$
-  over $F$: they both make $H$ true and decrease the value of $n(x)$. The abstract action $\bar{a'}=\abst{H}{\neg H}$ is also sound.
+  is sound in the generalized problem $\Q_{clear}$ which consists of all
+  the Blocksworld instances, with stack and unstack actions, whose goal
+  is $clear(x)$.
+  Indeed, if $s$ is a reachable state in an instance $P$ and $\bar{a}$ is
+  applicable in $\bar{s}$, the gripper must be empty in $s$ and some
+  block must be above $x$. Thus, if $A$ is the clear block above $x$ and
+  $on(A,B)$ is true in $s$, the action $a=unstack(A,B)$ is applicable in
+  $s$ and it has the same qualitative effects of $\bar{a}$ over $F$: they
+  both make $H$ true and both decrease the value of $n(x)$.
+  Likewise, the abstract action $\bar{a'}=\abst{H}{\neg H}$ is also sound. \qed
 \end{example}
@@
 For a set of abstract actions to be complete in this example, it should 
-should contain  abstracts that track the  effects of the other  actions
-on the features such as stacking a block above $x$, picking up a block that is not above $x$, and so on.
-It may well be possible that a particular feature set $F$ can't support a set of abstract actions
-$A_F$ that is both sound and complete relative to  $\Q$. Soundness, however,  is crucial for deriving general plans as it ensures
-that in any state of any instance of $\Q$,  an applicable  abstract action  can be ``instantiated'' 
-by   one or more concrete actions that have the same qualitative effects on the features.
+contain abstracts actions to represent all the concrete actions in the
+problem such as stacking a block above $x$, picking up a block that is
+not above $x$, and so on.
+It may well be possible that a particular feature set $F$ cannot support
+a set of abstract actions $A_F$ that is both sound and complete relative
+to $\Q$. Soundness, however, is crucial for deriving general plans as it
+ensures that in any state of any instance of $\Q$, an applicable abstract
+action  can be ``instantiated'' by one or more concrete actions that have
+the same qualitative effects on the features.
 
@@
 
-A solution to a  generalized problem $\Q$ given the  features $F$ and  abstract actions $A_F$
-is a partial  function $\pi$ that maps \emph{abstract states} into \emph{abstract actions} such that
-$\pi$ solves all instances $P$ in $\Q$. The plan or policy $\pi$ induces a  trajectory $s_0,a_0,s_1, \ldots, s_n$ in an instance $P$ of $\Q$
-iff a)~$s_0$ is the initial state in $P$, b)~$a_i$ is one of the actions that instantiate 
-$\pi(\bar{s_i})$ in $s_i$, c)~$s_{i+1}=f(a_i,s_i)$ if $a_i$ is applicable in $s_i$,
-and d)~$s_n$ is a goal  state of $P$,  $n=\infty$, or $\pi(\bar{s_n})$ is undefined, denotes an abstract action
-that is not applicable in $\bar{s_n}$, or no applicable action $a_n$ instantiates  $\pi(\bar{s_n})$.
-The policy $\pi$ solves $P$ iff all the induced trajectories reach the goal of $P$.
+A solution to a generalized problem $\Q$ given the features $F$ and abstract
+actions $A_F$ is a partial function $\pi$ that maps \emph{abstract states}
+into \emph{abstract actions} such that $\pi$ solves all instances $P$ in $\Q$.
+The plan or policy $\pi$ induces a trajectory $s_0,a_0,s_1, \ldots, s_n$ in
+an instance $P$ of $\Q$ iff
+1)~$s_0$ is the initial state in $P$,
+2)~$a_i$ is one of the actions that instantiate $\pi(\bar{s}_i)$ in $s_i$,
+3)~$a_i$ is applicable in $s_i$ and $s_{i+1}=f(a_i,s_i)$, and
+4)~$s_n$ is a goal state of $P$, $n=\infty$, or $\pi(\bar{s}_n)$ either is
+undefined, denotes an abstract action that is not applicable in $\bar{s_n}$,
+or no applicable action $a_n$ instantiates $\pi(\bar{s}_n)$.
+The policy $\pi$ solves $P$ iff all the induced trajectories reach the
+goal of $P$.
 
 \begin{example}
-A solution for the generalized problem $\Q_{clear}$  with feature set $F=\{H,n(x)\}$
-is the policy $\pi$  given by the rules $\Rule{\neg H, n(x)>0}{\bar{a}}$ and $\Rule{H, n(x)>0}{\bar{a}'}$
-for the abstract actions $\bar{a}$ and $\bar{a'}$ above. The policy picks blocks above $x$ and puts them aside
-(not above $x$) until $n(x)$ becomes zero.
+  A solution for $\Q_{clear}$ with feature set
+  $F=\{H,n(x)\}$ is the policy $\pi$ given by the rules
+  $\Rule{\neg H, n(x)>0}{\bar{a}}$ and $\Rule{H, n(x)>0}{\bar{a}'}$
+  for the abstract actions $\bar{a}$ and $\bar{a'}$ above.
+  The policy picks blocks above $x$ and puts them aside (not above $x$)
+  until $n(x)$ becomes zero. \qed
 \end{example}
@@
 
-The steps for obtaining general policies $\pi$ for a generalized problem $\Q$
-using the features $F$ and the  abstract actions $A_F$ are as follows \cite{bonet:ijcai2018}:
-
-\medskip
-
-\noindent 1. The state variables $V_F$ and the abstract actions $A_F$ are extended
-  with initial and goal  formulas $I_F$ and $G_F$  over $V_F$ to yield a   \emph{numerical, non-deterministic planning problem}
-  $Q_F=\tup{V_F,I_F,G_F,A_F}$. For soundness,  $I_F$ must be  such that the initial states of instances $P$ in $\Q$ all satisfy $I_F$,
-  while all states that satisfy $G_F$ must be goal states of $P$.\footnote{
-   A state $s$ is said to satisfy $I_F$ or $G_F$ if  $\bar{s}$ does.}
-
-\medskip
-
-\noindent 2. The problem $Q_F=\tup{V_F,I_F,G_F,A_F}$ is converted into a \emph{boolean FOND}
-  problem $Q'_F=\tup{V'_F,I'_F,G'_F,A'_F}$ by replacing 
-  the numerical variables $n\in N$ by the symbols $n=0$, the first-order literals $n=0$ by propositional literals  $n=0$, 
-  the effects $n\pplus$ by deterministic effects $n > 0$, and the effects $n\mminus$ by non-deterministic effects $n>0\,|\,n=0$.
-
-  \medskip
-
-\noindent 3. The strong cyclic solutions of  $Q'_F$ assume that the
-  non-deterministic effects $n>0\,|\,n=0$ are \emph{fair} \cite{strong-cyclic},  
-  yet  they are  \emph{conditionally fair} only: infinite decrements of $n$ ensure $n=0$  eventually but just without
-  infinite  increments. The problem of obtanining  solutions of $Q'_F$ that assume conditional fairness only, the  terminating
-  solutions in   \cite{srivastava:aaai2011}, is mapped into the problem of solving an  amended FOND $Q^+_F$
-  \cite{bonet:ijcai2017}.\footnote{For this, we use a translation made available by the authors that fixes bugs in the  description found in the paper
-     at https://github.com/bonetblai/qnp2fond.}
-
-\medskip
-
-The strong cyclic solutions of  $Q^+_F$ computed by FOND planners off-the-shelf   are solutions to the generalized problem $\Q_F$.
+The steps for obtaining policies $\pi$ for a generalized problem $\Q$
+using the features $F$ and abstract actions $A_F$ are as follows \cite{bonet:ijcai2018}:
+
+\begin{enumerate}[1.]
+  \item The state variables $V_F$ and the abstract actions $A_F$ are
+    extended with initial and goal formulas $I_F$ and $G_F$ over $V_F$
+    to yield a \emph{numerical and non-deterministic planning problem}
+    $Q_F=\tup{V_F,I_F,G_F,A_F}$. For soundness, $I_F$ must be such that
+    the initial states of instances $P$ in $\Q$ all satisfy $I_F$, while
+    all states that satisfy $G_F$ must be goal states of $P$.\footnote{A
+    state $s$ is said to satisfy $I_F$ or $G_F$ if  $\bar{s}$ does.}
+%
+  \item The problem $Q_F=\tup{V_F,I_F,G_F,A_F}$ is converted into a
+    \emph{boolean FOND} problem $Q'_F=\tup{V'_F,I'_F,G'_F,A'_F}$ by
+    replacing the numerical variables $n\in N$ by the symbols $n=0$,
+    the first-order literals $n=0$ by propositional literals  $n=0$, 
+    the effects $n\pplus$ by deterministic effects $n > 0$, and the
+    effects $n\mminus$ by non-deterministic effects $n>0\,|\,n=0$.
+%
+  \item The solutions computed by a FOND planner on $Q'_F$ are the
+    strong cyclic solutions of $Q'_F$ \cite{strong-cyclic}.
+    Such solutions however do not necessarily solve $Q_F$ because the
+    non-determistic effects $n>0\,|\,n=0$ in $Q'_F$ are \emph{conditionally fair}
+    instead of being just \emph{fair}: infinite decrements of $n$ ensure
+    $n=0$ eventually \emph{only} when number of increments of $n$ is finite.
+    The problem of obtanining solutions of $Q'_F$ that only assume conditional
+    fairness, so-called terminating solutions in \cite{srivastava:aaai2011},
+    is mapped into the problem of solving an amended FOND $Q^+_F$ that
+    only assumes (standard) fairness \cite{bonet:ijcai2017}.\footnote{The
+    translation to $Q^+_F$ is performed with software kindly supplied by the authors of \cite{bonet:ijcai2017}.}
+    %available by the authors that fixes bugs in the description found
+    %in the paper at https://github.com/bonetblai/qnp2fond.}
+\end{enumerate}
+
+%The strong cyclic solutions of  $Q^+_F$ computed by FOND planners off-the-shelf   are solutions to the generalized problem $\Q_F$.
 
@@
   Since  no action in $A_F$  increases $n(x)$,   $Q^+_F=Q'_F$, and the strong cyclic solution  of $Q'_F$,
-  that corresponds to the policy above, solves $\Q_{clear}$.
+  that corresponds to the policy above, solves $\Q_{clear}$. \qed
 \end{example}
@@
 \begin{definition}
- For a generalized problem $\Q$, a \emph{sample set}  $\cal S$ is a non-empty set of tuples  $(s,s',P)$
+ For a generalized problem $\Q$, a \emph{sample set}  $\S$ is a non-empty set of tuples  $(s,s',P)$
  such that 1)~$P$ is an instance of $\Q$, 2)~$s$ is a reachable state in $P$, 3)~$s'$
  is a possible succesor of $s$ in $P$; i.e., $s'=f(a,s)$ for some action $a$ applicable in $s$,
- 4)~$\cal S$ is closed in the sense  that if $(s,s',P)$ is in $\cal S$, then $(s,s'',P)$ is in $\cal S$  if $s''$ is another possible succesor of $s$ in $P$.
+ 4)~$\S$ is closed in the sense  that if $(s,s',P)$ is in $\S$, then $(s,s'',P)$ is in $\S$  if $s''$ is another possible succesor of $s$ in $P$.
 \end{definition}
@@
 By assuming that sampled states $s$ are tagged with the instance $P$, 
-we can abbreviate the tuples $(s,s',P)$ as $(s,s')$, and refer to
-the states $s'$ in  the pairs $(s,s') \in {\cal S}$ as the successors
-of $s$ in ${\cal S}$. Notice that the closure condition 4) ensures that
-states $s$ appearing \emph{first} in pairs $(s,s')$ must be fully expanded in ${\cal S}$;
+we abbreviate the tuples $(s,s',P)$ as $(s,s')$, and refer to the
+states $s'$ in  the pairs $(s,s') \in \S$ as the successors of
+$s$ in $\S$. Notice that the closure condition ensures that
+states $s$ appearing \emph{first} in pairs $(s,s')$ must be fully expanded in $\S$;
 namely, all possible transitions $(s,s'')$ must be in the sample as well. We call them
-the \emph{expanded} states in ${\cal S}$.
+the \emph{expanded} states in $\S$.
 
@@
 \begin{definition}
-  A set of abstract actions $A_F$ is sound \emph{relative to a  sample set} $\cal S$ for $\Q$
-  iff for  any  abstract action $\bar{a}$ in $A_F$ that is applicable in  a expanded state $s$ of ${\cal S}$,
-  there is a transition $(s,s')$ in $\cal S$ that has the same qualitative  effects over $F$ as $\bar{a}$.
+  A set of abstract actions $A_F$ is sound \emph{relative to a  sample set} $\S$ for $\Q$
+  iff for  any  abstract action $\bar{a}$ in $A_F$ that is applicable in  a expanded state $s$ of $\S$,
+  there is a transition $(s,s')$ in $\S$ that has the same qualitative  effects over $F$ as $\bar{a}$.
 \end{definition}
@@
 \begin{definition}
-A set of abstract actions $A_F$ is complete   \emph{relative to a sample set} $\cal S$ for $\Q$
-iff for each transition  $(s,s')$ in  $\cal S$, there is an abstract action $\bar{a}$ in $A_F$
+  A set of abstract actions $A_F$ is complete   \emph{relative to a sample set} $\S$ for $\Q$
+  iff for each transition  $(s,s')$ in  $\S$, there is an abstract action $\bar{a}$ in $A_F$
   with the same qualitative effects over in $F$ that is applicable in $s$.
@@
 
-For computing the features and abstract actions
-we define  a  formula  $T({\cal S,F})$ that  is \emph{satisfiable} 
-iff there is a set of features $F  \subseteq {\cal F}$ and a set of abstract actions $A_F$ over $F$
-that is sound and complete relative to the sample set ${\cal S}$.
-
-\Omit{The  sets  $F$ and $A_F$ can
-     then be read off  from any  satisfying assignment. Later we will add weights
-     to so
-     me of the clauses to establish a preference over these assignments
-     (e.g., for  minimizing $|F|$).}
+For computing the features and abstract actions we define a formula
+$T(\S,\F)$ that is \emph{satisfiable} iff there is a set of
+features $F \subseteq \F$ and a set of abstract actions $A_F$
+over $F$ that is sound and complete relative to the sample set
+$\S$.
+\textcolor{red}{\bf **** Need to say what $\F$ is ****}
+
+\Omit{The  sets  $F$ and $A_F$ can then be read off from any satisfying
+  assignment. Later we will add weights to so me of the clauses to establish
+  a preference over these assignments (e.g., for  minimizing $|F|$).
+}
 
@@
 
-For each transition  $(s, s') \in \mathcal{S}$ and each feature $f \in \mathcal{F}$,
-$\Delta_f(s, s') \in \{+, -, \uparrow, \downarrow, \bot\}$ will denote the \emph{qualitative change of value}
-of feature $f$ along the transition $(s, s')$, which can go from false to true ($+$), from true to false ($-$) or remain unchanged ($\bot$),
-for boolean features, and can increase ($\uparrow$), decrease ($\downarrow$), or remain unchanged ($\bot$), for numerical features.
-The \emph{propositional variables}  in $T({\cal S,F})$ are then:
-
-  \begin{itemize}
-  \item  $selected(f)$ for each $f \in {\cal F}$,  true iff  $f$  selected  (in  $F$).
-  \item $D_1(s,t)$ for  states   $s$ and $t$ expanded in ${\cal S}$,  true iff selected features distinguish  $s$ from $t$;
-      i.e.  if $s$ and $t$ disagree on the truth value of a selected boolean feature $p$ or atom $n=0$ for a selected numerical feature $n$.
-      % Since $D_1$ is symmetric, only symbols for  $s < t$ are created   for  some  arbitrary, fixed order over the states in ${\cal S}_1$.
-  \item $D_2(s, s', t, t')$ for  each $(s, s')$ and $(t, t')$ in $\mathcal{S}$,
-    true iff selected features $f$ distinguish the two transitions; i.e., if $\Delta_f(s, s')\not=\Delta_f(t,t')$. 
-%      As for $D_1$, the number of $D_2$ variables is cut in half due to symmetries.
-  \end{itemize}
+For each transition  $(s, s') \in \S$ and each feature $f \in \F$,
+$\Delta_f(s, s') \in \{+, -, \uparrow, \downarrow, \bot\}$ will denote the
+\emph{qualitative change of value} of feature $f$ along the transition $(s, s')$,
+which can go from false to true ($+$), from true to false ($-$) or remain
+unchanged ($\bot$), for boolean features, and can increase ($\uparrow$),
+decrease ($\downarrow$), or remain unchanged ($\bot$), for numerical features.
+The \emph{propositional variables} in $T(\S,\F)$ are then:
+\begin{enumerate}[{\small$\bullet$}]
+  \item $selected(f)$ for each $f \in \F$, true iff $f$ selected (in $F$).
+  \item $D_1(s,t)$ for states $s$ and $t$ expanded in $\S$,  true iff
+    selected features distinguish $s$ from $t$; i.e.\ if $s$ and $t$ disagree
+    on the truth value of a selected boolean feature $p$ or atom $n=0$ for a
+    selected numerical feature $n$.
+    %Since $D_1$ is symmetric, only symbols for $s < t$ are created for some
+    %arbitrary, fixed order over the states in $\S_1$.
+  \item $D_2(s, s', t, t')$ for each $(s, s')$ and $(t, t')$ in $\S$,
+    true iff selected features $f$ distinguish the two transitions; i.e., if
+    $\Delta_f(s, s')\not=\Delta_f(t,t')$. 
+    %As for $D_1$, the number of $D_2$ variables is cut in half due to symmetries.
+\end{enumerate}
 
 \medskip
-
-\noindent The  first  formulas in $T({\cal S,\cal F})$ capture the meaning of $D_1$:
+\noindent The first formulas in $T(\S,\F)$ capture the meaning of $D_1$:
 
 \begin{equation}
-  D_1(s, t) \Leftrightarrow \bigvee_{f} \, selected(f)
-  \label{d1}
+  \label{eq:d1}
+  D_1(s, t) \ \Leftrightarrow\ \textstyle \bigvee_{f}  selected(f)
 \end{equation}
 
-\noindent where $f$ ranges over features in $\mathcal{F}$ with  different qualitative values in $s$ and $t$; i.e.
-boolean features $p$ with different truth value in $s$ and $t$, and numerical features $n$ with different truth values
-for atom $n=0$ in $s$ and $t$.
+\noindent where $f$ ranges over features in $\F$ with  different qualitative
+values in $s$ and $t$; i.e.\ boolean features $p$ with different truth value
+in $s$ and $t$, and numerical features $n$ with different truth values for
+atom $n=0$ in $s$ and $t$.
 
@@
 \begin{equation}
-  D_2(s, s', t, t') \Leftrightarrow \bigvee_f \,  selected(f)
+  \label{eq:d2}
+  D_2(s, s', t, t') \ \Leftrightarrow\ \textstyle\bigvee_f  selected(f)
 \end{equation}
 
-\noindent where $f$  ranges over features $f \in \mathcal{F}$ that have the same qualitative values in $s$ and $t$
-but which change differently in the two transitions; i.e.,  $\Delta_f(s, s') \neq \Delta_f(t, t')$.
-
-The third  class of formulas relate $D_1$ and $D_2$ by  enforcing  \emph{soundness} and \emph{completeness} over  the samples in ${\cal S}$.
-Yet, since the abstract actions  are not given but are derived, the qualitative changes in each  transition $(s,s')$ in ${\cal S}$
+\noindent where $f$ ranges over features $f \in \F$ that have the same
+qualitative values in $s$ and $t$ but which change differently in the two
+transitions; i.e., $\Delta_f(s, s') \neq \Delta_f(t, t')$.
+
+The third class of formulas relate $D_1$ and $D_2$ by enforcing \emph{soundness}
+and \emph{completeness} over  the samples in $\S$.
+Yet, since the abstract actions  are not given but are derived, the qualitative changes in each  transition $(s,s')$ in $\S$
 are  taken as templates  of abstract actions. Thus, if $s$ and $t$ are not distinguished by the selected features, 
-for each transition $(s, s') \in \mathcal{S}$, there must be a transition $(t, t') \in \mathcal{S}$
+for each transition $(s, s') \in \S$, there must be a transition $(t, t') \in \S$
 such that that the two transitions are not distinguished by the selected features either.
@@
 \begin{equation}
-  \neg D_1(s, t) \, \Rightarrow \, \bigvee_{t'} \neg D_2(s, s', t, t')
-  \label{bridge1}
+  \label{eq:bridge1}
+  \neg D_1(s, t) \  \Rightarrow\ \textstyle\bigvee_{t'} \neg D_2(s, s', t, t')
 \end{equation}
 
-\noindent where $s$ and $t$ are expanded states in  ${\cal S}$, $s'$ is a successor of $s$ in ${\cal S}$,
-and $t'$ ranges over the successors of $t$ in ${\cal S}$.
-
-%%% --
+\noindent where $s$ and $t$ are expanded states in  $\S$, $s'$ is a successor of $s$ in $\S$,
+and $t'$ ranges over the successors of $t$ in $\S$.
+
 \Omit{ % left out: this is for efficiency only, and is not strictly needed. Also, doesn't fit as such with goal-marking
@@
 
-  such that $s < t$, and transition $(s,s')$ in ${\cal S}$. Due to the symmetry breaking caused by the constraint $s < t$,
+  such that $s < t$, and transition $(s,s')$ in $\S$. Due to the symmetry breaking caused by the constraint $s < t$,
   the following formulas are also needed
@@
 \begin{equation}
-  \neg D_1(s, t) \, \Rightarrow \, \bigvee_{s'} \neg D_2(s, s', t, t')
-    \label{bridge2}
+  \label{eq:bridge2}
+  \neg D_1(s, t) \, \rightarrow \, \bigvee_{s'} \neg D_2(s, s', t, t')
 \end{equation}
@@
 \noindent where the  iteration in the right hand disjunction goes over the different successor states $s'$ of the first argument
-$s$ of $D_1$ that are in the sample set ${\cal S}$, $s$ and $t$ are states in ${\cal S}_1$ as before with $s < t$, and $t'$ is a
-successor state of $t$ in ${\cal S}$; i.e. $(t,t') \in {\cal S}$.
+$s$ of $D_1$ that are in the sample set $\S$, $s$ and $t$ are states in $\S_1$ as before with $s < t$, and $t'$ is a
+successor state of $t$ in $\S$; i.e. $(t,t') \in \S$.
 }
@@
 
-The last  set of clauses force the selected features to distinguish goal from non-goal states as:
-
+The last set of clauses force the selected features to distinguish goal
+from non-goal states as:
 \begin{equation}
+  \label{eq:goal}
   D_1(s,t) 
-\label{goal}
 \end{equation}
 
-\noindent where  $s$ and $t$  are expanded states  ${\cal S}$ such that exactly one of them is a goal state. 
-For this, it is assumed that  states $s$ in   tuples $(s,s',P)$ are labeled as   goal states  of $P$ or non-goals.
-
-The theory $T({\cal S},{\cal F})$ given by the formulas  (\ref{d1})--(\ref{goal})
-has a number of atoms that is linear in the size of the feature pool $\cal F$,
-and quadratic in the number of expanded  states $s$  and transitions $(s,s')$ in ${\cal S}$.
-The resulting number of clauses is ....  , which can be roughly cut in half by  symmetry considerations.
-
-In the experiments we will consider also a smaller theory  that avoids the consideration of all pairs
-of expanded states or all pairs of transitions in $\cal S$. For this, some expanded states $s$ and transitions $(s,s')$ in
-tuples $(s,s',P)$ are marked as \emph{goal relevant}. This is achieved by computing some plans for some instances $P$ in $\Q$, 
-and by adding the resulting state transitions to ${\cal S}$ at the same time that the first
-arguments $s$ and $(s,s')$ in the  $D_1$ and $D_2$ atoms are restricted to be marked.
+\noindent where  $s$ and $t$  are expanded states  $\S$ such that exactly one of them is a goal state. 
+For this, it is assumed that expanded states $s$ in $\S$ are labeled as goal or non-goal states.
+
+The theory $T(\S,\F)$ given by the formulas  (\ref{eq:d1})--(\ref{eq:goal})
+has a number of atoms that is linear in the size of the feature pool $\F$,
+and quadratic in the number of expanded  states $s$  and transitions $(s,s')$ in $\S$.
+The resulting number of clauses is \textcolor{red}{\bf XXXX}, which can be
+roughly cut in half by  symmetry considerations.
+
+In the experiments we also consider a smaller theory that avoids the
+consideration of all pairs of expanded states or all pairs of transitions
+in $\S$. For this, some expanded states $s$ and transitions $(s,s')$ in
+tuples $(s,s',P)$ are marked as \emph{goal relevant}. This is achieved by
+computing some plans for some instances $P$ in $\Q$, and by adding the
+resulting state transitions to $\S$ at the same time that the first
+arguments $s$ and $(s,s')$ in the $D_1$ and $D_2$ atoms are restricted
+to be marked.
 If the set of marked states and transitions is much smaller than the set of unmarked ones, 
-then the resulting formula,  $T_G({\cal S},{\cal F})$ will be much smaller than $T({\cal S},{\cal F})$.
-The goal marking scheme that involves the theory $T_G({\cal S},{\cal F})$ in place of $T({\cal S},{\cal F})$
-amounts to enforcing soundness on  the sample set ${\cal S}$ but not completeness, yet this does
+then the resulting formula,  $T_G(\S,\F)$ will be much smaller than $T(\S,\F)$.
+The goal marking scheme that involves the theory $T_G(\S,\F)$ in place of $T(\S,\F)$
+amounts to enforcing soundness on  the sample set $\S$ but not completeness, yet this does
 not affect the resulting  formal guarantees that depend on soundness only.
@@
 
-If  $\sigma$ is  a satisfying assignment of the theory $T({\cal S},{\cal F})$, 
-we define $F_{\sigma}$ as the set of  features $f$ in $\cal F$ such that $selected(f)$ is true in $\sigma$,
-and $A_{\sigma}$ as the set of abstract actions that capture  the  transitions $(s,s')$ in ${\cal S}$
-with duplicates and redundancies removed. The abstract action $\bar{a}=\abst{Pre}{\Eff}$
-that captures the transition $(s,s')$ has the \emph{precondition} $p$ (resp. $\neg p$)
-if $p$ is a boolean feature in $F_{\sigma}$ that is true (resp. false) in $s$, and has
-the precondition $n=0$ (resp. $n > 0$) if $n$ is a numerical feature in $F_{\sigma}$
-such that $n=0$ is true (resp. false) in $s$. Similarly, $\bar{a}$ has the \emph{effect}
-$p$ (resp. $\neg p$) if $p$ is a boolean feature in $F_{\sigma}$ that is true (resp. false)
-in $s'$ but false (resp. true) in $s$, and the effect $n\mminus$ (resp. $n\pplus$) if $n$ is a numerical feature in $F_{\sigma}$
-whose values decreases (resp. increases) in the transition from $s$ to $s'$. 
-If two abstract actions $\bar{a}$ and $\bar{a}$ differ only in the sign of a precondition,
-the precondition can be dropped, removing the duplicates and iterating until  no more simplifications
-can be done. They key properties are:
+For a satisfying assignment $\sigma$ of the theory $T(\S,\F)$, let
+$F_{\sigma}$ be the set of features $f$ in $\F$ such that $selected(f)$
+is true in $\sigma$, and let $A_{\sigma}$ be the set of abstract actions
+that capture the transitions $(s,s')$ in $\S$, with duplicates and redundancies removed.
+The abstract action $\bar{a}=\abst{Pre}{\Eff}$ that captures the transition $(s,s')$
+has the \emph{precondition} $p$ (resp.\ $\neg p$) if $p$ is a boolean feature in
+$F_{\sigma}$ that is true (resp.\ false) in $s$, and has the precondition $n=0$
+(resp.\ $n > 0$) if $n$ is a numerical feature in $F_{\sigma}$ such that $n=0$
+is true (resp.\ false) in $s$. Similarly, $\bar{a}$ has the \emph{effect} $p$
+(resp.\ $\neg p$) if $p$ is a boolean feature in $F_{\sigma}$ that is true
+(resp.\ false) in $s'$ but false (resp.\ true) in $s$, and the effect $n\mminus$
+(resp.\ $n\pplus$) if $n$ is a numerical feature in $F_{\sigma}$ whose values
+decreases (resp.\ increases) in the transition from $s$ to $s'$. 
+If two abstract actions $\bar{a}$ and $\bar{a}$ differ only in the sign of a
+precondition, the precondition can be dropped, removing the duplicates and
+iterating until no more simplifications can be done. They key properties are:
 
 \begin{theorem}
-  The theory $T({\cal S},{\cal F})$ is satisfiable iff there is a set of features $F \subseteq {\cal F}$
-  and a set of abstract actions $A_F$ over $F$ such that $A_F$ is sound and complete relative to  ${\cal S}$.
+  The theory $T(\S,\F)$ is satisfiable iff there is a set of features $F \subseteq \F$
+  and a set of abstract actions $A_F$ over $F$ such that $A_F$ is sound and complete relative to  $\S$.
 \end{theorem}
@@
 \begin{theorem}
-  If $\sigma$ is a satisfying assigment of $T({\cal S},{\cal F})$, then the abstraction actions in $A_{\sigma}$ 
-  are  sound and complete  relative to  ${\cal S}$.
+  If $\sigma$ is a satisfying assigment of $T(\S,\F)$, the set of abstract
+  actions $A_{\sigma}$ is sound and complete relative to $\S$.
 \end{theorem}
@@
 \begin{theorem}
-  If $\sigma$ satisfies  $T_G({\cal S},{\cal F})$, then $A_{\sigma}$   are  sound  relative to  ${\cal S}$.
+  If $\sigma$ satisfies  $T_G(\S,\F)$, $A_{\sigma}$  is sound relative to $\S$.
 \end{theorem}
@@
 correspond to the abstract states over the selected features in $F_{\sigma}$ 
-that are false in all expanded states in ${\cal S}$ marked as non-goals
+that are false in all expanded states in $\S$ marked as non-goals
 and true in some expanded goal state. This DNF formula can   be simplified in standard ways. 
-Due to  (\ref{d1}) and (\ref{goal}) that force the selected features in $F_{\sigma}$
+Due to  (\ref{eq:d1}) and (\ref{eq:goal}) that force the selected features in $F_{\sigma}$
 to   distinguish goal  from non-goal states, we have that:
@@
 \begin{theorem}
-For a satisfying assigment $\sigma$  of $T({\cal S},{\cal F})$, 
-$s$ is expanded goal   state in ${\cal S}$  iff  $s$ satisfies  $G_F=G_{\sigma}$.
+For a satisfying assigment $\sigma$  of $T(\S,\F)$, 
+$s$ is expanded goal   state in $\S$  iff  $s$ satisfies  $G_F=G_{\sigma}$.
 \end{theorem}
 }
-
-Different  satisfying  assignments $\sigma$,  give rise to different  features $F_{\sigma}$
+Different satisfying assignments $\sigma$ give rise to different features $F_{\sigma}$
 and abstract actions $A_{\sigma}$. More meaningful features are found by looking for
 satisfying assignments $\sigma$ that minimizes a cost measure such as $|F_{\sigma}|$
-or  more generally $\sum_{f \in F_{\sigma}} cost(f)$. We will  achieve this  constrained
-optimization through the use  Min SAT or Max-Weighted SAT solvers,  where it suffices to give
-infinite weights  to the clauses resulting from $T({\cal S},{\cal F})$ and weights given by $cost(f)$ 
-to the ``soft''  unit clauses  $\neg selected(f)$.
+or more generally $\sum_{f \in F_{\sigma}} cost(f)$. We achieve this constrained
+optimization through the use Min SAT or Max-Weighted SAT solvers, where it suffices
+to give infinite weights to the clauses resulting from $T(\S,\F)$ and weights
+given by $cost(f)$ to the ``soft'' unit clauses $\neg selected(f)$.
 
@@
 
-The computation of the feature pool $\cal F$ used in $T({\cal S},{\cal F})$
+The computation of the feature pool $\F$ used in $T(\S,\F)$
 relies on the predicates used to encode the instances in $\Q$.
@@
 The key issue is how to define new predicates from old  ones. For this 
-we take a solution off-the-shelf: \emph{description logics (DL)},  called  also
+we take a solution off-the-shelf: \emph{description logics (DL)},  also called
 \emph{concept languages},  are grammars for doing exactly this.
@@
 Concept languages have been used before for  generating general policies using purely learning methods  \cite{martin:concept,fern:bias}.
-For details on the syntax and semantics of description logics; see \cite{dl-handbook}.
+\citeay{dl-handbook} give syntax and semantics of description logics.
 
@@
 
-The DL grammar is given by the rules:
-
-  \begin{itemize}
-  \item $C \ \leftarrow\ \ C_p, C_u, C_x$: primitive, universal,  nominal c's,\footnote{Denotation of $C_x$ for parameter $x$ is $\{x\}$.}
+The DL grammar is given by the rules: \textcolor{red}{\bf **** More details?}
+\begin{enumerate}[{\small$\bullet$}]
+  \item $C \ \leftarrow\ \ C_p, C_u, C_x$: primitive, universal, and nominal c's,\footnote{Denotation of $C_x$ for parameter $x$ is $\{x\}$.}
   \item $C \ \leftarrow\  \neg C, C \sqcap C'$: negation and conjunction, 
@@
 %     \item $R \leftarrow R:C$, denotes pairs $(x,y)$ in $R$ s.t. $C(y)$.
- \end{itemize}
+ \end{enumerate}
 
@@
 
-The set of all concepts and roles with complexity no larger than $k$ is denoted as ${\cal G}^k$,
-with their complexity given by the minimum number of rules for generating them.
-The size  of this set is reduced by incrementally pruning ``duplicate'' concepts,
-those whose denotation coincides with the denotation of a another concept over all
-the sampled states in ${\cal S}$.
-
-The pool of features ${\cal F}={\cal F}^k$  is defined from ${\cal G}^k$
-and contains the  features $f$ that represent the functions $\phi_f(s)$
-as follows:
-
-\begin{itemize} 
+The set of all concepts and roles with complexity no larger than $k$ is
+denoted as $\G^k$. The complexity of a concept or role is the minimum
+number of rules needed to generate it. The size of $\G^k$ is reduced by
+incrementally pruning ``duplicate'' concepts; namely, those whose denotation
+coincides with the denotation of another concept over all the sampled states.
+%in $\S$.
+
+The pool $\F=\F^k$  is defined from $\G^k$ and contains
+the  features $f$ that represent the functions $\phi_f(s)$ as follows:
+\begin{enumerate}[{\small$\bullet$}]
+\item Boolean $b_p$ for each primitive predicate $p$ of arity $0$,
+\item Boolean $b_C$ for each $C$ that are true in $s$ iff $|C^s| > 0$,
 \item Numerical $n_C$ for each $C$ that stand for cardinality $|C^s|$,
-\item Boolean  $b_C$ for each   $C$ that are  true in $s$ iff $|C^s| > 0$,
-\item Boolean   $b_p$ for each primitive predicate $p$ of arity $0$,
-\item Numerical $\textit{dist}(C_1,R:C,C_2)$,  that represent that smallest $n$ such that for $x_1, \ldots, x_n$, 
-$C_1^s(x_1)$, $C_2^s(x_{n})$, and $R:C(x_i,x_{i+1})$ for $i=1, \ldots, n$. 
-\end{itemize}
-
-The role $R:C$ is such that its denotation contains the pairs $(x,y)$ in $R$ such that $y$ is in $C$.
-Some pruning is done in this space of features as well. E.g., numerical features $n_C$ are pruned from the pool
-when there are  no states $s$ and $s'$ in the sample set for which $n_c(s)=0$ and $n_c(s') > 1$,
-while boolean features $n_b$  are pruned if uniformly true or false for all $s$, or if $n_C$ not pruned. 
-The measure  $cost(f)$ is set to the complexity of $C$ for $n_C$ and $b_C$, to $0$ for $b_p$
-and to the complexities of $C_1$, $R$, $C$, and $C_2$ added up for $\textit{dist}(C_1,R:C,C_2)$.
-Only features with cost bounded by $k$ are allowed in $\cal F$.
+\item Numerical $\textit{dist}(C_1,R:C,C_2)$, that represent that smallest
+  $n$ such that for $x_1, \ldots, x_n$, $C_1^s(x_1)$, $C_2^s(x_{n})$, and
+  $(R:C)(x_i,x_{i+1})$ for $i=1, \ldots, n$. 
+\end{enumerate}
+
+The role $R:C$ is such that its denotation contains the pairs $(x,y)$ in $R$
+such that $y$ is in $C$.
+Some pruning is done in this space of features as well. E.g., numerical
+features $n_C$ are pruned from the pool when there are no states $s$ and $s'$
+in the sample set for which $n_c(s)=0$ and $n_c(s') > 1$, while boolean
+features $n_b$ are pruned if uniformly true or false for all $s$, or if $n_C$
+not pruned. 
+The measure $cost(f)$ is set to the complexity of $C$ for $n_C$ and $b_C$,
+to $0$ for $b_p$ and to the complexities of $C_1$, $R$, $C$, and $C_2$ added
+up for $\textit{dist}(C_1,R:C,C_2)$.
+Only features with cost bounded by $k$ are allowed in $\F$.
 
@@
 In summary, the steps for computing general plans are:
-
 \begin{enumerate}
-\item a sample set ${\cal S}$ is computed from a few  instances $P$ of  a given generalized problem $\Q$ 
-\item a pool of features $\cal F$ is obtained from the predicates used in the instances $P$, the  grammar, and bound $k$,
-\item a satisfying assignment $\sigma$ of the theory $T({\cal S},{\cal F})$ or $T_G({\cal S},{\cal F})$
-that minimizes  $\sum_{f \in F_{\sigma}} cost(f)$ is obtained using a Weighted-MAX SAT solver,
-\item the features $F$ and abstract actions $A_F$  are extracted from $\sigma$ as $F_{\sigma}$ and $A_{\sigma}$,
-\item the problem $Q_F=\tup{V_F,I_F,G_F,A_F}$ is  defined with initial and goal conditions $I_F$ and $G_F$ provided by hand
-to match $\Q$, 
-\item the FONDP  $Q^+_F$, constructed from $Q_F$,  is solved by a FOND planner off-the-shelf
+  \item a sample set $\S$ is computed from a few instances $P$ of a given
+    generalized problem $\Q$,
+  \item a pool of features $\F$ is obtained from the predicates used in the
+    instances $P$, the grammar, the bound $k$, and the sample $\S$ (used for
+    pruning concepts and roles),
+  \item a satisfying assignment $\sigma$ of the theory $T(\S,\F)$ or $T_G(\S,\F)$
+    that minimizes $\sum_{f \in F_{\sigma}} cost(f)$ is obtained using a
+    Weighted-MAX SAT solver,
+  \item the features $F$ and abstract actions $A_F$ are extracted from $\sigma$
+    as $F_{\sigma}$ and $A_{\sigma}$,
+  \item the problem $Q_F=\tup{V_F,I_F,G_F,A_F}$ is defined with initial and goal
+    conditions $I_F$ and $G_F$ provided by hand to match $\Q$, 
+  \item the FONDP $Q^+_F$, constructed from $Q_F$, is solved by an off-the-shelf
+    FOND planner.
 \end{enumerate}
 
-Note that the  policy $\pi$ that results from the last step  deals with propositional symbols
-that correspond to atoms $p$ and $n=0$ for boolean and numerical features $p$ and $n$ in $F$. 
+Notice that the policy $\pi$ that results from the last step deals with
+propositional symbols that correspond to atoms $p$ and $n=0$ for boolean
+and numerical features $p$ and $n$ in $F$.
 When this policy is applied to an instance $P$ of $\Q$, however, such features denote state functions
@@
 
-If we write $A \models B$ to express that all the states that satisfy $A$, satisfy $B$, 
-then the formal guarantees can be expressed as:
+If we write $A \vDash B$ to express that all the states that satisfy $A$, satisfy $B$, 
+the formal guarantees can be expressed as:
 
 \begin{theorem}
-If the abstract actions $A_F$ that are sound relative to the sample set $\cal S$  are actually sound,
+If the abstract actions $A_F$ that are sound relative to the sample set $\S$  are actually sound,
 then a   policy $\pi$ that solves  $Q^+_F$  is guaranteed to solve  all instances $P$ of $\Q$
-with initial and goal situations $I$ and $G$ s.t. $I \models  I_F$ and $G_F \models G$.
+with initial and goal situations $I$ and $G$ such that $I \vDash  I_F$ and $G_F \vDash G$.
 \end{theorem}
@@
 This means that a policy $\pi$ obtained with a FOND planner from $Q^+_F$ 
-solves solves the  generalized problem $\Q$, provided that the  abstract
-actions learned  are  sound and  that $\Q$ is restrictedto the instances
-$P$ whose initial and goal conditions comply with the conditions on $I_F$ and $G_F$.
-Since we don't have a language for specifying $\Q$ formally, we can take the formulas
-$I_F$ and $G_F$ indeed as a partial formal specification of $\Q$. The theorem then says that
-if soundness over the samples, ensure formal soundness over $\Q$, the policy that solves the
-FONDP $Q^+_F$, solves $\Q$. The result holds whether $T$ or $T_G$ (goal marking) is used in step 3.
+solves the generalized problem $\Q$, provided that the abstract actions
+learned are sound and that $\Q$ is restricted to the instances $P$ whose
+initial and goal conditions comply with the conditions on $I_F$ and $G_F$.
+\textcolor{red}{\bf ** NEXT IS UNCLEAR **}
+Since we do not have a language for specifying $\Q$ formally, we can take
+the formulas $I_F$ and $G_F$ indeed as a partial formal specification of
+$\Q$. The theorem then says that if soundness over the samples, ensure
+formal soundness over $\Q$, the policy that solves the FONDP $Q^+_F$,
+solves $\Q$. The result holds whether $T$ or $T_G$ (goal marking) is used in step 3.
 In principle, we could learn the formulas $I_F$ and $G_F$ from the sample as well,  but with different guarantees.
@@
 % *Fill up general details in 1--6 that don't change from problem to problem
-%how ${\cal S}$ obtained
+%how $\S$ obtained
 The sample set $\mathcal{S}$ is obtained following one of two possible strategies.
 In the \emph{exhaustive} sampling strategy,
-$\mathcal{S}$ contains, for every one of the training instances $P$,
+$\S$ contains, for every one of the training instances $P$,
 all the states generated on a breadth-first search from the initial state in $P$
 that is run until $m$ expansions have been performed. If that set includes no goal state,
-an optimal plan for $P$ is computed and all of the states along the plan are included in $\mathcal{S}$ as well,
+an optimal plan for $P$ is computed and all of the states along the plan are included in $\S$ as well,
 as we need at least one goal state in the sample set.
@@
 one random goal state from the larger set is drawn at random and added into the sample.
+\textcolor{red}{\bf **** What if $M$ doesn't have a goal? ****}
 
@@
 
-All the results below are with goal marking using the theory $T_G({\cal S},{\cal F})$ that is
-smaller than $T({\cal S,F})$ and produces smaller abstract action sets  $A_F$, which are forced
-to be sound relative to the sample $\mathcal{S}$ but not complete, thus 
-avoiding abstract actions that may make  planning harder and but are not relevant to the goal.
-In general, the theory $T({\cal S},{\cal F})$ results in a significantly larger number of atoms and clauses,
-as seen in \ref{tab:exp-result-data}, which makes the MAXSAT solving process much harder.
-
-The goal marking scheme that  yields the smaller  theory $T_G({\cal S},{\cal F})$ is obtained an optimal plan over
+The goal marking scheme that  yields the smaller  theory $T_G(\S}, \F)$ is obtained an optimal plan over
 a few (always less than xxxx) instances $P$. ??
@@
 
- & $I$ & $\abs{\mathcal{S}}$ & $k$ & $\abs{\mathcal{F}^k}$ & \multicolumn{2}{c}{$T$} & \multicolumn{2}{c}{$T_G$} &  $C_{\text{SAT}}$ & $\abs{F}$ & $\abs{A_F}$ & $\abs{I_F}$  & $\abs{G_F}$ & $\abs{Q^+_F}$ & $C_{\text{FOND}}$ \\ %
+ & $I$ & $\abs{\S}$ & $k$ & $\abs{\F^k}$ & \multicolumn{2}{c}{$T$} & \multicolumn{2}{c}{$T_G$} &  $C_{\text{SAT}}$ & $\abs{F}$ & $\abs{A_F}$ & $\abs{I_F}$  & $\abs{G_F}$ & $\abs{Q^+_F}$ & $C_{\text{FOND}}$ \\ %
 
@@
   $I$: number of training instances; 
-  $\abs{\mathcal{S}}$: sample set size, in total number of transitions; 
-  $k$: max. feature comlexity; $\abs{\mathcal{F}^k}$: size of the pool of features;
-  $\#V$, $\#C$: Number of variables and clauses of the theories $T({\cal S},{\cal F})$ and $T_G({\cal S},{\cal F})$;
+  $\abs{\S}$: sample set size, in total number of transitions;
+  $k$: max. feature comlexity; $\abs{\F^k}$: size of the pool of features;
+  $\#V$, $\#C$: Number of variables and clauses of the theories $T(\S,\F)$ and $T_G(\S,\F)$;
   $C_{\text{SAT}}$, $C_{\text{FOND}}$: computation time (sec.) of the SAT and FOND solvers, resp.;
@@
   $\abs{Q^+_F}$: ??? .
-  All figures that depend on the SAT theory are given for $T_G({\cal S},{\cal F})$.
+  All figures that depend on the SAT theory are given for $T_G(\S,\F)$.
 }}
@@
 The generalized problem $\Q_{clear}$ aims at achieving the goal $clear(x)$ for any block $x$
-in the Blocks world. 
+in the Blocksworld.
 We take the standard STRIPS representation with primitive predicates $on(a, b)$, $clear(b)$, $ontable(b)$, $holding(b)$ and $handempty$.
@@
 The problem $\Q_{on}$ is aims at achieving a goal of the form $on(x,y)$
-in the Blocks world for any two blocks $x$ and $y$ that are initially not in the same tower.
+in the Blocksworld for any two blocks $x$ and $y$ that are initially not in the same tower.
 We take the same STRIPS representation as above. Based on 3 different instances with up to 14
@@
 
-We have introduced a scheme for computing general plans that mixes learning and planning: a learner infers
-a general abstraction made up of features and abstract actions by enforcing  a notion of soundness and completeness
-over a suitable set of samples, which is transformed and fed into a FOND planner.
-The number of samples required for obtaining correct general plans is very small,
-as the learner doesn't have to learn how to plan,  but just to identify the
-general features that are relevant for the planner. Moreover, unlike, purely
-learning approaches, the scope and correctness of the resulting general plans
-can be assessed.  We have carried out a number of experiments over several
-domains illustrating the versatility of the approach. In the future, we
-want to explore the use of the learned  features and actions
-in classical planning, the study of  formal characterizations
-of the generalized problem $\Q$ and  the possibilities
-of testing soudness over $\Q$ automatically, and ways
-for improving the scalability of the approach, for example,
-by collecting samples intelligently and incrementally,
-from the states where the computed  policy fails,
-and by reusing the features and actions obtained
-for some goal for more complex goals.
+We have introduced a scheme for computing general plans that mixes
+learning and planning: a learner infers a general abstraction made
+up of features and abstract actions by enforcing a notion of soundness
+and completeness over a suitable set of samples, which is transformed
+and fed into a FOND planner.
+The number of samples required for obtaining correct general plans is
+very small, as the learner does not have to learn how to plan, but
+just to identify the general features that are relevant for the planner.
+Moreover, unlike, purely learning approaches, the scope and correctness
+of the resulting general plans can be assessed.
+We have carried out a number of experiments over several domains to 
+illustrate the versatility of the approach. In the future, we want to
+explore the use of the learned features and actions in classical
+planning, to study of formal characterizations of the generalized
+problem $\Q$ and the possibilities of testing soudness over $\Q$
+automatically, and to improve the scalability of the approach,
+for example, by collecting samples intelligently and incrementally,
+from the states where the computed policy fails, and by reusing the
+features and actions obtained for some goal for more complex goals.
 
 \Omit{
-
 ** discussion  of experiments; things we do well and not, scope, lessons. limitations (expressive? scalability?)
-
 ** briefly related work, expanding paragraph in intro.
-
 ** summary and cconlusions

### 2018-09-05 doc/paper.tex
Draft sent to H

@@
 \newcommand{\Example}{\medskip\noindent\textbf{Example.}\xspace}
+
+\renewcommand\labelitemi{\raisebox{0.4ex}{\tiny$\bullet$}} %% Decrease size of items in enumerate and raise them to center them vertically
 
@@
 and places it  on the table, achieves the goal $clear(x)$
-in \emph{any} instance of the blocks world where the gripper is initially empty.
+in \emph{any} instance of the Blocksworld where the gripper is initially empty.
 Once this  general plan or  policy is derived it can be  applied to solve an infinite collection
@@
 The \emph{propositional variables} in $T(\S,\F)$ are then:
-\begin{enumerate}[{\small$\bullet$}]
+\begin{itemize}
   \item $selected(f)$ for each $f \in \F$, true iff $f$ selected (in $F$).
@@
     %As for $D_1$, the number of $D_2$ variables is cut in half due to symmetries.
-\end{enumerate}
+\end{itemize}
 
@@
 The DL grammar is given by the rules: \textcolor{red}{\bf **** More details?}
-\begin{enumerate}[{\small$\bullet$}]
+\begin{itemize}
   \item $C \ \leftarrow\ \ C_p, C_u, C_x$: primitive, universal, and nominal c's,\footnote{Denotation of $C_x$ for parameter $x$ is $\{x\}$.}
@@
 %     \item $R \leftarrow R:C$, denotes pairs $(x,y)$ in $R$ s.t. $C(y)$.
- \end{enumerate}
+ \end{itemize}
 
@@
 the  features $f$ that represent the functions $\phi_f(s)$ as follows:
-\begin{enumerate}[{\small$\bullet$}]
+\begin{itemize}
 \item Boolean $b_p$ for each primitive predicate $p$ of arity $0$,
@@
   $(R:C)(x_i,x_{i+1})$ for $i=1, \ldots, n$. 
-\end{enumerate}
+\end{itemize}
 
@@
 
-We evaluate the computational model described in steps 1--6 above over several domains.
-%
-
+We evaluate the computational model described in steps 1--6 above over 4 different domains.
+%
 % *Fill up general details in 1--6 that don't change from problem to problem
 %how $\S$ obtained
-The sample set $\mathcal{S}$ is obtained following one of two possible strategies.
-In the \emph{exhaustive} sampling strategy,
-$\S$ contains, for every one of the training instances $P$,
-all the states generated on a breadth-first search from the initial state in $P$
-that is run until $m$ expansions have been performed. If that set includes no goal state,
-an optimal plan for $P$ is computed and all of the states along the plan are included in $\S$ as well,
-as we need at least one goal state in the sample set.
-%
-In the \emph{random} sampling strategy, a large number $M$ of states are generated, potentially exhausting the full state space,
-and then a certain number $m \ll M$ of them is sampled uniformly at random. If no goal state is found among those $m$,
-one random goal state from the larger set is drawn at random and added into the sample.
-\textcolor{red}{\bf **** What if $M$ doesn't have a goal? ****}
+The sample set $\S$
+% s obtained following one of two possible strategies.
+% In the \emph{exhaustive} sampling strategy, $\S$
+contains, for every one of the training instances $P$ in $\Q$,
+the first $m$  states expanded by a standard breadth-first search from the initial state in $P$,
+plus all the states along an optimal plan $\pi^*$ which is separately computed to ensure that $\S$ 
+contains at least one goal state. 
+All successor states of states in $\S$ are also added to $\S$.
+%
+For the goal marking scheme that yields the smaller theory $T_G(\S, \F)$, the states that are marked as
+goal-relevant are those in $\pi^*$.
+
+%
+% In the \emph{random} sampling strategy, a large number $M$ of states are generated, potentially exhausting the full state space,
+% and then a certain number $m \ll M$ of them is sampled uniformly at random. If no goal state is found among those $m$,
+% one random goal state from the larger set is drawn at random and added into the sample.
+% \textcolor{red}{\bf **** What if $M$ doesn't have a goal? ****}
 
@@
 
-** Details on the FOND planner, QNP translation, etc. **
+\alert{Details on the FOND planner, QNP translation, etc.}
 
@@
 
-The goal marking scheme that  yields the smaller  theory $T_G(\S}, \F)$ is obtained an optimal plan over
-a few (always less than xxxx) instances $P$. ??
-
 \begin{table*}[t]
@@
 
-\newcommand{\rowspacing[1]}{&&&&&&&&&&&&&&&\\[\dimexpr-\normalbaselineskip+2pt]}
-
-\begin{tabular}{%
-@{}l%
-@{\extracolsep{1.5em}}r%@{}l%
-@{\extracolsep{1.5em}}r%
-@{\extracolsep{8pt}}r%
-%
-@{\extracolsep{1.5em}}r%
-@{\extracolsep{8pt}}r%
-@{\extracolsep{8pt}}r%
-@{\extracolsep{8pt}}r%
-%
-@{\extracolsep{1.5em}}r%
-@{\extracolsep{8pt}}r%
-@{\extracolsep{8pt}}r%
-@{\extracolsep{8pt}}r%
-%
-@{\extracolsep{1.5em}}r%
-@{\extracolsep{8pt}}r%
-@{\extracolsep{8pt}}r%
-@{\extracolsep{8pt}}r%
-%
-@{}}
-
+\newcommand{\rowspacing[1]}{&&&&&&&&&&&&&&&&\\[\dimexpr-\normalbaselineskip+2pt]}
+
+\begin{tabular}{rrrrrrrrrrrrrrrrr}
 % @{\extracolsep{3pt}}
 
- & $I$ & $\abs{\S}$ & $k$ & $\abs{\F^k}$ & \multicolumn{2}{c}{$T$} & \multicolumn{2}{c}{$T_G$} &  $C_{\text{SAT}}$ & $\abs{F}$ & $\abs{A_F}$ & $\abs{I_F}$  & $\abs{G_F}$ & $\abs{Q^+_F}$ & $C_{\text{FOND}}$ \\ %
+&$I$ & $\abs{\S}$ & $k$ & $\abs{\F^k}$ & \multicolumn{2}{c}{$T$} & \multicolumn{2}{c}{$T_G$} &  $T_{\text{SAT}}$ & $\abs{F}$ & $\abs{A_F}$ & $T_{\text{trans}}$  & \multicolumn{2}{c}{$Q^+_F$}  & $T_{\text{FOND}}$
+& $\abs{\pi}$ \\ %
 
 \rowspacing 
-\cline{6-7} \cline{8-9}%
+\cline{6-7} \cline{8-9} \cline{14-15}%
 \rowspacing
 
- & &     &      &         & $\#V$ & $\#C$           & $\#V$ & $\#C$             &     &&&&&& \\%
+ & &     &      &         & $\#V$ & $\#C$           & $\#V$ & $\#C$             &     &&&&  $\#Act$ & $\#At$&& \\%
 
@@
 \multirow{1}{*}{$\Q_{clear}$}
-& 1     &     927 & 8  & 322 & 535K & 59.6M & 7.7K & 767K   &  0.01 & 3 & 2   &  - & - & - & - \\% Total complexity of F: 8
+& 1     &     927 & 8  & 322 & 535K & 59.6M & 7.7K & 767K   &  0.01 & 3 & 2   &  - & - & - & - & - \\% Total complexity of F: 8
 
@@
 \multirow{1}{*}{$\Q_{on}$}
-& 3 &        420  & 8 & 657 & 128K & 25.8M & 18.3K &     3.3M    &  0.02&  5 & 7   &  - & - & - & - \\% Total complexity of F: 18
+& 3 &        420  & 8 & 657 & 128K & 25.8M & 18.3K &     3.3M    &  0.02&  5 & 7   &  - & - & - & - & -\\% Total complexity of F: 18
 % Max-sat problem has 128217 variables and 25825428 unique clauses (with repetitions: 25825428). Top weight is 4577
@@
 \multirow{1}{*}{$\Q_{gripper}$}    
-& 2 &         403 & 8 & 130 & 93K & 4.7M & 8.1K &   358K    &  0.01 & 4 & 5&  - & - & - & - \\% Total complexity of F: 10
+& 2 &         403 & 8 & 130 & 93K & 4.7M & 8.1K &   358K    &  0.01 & 4 & 5&  - & - & - & - & -\\% Total complexity of F: 10
 
@@
 \multirow{1}{*}{$\Q_{reward}$}
-& 2 &         568 & 8 & 280 & 184K & 11.9M & 15.9K & 1.2M   &  0.01 & 2 & 2 &  - & - & - & - \\% Total complexity of F: ?
+& 2 &         568 & 8 & 280 & 184K & 11.9M & 15.9K & 1.2M   &  0.01 & 2 & 2 &  - & - & - & - & -\\% Total complexity of F: ?
 \end{tabular}
@@
   $\#V$, $\#C$: Number of variables and clauses of the theories $T(\S,\F)$ and $T_G(\S,\F)$;
-  $C_{\text{SAT}}$, $C_{\text{FOND}}$: computation time (sec.) of the SAT and FOND solvers, resp.;
+  $T_{\text{SAT}}$, $T_{\text{FOND}}$: computation time (sec.) of the SAT and FOND solvers, resp.;
   $\abs{F}$: number of selected features;
   $\abs{A_F}$: number of abstract actions;
-  $\abs{I_F}$: ???;
-  $\abs{G_F}$: ??? ;
-  $\abs{Q^+_F}$: ??? .
+  $T_{\text{trans}}$: time (sec.) to generate the FOND problem $Q^+_F$;
+  $\#Act$, $\#At$: number if actions and atoms in $Q^+_F$;
+  $\abs{\pi}$: size of the policy.
+%  
   All figures that depend on the SAT theory are given for $T_G(\S,\F)$.
@@
 With one single instance $\Q_{clear}$ with 5 blocks, we are able to learn
-an abstract action model which features two actions:
-
-- \emph{pick-above-$x$}: \tuple{\neg X, \neg H, n(x) > 0; n(x) \downarrow},
-
-- \emph{place-aside}: \tuple{\neg X, H; \neg H},
-
-\noindent both described in terms of features $H: \abs{holding}$, $X: \abs{holding \sqcap \set{x}}$ and $n(x)\equiv \abs{\exists on^* . \set{x}}$, 
-that denote, respectively, whether any block is being held, whether block $x$ is being held,
-and the number of blocks above $x$.\footnote{
-Action and features names are assigned manually by us to increase readability, but correspond faithfully
-to the learnt entities.
+an abstract action space with features:
+
+\begin{itemize}
+\item $H: \abs{holding}$: whether any block is being held,
+\item $X: \abs{holding \sqcap \set{x}}$: whether block $x$ is being held,
+\item $n(x): \abs{\exists on^* . \set{x}}$: number of blocks above $x$.\footnote{
+Action and features names are assigned manually to increase readability, based on the denotations of the involved concepts.
 }
-%
-The model is provably correct, but not complete in general, as it lacks e.g.
-any action to pick block $x$.
-However, it is sufficient to obtain a policy with a single loop that will 
-alternatively \emph{pick-above-$x$} and \emph{place-aside} until block $x$ is clear,
-which provably \emph{achieves the goal from any configuration of blocks, for any number of blocks}.
-
-% Action 1:
+\end{itemize}
+
+\noindent The resulting action model has two actions:
+
+\begin{itemize}
+\item \emph{pick-above-$x$}: \tuple{\neg X, \neg H, n(x) > 0; n(x) \downarrow},
+\item \emph{place-aside}: \tuple{\neg X, H; \neg H},
+\end{itemize}
+
+\noindent 
+The model is provably correct with respect to the generalized $\Q_{clear}$, but incomplete, as it lacks e.g.
+an action to pick block $x$. However, it allows us to obtain the following policy from the FOND translation $Q^+_F$:
+
+\alert{Display policy here},
+
+\noindent which will alternatively \emph{pick-above-$x$} and \emph{place-aside} the picked block 
+until block $x$ is clear. This can be proven to \emph{achieve the goal from any starting block configuration, for any number of blocks} in 
+$\Q_{clear}$.
+
+% Action 1: PLACE-ASIDE
 %         PRE: NOT holding(a), bool[holding]
@@
 % 
-% Action 2:
+% Action 2: PICK-FROM-A
 %         PRE: NOT bool[holding], NOT holding(a), n(a) > 0
@@
 \paragraph{Stacking two blocks from different towers.}
-The problem $\Q_{on}$ is aims at achieving a goal of the form $on(x,y)$
-in the Blocksworld for any two blocks $x$ and $y$ that are initially not in the same tower.
-We take the same STRIPS representation as above. Based on 3 different instances with up to 14
-blocks, we are able to learn an abstract action model with 5 features and 7 actions.
-The set of features is made up of  $X$, $n(x)$ and $n(y)$, with the same syntax and semantics as given before,
-as well as $E: handempty$
-and $XY: [(\exists on^{-1} . \set{x})\sqcap \set{b}]$,
-the second of which precisely denotes that block $x$ is on block $y$.
-
-The set of actions includes actions to pick up the topmost block above $x$, to pick up the topmost block above $y$,
-to put a block being held \emph{aside}, i.e. not above $x$ nor above $y$, to pick $x$, 
-to place $x$ on $y$ and to place $x$ on the table.
-
-- \emph{pick-up-above-$x$}:
-
-- \emph{pick-up-above-$y$}:
+The problem $\Q_{on}$ aims at achieving a goal of the form $on(x,y)$
+in the Blocksworld for any two blocks $x$ and $y$.
+To simplify, we assume that $x$ and $y$ never start being in the same tower.
+We take the same STRIPS representation as above. From 3 different training instances with up to 14
+blocks, we are able to learn an abstract space based on the following 5 features:
+
+\begin{itemize}
+\item $X$, $n(x)$, $n(y)$: as in $\Q_{clear}$,
+\item $E: handempty$: whether gripper is empty,
+\item $XY: [(\exists on^{-1} . \set{x})\sqcap \set{b}]$: whether $x$ is on $y$.
+\end{itemize}
+
+\noindent Interestingly, the last feature is exactly a concept-based representation of the actual STRIPS goal $on(x,y)$.
+The corresponding action model $A_F$ is:
+
+\begin{itemize}
+\item \emph{pick-above-$x$}:  \tuple{\neg X, \neg XY, E, n(x) > 0, n(y)>0; n(x) \downarrow, \neg E  }
+\item \emph{pick-above-$y$}:  \tuple{\neg X, \neg XY, E, n(x) = 0, n(y)>0; n(y) \downarrow, \neg E  }
+\item \emph{put-$x$-on-$y$}:  \tuple{X, \neg XY, \neg E, n(x)=0, n(y)=0 ; n(y) \uparrow, \neg X, E, XY}
+\item \emph{pick-$x$}:  \tuple{;}
+\item \emph{put-$x$-on-table}:  \tuple{;}
+\item \emph{put-aside-1}:  \tuple{;}
+\item \emph{put-aside-2}:  \tuple{;}
+
+\end{itemize}
+
+% Action 1 [PUT-ASIDE-1]:
+%         PRE: NOT bool[handempty], NOT holding(a), NOT on(a, b)_2, n(a) = 0
+%         EFFS: bool[handempty]
+% 
+% Action 2 [PUT-ASIDE-2]:
+%         PRE: NOT bool[handempty], NOT holding(a), NOT on(a, b)_2, n(a) > 0, n(b) > 0
+%         EFFS: bool[handempty]
+% 
+% Action 3 [PICK-A]:
+%         PRE: NOT holding(a), NOT on(a, b)_2, bool[handempty], n(a) = 0, n(b) = 0
+%         EFFS: NOT bool[handempty], holding(a)
+% 
+% Action 4 [PUT-A-ONTABLE]:
+%         PRE: NOT bool[handempty], NOT on(a, b)_2, holding(a), n(a) = 0, n(b) > 0
+%         EFFS: NOT holding(a), bool[handempty]
+% 
+% Action 5 [PICK-FROM-ABOVE-A]:
+%         PRE: NOT holding(a), NOT on(a, b)_2, bool[handempty], n(a) > 0, n(b) > 0
+%         EFFS: DEC n(a), NOT bool[handempty]
+% 
+% Action 6 [PICK-FROM-ABOVE-B]:
+%         PRE: NOT holding(a), NOT on(a, b)_2, bool[handempty], n(a) = 0, n(b) > 0
+%         EFFS: DEC n(b), NOT bool[handempty]
+% 
+% Action 7 [PUT-A-ON-B]:
+%         PRE: NOT bool[handempty], NOT on(a, b)_2, holding(a), n(a) = 0, n(b) = 0
+%         EFFS: INC n(b), NOT holding(a), bool[handempty], on(a, b)_2
+
+\alert{ongoing}
 
@@
 
-DISCUSS POLICY WHEN AVAILABLE. / 
-DISCUSS WHY WOULDN'T WORK WHEN IN SAME TOWER.
+\alert{Display policy here}
 
@@
 
-- $C: \exists carry . \top$ (number of balls carried by the robot),
-
-- $X: at\_robby \sqcap \set{x}$ (whether robot in target room),
-
-- $G: \abs{free}$ (number of empty grippers), and
-
-- $B: \exists at . \neg \set{x}$ (number of balls not in target room),
+\begin{itemize}
+\item $C: \exists carry . \top$ (number of balls carried by the robot), 
+\item $X: at\_robby \sqcap \set{x}$ (whether robot in target room),
+\item $G: \abs{free}$ (number of empty grippers), and
+\item $B: \exists at . \neg \set{x}$ (number of balls not in target room),
+\end{itemize}
 
@@
 
-- \emph{drop-ball(s)-at-$x$}: \tuple{C>0, X;  C\downarrow, G \uparrow}
-
-- \emph{move-to-$x$-half-loaded}: \tuple{\neg X, B=0, C>0, G>0;   X}
-
-- \emph{move-to-$x$-fully-loaded}: \tuple{\neg X, C>0, G=0;   X}
-
-- \emph{pick-ball-not-in-$x$}: \tuple{\neg X, B > 0, G > 0;   B \downarrow, G \downarrow, C\uparrow}
-
-- \emph{drop-$x$}: \tuple{X, C=0, G > 0;   \neg X}
+\begin{itemize}
+\item \emph{drop-ball(s)-at-$x$}: \tuple{C>0, X;  C\downarrow, G \uparrow}
+\item \emph{move-to-$x$-half-loaded}: \tuple{\neg X, B=0, C>0, G>0;   X}
+\item \emph{move-to-$x$-fully-loaded}: \tuple{\neg X, C>0, G=0;   X}
+\item \emph{pick-ball-not-in-$x$}: \tuple{\neg X, B > 0, G > 0;   B \downarrow, G \downarrow, C\uparrow}
+\item \emph{drop-$x$}: \tuple{X, C=0, G > 0;   \neg X}
+\end{itemize}
+\noindent \alert{Action model to be checked}
 
@@
 
-The policy that results ... (!!??)
+\alert{Display policy here}
 
@@
 
-*** XXX ***
+\alert{*** TO BE FINISHED ***}
 

### 2018-09-05 doc/paper.tex
Draft sent to H&B

@@
 \begin{itemize}
-\item $H: \abs{holding}$: whether any block is being held,
-\item $X: \abs{holding \sqcap \set{x}}$: whether block $x$ is being held,
-\item $n(x): \abs{\exists on^* . \set{x}}$: number of blocks above $x$.\footnote{
+\item $H: \abs{holding}$, whether any block is being held,
+\item $X: \abs{holding \sqcap \set{x}}$, whether block $x$ is being held,
+\item $n(x): \abs{\exists on^* . \set{x}}$, number of blocks above $x$.\footnote{
 Action and features names are assigned manually to increase readability, based on the denotations of the involved concepts.
@@
 \begin{itemize}
-\item \emph{pick-above-$x$}: \tuple{\neg X, \neg H, n(x) > 0; n(x) \downarrow},
-\item \emph{place-aside}: \tuple{\neg X, H; \neg H},
+\item \emph{pick-above-$x$}: \hspace*{\fill} \tuple{\neg X, \neg H, n(x) > 0; X, n(x) \downarrow},
+\item \emph{place-aside}: \hspace*{\fill} \tuple{\neg X, H; \neg H},
 \end{itemize}
@@
 \begin{itemize}
-\item $X$, $n(x)$, $n(y)$: as in $\Q_{clear}$,
-\item $E: handempty$: whether gripper is empty,
-\item $XY: [(\exists on^{-1} . \set{x})\sqcap \set{b}]$: whether $x$ is on $y$.
+\item $X$, $n(x)$, $n(y)$, as in $\Q_{clear}$,
+\item $E: handempty$, whether gripper is empty,
+\item $XY: [(\exists on^{-1} . \set{x})\sqcap \set{b}]$, whether $x$ is on $y$.
 \end{itemize}
 
-\noindent Interestingly, the last feature is exactly a concept-based representation of the actual STRIPS goal $on(x,y)$.
+\noindent Interestingly, the last feature is a concept-based representation of the actual STRIPS goal $on(x,y)$.
 The corresponding action model $A_F$ is:
@@
 \begin{itemize}
-\item \emph{pick-above-$x$}:  \tuple{\neg X, \neg XY, E, n(x) > 0, n(y)>0; n(x) \downarrow, \neg E  }
-\item \emph{pick-above-$y$}:  \tuple{\neg X, \neg XY, E, n(x) = 0, n(y)>0; n(y) \downarrow, \neg E  }
-\item \emph{put-$x$-on-$y$}:  \tuple{X, \neg XY, \neg E, n(x)=0, n(y)=0 ; n(y) \uparrow, \neg X, E, XY}
-\item \emph{pick-$x$}:  \tuple{;}
-\item \emph{put-$x$-on-table}:  \tuple{;}
-\item \emph{put-aside-1}:  \tuple{;}
-\item \emph{put-aside-2}:  \tuple{;}
-
+% ATM I use two lines, as doesn't fit in one :-(
+\item \emph{pick-above-$x$}: \\\hspace*{\fill}  \tuple{\neg X, \neg XY, E, n(x) > 0, n(y)>0; n(x) \downarrow, \neg E  }
+
+\item \emph{pick-above-$y$}: \\\hspace*{\fill}  \tuple{\neg X, \neg XY, E, n(x) = 0, n(y)>0; n(y) \downarrow, \neg E  }
+
+\item \emph{put-$x$-on-$y$}: \\\hspace*{\fill} \tuple{X, \neg XY, \neg E, n(x)=n(y)=0 ; n(y) \uparrow, \neg X, E, XY}
+
+\item \emph{pick-$x$}:         \hspace*{\fill}  \tuple{\neg X, \neg XY, E, n(x)=0, n(y)=0; E, X}
+
+\item \emph{put-$x$-table}:                     \tuple{\neg E, \neg XY, X, n(x)=0, n(y) > 0; \neg X, E}
+
+\item \emph{put-aside-1}:      \hspace*{\fill} \tuple{\neg E, \neg X, \neg XY, n(x) = 0; E}
+
+\item \emph{put-aside-2}:      \hspace*{\fill} \tuple{\neg E, \neg X, \neg XY, n(x) > 0, n(y) > 0; E}
 \end{itemize}
@@
 
-\alert{ongoing}
-
-This last action will indeed be necessary in case block $x$ is initially being held
+Action \emph{put-$x$-table} will indeed be necessary only in case block $x$ starts being held
 but block $y$ is not clear.
-
+%
+The two last \emph{put-aside} actions show that when using the goal-oriented theory $T_G(\S,\F)$
+it is not always possible to completely merge similar actions.
+This is because some states (in this case, any state with $n(x) > 0 \land n(y) = 0$ and some block $b \not\in \set{x,y}$ being held)
+by chance did not appear on the optimal plans computed, which makes the resulting $A_F$ not complete in general.
+This limitation could be circumvented either by manually providing an additional 
+instance fulfilling that condition, or, more robustly, by using a larger set of optimal plans instead of just one.
+
+Interestingly, action \emph{pick-$x$} illustrates a different type of incompleteness, as it will
+only pick block $x$ whenever block $y$ is clear. No alternative optimal plan will possibly fix this,
+since (provided $x$ and $y$ are initially not in the same tower) no optimal plan for $on(x,y)$ will pick block $x$ before $y$ is clear.
+This incompleteness however will make the computation of the policies easier and the resulting policies likely more effective.
+
+The policy ...
 \alert{Display policy here}
@@
 %
-We feed our pipeline with 2 training instances of the problem with only 2 rooms and at most 4 balls.
-The resulting abstract action model is based on the following features:
+Using 2 training instances $P$ with only 2 rooms and at most 4 balls, we learn an abstract state space
+based on the following features:
 
 \begin{itemize}
-\item $C: \exists carry . \top$ (number of balls carried by the robot), 
-\item $X: at\_robby \sqcap \set{x}$ (whether robot in target room),
-\item $G: \abs{free}$ (number of empty grippers), and
-\item $B: \exists at . \neg \set{x}$ (number of balls not in target room),
+\item $C: \abs{\exists carry . \top}$, number of balls carried by the robot, 
+\item $X: at\_robby \sqcap \set{x}$, whether robot in target room,
+\item $G: \abs{free}$, number of empty grippers, and
+\item $B: \abs{\exists at . \neg \set{x}}$, number of balls not in target room,
 \end{itemize}
 
-\noindent and contains the following actions :
+\noindent and an abstract action model $A_F$:
 
 \begin{itemize}
-\item \emph{drop-ball(s)-at-$x$}: \tuple{C>0, X;  C\downarrow, G \uparrow}
-\item \emph{move-to-$x$-half-loaded}: \tuple{\neg X, B=0, C>0, G>0;   X}
-\item \emph{move-to-$x$-fully-loaded}: \tuple{\neg X, C>0, G=0;   X}
-\item \emph{pick-ball-not-in-$x$}: \tuple{\neg X, B > 0, G > 0;   B \downarrow, G \downarrow, C\uparrow}
-\item \emph{drop-$x$}: \tuple{X, C=0, G > 0;   \neg X}
+\item \emph{drop-ball-at-$x$}: \hspace*{\fill} \tuple{C>0, X;  C\downarrow, G \uparrow}
+\item \emph{move-to-$x$-fully-loaded}: \hspace*{\fill} \tuple{\neg X, C>0, G=0;   X}
+\item \emph{move-to-$x$-half-loaded}: \hspace*{\fill} \tuple{\neg X, B=0, C>0, G>0;   X}
+\item \emph{pick-ball-not-in-$x$}: \hspace*{\fill} \tuple{\neg X, B > 0, G > 0;   B \downarrow, G \downarrow, C\uparrow}
+\item \emph{leave-$x$}: \hspace*{\fill} \tuple{X, C=0, G > 0;   \neg X}
 \end{itemize}
-\noindent \alert{Action model to be checked}
-
-The model is not complete, as it e.g. features no action to pick a ball from room $x$,
-but is goal-oriented as a result of using the SAT theory $T_G$, in that it features all of the actions
-necessary to achieve the goal.
-This can be seen e.g. in the fact that the action to move \emph{move-to-$x$-half-loaded}
-can only be performed when no more balls remain in the other room.
-Additionally, the model is \emph{provably correct for any gripper instance with an arbitrary number of balls, 
-rooms and grippers}. CHECK !!
+
+The marking of only states in optimal plans as \emph{goal-relevant} introduces again a bias
+that can be seen e.g. in the fact that the robot is not allowed to leave room $x$ if carrying any ball,
+or that \emph{move-to-$x$-half-loaded} is only allowed in the last move, when no more balls remain in the other room.
+Although allowing moves without full load before that would help minimizing the action model and achieving completeness, it would certainly \emph{not}
+help obtaining better policies.
+%
+Despite this incompleteness, the action model is \emph{provably correct for any $\Q_{gripper}$ instance with an arbitrary number of balls, 
+rooms and grippers}.
 
@@
 % 
-% Action 2:
+% Action 2: [move-to-x-half-loaded]
 %         PRE: NOT robot-at-B, nballs-A = 0, ncarried > 0, nfree-grippers > 0
@@
 % 
-% Action 3:
+% Action 3: [move-to-x-fully-loaded]
 %         PRE: NOT robot-at-B, ncarried > 0, nfree-grippers = 0
@@
 respectively the position of the rewards, of the agent, of the blocked cells, and the grid topology.
-We use two single instances of the problem with size $5 \times 5$ and $4 \times 4$ and different distributions of blocks and rewards.
-We learn a set with 2 features:
-
-- $R: \abs{reward}$ (number of remaining rewards), and
-
-- $D: dist(at, adjacent:(\neg blocked), reward)$ (distance from current cell to closest cell with reward, traversing adjacent, unblocked cells only).
-
-\noindent The action model has also 2 actions:
-
-- \emph{pick-reward}: \tuple{D=0, R > 0;  R \downarrow, D \uparrow}
-
-- \emph{move-towards-closest-reward}: \tuple{R>0, D>0; D \downarrow}
-
-The model is not complete ... 
-...
-
-The policy ... 
-... EXPAND A BIT - TOO TIRED NOW.
-Discuss suboptimality of the policy.
-Perhaps discuss D=$\infty$ when no reward?
+This is the only problem where we test the $\textit{dist}(C_1,R:C,C_2)$ min-distance features introduced before.
+We use two single training instances with grid sizes $5 \times 5$ and $4 \times 4$ and different distributions of blocks and rewards,
+and learn an abstraction with 2 features:
+
+\begin{itemize}
+\item $R: \abs{reward}$, number of remaining rewards, and
+\item $D: dist(at, adjacent:(\neg blocked), reward)$, distance from current cell to closest cell with reward, traversing adjacent, unblocked cells only.
+\end{itemize}
+
+\noindent The action model has the following 2 actions:
+
+\begin{itemize}
+\item \emph{pick-reward}: \hspace*{\fill} \tuple{D=0, R > 0;  R \downarrow, D \uparrow}
+\item \emph{move-towards-closest-reward}: \hspace*{\fill} \tuple{R>0, D>0; D \downarrow}
+\end{itemize}
+
+The model is very simple, and can be shown to solve any $\Q_{reward}$ instance irrespective of 
+grid size and number and location of rewards and cell blocks, \emph{as long as the instance is solvable},
+i.e. all rewards are reachable.
+\alert{Interesting point here: how does our approach fare with unsolvability? And with problems with dead-end?}
+
+\alert{Display policy here}
+
+Discuss suboptimality of the policy?
 

### 2018-09-05 doc/paper2.tex
H's revision at around 23h

--- 2018-09-05 doc/paper.tex
@@
 \usepackage{graphicx}  %Required
+\usepackage{booktabs}  %Required
+
 \frenchspacing  %Required
@@
 
-\renewcommand\labelitemi{\raisebox{0.4ex}{\tiny$\bullet$}} %% Decrease size of items in enumerate and raise them to center them vertically
-
 %PDF Info Is Required (no accents, etc.!):
@@
   values into actions, and that they can often be computed with fully observable non-deterministic
-  (FOND) planners on suitable FOND problems. %after a sequence of transformations.
+  (FOND) planners. %  on suitable FOND problems. %after a sequence of transformations.
   The actions in such plans, however, are not the actions in the instances themselves, which are
-  not necessarily common to  other instances, but abstract actions that track the effect of the
+  not necessarily common to  other instances, but abstract actions that track the effect of 
   actions on the set of features. The formulation assumes that both the features and the abstract
   actions are given.
-  In this work, we address this limitation by showing how to automatically learn the features and
-  abstract actions. The resulting account of generalized planning combines learning and planning
+  In this work, we address this limitation by showing how to learn them automatically. 
+  %  the features and  abstract actions.
+  The resulting account of generalized planning combines learning and planning
   in a novel way: a \emph{learner}, based on a min SAT formulation, yields the features and abstract
-  actions from state transitions sampled from some problem instances, and a FOND \emph{planner}
+  actions from state transitions sampled from problem instances, and a FOND \emph{planner}
   uses this information, suitably transformed, to produce the general plans.
@@
 
-Generalized planning studies the computation of plans for  solving  multiple  instances at once
+Generalized planning studies the computation of plans  that   solve   multiple  instances
 \cite{srivastava08learning,bonet09automatic,srivastava:generalized,hu:generalized,BelleL16,anders:generalized}.
@@
 the primitive predicates used to define the instances and a given sample of state transitions.}
-For example,  the general policy for achieving $clear(x)$ of \citeay{bonet:ijcai2018}
-is obtained using a FOND planner on an abstracton that consists of the boolean
-feature $H$ that tracks whether the gripper holds a block and a numerical feature
+For example,  the general policy for achieving $clear(x)$ % of \citeay{bonet:ijcai2018}
+is obtained using a FOND planner on an abstraction that consists of  a  boolean
+feature $H$  that tracks whether the gripper holds a block,  a numerical feature
 $n(x)$ that counts the number of blocks above $x$, and two abstract actions:
@@
 (decrement $n(x)$), and the other with precondition $H$ and effect $\neg H$.
-In this work, we show how to obtain such policies without having to provide the
-features and abstract actions by hand.  The abstraction is learned instead from
-the domain predicates and state transitions sampled from some instances.
+In this work, we show how to obtain such policies from STRIPS instances only
+without having to provide the features and abstract actions by hand.
 
@@
 
-The paper is organized as follows. We provide the background first and show
+The paper is organized as follows. We provide the background  first and show then
 how features and abstract actions can be learned by enforcing soundness and
-completeness over a set of samples using a pool of features obtained the
-domain predicates. The full computational model is then summarized, followed
-by experimental results and a discussion.
+completeness over a set of samples  using features derived from the domain predicates.
+The full computational model is then summarized, followed by experimental results
+and a discussion.
 
@@
 if it results in a numerical value, that we assume to be a non-negative integer.
-The features $H$ and $n(x)$ represent two features in the Blocksworld:
+$H$ and $n(x)$ represent two features in the Blocksworld:
 the first is a boolean feature that tracks whether the gripper is empty, while
@@
 and $\neg p$, that abbreviate the  atoms $p=true$ and $p=false$, while  over  numerical state variables $n$,
-preconditions are of  the form $n=0$ and $n > 0$, and effects are  of the form $n\mminus$ (decrements) and
+preconditions are of  the form $n=0$ and $n > 0$, and effects  of the form $n\mminus$ (decrements) and
 $n\pplus$ (increments).
@@
 is the truth valuation that makes $p$ true iff $\phi_p(s)=true$, and $n=0$
-true iff $\phi_n(s) = 0$, and an abstract action $\bar{a}$ is applicable in
+true iff $\phi_n(s) = 0$. An abstract action $\bar{a}$ is applicable in
 a state $s$ if its preconditions are satisfied in $\bar{s}$.
@@
   is sound in the generalized problem $\Q_{clear}$ which consists of all
-  the Blocksworld instances, with stack and unstack actions, whose goal
-  is $clear(x)$.
+  BW instances with stack and unstack actions with goal $clear(x)$.
   Indeed, if $s$ is a reachable state in an instance $P$ and $\bar{a}$ is
@@
   $on(A,B)$ is true in $s$, the action $a=unstack(A,B)$ is applicable in
-  $s$ and it has the same qualitative effects of $\bar{a}$ over $F$: they
-  both make $H$ true and both decrease the value of $n(x)$.
+  $s$ and has the same qualitative effects of $\bar{a}$ over $F$: they
+  both make $H$ true and they both decrease the value of $n(x)$.
   Likewise, the abstract action $\bar{a'}=\abst{H}{\neg H}$ is also sound. \qed
@@
 
-For a set of abstract actions to be complete in this example, it should 
-contain abstracts actions to represent all the concrete actions in the
-problem such as stacking a block above $x$, picking up a block that is
+For a set of abstract actions to be complete in this example, it must
+contain actions to represent all concrete actions in the problem
+such as stacking a block above $x$, picking up a block that is
 not above $x$, and so on.
@@
     the initial states of instances $P$ in $\Q$ all satisfy $I_F$, while
-    all states that satisfy $G_F$ must be goal states of $P$.\footnote{A
-    state $s$ is said to satisfy $I_F$ or $G_F$ if  $\bar{s}$ does.}
+    all states that satisfy $G_F$ must be goal states of $P$.
+    % \footnote{A   state $s$ is said to satisfy $I_F$ or $G_F$ if  $\bar{s}$ does.}
+    % no longer needed
 %
@@
     Such solutions however do not necessarily solve $Q_F$ because the
-    non-determistic effects $n>0\,|\,n=0$ in $Q'_F$ are \emph{conditionally fair}
-    instead of being just \emph{fair}: infinite decrements of $n$ ensure
-    $n=0$ eventually \emph{only} when number of increments of $n$ is finite.
+    non-deterministic effects $n>0\,|\,n=0$ in $Q'_F$ are not \emph{fair} but
+    \emph{conditionally fair}: infinite decrements of $n$ ensure
+    $n=0$ eventually \emph{only} when the number of increments of $n$ is finite.
     The problem of obtanining solutions of $Q'_F$ that only assume conditional
-    fairness, so-called terminating solutions in \cite{srivastava:aaai2011},
+    fairness, the so-called terminating solutions in \cite{srivastava:aaai2011},
     is mapped into the problem of solving an amended FOND $Q^+_F$ that
-    only assumes (standard) fairness \cite{bonet:ijcai2017}.\footnote{The
-    translation to $Q^+_F$ is performed with software kindly supplied by the authors of \cite{bonet:ijcai2017}.}
+     assumes standard  fairness \cite{bonet:ijcai2017}.\footnote{The
+    translation to $Q^+_F$ is performed with software supplied by the authors.}
     %available by the authors that fixes bugs in the description found
@@
 features $F$ and  sound abstract actions $A_F$ provided by hand.
-The contribution of this work is to learn them both from samples:
+The contribution of this work is to learn them  from samples:
 
@@
   is a possible succesor of $s$ in $P$; i.e., $s'=f(a,s)$ for some action $a$ applicable in $s$,
- 4)~$\S$ is closed in the sense  that if $(s,s',P)$ is in $\S$, then $(s,s'',P)$ is in $\S$  if $s''$ is another possible succesor of $s$ in $P$.
+  4)~$\S$ is closed in the sense  that if $(s,s',P)$ is in $\S$, then $(s,s'',P)$ is in $\S$  if $s''$ is another
+  succesor of $s$ in $P$.
 \end{definition}
@@
 states $s'$ in  the pairs $(s,s') \in \S$ as the successors of
-$s$ in $\S$. Notice that the closure condition ensures that
+$s$ in $\S$. The  closure condition 4)~requires that
 states $s$ appearing \emph{first} in pairs $(s,s')$ must be fully expanded in $\S$;
@@
 
-For a sufficiently large sample set, these approximate notions of soundness and completeness
-converge to the exact notions.
+For a sufficiently large sample set, these approximate notions converge to their exact versions.
 
@@
 For computing the features and abstract actions we define a formula
-$T(\S,\F)$ that is \emph{satisfiable} iff there is a set of
+$T(\S,\F)$ where ${\cal S}$ is a sample set for $\Q$ and $\cal F$
+is a feature pool defined from the primitive domain predicates
+as described below. The formula is  \emph{satisfiable} iff there is a set of
 features $F \subseteq \F$ and a set of abstract actions $A_F$
@@
 $\S$.
-\textcolor{red}{\bf **** Need to say what $\F$ is ****}
-
-\Omit{The  sets  $F$ and $A_F$ can then be read off from any satisfying
-  assignment. Later we will add weights to so me of the clauses to establish
-  a preference over these assignments (e.g., for  minimizing $|F|$).
-}
 
@@
 The \emph{propositional variables} in $T(\S,\F)$ are then:
-\begin{itemize}
+\begin{enumerate}[{\small$\bullet$}]
   \item $selected(f)$ for each $f \in \F$, true iff $f$ selected (in $F$).
@@
     %As for $D_1$, the number of $D_2$ variables is cut in half due to symmetries.
-\end{itemize}
+\end{enumerate}
 
@@
 
-\noindent where $f$ ranges over features in $\F$ with  different qualitative
+\noindent where $f$ ranges over the features in $\F$ with  different qualitative
 values in $s$ and $t$; i.e.\ boolean features $p$ with different truth value
 in $s$ and $t$, and numerical features $n$ with different truth values for
-atom $n=0$ in $s$ and $t$.
+ $n=0$ in $s$ and $t$.
 
@@
 
-\noindent where $f$ ranges over features $f \in \F$ that have the same
+\noindent where $f$ ranges over the features $f \in \F$ that have the same
 qualitative values in $s$ and $t$ but which change differently in the two
@@
 The third class of formulas relate $D_1$ and $D_2$ by enforcing \emph{soundness}
-and \emph{completeness} over  the samples in $\S$.
-Yet, since the abstract actions  are not given but are derived, the qualitative changes in each  transition $(s,s')$ in $\S$
+and \emph{completeness} over  the samples.  Since the abstract actions  are not given,
+the qualitative changes in each  transition $(s,s')$ in $\S$
 are  taken as templates  of abstract actions. Thus, if $s$ and $t$ are not distinguished by the selected features, 
@@
 The theory $T(\S,\F)$ given by the formulas  (\ref{eq:d1})--(\ref{eq:goal})
-has a number of atoms that is linear in the size of the feature pool $\F$,
-and quadratic in the number of expanded  states $s$  and transitions $(s,s')$ in $\S$.
-The resulting number of clauses is \textcolor{red}{\bf XXXX}, which can be
-roughly cut in half by  symmetry considerations.
+has a number of atoms that is linear in the size $|{\cal F}|$  of the feature pool $\F$,
+and quadratic in the number $|{\cal S}_1$ of expanded  states $s$  and
+in the number $|{\cal S}|_2$ of transitions $(s,s')$ in $\S$.
+The resulting number of clauses grows   with
+$|{\cal S}_1|^2 \times ({\cal F} + {\cal S}_2)$, and 
+can be roughly cut in half by  symmetry considerations.
 
@@
 \begin{theorem}
-  If $\sigma$ is a satisfying assigment of $T(\S,\F)$, the set of abstract
-  actions $A_{\sigma}$ is sound and complete relative to $\S$.
+  If $\sigma$ is a satisfying assigment of $T(\S,\F)$, 
+  $A_{\sigma}$ is sound and complete relative to $\S$.
 \end{theorem}
@@
 }
+% 
 Different satisfying assignments $\sigma$ give rise to different features $F_{\sigma}$
@@
 satisfying assignments $\sigma$ that minimizes a cost measure such as $|F_{\sigma}|$
-or more generally $\sum_{f \in F_{\sigma}} cost(f)$. We achieve this constrained
-optimization through the use Min SAT or Max-Weighted SAT solvers, where it suffices
+or more generally $\sum_{f \in F_{\sigma}} cost(f)$. We achieve this
+optimization through the use of Min SAT or Max-Weighted SAT solvers, where it suffices
 to give infinite weights to the clauses resulting from $T(\S,\F)$ and weights
@@
 relies on the predicates used to encode the instances in $\Q$.
-A simple grammar allows  us   to construct new predicates from them, 
-and the boolean and numerical features in the pool
-are defined from the resulting \emph{unary predicates}. 
-The key issue is how to define new predicates from old  ones. For this 
-we take a solution off-the-shelf: \emph{description logics (DL)},  also called
-\emph{concept languages},  are grammars for doing exactly this.
+A simple grammar allows  us   to  construct new predicates 
+and to define  the boolean and numerical features in $\F$
+in terms of them. The key issue is how to the define
+new predicates. For this, we take a solution off-the-shelf: \emph{description logics (DL)},  also called
+\emph{concept languages},  are grammars for doing  this.
 Concepts $C$  refer  to predicates of arity 1 and roles $R$  to predicates of arity 2.
@@
 
-The DL grammar is given by the rules: \textcolor{red}{\bf **** More details?}
-\begin{itemize}
-  \item $C \ \leftarrow\ \ C_p, C_u, C_x$: primitive, universal, and nominal c's,\footnote{Denotation of $C_x$ for parameter $x$ is $\{x\}$.}
+We use a standard  DL grammar  given by the rules:
+\begin{enumerate}[{\small$\bullet$}]
+  \item $C \ \leftarrow\ \ C_p, C_u, C_x$: primitive, univ.,  nominal concepts,\footnote{Denotation of $C_x$ for parameter $x$ is $\{x\}$.}
   \item $C \ \leftarrow\  \neg C, C \sqcap C'$: negation and conjunction, 
@@
 %     \item $R \leftarrow R:C$, denotes pairs $(x,y)$ in $R$ s.t. $C(y)$.
- \end{itemize}
+ \end{enumerate}
 
 The  compositional  semantics   of these rules allows us to determine the denotations (extensions)   $C^s$ and $R^s$  of
-the resulting unary and binary predicates from the denotation of the primitive predicates in any  state $s$.
-
-** Check what is used here and not; see what needs to be explained; keep concise, improve lay out.
+the resulting unary and binary predicates $C^s$ and $R^s$ from the denotation of the primitive predicates
+in any  state $s$. We will see examples of derived concepts below such as ${\exists on^* . C_x}$,
+whose denotation in a state $s$ of the Blockworld is the set of blocks that are above $x$ in $s$.
+The numerical features $n_C$ will be associated with the cardinality of the extension $C^s$.
 
@@
 the  features $f$ that represent the functions $\phi_f(s)$ as follows:
-\begin{itemize}
+\begin{enumerate}[{\small$\bullet$}]
 \item Boolean $b_p$ for each primitive predicate $p$ of arity $0$,
@@
   $(R:C)(x_i,x_{i+1})$ for $i=1, \ldots, n$. 
-\end{itemize}
+\end{enumerate}
 
@@
 features $n_b$ are pruned if uniformly true or false for all $s$, or if $n_C$
-not pruned. 
+is not pruned. 
 The measure $cost(f)$ is set to the complexity of $C$ for $n_C$ and $b_C$,
@@
 In summary, the steps for computing general plans are:
-\begin{enumerate}
+\begin{enumerate}[1.]
   \item a sample set $\S$ is computed from a few instances $P$ of a given
@@
   \item a pool of features $\F$ is obtained from the predicates used in the
-    instances $P$, the grammar, the bound $k$, and the sample $\S$ (used for
-    pruning concepts and roles),
+    instances $P$, the grammar, a  bound $k$, and the sample $\S$ (used for
+    pruning),
   \item a satisfying assignment $\sigma$ of the theory $T(\S,\F)$ or $T_G(\S,\F)$
@@
     conditions $I_F$ and $G_F$ provided by hand to match $\Q$, 
-  \item the FONDP $Q^+_F$, constructed from $Q_F$, is solved by an off-the-shelf
+  \item the FONDP $Q^+_F$, constructed automatically from $Q_F$, is solved by an off-the-shelf
     FOND planner.
@@
 When this policy is applied to an instance $P$ of $\Q$, however, such features denote state functions
-which are determined by the  semantics of the concepts and roles used in  the features.
+which are determined by the  semantics of the concepts and roles involved
 (e.g., $n_C$ stands for feature function $\phi_{n_c}(s)=|C^s|$).
@@
 initial and goal conditions comply with the conditions on $I_F$ and $G_F$.
-\textcolor{red}{\bf ** NEXT IS UNCLEAR **}
-Since we do not have a language for specifying $\Q$ formally, we can take
-the formulas $I_F$ and $G_F$ indeed as a partial formal specification of
+% \textcolor{red}{\bf ** NEXT IS UNCLEAR **}
+% ok, I removed this. It's not strictly needed and we don't use it
+\Omit{
+Since we do not use a language for specifying $\Q$ formally, we will 
+take the formulas $I_F$ and $G_F$  as a partial formal specification of
 $\Q$. The theorem then says that if soundness over the samples, ensure
@@
 In principle, we could learn the formulas $I_F$ and $G_F$ from the sample as well,  but with different guarantees. 
+}
 
@@
 
-We evaluate the computational model described in steps 1--6 above over 4 different domains.
+We evaluate the computational model on 4 different generalizd planning problems.
+%We evaluate the computational model described in steps 1--6 above over 4 different domains.
 %
@@
 %how $\S$ obtained
-The sample set $\S$
+%The sample set $\S$
 % s obtained following one of two possible strategies.
 % In the \emph{exhaustive} sampling strategy, $\S$
-contains, for every one of the training instances $P$ in $\Q$,
-the first $m$  states expanded by a standard breadth-first search from the initial state in $P$,
-plus all the states along an optimal plan $\pi^*$ which is separately computed to ensure that $\S$ 
-contains at least one goal state. 
-All successor states of states in $\S$ are also added to $\S$.
-%
-For the goal marking scheme that yields the smaller theory $T_G(\S, \F)$, the states that are marked as
-goal-relevant are those in $\pi^*$.
+For each generalized problem $\Q$, a number of training instances $P$ in $\Q$
+are selected and explored in order to generate the sample set $\S$. %that is used to learn the features and actions for $\Q$.
+For each such $P$, $\S$ contains the first $m$ states that are expanded by
+breadth-first search from the initial state in $P$, plus all the states along
+an optimal plan $\pi^*$ from the initial state of each such $P$ \textcolor{red}{**** for each $P$? ****}
+In addition, $\S$ is closed by adding the transitions $(s,s')$ for all states
+$s$ in $\S$ such that $s'$ not in $\S$.
+For the goal marking scheme that yields the smaller theory $T_G(\S, \F)$,
+the states that are marked as goal-relevant are the states reached by
+the optimal plan $\pi^*$.
 
@@
 
-The complexity bound $k$ in the feature pool $\mathcal{F} = \mathcal{F}^k$ is set to a value of $k=8$,
-with just one difference in the concept generation process: distance features are only generated for the last example,
-meant to illustrate their usefulness.
+The process to generate the pool of features is the same across all the
+experiments except that distance features are generated only for the
+last example. A complexity bound $k=8$ is used to limit the size of the
+pool.
+The resulting theories are then solved with the Open-WBO MAXSAT solver
+\cite{martins2014open}.
+
+The learnt abstractions are passed through the provided QNP translator 
+to generate FOND problems which are solved using a SAT-based FOND
+planner \cite{tomas:fond-sat}. The translation is in all cases very fast
+(i.e., $\sim 0.01$ seconds) as it runs in low polynomial time in the size
+of the abstraction. Solving the FOND problem, on the other hand, is
+more demanding.
+
+%the value $k=8$ in all the expetimenin the feature pool $\mathcal{F} = \mathcal{F}^k$ is set to a value of $k=8$,
+%with just one difference in the concept generation process: distance features are only generated for the last example,
+%meant to illustrate their usefulness.
 % In our experiments, we increased this value to $k=16$ if no model was found
@@
 % We have considered strategies to increase the bound adaptively
-The resulting theories are then solved with the Open-WBO MAXSAT solver \cite{martins2014open}.
-
-\alert{Details on the FOND planner, QNP translation, etc.}
-
-The whole computational pipeline made up of steps 1--6 is processed on Intel Xeon E5-2660 CPUs,
+
+%\alert{Details on the FOND planner, QNP translation, etc.}
+
+Very few training instances $P$ are used to generate the sample sets,
+and each such instance is also small. The instances were manually
+chosen in order to display interesting situations, yet we expect
+similar results can be obtained in a fully automatic manner by 
+increasing and diversifying the set of training instances.
+
+The whole computational pipeline is processed on Intel Xeon E5-2660 CPUs
 with time and memory cutoffs of 1h and 32GB.
-
-The number of instances $P$ used to generate the samples was kept small
-by choosing the instances $P$ manually, like a teacher.
-Very few instances turn out to be required for generating general plans.
-The learning scheme is not supervised, as the training set are state
-transitions, not state action pairs to be generalized; but is not fully unsupervised either as currently
-implemented, with the instances $P$ from $\Q$ selected by hand.
-We do not expect the results to be different if such instances
-were to be sampled at random, but the number of instances required for
-producing sound abstract actions and valid general plans, as well as the size of the
-theories, would likely be higher.
+Table~\ref{tab:exp-result-data} summarizes the relevant data for each
+of the benchmarks.
+
+%by choosing the training instances $P$ manually, like a teacher. 
+%Very few instances turn out to be required for generating general plans.
+%The learning scheme is not supervised, as the training set are state
+%transitions, not state action pairs to be generalized; but is not fully
+%unsupervised either as currently implemented, with the instances $P$ from
+%$\Q$ selected by hand.
+%We do not expect the results to be different if such instances were to be
+%sampled at random, but the number of instances required for producing sound
+%abstract actions and valid general plans, as well as the size of the
+%theories, would likely be higher.
+
+\newcommand{\rowspacing[1]}{&&&&&&&&&&&&&&&&\\[\dimexpr-\normalbaselineskip+2pt]}
 
@@
   \centering
-% \addtolength{\tabcolsep}{-1pt}
-{
-% \def\arraystretch{1}%  1 is the default
-\small
-
-\newcommand{\rowspacing[1]}{&&&&&&&&&&&&&&&&\\[\dimexpr-\normalbaselineskip+2pt]}
-
-\begin{tabular}{rrrrrrrrrrrrrrrrr}
-% @{\extracolsep{3pt}}
-
-&$I$ & $\abs{\S}$ & $k$ & $\abs{\F^k}$ & \multicolumn{2}{c}{$T$} & \multicolumn{2}{c}{$T_G$} &  $T_{\text{SAT}}$ & $\abs{F}$ & $\abs{A_F}$ & $T_{\text{trans}}$  & \multicolumn{2}{c}{$Q^+_F$}  & $T_{\text{FOND}}$
-& $\abs{\pi}$ \\ %
-
-\rowspacing 
-\cline{6-7} \cline{8-9} \cline{14-15}%
-\rowspacing
-
- & &     &      &         & $\#V$ & $\#C$           & $\#V$ & $\#C$             &     &&&&  $\#Act$ & $\#At$&& \\%
-
-\rowspacing \hline \rowspacing
-\multirow{1}{*}{$\Q_{clear}$}
-& 1     &     927 & 8  & 322 & 535K & 59.6M & 7.7K & 767K   &  0.01 & 3 & 2   &  - & - & - & - & - \\% Total complexity of F: 8
-
-\rowspacing \hline \rowspacing
-\multirow{1}{*}{$\Q_{on}$}
-& 3 &        420  & 8 & 657 & 128K & 25.8M & 18.3K &     3.3M    &  0.02&  5 & 7   &  - & - & - & - & -\\% Total complexity of F: 18
-% Max-sat problem has 128217 variables and 25825428 unique clauses (with repetitions: 25825428). Top weight is 4577
-
-\rowspacing \hline \rowspacing
-\multirow{1}{*}{$\Q_{gripper}$}    
-& 2 &         403 & 8 & 130 & 93K & 4.7M & 8.1K &   358K    &  0.01 & 4 & 5&  - & - & - & - & -\\% Total complexity of F: 10
-
-\rowspacing \hline \rowspacing
-\multirow{1}{*}{$\Q_{reward}$}
-& 2 &         568 & 8 & 280 & 184K & 11.9M & 15.9K & 1.2M   &  0.01 & 2 & 2 &  - & - & - & - & -\\% Total complexity of F: ?
+  %\resizebox{.95\textwidth}{!}{
+    \begin{tabular}{lrrrrr@{}rrrrrrrr}
+      \toprule
+                     &     &            &            & \multicolumn{2}{c}{$T(\S,\F)$} && \multicolumn{2}{c}{$T_G(\S,\F)$} \\
+      \cmidrule{5-6}
+      \cmidrule{8-9}
+                     & $n$ & $\abs{\S}$ & $\abs{\F}$ &   $np$ &   $nc$ &&   $np$ &   $nc$ & $t_{\text{SAT}}$ & $\abs{F}$ & $\abs{A_F}$ & $t_{\text{FOND}}$ & $\abs{\pi}$ \\
+      \midrule
+      $\Q_{clear}$   &   1 &        927 &        322 &   535K &  59.6M &&   7.7K &   767K &             0.01 &         3 &           2 &              0.46 &           5 \\
+      $\Q_{on}$      &   3 &        420 &        657 &   128K &  25.8M &&  18.3K &   3.3M &             0.02 &         5 &           7 &              7.56 &          12 \\
+      $\Q_{gripper}$ &   2 &        403 &        130 &    93K &   4.7M &&   8.1K &   358K &             0.01 &         4 &           5 &            171.92 &          14 \\
+      $\Q_{reward}$  &   2 &        568 &        280 &   184K &  11.9M &&  15.9K &   1.2M &             0.01 &         2 &           2 &              1.36 &           7 \\
+      \bottomrule
     \end{tabular}
-
+  %}
+  \caption{Results on 4 different benchmarks: $n$ is number of training instances,
+    $\abs{\S}$ is number of transitions in $\S$,
+    %$k$ is max.\ feature comlexity,
+    $\abs{\F}$ is size of feature pool,
+    $np$ and $nc$ are the number of propositions and clauses in theories $T(\S,\F)$ and $T_G(\S,\F)$,
+    $t_{\text{SAT}}$ is time for SAT solver on theory, 
+    $\abs{F}$ is number of selected features,
+    $\abs{A_F}$ is number of abstract actions,
+    $t_{\text{FOND}}$ is time for the FOND planner, and
+    $\abs{\pi}$ is the size of the found solution.
+    %$time_{\text{SAT}}$, $T_{\text{FOND}}$: computation time (sec.) of the SAT and FOND solvers, resp.;
+    %$T_{\text{trans}}$: time (sec.) to generate the FOND problem $Q^+_F$;
+    %$\#Act$, $\#At$: number if actions and atoms in $Q^+_F$;
+    All times are in seconds. %figures that depend on the SAT theory are given for $T_G(\S,\F)$.
   }
-\caption{{
-  Experimental Results on 4 different generalized planning problems. 
-  $I$: number of training instances; 
-  $\abs{\S}$: sample set size, in total number of transitions;
-  $k$: max. feature comlexity; $\abs{\F^k}$: size of the pool of features;
-  $\#V$, $\#C$: Number of variables and clauses of the theories $T(\S,\F)$ and $T_G(\S,\F)$;
-  $T_{\text{SAT}}$, $T_{\text{FOND}}$: computation time (sec.) of the SAT and FOND solvers, resp.;
-  $\abs{F}$: number of selected features;
-  $\abs{A_F}$: number of abstract actions;
-  $T_{\text{trans}}$: time (sec.) to generate the FOND problem $Q^+_F$;
-  $\#Act$, $\#At$: number if actions and atoms in $Q^+_F$;
-  $\abs{\pi}$: size of the policy.
-%  
-  All figures that depend on the SAT theory are given for $T_G(\S,\F)$.
-}}
   \label{tab:exp-result-data}
 \end{table*}
-% 2018-09-05 00:54:08 INFO     Selected (states / goals / optimal tx): 220 / 34 / 20
-
-Table~\ref{tab:exp-result-data} summarizes the relevant data for each of the test benchmarks.
 
@@
 \paragraph{Clearing a block.}
-The generalized problem $\Q_{clear}$ aims at achieving the goal $clear(x)$ for any block $x$
-in the Blocksworld.
-We take the standard STRIPS representation with primitive predicates $on(a, b)$, $clear(b)$, $ontable(b)$, $holding(b)$ and $handempty$.
-%
-With one single instance $\Q_{clear}$ with 5 blocks, we are able to learn
-an abstract action space with features:
-
-\begin{itemize}
-\item $H: \abs{holding}$, whether any block is being held,
-\item $X: \abs{holding \sqcap \set{x}}$, whether block $x$ is being held,
-\item $n(x): \abs{\exists on^* . \set{x}}$, number of blocks above $x$.\footnote{
-Action and features names are assigned manually to increase readability, based on the denotations of the involved concepts.
-}
-\end{itemize}
-
-\noindent The resulting action model has two actions:
-
-\begin{itemize}
-\item \emph{pick-above-$x$}: \hspace*{\fill} \tuple{\neg X, \neg H, n(x) > 0; X, n(x) \downarrow},
-\item \emph{place-aside}: \hspace*{\fill} \tuple{\neg X, H; \neg H},
-\end{itemize}
-
-\noindent 
-The model is provably correct with respect to the generalized $\Q_{clear}$, but incomplete, as it lacks e.g.
-an action to pick block $x$. However, it allows us to obtain the following policy from the FOND translation $Q^+_F$:
-
-\alert{Display policy here},
-
-\noindent which will alternatively \emph{pick-above-$x$} and \emph{place-aside} the picked block 
-until block $x$ is clear. This can be proven to \emph{achieve the goal from any starting block configuration, for any number of blocks} in 
-$\Q_{clear}$.
+The problem $\Q_{clear}$ aims at reaching the goal $clear(x)$ for a
+block $x$ in Blocksworld. We use the standard encoding that involves
+the primitive predicates $on(\cdot,\cdot)$, $clear(\cdot)$,
+$ontable(\cdot)$, $holding(\cdot)$, and $handempty$, and stack/unstack
+actions.
+Using one single training instance of 5 blocks, we are able to learn
+a sound (but incomplete) abstract model with the two abstract
+actions:\footnote{The action names are not learnt but provided by us
+  to make their meaning explicit.}
+\begin{enumerate}[--]
+  \item $\text{put-aside} = \abst{\neg X, H}{\neg H}$,
+  \item $\text{pick-above-$x$} = \abst{\neg X, \neg H, n(x) > 0}{H, n(x)\mminus}$.
+\end{enumerate}
+The abstract model contains two boolean features $X$ and $H$, and
+a single numerical feature $n(x)$. 
+The first feature tracks whether the gripper holds block $x$, the
+second whether the gripper holds any block (including $x$), and the
+last tracks the number of blocks above $x$.
+These features are expressed in DL as:
+\begin{enumerate}[--]
+  \item $H: \abs{holding}$, \textcolor{red}{Looks numerical because card..}
+  \item $X: \abs{holding \sqcap \set{x}}$,
+  \item $n(x): \abs{\exists\,on^* . \set{x}}$. \textcolor{red}{Explain?}
+\end{enumerate}
+
+The translator produces a FOND problem in 0.01 seconds that is solved
+by the planner in 0.46 seconds.
+The resulting plan makes a loop that applies pick-above-$x$ followed
+by put-aside until $x$ becomes clear, yet in the last iteration the
+held block is not put aside. The initial and goal situations are
+$\neg X,\neg H,n(x)>0$ and $\neg X,n(x)=0$ respectively.
 
@@
 \paragraph{Stacking two blocks from different towers.}
-The problem $\Q_{on}$ aims at achieving a goal of the form $on(x,y)$
-in the Blocksworld for any two blocks $x$ and $y$.
-To simplify, we assume that $x$ and $y$ never start being in the same tower.
-We take the same STRIPS representation as above. From 3 different training instances with up to 14
-blocks, we are able to learn an abstract space based on the following 5 features:
-
+The problem $\Q_{on}$ consists in achieving the goal $on(x,y)$
+in Blocksworld for any two blocks $x$ and $y$ that are not
+\emph{initially} not in the same tower.
+Using the above Blocksworld encoding and only 3 different training
+instances with up to 14 blocks, an abstract model involving 3 boolean
+features, 2 numerical features, and 7 actions is learnt.
+The boolean features are $E$, $X$ and $on(X,Y)$ that stand for the gripper
+being empty, the gripper holding block $X$, and the block $x$ being directly on
+block $y$, while the learnt numerical features are $n(x)$ and $n(y)$ that stand
+for the number of blocks above $x$ and $y$ respectively.
+The features $X$, $n(x)$, and $n(y)$ are specified in DL as above,
+while $E: handempty$ and $on(X,Y): \abs{(\exists\,on^{-1}.\{x\}) \sqcap \{y\}}$.
+Interestingly, this last feature is the concept-based representation
+of the goal $on(x,y)$.
+
+%% blocks
+%% 5 bool[handempty] 0 n(a) 1 n(b) 1 holding(a) 0 on(a,b)_2 0
+%% 5 bool[handempty] 1 holding(a) 0 n(a) 1 n(b) 1 on(a,b)_2 0
+%% 4 holding(a) 0 n(a) 0 n(b) 1 on(a,b)_2 1
+%% 7
+%% action_1 ; put held block (diff. from a and b) aside
+%% 4 bool[handempty] 0 holding(a) 0 n(a) 0 on(a,b)_2 0
+%% 1 bool[handempty] 1
+%% action_2 ; put aside
+%% 5 bool[handempty] 0 holding(a) 0 n(a) 1 n(b) 1 on(a,b)_2 0
+%% 1 bool[handempty] 1
+%% action_3 ; pick block a
+%% 5 bool[handempty] 1 holding(a) 0 n(a) 0 n(b) 0 on(a,b)_2 0
+%% 2 bool[handempty] 0 holding(a) 1
+%% action_4 ; put a (being held) aside
+%% 5 bool[handempty] 0 holding(a) 1 n(a) 0 n(b) 1 on(a,b)_2 0
+%% 2 bool[handempty] 1 holding(a) 0
+%% action_5 ; pick block different from b that is above a
+%% 5 bool[handempty] 1 holding(a) 0 n(a) 1 n(b) 1 on(a,b)_2 0
+%% 2 bool[handempty] 0 n(a) 0
+%% action_6 ; pick block different from a that is above b
+%% 5 bool[handempty] 1 holding(a) 0 n(a) 0 n(b) 1 on(a,b)_2 0
+%% 2 bool[handempty] 0 n(b) 0
+%% action_7 ; put a directly on b
+%% 5 bool[handempty] 0 holding(a) 1 n(a) 0 n(b) 0 on(a,b)_2 0
+%% 4 bool[handempty] 1 holding(a) 0 n(b) 1 on(a,b)_2 1
+
+The set of 7 abstract actions includes actions to pick up the clear
+block above $x$ or $y$, to put aside the block being held (not above
+$x$ nor $y$), to pick block $x$, to place $x$ on $y$, and to place
+$x$ on the table.
+
+The translator takes 0.01 seconds and the planner computes a
+solution in 7.56 seconds. The policy loops to clear block $x$, then
+loops to clear block $y$, and finishes by picking $x$ and
+placing it on block $y$.
+
+\Omit{ % CONTAINS FEATURES AND ACTIONS
 \begin{itemize}
@@
 \alert{Display policy here}
+}
 
 \paragraph{Gripper.}
+The problem is based on the classical Gripper problem where a robot with
+two grippers needs to move a number of balls from one room into another
+\emph{target} room $x$. Each gripper of the robot may carry one ball 
+at a time, and the number of balls to move depends on the instance of
+the problem.
+We use the standard encoding with the primitive predicates $at\text{-}robby(l)$,
+$at\text{-}ball(b, l)$, $free(g)$, $carry(b, g)$ that denote, respectively,
+whether the robot (the ball) is at location $l$, whether gripper $g$ is free,
+and whether gripper $g$ carries ball $b$, plus unary type predicates $room$,
+$ball$, and $gripper$.
+
+The pipeline is fed with 2 training instances with 2 rooms and at most
+4 balls. The resulting abstract model contains the features:
+\begin{enumerate}[--]
+  \item $X: at\_robby \sqcap \set{x}$ (whether robby is in target room),
+  \item $B: \exists\,at . \neg \set{x}$ (number of balls not in target room),
+  \item $C: \exists\,carry . \top$ (number of balls carried by robby),
+  \item $G: \abs{free}$ (number of empty grippers).
+\end{enumerate}
+The abstract actions in the learnt model are:
+\begin{enumerate}[--]
+  \item $\text{drop-ball-at-$x$}=\abst{C>0, X}{C\mminus, G \pplus}$,
+  \item $\text{move-to-$x$-half-loaded}\! =\!  \abst{\neg X, B=0, C>0, G\!>\!0}{X}$,
+  \item $\text{move-to-$x$-fully-loaded} = \abst{\neg X, C>0, G=0}{X}$,
+  \item $\text{pick-ball-not-in-$x$} = \abst{\neg X, B > 0, G > 0}{B\mminus, G\mminus, C\pplus}$,
+  \item $\text{leave-$x$} = \abst{X, C=0, G > 0}{\neg X}$.
+\end{enumerate}
+
+As before, the model is sound and incomplete, but it remains sound
+for versions of the problem that involve more than two grippers,
+and the found solution also solves such problems.
+
+The problem is translated into FOND in 0.01 seconds and the
+planner finds a solution in 171.92 seconds.
+
+\textcolor{red}{\bf *** NEED POLICY ***}
+
+\Omit{%BLAI
 The problem $\Q_{gripper}$ is based on the classical Gripper problem, where a robot with two grippers needs to 
@@
 \alert{Display policy here}
+}
 
@@
 \paragraph{Collecting rewards in a maze.}
+The problem consists in an agent that needs to navigate a rectangular
+grid to pick up rewards that are spread throughout the grid, while 
+avoiding cells that are considered as \emph{blocked}.
+The encoding has primitive predicates $reward(c)$, $at(c)$, $blocked(c)$
+and $adjacent(c_1, c_2)$ that denote the position of the rewards, of
+the agent, of the blocked cells, and the grid topology respectively.
+The computational pipeline is fed with two instances of sizes $4 \times 4$
+and $5\times 5$ and given distributions of blocks and rewards.
+A set featuring two numerical features is learnt:
+\begin{enumerate}[--]
+  \item $R: \abs{reward}$ (number of remaining rewards), and
+  \item $D: dist(at, adjacent:(\neg blocked), reward)$ (distance from current
+    cell to closest cell with reward, traversing adjacent, unblocked cells only).
+\end{enumerate}
+The learnt action model is:
+\begin{enumerate}[--]
+  \item $\text{pick-reward} = \abst{D=0, R > 0}{R\mminus, D\pplus}$,
+  \item $\text{move-towards-closest-reward} = \abst{R>0, D>0}{D\mminus}$.
+\end{enumerate}
+
+The problem is translated into FOND in 0.01 seconds and the
+planner finds a solution in 1.36 seconds.
+
+\textcolor{red}{\bf *** NEED POLICY ***}
+
+\Omit{
 The objective in the problem $\Q_{reward}$ is to move in an $n \times m$ grid to pick up certain
@@
 Discuss suboptimality of the policy?
+}
 
@@
 We have introduced a scheme for computing general plans that mixes
-learning and planning: a learner infers a general abstraction made
+learning and planning: a \emph{learner} infers a general abstraction made
 up of features and abstract actions by enforcing a notion of soundness
-and completeness over a suitable set of samples, which is transformed
-and fed into a FOND planner.
+and completeness over a  set of samples, and a \emph{planner}
+uses the abstraction, suitably transformed, to compute general plans. 
 The number of samples required for obtaining correct general plans is
-very small, as the learner does not have to learn how to plan, but
-just to identify the general features that are relevant for the planner.
+very small as the learner does not have to learn how to plan, but
+just identify the features that are relevant for the  planner to track.
 Moreover, unlike, purely learning approaches, the scope and correctness
@@
 We have carried out a number of experiments over several domains to 
-illustrate the versatility of the approach. In the future, we want to
-explore the use of the learned features and actions in classical
-planning, to study of formal characterizations of the generalized
-problem $\Q$ and the possibilities of testing soudness over $\Q$
-automatically, and to improve the scalability of the approach,
-for example, by collecting samples intelligently and incrementally,
+illustrate the versatility of the approach. In the future, we want 
+% explore the use of the learned features and actions in classical planning, to
+to study of formal characterization  of the generalized problem $\Q$ and the possibilities of
+testing soudness over $\Q$ automatically, and to improve the scalability of the approach,
+for example, by collecting the samples intelligently and incrementally
 from the states where the computed policy fails, and by reusing the
 features and actions obtained for some goal for more complex goals.
+The learned features may also have other uses.
 

### 2018-09-05 doc/paper.tex
Blai's iteration at 1AM

--- 2018-09-05 doc/paper.tex
@@
 
-\title{Learning features and abstract actions for computing generalized plans}
+\title{Learning Features and Abstract Actions for Computing Generalized Plans}
 
@@
 \author{\textbf{Submission \#5695}}
+\nocopyright
 
@@
   is sound in the generalized problem $\Q_{clear}$ which consists of all
-  the Blocksworld instances, with stack and unstack actions, whose goal
-  is $clear(x)$.
+  Blocksworld instances with stack and unstack actions with goal $clear(x)$.
   Indeed, if $s$ is a reachable state in an instance $P$ and $\bar{a}$ is
@@
     Such solutions however do not necessarily solve $Q_F$ because the
-    non-determistic effects $n>0\,|\,n=0$ in $Q'_F$ are \emph{conditionally fair}
-    instead of being just \emph{fair}: infinite decrements of $n$ ensure
-    $n=0$ eventually \emph{only} when number of increments of $n$ is finite.
-    The problem of obtanining solutions of $Q'_F$ that only assume conditional
-    fairness, so-called terminating solutions in \cite{srivastava:aaai2011},
+    non-deterministic effects $n>0\,|\,n=0$ in $Q'_F$ are not \emph{fair} but
+    \emph{conditionally fair}: infinite decrements of $n$ ensure
+    $n=0$ eventually \emph{only} when the number of increments of $n$ is finite.
+    The problem of obtaining solutions of $Q'_F$ that only assume conditional
+    fairness, the so-called terminating solutions in \cite{srivastava:aaai2011},
     is mapped into the problem of solving an amended FOND $Q^+_F$ that
-    only assumes (standard) fairness \cite{bonet:ijcai2017}.\footnote{The
-    translation to $Q^+_F$ is performed with software kindly supplied by the authors of \cite{bonet:ijcai2017}.}
+     assumes standard  fairness \cite{bonet:ijcai2017}.\footnote{The
+    translation to $Q^+_F$ is performed with software supplied by the authors.}
     %available by the authors that fixes bugs in the description found
@@
   such that 1)~$P$ is an instance of $\Q$, 2)~$s$ is a reachable state in $P$, 3)~$s'$
- is a possible succesor of $s$ in $P$; i.e., $s'=f(a,s)$ for some action $a$ applicable in $s$,
- 4)~$\S$ is closed in the sense  that if $(s,s',P)$ is in $\S$, then $(s,s'',P)$ is in $\S$  if $s''$ is another possible succesor of $s$ in $P$.
+  is a possible successor of $s$ in $P$; i.e., $s'=f(a,s)$ for some action $a$ applicable in $s$,
+  4)~$\S$ is closed in the sense  that if $(s,s',P)$ is in $\S$, then $(s,s'',P)$ is in $\S$  if $s''$ is another
+  successor of $s$ in $P$.
 \end{definition}
@@
 
-For a sufficiently large sample set, these approximate notions of soundness and completeness
-converge to the exact notions.
-
-\section{Learning  features and  abstract actions}
+For a sufficiently large sample set, these approximate notions converge to their exact versions.
+
+\section{Learning Features and Abstract Actions}
 
 For computing the features and abstract actions we define a formula
-$T(\S,\F)$ that is \emph{satisfiable} iff there is a set of
+$T(\S,\F)$ where ${\cal S}$ is a sample set for $\Q$ and $\cal F$
+is a feature pool defined from the primitive domain predicates
+as described below. The formula is  \emph{satisfiable} iff there is a set of
 features $F \subseteq \F$ and a set of abstract actions $A_F$
@@
 \begin{theorem}
-  If $\sigma$ is a satisfying assigment of $T(\S,\F)$, the set of abstract
-  actions $A_{\sigma}$ is sound and complete relative to $\S$.
+  If $\sigma$ is a satisfying assignment of $T(\S,\F)$, 
+  $A_{\sigma}$ is sound and complete relative to $\S$.
 \end{theorem}
@@
 The goal expression $G_F$ can be provided explicitly  in terms of the primitive predicates,
-but  more generally can be extracted  from the satisfying assigment $\sigma$ as well.
+but  more generally can be extracted  from the satisfying assignment $\sigma$ as well.
 For this, $G_F = G_{\sigma}$ is defined as the DNF formula whose terms 
@@
 \begin{theorem}
-For a satisfying assigment $\sigma$  of $T(\S,\F)$, 
+For a satisfying assignment $\sigma$  of $T(\S,\F)$, 
 $s$ is expanded goal   state in $\S$  iff  $s$ satisfies  $G_F=G_{\sigma}$.
@@
 
-The DL grammar is given by the rules: \textcolor{red}{\bf **** More details?}
+We use a standard  DL grammar  given by the rules:
 \begin{enumerate}[{\small$\bullet$}]
-  \item $C \ \leftarrow\ \ C_p, C_u, C_x$: primitive, universal, and nominal c's,\footnote{Denotation of $C_x$ for parameter $x$ is $\{x\}$.}
+  \item $C \ \leftarrow\ \ C_p, C_u, C_x$: primitive, univ.,  nominal concepts,\footnote{Denotation of $C_x$ for parameter $x$ is $\{x\}$.}
   \item $C \ \leftarrow\  \neg C, C \sqcap C'$: negation and conjunction, 
-  \item $C \ \leftarrow\  \exists R.C , \forall R.C$: concepts that use of  roles: first denotes $x$ s.t. for some  $y$,  $R(x,y) \land C(y)$.
-  Second,  $x$ s.t. for all $y$, $R(x,y) \rightarrow C(y)$,
-  \item $C \ \leftarrow\  R=R'$: stands for $x$ s.t. set of $y$'s  s.t. $R(x,y)$ equal to set of $y$'s s.t. $R'(x,y)$,
+  \item $C \ \leftarrow\  \exists R.C , \forall R.C$: concepts that use of  roles: first denotes $x$ s.t.\ for some  $y$,  $R(x,y) \land C(y)$.
+  Second,  $x$ s.t.\ for all $y$, $R(x,y) \rightarrow C(y)$,
+  \item $C \ \leftarrow\  R=R'$: stands for $x$ s.t.\ set of $y$'s  s.t. $R(x,y)$ equal to set of $y$'s s.t. $R'(x,y)$,
     \item $R \leftarrow R_p, R^{-1}, R^*$: primitive, inverse, transitive closure.
@@
 The  compositional  semantics   of these rules allows us to determine the denotations (extensions)   $C^s$ and $R^s$  of
-the resulting unary and binary predicates from the denotation of the primitive predicates in any  state $s$.
-
-** Check what is used here and not; see what needs to be explained; keep concise, improve lay out.
+the resulting unary and binary predicates $C^s$ and $R^s$ from the denotation of the primitive predicates
+in any  state $s$. We will see examples of derived concepts below such as ${\exists on^* . C_x}$,
+whose denotation in a state $s$ of the Blockworld is the set of blocks that are above $x$ in $s$.
+The numerical features $n_C$ will be associated with the cardinality of the extension $C^s$.
 
@@
 \begin{enumerate}[{\small$\bullet$}]
-\item Boolean $b_p$ for each primitive predicate $p$ of arity $0$,
+\item Boolean $b_p$ for each primitive predicate $p$ of arity $0$ that is true iff $p^s$ is true,
 \item Boolean $b_C$ for each $C$ that are true in $s$ iff $|C^s| > 0$,
@@
     conditions $I_F$ and $G_F$ provided by hand to match $\Q$, 
-  \item the FONDP $Q^+_F$, constructed from $Q_F$, is solved by an off-the-shelf
+  \item the FOND problem $Q^+_F$, constructed automatically from $Q_F$, is solved by an off-the-shelf
     FOND planner.
@@
 When this policy is applied to an instance $P$ of $\Q$, however, such features denote state functions
-which are determined by the  semantics of the concepts and roles used in  the features.
+which are determined by the  semantics of the concepts and roles involved
 (e.g., $n_C$ stands for feature function $\phi_{n_c}(s)=|C^s|$).
-The exception are the boolean features $b_p$ associated with
-predicates $p$ of zero arity which are treated as fluents.
 
@@
 solves the generalized problem $\Q$, provided that the abstract actions
-learned are sound and that $\Q$ is restricted to the instances $P$ whose
-initial and goal conditions comply with the conditions on $I_F$ and $G_F$.
-\textcolor{red}{\bf ** NEXT IS UNCLEAR **}
-Since we do not have a language for specifying $\Q$ formally, we can take
-the formulas $I_F$ and $G_F$ indeed as a partial formal specification of
+learned are sound relative to $\Q$ and that $\Q$ is restricted to the instances $P$ whose
+initial and goal conditions comply with   $I_F$ and $G_F$.
+However,  for a policy to be correct, the abstract actions do
+not have to be sound relative to all reachable states $s$ in the instances $P$ of $\Q$, it  suffices
+if they are sound on the states $s$ that are reachable with the policy. We will see an example of this below.
+
+% \textcolor{red}{\bf ** NEXT IS UNCLEAR **}
+% ok, I removed this. It's not strictly needed and we don't use it
+\Omit{
+Since we do not use a language for specifying $\Q$ formally, we will 
+take the formulas $I_F$ and $G_F$  as a partial formal specification of
 $\Q$. The theorem then says that if soundness over the samples, ensure
@@
 
-We evaluate the computational model described in steps 1--6 above over several domains.
-%
-
-% *Fill up general details in 1--6 that don't change from problem to problem
-%how $\S$ obtained
-The sample set $\mathcal{S}$ is obtained following one of two possible strategies.
-In the \emph{exhaustive} sampling strategy,
-$\S$ contains, for every one of the training instances $P$,
-all the states generated on a breadth-first search from the initial state in $P$
-that is run until $m$ expansions have been performed. If that set includes no goal state,
-an optimal plan for $P$ is computed and all of the states along the plan are included in $\S$ as well,
-as we need at least one goal state in the sample set.
-%
-In the \emph{random} sampling strategy, a large number $M$ of states are generated, potentially exhausting the full state space,
-and then a certain number $m \ll M$ of them is sampled uniformly at random. If no goal state is found among those $m$,
-one random goal state from the larger set is drawn at random and added into the sample.
-\textcolor{red}{\bf **** What if $M$ doesn't have a goal? ****}
-
-The complexity bound $k$ in the feature pool $\mathcal{F} = \mathcal{F}^k$ is set to a value of $k=8$,
-with just one difference in the concept generation process: distance features are only generated for the last example,
-meant to illustrate their usefulness.
-% In our experiments, we increased this value to $k=16$ if no model was found
-% for the logical theories $T$ and $T_G$, and although it would be more effective
-% to devise an adaptive strategy to do so incrementally or through binary search,
-% that is beyond the scope of this work.
-% theory $T({\cal S},{\cal F})$, in which case it is increased to $k=20$.
-% We have considered strategies to increase the bound adaptively
-The resulting theories are then solved with the Open-WBO MAXSAT solver \cite{martins2014open}.
-
-** Details on the FOND planner, QNP translation, etc. **
-
-The whole computational pipeline made up of steps 1--6 is processed on Intel Xeon E5-2660 CPUs,
-with time and memory cutoffs of 1h and 32GB.
-
-The number of instances $P$ used to generate the samples was kept small
-by choosing the instances $P$ manually, like a teacher.
-Very few instances turn out to be required for generating general plans.
-The learning scheme is not supervised, as the training set are state
-transitions, not state action pairs to be generalized; but is not fully unsupervised either as currently
-implemented, with the instances $P$ from $\Q$ selected by hand.
-We do not expect the results to be different if such instances
-were to be sampled at random, but the number of instances required for
-producing sound abstract actions and valid general plans, as well as the size of the
-theories, would likely be higher.
-
-The goal marking scheme that  yields the smaller  theory $T_G(\S}, \F)$ is obtained an optimal plan over
-a few (always less than xxxx) instances $P$. ??
+We evaluate the computational model on four generalized problems $\Q$. For each $\Q$, we have selected  a few and small  ``training''  instances $P$ in $\Q$ by hand, 
+from which we draw  the sample sets $\cal S$.  $\S$ is constructed by collecting  the first $m$ states
+generated by  a breadth-first search  along with  the states generated in an optimal plan. This is done for each
+training instance  $P$. The plans ensure that $\cal S$ contains some goal states, and they are used to mark states and transitions for
+constructing  the theory $T_G({\cal S,\cal F})$, which is the one used in the experiments. $\S$ is closed by fully expanding the states selected
+while the bound $k$ for generating  $\cal F$ is set to $8$. The  distance features are used only in the last problem.
+The Weighted-Max Solver is Open-WBO \cite{martins2014open} and the FOND planner is SAT-FOND \cite{tomas:fond-sat}.
+The translation from $Q'_F$ to $Q^+_F$ is very fast, in  the order of $0.01$ seconds in all cases.  
+The whole computational pipeline summarized by the steps 1--6 above is processed on Intel Xeon E5-2660 CPUs
+with time and memory cutoffs of 1h and 32GB. Table~\ref{tab:exp-result-data} summarizes the relevant data for 
+the problems.
 
@@
   \centering
-% \addtolength{\tabcolsep}{-1pt}
-{
-% \def\arraystretch{1}%  1 is the default
-\small
-
-\newcommand{\rowspacing[1]}{&&&&&&&&&&&&&&&\\[\dimexpr-\normalbaselineskip+2pt]}
-
-\begin{tabular}{%
-@{}l%
-@{\extracolsep{1.5em}}r%@{}l%
-@{\extracolsep{1.5em}}r%
-@{\extracolsep{8pt}}r%
-%
-@{\extracolsep{1.5em}}r%
-@{\extracolsep{8pt}}r%
-@{\extracolsep{8pt}}r%
-@{\extracolsep{8pt}}r%
-%
-@{\extracolsep{1.5em}}r%
-@{\extracolsep{8pt}}r%
-@{\extracolsep{8pt}}r%
-@{\extracolsep{8pt}}r%
-%
-@{\extracolsep{1.5em}}r%
-@{\extracolsep{8pt}}r%
-@{\extracolsep{8pt}}r%
-@{\extracolsep{8pt}}r%
-%
-@{}}
-
-% @{\extracolsep{3pt}}
-
- & $I$ & $\abs{\S}$ & $k$ & $\abs{\F^k}$ & \multicolumn{2}{c}{$T$} & \multicolumn{2}{c}{$T_G$} &  $C_{\text{SAT}}$ & $\abs{F}$ & $\abs{A_F}$ & $\abs{I_F}$  & $\abs{G_F}$ & $\abs{Q^+_F}$ & $C_{\text{FOND}}$ \\ %
-
-\rowspacing 
-\cline{6-7} \cline{8-9}%
-\rowspacing
-
- & &     &      &         & $\#V$ & $\#C$           & $\#V$ & $\#C$             &     &&&&&& \\%
-
-\rowspacing \hline \rowspacing
-\multirow{1}{*}{$\Q_{clear}$}
-& 1     &     927 & 8  & 322 & 535K & 59.6M & 7.7K & 767K   &  0.01 & 3 & 2   &  - & - & - & - \\% Total complexity of F: 8
-
-\rowspacing \hline \rowspacing
-\multirow{1}{*}{$\Q_{on}$}
-& 3 &        420  & 8 & 657 & 128K & 25.8M & 18.3K &     3.3M    &  0.02&  5 & 7   &  - & - & - & - \\% Total complexity of F: 18
-% Max-sat problem has 128217 variables and 25825428 unique clauses (with repetitions: 25825428). Top weight is 4577
-
-\rowspacing \hline \rowspacing
-\multirow{1}{*}{$\Q_{gripper}$}    
-& 2 &         403 & 8 & 130 & 93K & 4.7M & 8.1K &   358K    &  0.01 & 4 & 5&  - & - & - & - \\% Total complexity of F: 10
-
-\rowspacing \hline \rowspacing
-\multirow{1}{*}{$\Q_{reward}$}
-& 2 &         568 & 8 & 280 & 184K & 11.9M & 15.9K & 1.2M   &  0.01 & 2 & 2 &  - & - & - & - \\% Total complexity of F: ?
+  %\resizebox{.95\textwidth}{!}{
+    \begin{tabular}{lrrrrr@{}rrrrrrrr}
+      \toprule
+                     &     &            &            & \multicolumn{2}{c}{$T(\S,\F)$} && \multicolumn{2}{c}{$T_G(\S,\F)$} \\
+      \cmidrule{5-6}
+      \cmidrule{8-9}
+                     & $n$ & $\abs{\S}$ & $\abs{\F}$ &   $np$ &   $nc$ &&   $np$ &   $nc$ & $t_{\text{SAT}}$ & $\abs{F}$ & $\abs{A_F}$ & $t_{\text{FOND}}$ & $\abs{\pi}$ \\
+      \midrule
+      $\Q_{clear}$   &   1 &        927 &        322 &   535K &  59.6M &&   7.7K &   767K &             0.01 &         3 &           2 &              0.46 &           5 \\
+      $\Q_{on}$      &   3 &        420 &        657 &   128K &  25.8M &&  18.3K &   3.3M &             0.02 &         5 &           7 &              7.56 &          12 \\
+      $\Q_{gripper}$ &   2 &        403 &        130 &    93K &   4.7M &&   8.1K &   358K &             0.01 &         4 &           5 &            171.92 &          14 \\
+      $\Q_{reward}$  &   2 &        568 &        280 &   184K &  11.9M &&  15.9K &   1.2M &             0.01 &         2 &           2 &              1.36 &           7 \\
+      \bottomrule
     \end{tabular}
-
+  %}
+  \caption{Results on 4 different benchmarks: $n$ is number of training instances $P$,
+    $\abs{\S}$ is number of transitions in $\S$,
+    %$k$ is max.\ feature comlexity,
+    $\abs{\F}$ is size of feature pool,
+    $np$ and $nc$ are the number of propositions and clauses in theories $T(\S,\F)$ and $T_G(\S,\F)$,
+    $t_{\text{SAT}}$ is time for SAT solver on theory, 
+    $\abs{F}$ and  $\abs{A_F}$ are the resulting number of 
+    selected features and abstract actions, 
+    $t_{\text{FOND}}$ is time for the FOND planner, and
+    $\abs{\pi}$ is the size of the resulting policy.
+    %$time_{\text{SAT}}$, $T_{\text{FOND}}$: computation time (sec.) of the SAT and FOND solvers, resp.;
+    %$T_{\text{trans}}$: time (sec.) to generate the FOND problem $Q^+_F$;
+    %$\#Act$, $\#At$: number if actions and atoms in $Q^+_F$;
+    All times are in seconds. %figures that depend on the SAT theory are given for $T_G(\S,\F)$.
   }
-\caption{{
-  Experimental Results on 4 different generalized planning problems. 
-  $I$: number of training instances; 
-  $\abs{\S}$: sample set size, in total number of transitions;
-  $k$: max. feature comlexity; $\abs{\F^k}$: size of the pool of features;
-  $\#V$, $\#C$: Number of variables and clauses of the theories $T(\S,\F)$ and $T_G(\S,\F)$;
-  $C_{\text{SAT}}$, $C_{\text{FOND}}$: computation time (sec.) of the SAT and FOND solvers, resp.;
-  $\abs{F}$: number of selected features;
-  $\abs{A_F}$: number of abstract actions;
-  $\abs{I_F}$: ???;
-  $\abs{G_F}$: ??? ;
-  $\abs{Q^+_F}$: ??? .
-  All figures that depend on the SAT theory are given for $T_G(\S,\F)$.
-}}
   \label{tab:exp-result-data}
 \end{table*}
-% 2018-09-05 00:54:08 INFO     Selected (states / goals / optimal tx): 220 / 34 / 20
-
-Table~\ref{tab:exp-result-data} summarizes the relevant data for each of the test benchmarks.
 
@@
 \paragraph{Clearing a block.}
-The generalized problem $\Q_{clear}$ aims at achieving the goal $clear(x)$ for any block $x$
-in the Blocksworld.
-We take the standard STRIPS representation with primitive predicates $on(a, b)$, $clear(b)$, $ontable(b)$, $holding(b)$ and $handempty$.
-%
-With one single instance $\Q_{clear}$ with 5 blocks, we are able to learn
-an abstract action model which features two actions:
-
-- \emph{pick-above-$x$}: \tuple{\neg X, \neg H, n(x) > 0; n(x) \downarrow},
-
-- \emph{place-aside}: \tuple{\neg X, H; \neg H},
-
-\noindent both described in terms of features $H: \abs{holding}$, $X: \abs{holding \sqcap \set{x}}$ and $n(x)\equiv \abs{\exists on^* . \set{x}}$, 
-that denote, respectively, whether any block is being held, whether block $x$ is being held,
-and the number of blocks above $x$.\footnote{
-Action and features names are assigned manually by us to increase readability, but correspond faithfully
-to the learnt entities.
-}
-%
-The model is provably correct, but not complete in general, as it lacks e.g.
-any action to pick block $x$.
-However, it is sufficient to obtain a policy with a single loop that will 
-alternatively \emph{pick-above-$x$} and \emph{place-aside} until block $x$ is clear,
-which provably \emph{achieves the goal from any configuration of blocks, for any number of blocks}.
-
-% Action 1:
+$\Q_{clear}$ contains the Blocksworld instances  with goals of the form  $clear(x)$
+and stack/unstack actions. The primitive predicates, i.e., those appearing in the instances of $\Q_{clear}$, 
+are assumed to use the predicates $on(\cdot,\cdot)$, $clear(\cdot)$,
+$ontable(\cdot)$, $holding(\cdot)$, and $handempty$. 
+For this problem, a  single training instance $P$  with  5 blocks
+sufficed to learn an abstract model from which a general plan was computed.
+The  set of  features $F$ obtained from the theory $T_G$ is:
+
+\begin{enumerate}[--]
+  \item $H: holding$ (holding a block),
+  \item $X: \abs{holding \sqcap C_x}$ (holding block $x$),
+  \item $n(x): \abs{\exists\,on^* . \set{x}}$, (number of blocks above block $x$).
+\end{enumerate}
+
+\noindent The set of abstract actions obtained from $T_G$ is:\footnote{
+Feature and action names are provided   to make their intuitive meaning explicit.
+Their formal meaning is given by their definitions.}
+
+\begin{enumerate}[--]
+  \item $\text{put-aside} = \abst{\neg X, H}{\neg H}$,
+  \item $\text{pick-above-$x$} = \abst{\neg X, \neg H, n(x) > 0}{H, n(x)\mminus}$.
+\end{enumerate}
+
+The extra precondition $\neg X$ is not needed and would  be removed by drawing more samples,
+yet it does not prevent the computation of a solution to $\Q_{clear}$ using  $I_F = \{\neg H, \neg X, n(x) > 0\}$ and $G_F=\{n(x)=0\}$.
+The translator produces a FOND problem in 0.01 seconds that is solved
+by the FOND  planner in 0.46 seconds. The resulting plan implements  a loop that applies
+the pick-above-$x$ action followed by put-aside,  until $x$ becomes clear.
+
+% Action 1: PLACE-ASIDE
 %         PRE: NOT holding(a), bool[holding]
@@
 
-\paragraph{Stacking two blocks from different towers.}
-The problem $\Q_{on}$ is aims at achieving a goal of the form $on(x,y)$
-in the Blocksworld for any two blocks $x$ and $y$ that are initially not in the same tower.
-We take the same STRIPS representation as above. Based on 3 different instances with up to 14
-blocks, we are able to learn an abstract action model with 5 features and 7 actions.
-The set of features is made up of  $X$, $n(x)$ and $n(y)$, with the same syntax and semantics as given before,
-as well as $E: handempty$
-and $XY: [(\exists on^{-1} . \set{x})\sqcap \set{b}]$,
-the second of which precisely denotes that block $x$ is on block $y$.
-
-The set of actions includes actions to pick up the topmost block above $x$, to pick up the topmost block above $y$,
-to put a block being held \emph{aside}, i.e. not above $x$ nor above $y$, to pick $x$, 
-to place $x$ on $y$ and to place $x$ on the table.
-
-- \emph{pick-up-above-$x$}:
-
-- \emph{pick-up-above-$y$}:
-
-This last action will indeed be necessary in case block $x$ is initially being held
-but block $y$ is not clear.
-
-DISCUSS POLICY WHEN AVAILABLE. / 
-DISCUSS WHY WOULDN'T WORK WHEN IN SAME TOWER.
+\paragraph{Stacking two blocks.}
+$\Q_{on}$ consists  of Blocksworld instances with goals of the form $on(x,y)$,
+and initial situations where the blocks $x$ and $y$ are not in the same tower.
+The primitive predicates are the same as those above.
+For this problem, three training instances $P$ were used.
+The set of $F$ obtained from $T_G$ contains
+% 
+% Using the above Blocksworld encoding and only 3 different training
+% instances with up to 14 blocks, an abstract model involving 3 boolean
+% features, 2 numerical features, and 7 actions is learnt.
+the boolean features  $E$, $X$ and $G$,  that stand for the gripper
+being empty, the gripper holding block $x$, and the block $x$ being
+on block $y$, and the numerical features  $n(x)$ and $n(y)$ that stand
+for the number of blocks above $x$ and $y$ respectively.
+The DL  definition of the  features $X$, $n(x)$, and $n(y)$ are as above
+while $E$ is $handempty$ and $G$ stands for $(\exists\,on^{-1}.C_x)  \sqcap C_y$.
+% Interestingly, this last feature is the concept-based representation
+% of the goal $on(x,y)$.
+%
+%% blocks
+%% 5 bool[handempty] 0 n(a) 1 n(b) 1 holding(a) 0 on(a,b)_2 0
+%% 5 bool[handempty] 1 holding(a) 0 n(a) 1 n(b) 1 on(a,b)_2 0
+%% 4 holding(a) 0 n(a) 0 n(b) 1 on(a,b)_2 1
+%% 7
+%% action_1 ; put held block (diff. from a and b) aside
+%% 4 bool[handempty] 0 holding(a) 0 n(a) 0 on(a,b)_2 0
+%% 1 bool[handempty] 1
+%% action_2 ; put aside
+%% 5 bool[handempty] 0 holding(a) 0 n(a) 1 n(b) 1 on(a,b)_2 0
+%% 1 bool[handempty] 1
+%% action_3 ; pick block a
+%% 5 bool[handempty] 1 holding(a) 0 n(a) 0 n(b) 0 on(a,b)_2 0
+%% 2 bool[handempty] 0 holding(a) 1
+%% action_4 ; put a (being held) aside
+%% 5 bool[handempty] 0 holding(a) 1 n(a) 0 n(b) 1 on(a,b)_2 0
+%% 2 bool[handempty] 1 holding(a) 0
+%% action_5 ; pick block different from b that is above a
+%% 5 bool[handempty] 1 holding(a) 0 n(a) 1 n(b) 1 on(a,b)_2 0
+%% 2 bool[handempty] 0 n(a) 0
+%% action_6 ; pick block different from a that is above b
+%% 5 bool[handempty] 1 holding(a) 0 n(a) 0 n(b) 1 on(a,b)_2 0
+%% 2 bool[handempty] 0 n(b) 0
+%% action_7 ; put a directly on b
+%% 5 bool[handempty] 0 holding(a) 1 n(a) 0 n(b) 0 on(a,b)_2 0
+%% 4 bool[handempty] 1 holding(a) 0 n(b) 1 on(a,b)_2 1
+The set $A_F$ includes actions to pick up the clear
+block above $x$ or $y$, two actions to put aside the block being held,
+and actions to pick block $x$, to place $x$ on $y$, and to place
+$x$ on the table:
+\begin{enumerate}[--]
+  \item $\text{pick-ab-$x$}\!=\!\abst{E,\neg X,\neg G,n(x)\!\!>\!0,n(y)\!>\!0}{\neg E,n(x)\mminus}$,
+  \item $\text{pick-ab-$y$}\!=\!\abst{E,\neg X,\neg G,n(x)\!=\!0,n(y)\!\!>\!0}{\neg E,n(y)\mminus}$,
+  \item $\text{put-aside-1}=\abst{\neg E,\neg X,\neg G,n(x)=0}{E}$,
+  \item $\text{put-aside-2}=\abst{\neg E,\neg X,\neg G,n(x)>0,n(y)>0}{E}$,
+  \item $\text{pick-$x$}=\abst{E,\neg X,\neg G,n(x)=0,n(y)=0}{\neg E,X}$,
+  \item $\text{put-$x$-aside}\!=\!\abst{\neg E,X,\neg G,n(x)=0,n(y)>0}{E,\neg X}$,
+  \item $\text{put-$x$-on-$y$} = \langle \neg E, X, \neg G, n(x)=0, n(y)=0; E, \neg X,$ $G, n(y)\pplus \rangle$.
+\end{enumerate}
+
+The resulting abstraction with $I_F\!=\!\{ E, \neg X, \neg G, n(x)\!>\!0, n(y)\!>\!0\}$
+and $G_F = \{ \neg X, G, n(x)=0, n(y)\!>\!0\}$, is translated into the FOND problem
+$Q_F^+$  in 0.01 seconds,
+which is solved by the FOND planner  in 7.56 seconds.
+The policy implements a loop  to clear block $x$ as above, 
+followed by a loop to clear block $y$, and finishes by picking
+$x$ and placing it on $y$.
 
 \paragraph{Gripper.}
-The problem $\Q_{gripper}$ is based on the classical Gripper problem, where a robot with two grippers needs to 
-move a number of blocks from one room into another \emph{target} room $x$, which we take to be a parameter of the problem,
-each gripper of the robot being able to carry only one ball at a time.
-We use the standard STRIPS representation with primitive predicates $at\text{-}robby(l)$, $at\text{-}ball(b, l)$, $free(g)$, $carry(b, g)$ that denote,
-respectively, whether the robot (the ball) is at location $l$, whether gripper $g$ is free, and whether gripper $g$ carries ball $b$,
-plus unary type predicates $room$, $ball$, and $gripper$.
-%
-We feed our pipeline with 2 training instances of the problem with only 2 rooms and at most 4 balls.
-The resulting abstract action model is based on the following features:
-
-- $C: \exists carry . \top$ (number of balls carried by the robot),
-
-- $X: at\_robby \sqcap \set{x}$ (whether robot in target room),
-
-- $G: \abs{free}$ (number of empty grippers), and
-
-- $B: \exists at . \neg \set{x}$ (number of balls not in target room),
-
-\noindent and contains the following actions :
-
-- \emph{drop-ball(s)-at-$x$}: \tuple{C>0, X;  C\downarrow, G \uparrow}
-
-- \emph{move-to-$x$-half-loaded}: \tuple{\neg X, B=0, C>0, G>0;   X}
-
-- \emph{move-to-$x$-fully-loaded}: \tuple{\neg X, C>0, G=0;   X}
-
-- \emph{pick-ball-not-in-$x$}: \tuple{\neg X, B > 0, G > 0;   B \downarrow, G \downarrow, C\uparrow}
-
-- \emph{drop-$x$}: \tuple{X, C=0, G > 0;   \neg X}
-
-The model is not complete, as it e.g. features no action to pick a ball from room $x$,
-but is goal-oriented as a result of using the SAT theory $T_G$, in that it features all of the actions
-necessary to achieve the goal.
-This can be seen e.g. in the fact that the action to move \emph{move-to-$x$-half-loaded}
-can only be performed when no more balls remain in the other room.
-Additionally, the model is \emph{provably correct for any gripper instance with an arbitrary number of balls, 
-rooms and grippers}. CHECK !!
-
-The policy that results ... (!!??)
-
-% Action 1:
-%         PRE: ncarried > 0, robot-at-B
-%         EFFS: DEC ncarried, INC nfree-grippers
-% 
-% Action 2:
-%         PRE: NOT robot-at-B, nballs-A = 0, ncarried > 0, nfree-grippers > 0
-%         EFFS: robot-at-B
-% 
-% Action 3:
-%         PRE: NOT robot-at-B, ncarried > 0, nfree-grippers = 0
-%         EFFS: robot-at-B
-% 
-% Action 4:
-%         PRE: NOT robot-at-B, nballs-A > 0, nfree-grippers > 0
-%         EFFS: DEC nballs-A, DEC nfree-grippers, INC ncarried
-% 
-% Action 5:
-%         PRE: ncarried = 0, nfree-grippers > 0, robot-at-B
-%         EFFS: NOT robot-at-B
-
-\paragraph{Collecting rewards in a maze.}
-The objective in the problem $\Q_{reward}$ is to move in an $n \times m$ grid to pick up certain
-rewards that are spread out in certain grid cells, while avoiding other cells which are considered as \emph{blocked}.
-This example is from recent work in symbolic deep
-reinforcement learning, where there is a penalty associated for
-stepping in blocked cells instead \cite{garnelo2016towards}.
-%
-We represent the instances $P$ with primitive predicates $reward(c)$, $at(c)$, $blocked(c)$ and $adjacent(c_1, c_2)$ that denote
-respectively the position of the rewards, of the agent, of the blocked cells, and the grid topology.
-We use two single instances of the problem with size $5 \times 5$ and $4 \times 4$ and different distributions of blocks and rewards.
-We learn a set with 2 features:
-
-- $R: \abs{reward}$ (number of remaining rewards), and
-
-- $D: dist(at, adjacent:(\neg blocked), reward)$ (distance from current cell to closest cell with reward, traversing adjacent, unblocked cells only).
-
-\noindent The action model has also 2 actions:
-
-- \emph{pick-reward}: \tuple{D=0, R > 0;  R \downarrow, D \uparrow}
-
-- \emph{move-towards-closest-reward}: \tuple{R>0, D>0; D \downarrow}
-
-The model is not complete ... 
-...
-
-The policy ... 
-... EXPAND A BIT - TOO TIRED NOW.
-Discuss suboptimality of the policy.
-Perhaps discuss D=$\infty$ when no reward?
-
-% Action 1 [PICK-REWARD]:
-% 	PRE: min-dist[at, Restrict(adjacent,Not(blocked)), reward] = 0, num-rewards > 0
-% 	EFFS: DEC num-rewards, INC min-dist[at, Restrict(adjacent,Not(blocked)), reward]
-% 
-% Action 2 [MOVE-TO-CLOSEST-REWARD]:
-% 	PRE: min-dist[at, Restrict(adjacent,Not(blocked)), reward] > 0, num-rewards > 0
-% 	EFFS: DEC min-dist[at, Restrict(adjacent,Not(blocked)), reward]
-
-\paragraph{Observations.}
-We conclude the section with a few general observations:
-
-\begin{enumerate}
-\item
-\emph{Soundness}: interestingly, abstract actions obtained from a small number of samples in
-$\mathcal{S}$ turn out to be sound not just for $\mathcal{S}$ but for the generalized problem $\Q$,
-as discussed above. Yet this does not need to be true for the resulting policies to be
-correct for $\Q$.
-%
-Abstract actions $A_F$ do not actually have to be sound for a policy $\pi$
-obtained from them to be sound in $\Q$.
-Indeed, for the correctness of $\pi$ for solving $\Q$,
-it suffices for $A_F$ to be sound \emph{over the states reachable by $\pi$}
-in all the instances $P$ of $\Q$.
-
-For example, the abstract action \emph{pick-up-above-$y$} in $\Q_{on}$
-is sound relative to the sample set $\mathcal{S}$ used, but not relative to the general
-$\Q_{on}$. In any state $s$ where $x$ is a few blocks above $y$ and clear,
-there is no concrete action that instantiates \emph{pick-up-above-$y$},
-as picking up the topmost block above $y$ would necessarily make feature $X \equiv holding(x)$
-true, which the abstract action does not.
-%
-Keeping track of the information necessary to ensure general correctness would require indeed
-a feature set $F$ with higher complexity.
-Yet if we start in states where $x$ and $y$ are in different towers,
-states as $s$ will never be reached, and hence the policy is correct and solves $\Q_{on}$.
-
-\item
-\emph{Grammar and Feature Pool}: the general DL grammar we have described appears to be
-adequate to generate a sufficiently expressive pool of features for the examples we use,
-but this does not have to be always so.
-To come up with a general strategy in the Towers of Hanoi, for example,
-the first action to be taken  (whether to move the smallest disk to target or buffer peg)
-depends on the number of disks being odd or even,
-which cannot be easily expressed with the type of features generated by our grammar.
-
-\item
-\emph{Expressive and Computational Limitations}:
-The current approach
-
-*** XXX ***
-
-The use of primitive \emph{goal predicates} like $on_g$ with denotation
-given by the goal conjunctions in the concrete instances, as done in
-\cite{martin-geffner:generalized}, could potentially alleviate this
-e.g. to be able to generate general policies for all BW problems.
-
-Why pipeline cannot be used to solve arbitrary problems. Eg, logistics.
-Que partes explotan? Sizes k and $\abs{\mathcal{S}}$?
-
-* Goal in $\Q$ needs to be expressible in generic terms (but this is not well formalized in our formulation)
-* Increasing number of domain parameters kills scalability.
-* How to select which are the good parameters if we want arbitrary goals? This is also related to the proper formalization
-  of the approach, which we havent done.
-
-%  \item Running times of the SAT solver are negligible.
-%  \item Minimization of precondition DNFs
-%  \item Theory T
-\end{enumerate}
+$\Q_{gripper}$ contains classical Gripper instances  where a robot with
+some grippers needs to move a number of balls from one room into another
+\emph{target} room $x$. Each gripper may carry one ball
+at a time.
+The standard STRIPS predicates are $at\text{-}robby(l)$, $at\text{-}ball(b,l)$,
+$free(g)$, $carry(b,g)$ that denote, respectively, whether the robot
+(the ball) is at location $l$, whether gripper $g$ is free, and whether
+gripper $g$ carries ball $b$, plus unary type predicates $room$, $ball$,
+and $gripper$. The type predicates are also included.
+
+Pipeline 1--6 is fed  2 training instances $P$ from $\Q_{gripper}$  with 2
+rooms each, one with 4 balls and the other with 5. The set of features that
+is obtained from the theory $T_G$ is:
+\begin{enumerate}[--]
+  \item $X: at\_robby \sqcap C_x$ (whether robby is in target room),
+  \item $B: \exists\,at . \neg C_x$ (whether balls not in target room),
+  \item $C: |\exists\,carry . \top|$ (number of balls carried),
+  \item $G: \abs{free}$ (number of empty grippers).
+\end{enumerate}
+
+\noindent The resulting abstract actions are:
+
+\begin{enumerate}[--]
+  \item $\text{drop-ball-at-$x$}=\abst{C>0, X}{C\mminus, G \pplus}$,
+  \item $\text{move-to-$x$-half-loaded}\! =\!  \abst{\neg X, B=0, C>0, G\!>\!0}{X}$,
+  \item $\text{move-to-$x$-fully-loaded} = \abst{\neg X, C>0, G=0}{X}$,
+  \item $\text{pick-ball-not-in-$x$} = \abst{\neg X, B > 0, G > 0}{B\mminus, G\mminus, C\pplus}$,
+  \item $\text{leave-$x$} = \abst{X, C=0, G > 0}{\neg X}$.
+\end{enumerate}
+
+Once again, although this set of abstract actions  is not complete,
+it is sufficient for generating a general plan. The abstraction with
+$I_F = \{ B>0, \neg X, G>0, C=0\}$ and $G_F = \{ B=0, G>0, C=0 \}$ is
+translated into the FOND problem $Q^+_F$, which is solved by the
+planner in 171.92 seconds. 
+The resulting policy picks the balls in the source room $y$, one by one, 
+until the grippers are full or there are no more balls in $y$.
+It then moves to room $x$, with one of the two move actions above
+according to whether the grippers are fully loaded or not,
+and then drops the balls in room $x$ one by one, until the grippers
+become all empty. It then moves to room $y$ again if there are balls
+still to be moved, and the process repeats. Like before, the extra conditions
+that are not strictly required (in the move actions), do not preclude
+the solution of the problem, that results in a general plan that 
+works  for robots with any  number of grippers and  any number of balls.
+
+\paragraph{Collecting rewards.}
+$\Q_{reward}$ contains instances where  an agent needs to navigate a rectangular
+grid to pick up  rewards spread in a grid  while 
+avoiding \emph{blocked} cells. A variation of the example
+has been used recently in \cite{garnelo2016towards}.
+The STRIPS instances have the primitive predicates $reward(c)$, $at(c)$, $blocked(c)$
+and $adjacent(c_1, c_2)$ that denote the position of the rewards, of
+the agent, of the blocked cells, and the grid topology respectively.
+Pipeline 1--6  is fed with two instances of the problem of sizes $4 \times 4$
+and $5\times 5$, and two different  given distributions of blocked cells  and rewards.
+From the theory $T_G$, two numerical  features are learned:
+\begin{enumerate}[--]
+  \item $R: \abs{reward}$ (number of remaining rewards), and
+  \item $D: dist(at, adjacent:(\neg blocked), reward)$ (distance from current
+    cell to closest cell with reward, traversing adjacent, unblocked cells only).
+\end{enumerate}
+Likewise, two abstract actions are obtained:
+\begin{enumerate}[--]
+  \item $\text{collect-reward} = \abst{D=0, R > 0}{R\mminus, D\pplus}$,
+  \item $\text{move-to-closest-reward} = \abst{R>0, D>0}{D\mminus}$.
+\end{enumerate}
+
+The abstraction with $I_F = \{ R>0, D>0 \}$ and $G_F=\{R=0,D>0\}$
+is translated into the  FOND problem $Q^+_F$ which yields
+a policy  in 1.36 seconds. The policy moves the agent
+one step at a time, towards  the uncollected  reward
+that is closest, as measured by the numerical feature $D$.
+Once the reward is reached, the reward is consumed, and
+the process repeats until there are no more rewards to be collected.
+
+\subsection{Observations}
+
+We conclude the section with few observations.
+
+\paragraph{Soundness.} The abstract action \text{pick-ab-$y$} in $\Q_{on}$
+is sound relative to the sample set $\S$ used  but is not sound relative
+to $\Q_{on}$. In any state $s$ where $x$ is a few blocks above $y$ and clear,
+there is no concrete action that instantiates \text{pick-ab-$y$},
+as picking up the topmost block above $y$ would necessarily make feature $X$ (holding block $x$)
+true, which the abstract action does not. Still, the resulting policy is correct
+for $\Q_{on}$ because such states $s$ are not  visited by the policy
+from the given initial conditions.
+
+\paragraph{Expressivity.} The general DL grammar that we have described
+is adequate for many  domains but may fall short  on others. 
+For example,  in Towers of Hanoi, the first action to be taken
+(whether to move the smallest disk to target or buffer peg) depends
+on the number of disks begin even or  odd, which is not a feature
+that  would result  from the primitive predicates used in the STRIPS
+instances and the general grammar. Yet, for some reason,
+Tower of Hanoi is regarded as a puzzle.  For other problems, the  use of \emph{goal predicates};
+namely, copies of some primitive predicates  evaluated in the goal
+(e.g., for solving arbitrary Blocksworld instances) would be  necessary  \cite{martin-geffner:generalized}.
+
+\paragraph{Scalability.} The current bottleneck of the approach is 
+in the size of the SAT  theories used to derive the features
+and the actions. This puts a limit on the complexity of the
+feature grammars, the bound $k$, and the number of state
+transitions in the sample, as they  result otherwise
+in SAT theories that are just too large (see Table~\ref{tab:exp-result-data}).
+This is indeed the reason why have chosen to use the theories $T_G$
+with goal marking. Additional ideas, however, are required
+to improve scalability and to make the computational model
+captured by  steps 1--6 more robust.
 
@@
 We have carried out a number of experiments over several domains to 
-illustrate the versatility of the approach. In the future, we want to
-explore the use of the learned features and actions in classical
-planning, to study of formal characterizations of the generalized
-problem $\Q$ and the possibilities of testing soudness over $\Q$
-automatically, and to improve the scalability of the approach,
-for example, by collecting samples intelligently and incrementally,
+illustrate the versatility of the approach. In the future, we want 
+to study of formal characterization  of the generalized problem $\Q$ and the possibilities of
+testing soundness over $\Q$ automatically, and to improve the scalability of the approach,
+for example, by collecting the samples intelligently and incrementally
 from the states where the computed policy fails, and by reusing the
 features and actions obtained for some goal for more complex goals.
-
-\Omit{
-** discussion  of experiments; things we do well and not, scope, lessons. limitations (expressive? scalability?)
-** briefly related work, expanding paragraph in intro.
-** summary and cconlusions
-}
+The learned features may also have other uses.
 

### 2018-09-06 doc/paper.tex
Submitted version Thu 10AM

@@
 \begin{abstract}
-  Generalized planning is concerned with the computation of plans that solve multiple instances.
+  Generalized planning is concerned with the computation of plans that solve not one but multiple instances
+  of a class of planning problems.
   Recently, it has been shown that generalized plans can be expressed as mappings of feature
@@
   The actions in such plans, however, are not the actions in the instances themselves, which are
-  not necessarily common to  other instances, but abstract actions that track the effect of 
-  actions on the set of features. The formulation assumes that both the features and the abstract
+  not necessarily common to  other instances, but \emph{abstract} actions that are defined on a set 
+  of \emph{common} features. The formulation assumes that both the features and the abstract
   actions are given.
@@
 The resulting generalized plans  map feature values into abstract actions, and soundness
-ensures that the application  of an   abstract action  results in the application of
-a concrete action with a similar  effect  on  the features. Moreover, the form of the abstract actions
+ensures that the application of an abstract action can be mapped back into the application of
+a concrete action with analog effect  on  the features. Moreover, the form of the abstract actions
 ensures that generalized plans can be computed using fully observable non-deterministic (FOND) planners
@@
 
-Bonet and Geffner's  formulation of generalized planning assumes that the
+Bonet and Geffner's generalized planning formulation assumes that the
 features and abstract actions are given. In this work we address this limitation
@@
 (decrement $n(x)$), and the other with precondition $H$ and effect $\neg H$.
-In this work, we show how to obtain such policies from STRIPS instances only
+In this work, we show how to obtain such policies from STRIPS instances alone,
 without having to provide the features and abstract actions by hand.
@@
 learning and closer in aims to work on learning general policies from examples or
-experience \cite{martin:concept,fern:bias,mazebase,general-drl}.
+experience \cite{martin-geffner:generalized,fern:bias,mazebase,general-drl}.
 Generalized planning has also been formulated as a problem in first-order logic
-\cite{srivastava:generalized} and general plans over finite horizons have been
+\cite{srivastava:generalized}, and general plans over finite horizons have been
 derived using first-order regression \cite{boutilier2001symbolic,wang2008first,van2012solving}.
@@
 A \emph{generalized planning problem} $\Q$ is a collection of planning instances $P$.
-An instance $P$ is a  classical planning problems expressed in some compact language
+An instance $P$ is a  classical planning problem expressed in some compact language
 as a tuple $P=\tup{V,I,G,A}$ where $V$ is a set of state variables that can take a
@@
   is sound in the generalized problem $\Q_{clear}$ which consists of all
-  Blocksworld instances with stack and unstack actions with goal $clear(x)$.
+  Blocksworld instances with stack and unstack actions and goal $clear(x)$.
   Indeed, if $s$ is a reachable state in an instance $P$ and $\bar{a}$ is
@@
 for each transition $(s, s') \in \S$, there must be a transition $(t, t') \in \S$
-such that that the two transitions are not distinguished by the selected features either.
+such that the two transitions are not distinguished by the selected features either.
 This is expressed as
@@
 
-The theory $T(\S,\F)$ given by the formulas  (\ref{eq:d1})--(\ref{eq:goal})
+The theory $T(\S,\F)$ given by formulas (\ref{eq:d1}--\ref{eq:goal})
 has a number of atoms that is linear in the size $|{\cal F}|$  of the feature pool $\F$,
-and quadratic in the number $|{\cal S}_1$ of expanded  states $s$  and
-in the number $|{\cal S}|_2$ of transitions $(s,s')$ in $\S$.
-The resulting number of clauses grows   with
-$|{\cal S}_1|^2 \times ({\cal F} + {\cal S}_2)$, and 
-can be roughly cut in half by  symmetry considerations.
-
-In the experiments we also consider a smaller theory that avoids the
-consideration of all pairs of expanded states or all pairs of transitions
-in $\S$. For this, some expanded states $s$ and transitions $(s,s')$ in
-tuples $(s,s',P)$ are marked as \emph{goal relevant}. This is achieved by
-computing some plans for some instances $P$ in $\Q$, and by adding the
-resulting state transitions to $\S$ at the same time that the first
-arguments $s$ and $(s,s')$ in the $D_1$ and $D_2$ atoms are restricted
-to be marked.
-If the set of marked states and transitions is much smaller than the set of unmarked ones, 
-then the resulting formula,  $T_G(\S,\F)$ will be much smaller than $T(\S,\F)$.
-The goal marking scheme that involves the theory $T_G(\S,\F)$ in place of $T(\S,\F)$
-amounts to enforcing soundness on  the sample set $\S$ but not completeness, yet this does
-not affect the resulting  formal guarantees that depend on soundness only.
+and quadratic in the number $m_1$ of expanded states $s$  and
+in the number $m_2$ of transitions $(s,s')$ in $\S$.
+The number of clauses grows with
+$m_1^2 \times (\abs{{\cal F}} + m_2)$, and 
+can be roughly cut in half by eliminating symmetries.
+
+We also consider an alternative theory $T_G(\S,\F)$
+that trades off completeness guarantees for a much smaller theory size, which
+results from not having to consider
+all pairs of expanded states or transitions in $\S$.
+%
+% No need to mark states as well - we just mark transitions, and states are marked implicitly
+%
+For this, some of the transitions $(s,s')$ in $\S$  are marked as being \emph{goal-relevant},
+and completeness is sought \emph{only with respect to these transitions}.
+These goal-relevant transitions are selected for being part of a concrete plan
+that is computed for each instance $P$ in $\Q$. The theory  $T_G(\S,\F)$ is then as
+ $T(\S,\F)$, with the sole difference that the $s$ and $s'$ arguments in the $D_1$ and $D_2$ atoms 
+need to be part of some goal-relevant transition.
+% This is achieved by
+% computing some plans for some instances $P$ in $\Q$, and by adding the
+% resulting state transitions to $\S$ at the same time that the first
+% arguments $s$ and $(s,s')$ in the $D_1$ and $D_2$ atoms are restricted
+% to be marked.
+If the set of marked transitions is much smaller than the full sample set, as will typically be the case,
+then $T_G(\S,\F)$ will be much smaller than $T(\S,\F)$.
+$T_G(\S,\F)$ will guarantee soundness relative to $\S$
+but completeness only relative to the the subset of goal-relevant transitions in $\S$. Yet this does not affect
+the resulting formal guarantees, which depend only on soundness.
 
@@
 
-For a satisfying assignment $\sigma$ of the theory $T(\S,\F)$, let
+For a satisfying assignment $\sigma$ of the theories $T(\S,\F)$ and $T_G(\S,\F)$, let
 $F_{\sigma}$ be the set of features $f$ in $\F$ such that $selected(f)$
 is true in $\sigma$, and let $A_{\sigma}$ be the set of abstract actions
-that capture the transitions $(s,s')$ in $\S$, with duplicates and redundancies removed.
+that \emph{capture} all transitions $(s,s')$ in $\S$, in the case of theory $T$,
+and only the goal-relevant transitions $(s,s')$ in $\S$, in the case of $T_G$.
+% with duplicates and redundancies removed.
 The abstract action $\bar{a}=\abst{Pre}{\Eff}$ that captures the transition $(s,s')$
@@
 precondition, the precondition can be dropped, removing the duplicates and
-iterating until no more simplifications can be done. They key properties are:
+iterating until no more simplifications can be done. Their key properties are:
 
@@
 and abstract actions $A_{\sigma}$. More meaningful features are found by looking for
-satisfying assignments $\sigma$ that minimizes a cost measure such as $|F_{\sigma}|$
-or more generally $\sum_{f \in F_{\sigma}} cost(f)$. We achieve this
-optimization through the use of Min SAT or Max-Weighted SAT solvers, where it suffices
-to give infinite weights to the clauses resulting from $T(\S,\F)$ and weights
-given by $cost(f)$ to the ``soft'' unit clauses $\neg selected(f)$.
+satisfying assignments $\sigma$ that minimize a cost measure such as $|F_{\sigma}|$
+or, more generally, $\sum_{f \in F_{\sigma}} cost(f)$. We achieve this
+optimization by casting the problem as a 
+% through the use of Min SAT or 
+a Weighted Max-SAT problem where clauses resulting from the above theories are taken as \emph{hard} clauses
+and additional \emph{soft} unit clauses $\neg selected(f)$ are given weight $cost(f)$, for $f \in \mathcal{F}$.
 
@@
 A simple grammar allows  us to  construct new predicates 
-and to define  the boolean and numerical features in $\F$
-in terms of them. The key issue is how to the define
-new predicates. For this, we take a solution off-the-shelf: \emph{description logics (DL)},  also called
-\emph{concept languages},  are grammars for doing  this.
-Concepts $C$  refer  to predicates of arity 1 and roles $R$  to predicates of arity 2.
-Concept languages have been used before for  generating general policies using purely learning methods  \cite{martin:concept,fern:bias}.
+and to define boolean and numerical features in $\F$
+from them. 
+In order to construct these predicates, we use an off-the-shelf solution:
+\emph{description logics} (DL),  also called \emph{concept languages}, which are well suited to this task.
+In DL, \emph{concepts} $C$  refer to unary predicates, and \emph{roles} $R$  to binary predicates.
+Concept languages have been used before for  generating general policies using purely learning methods  \cite{martin-geffner:generalized,fern:bias}.
 \citeay{dl-handbook} give syntax and semantics of description logics.
@@
 
-We use a standard  DL grammar  given by the rules:
+We use a standard  DL grammar given by the following rules:
 \begin{enumerate}[{\small$\bullet$}]
-  \item $C \ \leftarrow\ \ C_p, C_u, C_x$: primitive, univ.,  nominal concepts,\footnote{Denotation of $C_x$ for parameter $x$ is $\{x\}$.}
-  \item $C \ \leftarrow\  \neg C, C \sqcap C'$: negation and conjunction, 
+  \item $C \ \leftarrow\ \ p, C_u, C_x$: primitive, universal, nominals: denotation of $C_u$ is whole universe, denotation of $C_x$, for parameter $x$, is $\set{x}$.
+  \item $C \ \leftarrow\  \neg C, C \sqcap C'$: negation and conjunction.
   \item $C \ \leftarrow\  \exists R.C , \forall R.C$: concepts that use of  roles: first denotes $x$ s.t.\ for some  $y$,  $R(x,y) \land C(y)$.
-  Second,  $x$ s.t.\ for all $y$, $R(x,y) \rightarrow C(y)$,
-  \item $C \ \leftarrow\  R=R'$: stands for $x$ s.t.\ set of $y$'s  s.t. $R(x,y)$ equal to set of $y$'s s.t. $R'(x,y)$,
+  Second,  $x$ s.t.\ for all $y$, $R(x,y) \rightarrow C(y)$.
+  \item $C \ \leftarrow\  R=R'$: stands for $x$ s.t.\ set of $y$'s  s.t. $R(x,y)$ equal to set of $y$'s s.t. $R'(x,y)$.
   \item $R \leftarrow R_p, R^{-1}, R^*$: primitive, inverse, transitive closure.
@@
 
-The  compositional  semantics   of these rules allows us to determine the denotations (extensions)   $C^s$ and $R^s$  of
-the resulting unary and binary predicates $C^s$ and $R^s$ from the denotation of the primitive predicates
-in any  state $s$. We will see examples of derived concepts below such as ${\exists on^* . C_x}$,
-whose denotation in a state $s$ of the Blockworld is the set of blocks that are above $x$ in $s$.
-The numerical features $n_C$ will be associated with the cardinality of the extension $C^s$.
+The primitive predicates $p$ used in the grammar must be common to all instances $P$ of the generalized problem $\Q$.
+They will usually be common domain predicates. Given that parameters $x$ are common as well,
+the compositional semantics of DL defines, for any concept $C$ and role $R$ resulting from these rules and 
+any state $s$ of $P$, a denotation $C^s$ ($R^s$) \emph{that can be determined from the denotation of the primitive predicates
+in $s$ alone}. We will below see examples of derived concepts such as $\exists\, on^* . C_x$,
+whose denotation in any Blocksworld state $s$ is the set of blocks that are above $x$ in $s$.
+Our numerical features $n_C$ will be defined by the cardinality of these denotations $C^s$.
 
@@
 
-The set of all concepts and roles with complexity no larger than $k$ is
-denoted as $\G^k$. The complexity of a concept or role is the minimum
-number of rules needed to generate it. The size of $\G^k$ is reduced by
-incrementally pruning ``duplicate'' concepts; namely, those whose denotation
-coincides with the denotation of another concept over all the sampled states.
-%in $\S$.
-
-The pool $\F=\F^k$  is defined from $\G^k$ and contains
-the  features $f$ that represent the functions $\phi_f(s)$ as follows:
+We define the \emph{complexity} of a concept or role as the minimum
+number of the above rules needed to generate it. We denote by $\G^k$
+the set of all concepts and roles with complexity no larger than $k$.
+We prune from $\G^k$ \emph{redundant} concepts whose denotation
+coincides with that of another concept over all sampled states
+in $\S$. 
+%
+The pool of features $\F=\F^k$  is defined from concepts $C \in \G^k$, and contains
+the following features $f$ that represent functions $\phi_f(s)$:
 \begin{enumerate}[{\small$\bullet$}]
-\item Boolean $b_p$ for each primitive predicate $p$ of arity $0$ that is true iff $p^s$ is true,
-\item Boolean $b_C$ for each $C$ that are true in $s$ iff $|C^s| > 0$,
-\item Numerical $n_C$ for each $C$ that stand for cardinality $|C^s|$,
-\item Numerical $\textit{dist}(C_1,R:C,C_2)$, that represent that smallest
+\item For each nullary primitive predicate $p$, a \emph{boolean} feature $b_p$ that is true in $s$ iff $p^s$ is true.
+
+\item For each concept $C$, 
+a \emph{boolean} feature $b_C$, if $|C^s| \in \set{0,1}$ for all sampled states $s$, 
+and a \emph{numerical} feature $n_C$ otherwise. The value of $b_C$ in $s$ is true iff $|C^s| = 1$;
+the value of $n_C$ is $|C^s|$.
+
+\item Numerical features $\textit{dist}(C_1,R{:}C,C_2)$, that represent the smallest
   $n$ such that for $x_1, \ldots, x_n$, $C_1^s(x_1)$, $C_2^s(x_{n})$, and
-  $(R:C)(x_i,x_{i+1})$ for $i=1, \ldots, n$. 
-\end{enumerate}
-
-The role $R:C$ is such that its denotation contains the pairs $(x,y)$ in $R$
-such that $y$ is in $C$.
-Some pruning is done in this space of features as well. E.g., numerical
-features $n_C$ are pruned from the pool when there are no states $s$ and $s'$
-in the sample set for which $n_c(s)=0$ and $n_c(s') > 1$, while boolean
-features $n_b$ are pruned if uniformly true or false for all $s$, or if $n_C$
-is not pruned. 
+  $(R{:}C)(x_i,x_{i+1})$ for $i=1, \ldots, n$. 
+\end{enumerate}
+
+The denotation $(R{:}C)^s$ of role $R{:}C$ contains all pairs $(x,y)$ in $R^s$
+s.t. $y$ is in $C^s$.
+We also prune from this space of features all features whose denotation is either constant or coincident with that of another feature over all states in $\S$.
 The measure $cost(f)$ is set to the complexity of $C$ for $n_C$ and $b_C$,
-to $0$ for $b_p$ and to the complexities of $C_1$, $R$, $C$, and $C_2$ added
-up for $\textit{dist}(C_1,R:C,C_2)$.
+to $0$ for $b_p$ and to the sum of complexities of $C_1$, $R$, $C$, and $C_2$, for $\textit{dist}(C_1,R{:}C,C_2)$.
 Only features with cost bounded by $k$ are allowed in $\F$.
@@
 $\Q$. The theorem then says that if soundness over the samples, ensure
-formal soundness over $\Q$, the policy that solves the FONDP $Q^+_F$,
+formal soundness over $\Q$, the policy that solves the FONDP problem $Q^+_F$,
 solves $\Q$. The result holds whether $T$ or $T_G$ (goal marking) is used in step 3.
@@
 from which we draw  the sample sets $\cal S$.  $\S$ is constructed by collecting  the first $m$ states
-generated by  a breadth-first search  along with  the states generated in an optimal plan. This is done for each
+generated by  a breadth-first search (where $m$ is typically never larger than a few hundreds)
+along with  the states generated in an \emph{optimal} plan. This is done for each
 training instance  $P$. The plans ensure that $\cal S$ contains some goal states, and they are used to mark states and transitions for
@@
     $\abs{\S}$ is number of transitions in $\S$,
-    %$k$ is max.\ feature comlexity,
+    %$k$ is max.\ feature complexity,
     $\abs{\F}$ is size of feature pool,
     $np$ and $nc$ are the number of propositions and clauses in theories $T(\S,\F)$ and $T_G(\S,\F)$,
-    $t_{\text{SAT}}$ is time for SAT solver on theory, 
+    $t_{\text{SAT}}$ is time for SAT solver on theory $T_G$, 
     $\abs{F}$ and  $\abs{A_F}$ are the resulting number of 
@@
 $\Q_{clear}$ contains the Blocksworld instances  with goals of the form  $clear(x)$
-and stack/unstack actions. The primitive predicates, i.e., those appearing in the instances of $\Q_{clear}$, 
-are assumed to use the predicates $on(\cdot,\cdot)$, $clear(\cdot)$,
+and stack/unstack actions. The primitive predicates, i.e., those appearing in all instances $P$ of $\Q_{clear}$, 
+are $on(\cdot,\cdot)$, $clear(\cdot)$,
 $ontable(\cdot)$, $holding(\cdot)$, and $handempty$. 
@@
 \begin{enumerate}[--]
-  \item $H: holding$ (holding a block),
-  \item $X: \abs{holding \sqcap C_x}$ (holding block $x$),
-  \item $n(x): \abs{\exists\,on^* . \set{x}}$, (number of blocks above block $x$).
+  \item $H: holding$ (whether some block is being held),
+  \item $X: holding \sqcap C_x$ (whether block $x$ is being held),
+  \item $n(x): \abs{\exists\,on^* . C_x}$, (number of blocks above block $x$).
 \end{enumerate}
@@
 and initial situations where the blocks $x$ and $y$ are not in the same tower.
-The primitive predicates are the same as those above.
+The primitive predicates are the same as in $\Q_{clear}$.
 For this problem, three training instances $P$ were used.
-The set of $F$ obtained from $T_G$ contains
+The set $F$ obtained from $T_G$ contains
 % 
@@
 The DL  definition of the  features $X$, $n(x)$, and $n(y)$ are as above
-while $E$ is $handempty$ and $G$ stands for $(\exists\,on^{-1}.C_x)  \sqcap C_y$.
+while $E$ is $handempty$ and $G$ stands for $(\exists\,on.C_y)  \sqcap C_x$.
 % Interestingly, this last feature is the concept-based representation
@@
 %% 4 bool[handempty] 1 holding(a) 0 n(b) 1 on(a,b)_2 1
-The set $A_F$ includes actions to pick up the clear
+The learned set $A_F$ includes actions to pick up the clear
 block above $x$ or $y$, two actions to put aside the block being held,
@@
   \item $X: at\_robby \sqcap C_x$ (whether robby is in target room),
-  \item $B: \exists\,at . \neg C_x$ (whether balls not in target room),
-  \item $C: |\exists\,carry . \top|$ (number of balls carried),
+  \item $B: |\exists\,at . \neg C_x|$ (number of balls not in target room),
+  \item $C: |\exists\,carry . C_u|$ (number of balls carried),
   \item $G: \abs{free}$ (number of empty grippers).
@@
 has been used recently in \cite{garnelo2016towards}.
-The STRIPS instances have the primitive predicates $reward(c)$, $at(c)$, $blocked(c)$
-and $adjacent(c_1, c_2)$ that denote the position of the rewards, of
+The STRIPS instances have the primitive predicates $reward(\cdot)$, $at(\cdot)$, $blocked(\cdot)$
+and $adjacent(\cdot,\cdot)$ that denote the position of the rewards, of
 the agent, of the blocked cells, and the grid topology respectively.
@@
 \begin{enumerate}[--]
-  \item $R: \abs{reward}$ (number of remaining rewards), and
-  \item $D: dist(at, adjacent:(\neg blocked), reward)$ (distance from current
+  \item $R: \abs{reward}$ (number of remaining rewards), 
+  \item $D: dist(at, adjacent{:}(\neg blocked), reward)$ (distance from current
     cell to closest cell with reward, traversing adjacent, unblocked cells only).
@@
 
-We conclude the section with few observations.
+We conclude the section with a few observations.
 
@@
 (whether to move the smallest disk to target or buffer peg) depends
-on the number of disks begin even or  odd, which is not a feature
+on the number of disks being even or  odd, which is not a feature
 that  would result  from the primitive predicates used in the STRIPS
@@
 just identify the features that are relevant for the  planner to track.
-Moreover, unlike, purely learning approaches, the scope and correctness
+Moreover, unlike purely learning approaches, the scope and correctness
 of the resulting general plans can be assessed.
@@
 illustrate the versatility of the approach. In the future, we want 
-to study of formal characterization  of the generalized problem $\Q$ and the possibilities of
+to study the formal characterization  of the generalized problem $\Q$ and the possibilities of
 testing soundness over $\Q$ automatically, and to improve the scalability of the approach,

### 2018-09-06 doc/paper.tex
Blai's last changes

@@
 3)~$a_i$ is applicable in $s_i$ and $s_{i+1}=f(a_i,s_i)$, and
-4)~$s_n$ is a goal state of $P$, $n=\infty$, or $\pi(\bar{s}_n)$ either is
+4)~$s_n$ is a goal state of $P$, $n=\infty$, or $\pi(\bar{s}_n)$ is either
 undefined, denotes an abstract action that is not applicable in $\bar{s_n}$,
@@
 \noindent The first formulas in $T(\S,\F)$ capture the meaning of $D_1$:
-
+%  
 \begin{equation}
@@
 
-\noindent The second formulas encode the meaning of $D_2$:
-
+The second formulas encode the meaning of $D_2$:
+%
 \begin{equation}
@@
 This is expressed as
-
+%
 \begin{equation}
@@
 
-The theory $T(\S,\F)$ given by formulas (\ref{eq:d1}--\ref{eq:goal})
+The theory $T(\S,\F)$ given by formulas \eqref{eq:d1}--\eqref{eq:goal}
 has a number of atoms that is linear in the size $|{\cal F}|$  of the feature pool $\F$,
@@
 % to be marked.
-If the set of marked transitions is much smaller than the full sample set, as will typically be the case,
-then $T_G(\S,\F)$ will be much smaller than $T(\S,\F)$.
-$T_G(\S,\F)$ will guarantee soundness relative to $\S$
-but completeness only relative to the the subset of goal-relevant transitions in $\S$. Yet this does not affect
-the resulting formal guarantees, which depend only on soundness.
+If the set of marked transitions is much smaller than the full sample set, %as will typically be the case,
+$T_G(\S,\F)$ is much smaller than $T(\S,\F)$.
+$T_G(\S,\F)$ guarantees soundness relative to $\S$
+but completeness only relative to the the subset of goal-relevant transitions in $\S$.
+Yet this does not affect the resulting formal guarantees, which depend only on soundness.
 
@@
 and true in some expanded goal state. This DNF formula can   be simplified in standard ways. 
-Due to  (\ref{eq:d1}) and (\ref{eq:goal}) that force the selected features in $F_{\sigma}$
+Due to  \eqref{eq:d1} and \eqref{eq:goal} that force the selected features in $F_{\sigma}$
 to   distinguish goal  from non-goal states, we have that:
@@
 The primitive predicates $p$ used in the grammar must be common to all instances $P$ of the generalized problem $\Q$.
-They will usually be common domain predicates. Given that parameters $x$ are common as well,
-the compositional semantics of DL defines, for any concept $C$ and role $R$ resulting from these rules and 
-any state $s$ of $P$, a denotation $C^s$ ($R^s$) \emph{that can be determined from the denotation of the primitive predicates
-in $s$ alone}. We will below see examples of derived concepts such as $\exists\, on^* . C_x$,
-whose denotation in any Blocksworld state $s$ is the set of blocks that are above $x$ in $s$.
-Our numerical features $n_C$ will be defined by the cardinality of these denotations $C^s$.
+%They will usually be common domain predicates.
+Given that parameters $x$ are common as well,
+the compositional semantics of DL defines, for any concept $C$ and role $R$ resulting
+from these rules and any state $s$ of $P$, a denotation $C^s$ ($R^s$)
+\emph{that can be determined from the denotation of the primitive predicates in $s$ alone}.
+Below we see examples of derived concepts such as $\exists\, on^* . C_x$, whose denotation
+in any Blocksworld state $s$ is the set of blocks that are above $x$ in $s$.
+The numerical features $n_C$ are defined by the cardinality of such denotations $C^s$.
 
@@
 not have to be sound relative to all reachable states $s$ in the instances $P$ of $\Q$, it  suffices
-if they are sound on the states $s$ that are reachable with the policy. We will see an example of this below.
+if they are sound on the states $s$ that are reachable with the policy. We see an example of this below.
 

### 2018-11-06 doc/camera-ready/paper.tex
Add Hector's first rewriting for the camera-ready version

--- 2018-09-06 doc/paper.tex
@@
 \pdfinfo{
-/Title ()
-/Author ()}
+/Title (Learning Features and Abstract Actions for Computing Generalized Plans)
+/Author (Blai Bonet, Guillem Frances, Hector Geffner)}
 
@@
 
-% \author{B. Bonet, G. Frances, H. Geffner} % Alphabetical
-% 
-\author{\textbf{Submission \#5695}}
+% \author{Blai  Bonet, G. Frances, H. Geffner} % Alphabetical
+
+\author{
+Blai Bonet$^1$,
+Guillem Franc\`es$^2$,
+Hector Geffner$^3$
+\\
+$^1$ Universidad Sim\'on Bol\'{\i}var, Caracas, Venezuela\\
+$^2$ University of Basel, Basel, Switzerland \\
+$^3$ ICREA \& Universitat Pompeu Fabra, Barcelona, Spain\\
+%
+bonet@usb.ve,
+guillem.frances@unibas.ch,
+hector.geffner@upf.edu
+}
+
+% 
+% \author{\textbf{Submission \#5695}}
+
 \nocopyright
@@
   Generalized planning is concerned with the computation of plans that solve not one but  multiple instances
-  of a class of planning problems.
-  Recently, it has been shown that generalized plans can be expressed as mappings of feature
+  of a planning domain. Recently, it has been shown that generalized plans can be expressed as mappings of feature
   values into actions, and that they can often be computed with fully observable non-deterministic
@@
   not necessarily common to  other instances, but \emph{abstract} actions that are defined on a set 
-  of \emph{common} features. The formulation assumes that both the features and the abstract
+  of \emph{common} features. The formulation assumes that the features and the abstract
   actions are given.
@@
   The resulting account of generalized planning combines learning and planning
-  in a novel way: a \emph{learner}, based on a min SAT formulation, yields the features and abstract
-  actions from state transitions sampled from problem instances, and a FOND \emph{planner}
+  in a novel way: a \emph{learner}, based on a Max SAT formulation, yields the features and abstract
+  actions from sampled state transitions, and a FOND \emph{planner}
   uses this information, suitably transformed, to produce the general plans.
@@
 Generalized planning studies the computation of plans  that   solve   multiple  instances
-\cite{srivastava08learning,bonet09automatic,srivastava:generalized,hu:generalized,BelleL16,anders:generalized}.
+%\cite{srivastava08learning,bonet09automatic,srivastava:generalized,hu:generalized,BelleL16,anders:generalized}.
+\cite{srivastava08learning,bonet:icaps2009,hu:generalized,BelleL16,anders:generalized}.
 For example, the plan  that iteratively  picks  a  clear block above $x$ 
@@
 Once this  general plan or  policy is derived it can be  applied to solve an infinite collection
-of instances that involve different initial states, different objects, and hence  different (ground)  actions.
+of instances that involve different initial states, different objects, and  different (ground)  actions.
 
@@
 ensures that the application of an abstract action can be mapped back into the application of
-a concrete action with analog effect  on  the features. Moreover, the form of the abstract actions
+a concrete action with the same effect  over the features. Moreover, the form of the abstract actions
 ensures that generalized plans can be computed using fully observable non-deterministic (FOND) planners
@@
 by  \emph{showing  how the features and abstract actions  can be learned from
-the primitive predicates used to define the instances and a given sample of state transitions.}
+the primitive predicates used to define the instances and sampled state transitions.}
 For example,  the general policy for achieving $clear(x)$ % of \citeay{bonet:ijcai2018}
@@
 In this work, we show how to obtain such policies from STRIPS instances alone,
-without having to provide the features and abstract actions by hand.
+without having to provide the features and the abstract actions by hand.
 
@@
 representation, and machine learning. We make use of SAT solvers and description
-logics for learning features and abstract actions \cite{sat-handbook,dl-handbook}.
+logics for learning features and abstract actions. %\cite{sat-handbook,dl-handbook}.
 The abstract actions provide a model from which the plans are obtained via
@@
 planners, and from purely learning approaches, in the \emph{formal guarantees} that
-are characteristic of planning: if the learned abstract actions are actually sound,
-the resulting general plans are correct.
+are characteristic of planning: if the learned abstract actions are  sound,
+the resulting general plans must be correct.
 
@@
 how features and abstract actions can be learned by enforcing soundness and
-completeness over a set of samples  using features derived from the domain predicates.
-The full computational model is then summarized, followed by experimental results
+completeness over a set of samples, from a pool of candidate features derived from the domain
+predicates. The full computational model is then summarized, followed by experimental results
 and a discussion.
@@
 
-We review generalized planning, abstract actions, and computation of solutions 
-following  \citeay{bonet:ijcai2018}.
+We review generalized planning,   abstract actions, and solutions following  \citeay{bonet:ijcai2018}.
 
@@
 A \emph{feature} $f$ for a class $\Q$ of problems represents a function $\phi_f$
-that takes an instance $P$ from $\Q$ and a state $s$ reachable in $P$, and results
+that takes an   instance $P$ from $\Q$ and an  state $s$ reachable in $P$, and results
 in a value denoted as $\phi_f(s)$. 
@@
 Symbols like $x$ denote \emph{parameters} whose value depends on the
-instance $P$ in $\Q$. For example, if $\Q_{clear}$ denotes the Blocksworld
-instances with goals of the form $clear(x)$, the value of $x$ in an instance
-$P$ with goal $clear(a)$ is $a$.
+instance $P$ in $\Q$. For example, if $\Q_{clear}$ denotes the Blocksworld (BW)
+instances with goals of the form $clear(x)$, then a problem 
+$P$ with goal $clear(a)$ will belong to $\Q_{clear}$ and 
+the value of $x$ in $P$ will be $a$.
+
+While instances $P$ in a generalized problem $\Q$ would normally share some structure \cite{bonet:ijcai2015},
+like the same planning domain and the same predicate symbols, this is not strictly  necessary.
+On the other hand,  the features $f$  must be common to all the instances  in $\Q$ and 
+represent functions $\phi_f(s)$ that are well defined over all reachable states $s$.
 
@@
 
-An \emph{abstract action}  for a generalized problem $\Q$ and a set of features $F$
+An \emph{abstract action}  for a generalized problem $\Q$ and a set $F$ of features
 is a pair $\bar{a}=\abst{Pre}{\Eff}$ where $Pre$ and $\Eff$ are the  preconditions
-and effects expressed in terms of a  set $V_F$ of boolean and numerical \emph{state variables}
-associated with the features in $F$. Preconditions and effects over boolean state variables $p$ are literals of the form $p$
-and $\neg p$, that abbreviate the  atoms $p=true$ and $p=false$, while  over  numerical state variables $n$,
-preconditions are of  the form $n=0$ and $n > 0$, and effects  of the form $n\mminus$ (decrements) and
-$n\pplus$ (increments).
-
-Features $f$ refer to \emph{state functions} $\phi_f(s)$ in the instances
-of the generalized problem $\Q$, but to \emph{state variables} in the
-abstraction. Likewise, an \emph{abstract state} is  a truth valuation over
-the state variables $p$ associated with the boolean features and the atoms
-$n=0$ over the state variables $n$ associated with the numerical features.
+and effects expressed in terms of the  set $V_F$ of boolean and numerical \emph{state variables}
+associated with the features. Preconditions and effects over boolean state variables $p$ are literals of the form $p$
+and $\neg p$, that abbreviate the  atoms $p=true$ and $p=false$, while
+preconditions over  numerical state variables $n$ are of  the form $n=0$ and $n > 0$,
+and effects of the form  $n\mminus$ (decrements) and  $n\pplus$ (increments).
+The language for the abstraction,  that combines boolean with numerical variables
+that can be decreased or increased by unspecified amounts, is the language of the \emph{qualitative numerical problems (QNPs)},
+that unlike standard numerical planning problems \cite{helmert:numerical-undecidable}
+are decidable and can be solved effectively by means of  FOND planners \cite{srivastava:aaai2011,bonet:ijcai2017}.
+
+Features $f$ refer to \emph{state functions} $\phi_f(s)$ in the instances $P$ 
+of the generalized problem $\Q$, but to \emph{state variables} in  the abstraction.
+Likewise, an \emph{abstract state} is  a \emph{truth valuation} over
+the atoms $p$ and $n=0$ defined over the state variables $p$ and $n$
+associated with the boolean and  numerical features.
 The abstract state $\bar{s}$ that \emph{corresponds} to a concrete state $s$
-is the truth valuation that makes $p$ true iff $\phi_p(s)=true$, and $n=0$
-true iff $\phi_n(s) = 0$. An abstract action $\bar{a}$ is applicable in
-a state $s$ if its preconditions are satisfied in $\bar{s}$.
-
-A set of abstract actions $A_F$ is \emph{sound} relative to a generalized
-problem $\Q$ iff for any reachable state $s$ over an instance $P$ in $\Q$
-and any abstract action $\bar{a}$ in $A_F$ that is applicable in $\bar{s}$,
-there is a concrete action $a$ in $P$ that is applicable in $s$ and that
-has the \emph{same qualitative effects as} $\bar{a}$ over the features.
-The action $a$ \emph{represents} or \emph{instantiates} the abstract action
-$\bar{a}$ on $s$. 
-%
-Similarly, a set of abstract actions $A_F$ is \emph{complete} iff for any
-reachable state $s$ over an instance $P$ in $\Q$ and any concrete action $a$
+in an instance $P$ of $\Q$ is the truth valuation that makes $p$ true iff $\phi_p(s)=true$,
+and $n=0$ true iff $\phi_n(s) = 0$. An abstract action $\bar{a}$ is applicable in  $s$
+if its preconditions
+are true in $\bar{s}$.
+
+An action $a$ and an abstract action $\bar{a}=\abst{Pre}{\Eff}$ are said to \emph{have  the same qualitative effects over the features $F$
+in a state $s$}, when they are both applicable in $s$,  and have the same effects on the boolean features $p$ and
+the same \emph{qualitative effects} on the numerical features $n$. Namely, if 1)~$p$ becomes true in the state $s'$
+that results from $a$ in $s$, i.e., $\phi_p(s')$ is true and $\phi_p(s)$  is false,
+iff $p \in \Eff$ and $\neg p \in Pre$, 2)~$p$ becomes false in $s'$, i.e., $\phi_p(s')$ is false and $\phi_p(s)$  is true,
+iff $\neg p \in \Eff$ and $p \in Pre$, 3)~$n$ increases in $s'$, i.e., $\phi_n(s') > \phi_n(s)$, iff $n\pplus \in \Eff$,
+and 4)~$n$ decreases  in $s'$, i.e., $\phi_n(s') < \phi_n(s)$, iff $n\mminus \in \Eff$.
+
+\begin{example}
+Let $\Q_{clear}$ stand for  all  Blocksworld instances with stack and unstack actions and goal $clear(x)$,
+and let $F$ represent the two features above,  $F=\{H,n(x)\}$.  If $A$ is the clear block above $x$
+ and the gripper is empty in a reachable state $s$ of some instance $P$ in $\Q_{clear}$, 
+the abstract action $\bar{a}=\abst{\neg H,n(x)>0}{H,n(x)\mminus}$ has the same qualitative
+effects over the features as the action   $a=unstack(A,B)$ in the state $s$:
+$\bar{a}$ and $a$  are both applicable in $s$, and while $\abs{a}$ makes the variable $H$ true and decreases
+$n(x)$, the action $a$ results in a state $s'$ where $\phi_{H}(s')$ becomes true and $\phi_{n(x)}(s')$ is
+smaller than $\phi_{n(x)}(s)$. \qed
+\end{example}
+
+\subsection{Sound and Complete Abstractions}
+
+The notions  of sound and complete sets of actions are key:
+
+\begin{definition}
+A set of abstract actions $A_F$ over the features $F$ is \emph{sound} relative to 
+$\Q$ iff for any reachable state $s$ over an   instance $P$ in $\Q$,
+if an abstract action $\bar{a}$ in $A_F$ is applicable in $\bar{s}$,
+there is an  action $a$ in $P$ that is applicable in $s$ and
+has the \emph{same qualitative effects} over the features as $\bar{a}$.
+\end{definition}
+
+\begin{definition}
+A set of abstract actions $A_F$ is \emph{complete} iff for any
+reachable state $s$ over an instance $P$ in $\Q$ and any  action $a$
 in $P$ that is applicable in $s$, there is an abstract action $\bar{a}$ in
-$A_F$ that is applicable in $\bar{s}$ and that has the \emph{same qualitative
+$A_F$ that is applicable in $\bar{s}$ and  has the \emph{same qualitative
 effects over the features as $a$.}
-
-The actions $a$ and $\bar{a}$ \emph{have the same qualitative effects} in
-a state $s$ when they have the same effects on the boolean features $p$,
-and the same \emph{qualitative effects} on the numerical features $n$; i.e.,
-$a$ decreases (resp.\ increases) the value of the function $\phi_n(s)$ in
-$s$ iff $\bar{a}$ decreases (resp.\ increases) the value of the state
-variable  $n$.
+\end{definition}
+
+\Omit{
+If we say that  action $a$ of an  instance $P$ in $\Q$   \emph{represents} or \emph{instantiates}
+an abstract action $\bar{a}$ in a state $s$ of $P$ when $a$ and $\bar{a}$ are both applicable in $s$
+and have  the \emph{same qualitative effects} over the features, then soundness means that an   applicable abstract action
+can always be instantiated by
+a  concrete action, while completeness  means that
+an applicable concrete actions  is always the instantiation of an abstract action.
+}
+
+Soundness and completeness are the two key  properties that  enable us to reason about the collection of
+instances $P$ in $\Q$ in terms of abstract actions that operate at the level of the  features directly.
 
 \begin{example}
-  For the features $F=\{H,n(x)\}$, the abstract action $\bar{a}=\abst{\neg H,n(x)>0}{H,n(x)\mminus}$ 
-  is sound in the generalized problem $\Q_{clear}$ which consists of all
-  Blocksworld instances with stack and unstack actions and goal $clear(x)$.
-  Indeed, if $s$ is a reachable state in an instance $P$ and $\bar{a}$ is
-  applicable in $\bar{s}$, the gripper must be empty in $s$ and some
-  block must be above $x$. Thus, if $A$ is the clear block above $x$ and
-  $on(A,B)$ is true in $s$, the action $a=unstack(A,B)$ is applicable in
-  $s$ and has the same qualitative effects of $\bar{a}$ over $F$: they
-  both make $H$ true and they both decrease the value of $n(x)$.
-  Likewise, the abstract action $\bar{a'}=\abst{H}{\neg H}$ is also sound. \qed
+  For the features  $F=\{H,n(x)\}$ and  $\Q_{clear}$ above, 
+  the abstract actions    $\bar{a}=\abst{\neg H,n(x)>0}{H,n(x)\mminus}$
+  and $\bar{a'}=\abst{H}{\neg H}$ are  sound but not complete.
+  For this set to be complete,  it will have to contain abstract actions to represent
+  all concrete actions in the problems such as stacking a block above $x$, picking up a block that is not above $x$, and so on.
 \end{example}
 
-For a set of abstract actions to be complete in this example, it must
-contain actions to represent all concrete actions in the problem
-such as stacking a block above $x$, picking up a block that is
-not above $x$, and so on.
-It may well be possible that a particular feature set $F$ cannot support
-a set of abstract actions $A_F$ that is both sound and complete relative
-to $\Q$. Soundness, however, is crucial for deriving general plans as it
-ensures that in any state of any instance of $\Q$, an applicable abstract
-action  can be ``instantiated'' by one or more concrete actions that have
-the same qualitative effects on the features.
+It may well be possible that a particular feature set $F$ does not  support a set of abstract actions $A_F$
+that is both sound and complete relative to $\Q$. Soundness, however, will be  crucial for deriving valid general plans.
 
@@
 3)~$a_i$ is applicable in $s_i$ and $s_{i+1}=f(a_i,s_i)$, and
-4)~$s_n$ is a goal state of $P$, $n=\infty$, or $\pi(\bar{s}_n)$ is either
-undefined, denotes an abstract action that is not applicable in $\bar{s_n}$,
-or no applicable action $a_n$ instantiates $\pi(\bar{s}_n)$.
-The policy $\pi$ solves $P$ iff all the induced trajectories reach the
-goal of $P$.
+4)~$s_n$ is a goal state of $P$  or  $\pi(\bar{s}_n)$ is 
+undefined,  not applicable in $\bar{s_n}$, or no applicable action $a_n$ in $P$ instantiates $\pi(\bar{s}_n)$.
+The policy $\pi$ solves $P$ iff all trajectories induced by $\pi$  reach a goal state of $P$.
 
 \begin{example}
-  A solution for $\Q_{clear}$ with feature set
-  $F=\{H,n(x)\}$ is the policy $\pi$ given by the rules
+  A solution for $\Q_{clear}$ with features  $F=\{H,n(x)\}$ is the policy $\pi$ given by the rules
   $\Rule{\neg H, n(x)>0}{\bar{a}}$ and $\Rule{H, n(x)>0}{\bar{a}'}$
   for the abstract actions $\bar{a}$ and $\bar{a'}$ above.
-  The policy picks blocks above $x$ and puts them aside (not above $x$)
+  The policy picks blocks above $x$ and puts them aside (never  above $x$)
   until $n(x)$ becomes zero. \qed
@@
     extended with initial and goal formulas $I_F$ and $G_F$ over $V_F$
-    to yield a \emph{numerical and non-deterministic planning problem}
-    $Q_F=\tup{V_F,I_F,G_F,A_F}$. For soundness, $I_F$ must be such that
+    to yield the \emph{abstraction}  $Q_F=\tup{V_F,I_F,G_F,A_F}$,
+    which is a QNP. For soundness, $I_F$ must be such that
     the initial states of instances $P$ in $\Q$ all satisfy $I_F$, while
     all states that satisfy $G_F$ must be goal states of $P$.
-    % \footnote{A   state $s$ is said to satisfy $I_F$ or $G_F$ if  $\bar{s}$ does.}
-    % no longer needed
-%
-  \item The problem $Q_F=\tup{V_F,I_F,G_F,A_F}$ is converted into a
+%
+  \item The abstraction $Q_F=\tup{V_F,I_F,G_F,A_F}$ is converted into a
     \emph{boolean FOND} problem $Q'_F=\tup{V'_F,I'_F,G'_F,A'_F}$ by
@@
     the first-order literals $n=0$ by propositional literals  $n=0$, 
-    the effects $n\pplus$ by deterministic effects $n > 0$, and the
+    the effects $n\pplus$ by effects $n > 0$, and the
     effects $n\mminus$ by non-deterministic effects $n>0\,|\,n=0$.
@@
     non-deterministic effects $n>0\,|\,n=0$ in $Q'_F$ are not \emph{fair} but
-    \emph{conditionally fair}: infinite decrements of $n$ ensure
-    $n=0$ eventually \emph{only} when the number of increments of $n$ is finite.
+    \emph{conditionally fair}: infinite decrements of $n$ ensure that 
+    $n=0$ is true eventually but \emph{only} when  the number of increments of $n$ is finite.
     The problem of obtaining solutions of $Q'_F$ that only assume conditional
-    fairness, the so-called terminating solutions in \cite{srivastava:aaai2011},
+    fairness, the so-called terminating solutions \cite{srivastava:aaai2011},
     is mapped into the problem of solving an amended FOND $Q^+_F$ that
-     assumes standard  fairness \cite{bonet:ijcai2017}.\footnote{The
-    translation to $Q^+_F$ is performed with software supplied by the authors.}
-    %available by the authors that fixes bugs in the description found
+    assumes standard  fairness \cite{bonet:ijcai2017}.\footnote{Translator
+    available at  https://github.com/bonetblai/qnp2fond.}
+    % translation to $Q^+_F$ is performed with software supplied by the authors.}
+    %available by the authors that fixes bugs in the description found]
        %in the paper at https://github.com/bonetblai/qnp2fond.}
@@
 
-%The strong cyclic solutions of  $Q^+_F$ computed by FOND planners off-the-shelf   are solutions to the generalized problem $\Q_F$.
+\begin{theorem}[\cite{bonet:ijcai2018}]
+If the abstract actions  $A_F$ are sound relative to the generalized problem $\Q$,   solutions  to $Q^+_F$
+computed by FOND planners off-the-shelf  are solutions to $\Q$.
+\end{theorem}
 
 \begin{example}
-  Let restrict  $\Q_{clear}$  to those  instances where  the gripper is initially empty
+  Let us restrict  $\Q_{clear}$  to those  instances where  the gripper is initially empty
   and there are blocks on top of $x$. For $F=\{H,n(x)\}$  and $A_F=\{\bar{a},\bar{a'}\}$ as above, 
@@
   $Q'_F$ is like $Q_F$ but with $n(x)=0$ regarded as a propositional symbol, $n(x) > 0$
-  as its negation, and the effect $n(x)\mminus$ of  $\bar{a}$ replaced by  $\neg n(x)=0 \, | \, n(x)=0$.
+  as its negation, and the effect $n(x)\mminus$ of  $\bar{a}$ replaced by  $n(x) > 0 \, | \, n(x)=0$.
   Since  no action in $A_F$  increases $n(x)$,   $Q^+_F=Q'_F$, and the strong cyclic solution  of $Q'_F$,
@@
 
-Bonet and Geffner's formulation and computational model rest on 
-features $F$ and  sound abstract actions $A_F$ provided by hand.
-The contribution of this work is to learn them  from samples:
+The formulation and   computational model above, from  \cite{bonet:ijcai2018}, 
+rests  on features $F$ and  abstract actions $A_F$ that are provided by hand.
+The contribution of this work is a \emph{method for  learning   them automatically
+by enforcing a form of soundness and completeness over a set of samples:}
 
@@
 states $s$ appearing \emph{first} in pairs $(s,s')$ must be fully expanded in $\S$;
-namely, all possible transitions $(s,s'')$ must be in the sample as well. We call them
+namely, all possible transitions $(s,s'')$ must be also in the sample. We call them
 the \emph{expanded} states in $\S$.
 
-The features and abstract actions for $\Q$ are  learned by
-enforcing an approximate form of soundness and completeness
-over the samples. Like before,  we take a transition $(s,s')$ and an abstract action $\bar{a}$
-to have  the same qualitative effects over a boolean  feature  iff they both make it true , they both make it false,  or
-they both leave it unchanged,  and over a numerical feature  iff they both decrease it,  they both increase it,  or they both leave it unchanged.
+For defining soundness and completeness \emph{over a sample set}, we say that  a transition $(s,s')$ and an abstract action $\bar{a}$
+have  the same qualitative effects over the features,  exactly when  the action $a$ that maps $s$ into $s'$ and the abstract action
+$\bar{a}$ have the same qualitative effects over the features in $s$:
 
@@
   iff for each transition  $(s,s')$ in  $\S$, there is an abstract action $\bar{a}$ in $A_F$
-  with the same qualitative effects over in $F$ that is applicable in $s$.
+  with the same qualitative effects over $F$ that is applicable in $s$.
 \end{definition}
 
-For a sufficiently large sample set, these approximate notions converge to their exact versions.
+Clearly, for a sufficiently large sample set, the approximate and exact  soundness and completeness notions  converge.
 
@@
 
-For computing the features and abstract actions we define a formula
-$T(\S,\F)$ where ${\cal S}$ is a sample set for $\Q$ and $\cal F$
-is a feature pool defined from the primitive domain predicates
-as described below. The formula is  \emph{satisfiable} iff there is a set of
-features $F \subseteq \F$ and a set of abstract actions $A_F$
-over $F$ that is sound and complete relative to the sample set
-$\S$.
-
-\subsection{SAT Encoding}
+For learning the features and abstract actions from a sample set $\S$ for $\Q$, 
+we define a propositional formula $T(\S,\F)$ where  $\cal F$ represents a large
+pool of candidate features. The formula is  \emph{satisfiable} iff there is
+a set of abstract actions $A_F$ over  $F \subseteq \F$ that is sound and complete
+relative to $\S$.
+
+\subsection{SAT Encoding: $T(\S,\F)$ and $T_G(\S,\F)$}
 
 For each transition  $(s, s') \in \S$ and each  feature $f \in \F$,
-$\Delta_f(s, s') \in \{+, -, \uparrow, \downarrow, \bot\}$ will denote the
+$\Delta_f(s, s') \in \{+, -, \uparrow, \downarrow, \bot\}$  denotes the
 \emph{qualitative change of value} of feature $f$ along the transition $(s, s')$,
@@
     selected features distinguish $s$ from $t$; i.e.\ if $s$ and $t$ disagree
-    on the truth value of a selected boolean feature $p$ or atom $n=0$ for a
-    selected numerical feature $n$.
+    on  the truth value of atom  $p$ or  $n=0$ for   selected  feature $p$ or $n$.
     %Since $D_1$ is symmetric, only symbols for $s < t$ are created for some
@@
 \noindent where $f$ ranges over the features in $\F$ with  different qualitative
-values in $s$ and $t$; i.e.\ boolean features $p$ with different truth value
-in $s$ and $t$, and numerical features $n$ with different truth values for
- $n=0$ in $s$ and $t$.
-
-The second formulas encode the meaning of $D_2$:
+values in $\bar{s}$ and $\bar{t}$; namely, the  boolean features $p$ with different
+truth values in $s$ and $t$, and the numerical features $n$ for which 
+the atom $n=0$ has different truth values in $s$ and $t$.
+
+The second class of  formulas encode the meaning of $D_2$:
 %
@@
 The third class of formulas relate $D_1$ and $D_2$ by enforcing \emph{soundness}
-and \emph{completeness} over  the samples.  Since the abstract actions  are not given,
+and \emph{completeness} over  the samples, expressed as
+\Omit{
+Since the abstract actions  are not given,
 the qualitative changes in each  transition $(s,s')$ in $\S$
@@
 This is expressed as
+}
 %
@@
 \noindent where $s$ and $t$ are expanded states in  $\S$, $s'$ is a successor of $s$ in $\S$,
-and $t'$ ranges over the successors of $t$ in $\S$.
-
-\Omit{ % left out: this is for efficiency only, and is not strictly needed. Also, doesn't fit as such with goal-marking
-  Refer to this as an ``efficiency optimization'' ..
-
-  such that $s < t$, and transition $(s,s')$ in $\S$. Due to the symmetry breaking caused by the constraint $s < t$,
-  the following formulas are also needed
-
-\begin{equation}
-  \label{eq:bridge2}
-  \neg D_1(s, t) \, \rightarrow \, \bigvee_{s'} \neg D_2(s, s', t, t')
-\end{equation}
-
-\noindent where the  iteration in the right hand disjunction goes over the different successor states $s'$ of the first argument
-$s$ of $D_1$ that are in the sample set $\S$, $s$ and $t$ are states in $\S_1$ as before with $s < t$, and $t'$ is a
-successor state of $t$ in $\S$; i.e. $(t,t') \in \S$.
-}
-%% --
-
-The last set of clauses force the selected features to distinguish goal
-from non-goal states as:
+and $t'$ ranges over the successors of $t$ in $\S$. This formula is crucial. 
+It says that \emph{if the selected features don't distinguish the state $s$ from $t$,
+then for each action $a$ that maps  $s$ into $s'$, there must be an action $b$ that maps the state
+$t$ into a state $t'$ such that the two transitions $(s,s')$ and $(t,t')$ affect the selected features
+in the same way.} The formula does not mention the actions $a$ and $b$ because their identity does not matter;
+it is just the state transitions that count. In addition, the formula does not involve abstract actions as they
+will be obtained from the transitions and  the satisfying assignment. Indeed, for each transition $(s,s')$ in the sample, there will be
+an abstract action $\bar{a}$ that  accounts for the transition; i.e., with the  same qualitative effects over the selected features.
+Moreover, if $D_1(s,t)$ and $D_2(s,s',t,t')$ are both false in the satisfying assignment, the two transitions $(s,s')$ and $(t,t')$
+will be captured by the same abstract action.
+
+The fourth and  last set of formulas force the selected features to distinguish goal from non-goal states as:
 \begin{equation}
@@
 \noindent where  $s$ and $t$  are expanded states  $\S$ such that exactly one of them is a goal state. 
-For this, it is assumed that expanded states $s$ in $\S$ are labeled as goal or non-goal states.
+% For this, it is assumed that expanded states $s$ in $\S$ are labeled as goal or non-goal states.
+For this, it is assumed that  states $s$ sampled from an instance $P$, are labeled as goal or non-goal states
+of $P$.
+
+\medskip
 
 The theory $T(\S,\F)$ given by formulas \eqref{eq:d1}--\eqref{eq:goal}
-has a number of atoms that is linear in the size $|{\cal F}|$  of the feature pool $\F$,
-and quadratic in the number $m_1$ of expanded states $s$  and
-in the number $m_2$ of transitions $(s,s')$ in $\S$.
-The number of clauses grows with
-$m_1^2 \times (\abs{{\cal F}} + m_2)$, and 
-can be roughly cut in half by eliminating symmetries.
-
-We also consider an alternative theory $T_G(\S,\F)$
-that trades off completeness guarantees for a much smaller theory size, which
-results from not having to consider
-all pairs of expanded states or transitions in $\S$.
-%
-% No need to mark states as well - we just mark transitions, and states are marked implicitly
-%
-For this, some of the transitions $(s,s')$ in $\S$  are marked as being \emph{goal-relevant},
-and completeness is sought \emph{only with respect to these transitions}.
-These goal-relevant transitions are selected for being part of a concrete plan
-that is computed for each instance $P$ in $\Q$. The theory  $T_G(\S,\F)$ is then as
- $T(\S,\F)$, with the sole difference that the $s$ and $s'$ arguments in the $D_1$ and $D_2$ atoms 
-need to be part of some goal-relevant transition.
-% This is achieved by
-% computing some plans for some instances $P$ in $\Q$, and by adding the
-% resulting state transitions to $\S$ at the same time that the first
-% arguments $s$ and $(s,s')$ in the $D_1$ and $D_2$ atoms are restricted
-% to be marked.
+has a number of atoms given by $m^2 * (b^2 + 1) + |\F|$,  where $m$ is the number of expanded
+states in $\S$ and $b$ is their average branching factor (transitions per state),
+and a number of clauses that grows with $m^2 * b^2 * |\F|$.
+%
+\Omit{
+This means  that  $50$ expanded states in the sample set
+with an average branching factor of $10$, and a set $\F$
+with $400$ candidate features, can generate up to $10^8$ clauses.
+While this is a loose upper bound for the number of clauses that
+follow from the implication   $D_2(s, s', t, t') \Leftharrow\ \textstyle\bigvee_f  selected(f)$
+in (\label{eq:d2}) and the numbers are further reduced by symmetry consideration,
+this is still a large number. ..
+}
+%
+
+We also consider an alternative theory $T_G(\S,\F)$ that is similar to $T(\S,\F)$
+but  smaller. It is obtained by marking  some of the transitions $(s,s')$ in $\S$  as  \emph{goal-relevant}.
+Then rather than creating abstract  actions to account for all the transitions, we only create
+abstract actions to account for the marked transitions. This is achieved by drawing the states
+$s$ and the transitions $(s,s')$ in formulas (\ref{eq:d1}) and (\ref{eq:d2}) from the
+set of marked transitions in $\S$. The states $t$ and the transitions $(t,t')$, on the other hand,
+are drawn from the sample set $\S$ exactly as before. This simplification preserves soundness over $\S$
+but the completeness is not over $\S$ but over the marked transitions. 
+The   transitions marked as goal relevant in $\S$  are obtained by computing one plan
+for each sampled instance $P$ from $\Q$.
+
+\Omit{
 If the set of marked transitions is much smaller than the full sample set, %as will typically be the case,
@@
 Yet this does not affect the resulting formal guarantees, which depend only on soundness. 
+}
 
@@
 decreases (resp.\ increases) in the transition from $s$ to $s'$.
-If two abstract actions $\bar{a}$ and $\bar{a}$ differ only in the sign of a
-precondition, the precondition can be dropped, removing the duplicates and
-iterating until no more simplifications can be done. Their key properties are:
+Duplicate abstract actions are removed, and if two abstract actions $\bar{a}$ and $\bar{a}'$
+differ only in the sign of a precondition, the two abstract actions are    merged into one with
+such a  precondition dropped. The key properties of $T(\S,\F)$ and $T_G(\S,\F)$ are:
 
@@
 \begin{theorem}
-  If $\sigma$ is a satisfying assignment of $T(\S,\F)$, 
-  $A_{\sigma}$ is sound and complete relative to $\S$.
+  If $\sigma$ is a satisfying assignment of $T(\S,\F)$,   $A_{\sigma}$ is sound and complete relative to $\S$.
 \end{theorem}
@@
 \begin{theorem}
-  If $\sigma$ satisfies  $T_G(\S,\F)$, $A_{\sigma}$  is sound relative to $\S$.
+  If $\sigma$ satisfies  $T_G(\S,\F)$, $A_{\sigma}$  is sound relative to $\S$ and complete relative to the marked
+  transitions in $\S$. 
 \end{theorem}
 
+While the initial and goal expressions $I_F$ and $G_F$ of the abstract problem can be learned 
+from the satisfying assignment $\sigma$ as well, for simplicity, in the examples below, they are
+provided by hand.\footnote{$G_F = G_{\sigma}$ can be defined as the DNF formula
+whose terms  correspond to the abstract states over the selected features in $F_{\sigma}$ 
+that are true in some  goal state of the sample  and false in all non-goal states.
+This DNF formula can also  be simplified in standard ways.}
+
 \Omit{
-%% commented out
-The goal expression $G_F$ can be provided explicitly  in terms of the primitive predicates,
-but  more generally can be extracted  from the satisfying assignment $\sigma$ as well.
-For this, $G_F = G_{\sigma}$ is defined as the DNF formula whose terms 
-correspond to the abstract states over the selected features in $F_{\sigma}$ 
-that are false in all expanded states in $\S$ marked as non-goals
-and true in some expanded goal state. This DNF formula can   be simplified in standard ways. 
 Due to  \eqref{eq:d1} and \eqref{eq:goal} that force the selected features in $F_{\sigma}$
@@
 %
-Different satisfying assignments $\sigma$ give rise to different features $F_{\sigma}$
+
+Different assignments $\sigma$ give rise to different features $F_{\sigma}$
 and abstract actions $A_{\sigma}$. More meaningful features are found by looking for
 satisfying assignments $\sigma$ that minimize a cost measure such as $|F_{\sigma}|$
-or, more generally, $\sum_{f \in F_{\sigma}} cost(f)$. We achieve this
-optimization by casting the problem as a 
-% through the use of Min SAT or 
-a Weighted Max-SAT problem where clauses resulting from the above theories are taken as \emph{hard} clauses
-and additional \emph{soft} unit clauses $\neg selected(f)$ are given weight $cost(f)$, for $f \in \mathcal{F}$.
+or, more generally, $\sum_{f \in F_{\sigma}} cost(f)$. We achieve this by casting our problem
+into a constrained optimization problem, and more especifically, a Weighted Max-SAT problem,
+where clauses resulting from the theories $T$ or $T_G$ are taken as \emph{hard} clauses
+and \emph{soft} unit clauses $\neg selected(f)$ are added with  positive weights $cost(f)$
+for all $f \in \F$.
 
@@
 
-The computation of the feature pool $\F$ used in $T(\S,\F)$
-relies on the predicates used to encode the instances in $\Q$.
-A simple grammar allows  us to  construct new predicates 
-and to define boolean and numerical features in $\F$
-from them. 
-In order to construct these predicates, we use an off-the-shelf solution:
-\emph{description logics} (DL),  also called \emph{concept languages}, which are well suited to this task.
-In DL, \emph{concepts} $C$  refer to unary predicates, and \emph{roles} $R$  to binary predicates.
-Concept languages have been used before for  generating general policies using purely learning methods  \cite{martin-geffner:generalized,fern:bias}.
-\citeay{dl-handbook} give syntax and semantics of description logics.
-
-\subsection{Grammar}
-
-We use a standard  DL grammar given by the following rules:
+The feature pool $\F$ used in $T(\S,\F)$ is obtained 
+from  predicates used to encode the instances in $\Q$.
+For this, these predicates are assumed to be common to
+all such instances, and to have arity no greater than 2.
+From these \emph{primitive predicates} and some composition rules,
+we define a large set of \emph{derived predicates $r$ of arity $1$}
+whose denotation $r^s$ in a state $s$ over an instance $P$
+refers to the set of constants (objects) $c$ in $P$ that
+have the property $r$ in $s$; i.e., $r^s=\{c \, | \, r(c) \hbox{ is true in $s$}\}$.
+Boolean and numerical features $p_r$ and $n_r$ can then be defined from  $r$,
+denoting the functions $\phi_{p_r}(s)$ and $\phi_{n_r}(s)$: the first is $0$
+if $|r^s|=0$ and else is $1$, the second represents the cardinality of $r$ in $s$,
+i.e., $|r^s|$.
+
+New unary predicates $r(x)$ can be created  defined by definitions of the form $r(x)$ iff $\exists y. q(x,y) \land t(y)$
+where $q$ and $t$ are primitive predicates of arity $2$ and $1$. we  use   instead the syntax and  grammar rules
+of  description logics \cite{dl-handbook}.  In description logics, unary and binary predicates
+are referred to as \emph{concepts} $C$  and \emph{roles} $R$, and they have been used before
+for capturing general policies using purely learning methods  \cite{martin-geffner:generalized,fern:bias}.
+
+\subsection{Concepts}
+
+The concepts and roles denoted as $C_p$ and $R_p$ represent the primitive predicates of arity 1 and 2 respectively.
+The grammar for generating new concepts and roles from them is:
+
 \begin{enumerate}[{\small$\bullet$}]
-  \item $C \ \leftarrow\ \ p, C_u, C_x$: primitive, universal, nominals: denotation of $C_u$ is whole universe, denotation of $C_x$, for parameter $x$, is $\set{x}$.
-  \item $C \ \leftarrow\  \neg C, C \sqcap C'$: negation and conjunction.
-  \item $C \ \leftarrow\  \exists R.C , \forall R.C$: concepts that use of  roles: first denotes $x$ s.t.\ for some  $y$,  $R(x,y) \land C(y)$.
-  Second,  $x$ s.t.\ for all $y$, $R(x,y) \rightarrow C(y)$.
-  \item $C \ \leftarrow\  R=R'$: stands for $x$ s.t.\ set of $y$'s  s.t. $R(x,y)$ equal to set of $y$'s s.t. $R'(x,y)$.
-  \item $R \leftarrow R_p, R^{-1}, R^*$: primitive, inverse, transitive closure.
+  \item $C_A \ \leftarrow\ \ C_p, C_u, C_x$,  primitive, universal, nominals: $C_u$ denotes whole universe,
+   $C_x$ denotes $\set{x}$ for parameter $x$ if any,
+  \item $C \ \leftarrow\  \neg C_A$,   negation on primitive, universal, nominals,
+  \item $C \ \leftarrow\   C \sqcap C'$,   conjunctions,
+  \item $C \ \leftarrow\  \exists R.C, \forall R.C$, existential and universal roles:  first denotes $x$'s  s.t.\  $R(x,y) \land C(y)$ for some  $y$, 
+  second denotes $x$'s  s.t. $R(x,y) \rightarrow C(y)$ for all $y$,
+  \item $C \ \leftarrow\  R=R'$, stands for $x$'s  s.t. $R(x,y)$ iff $R'(x,y)$ for all $y$,
+  %  set of $y$'s  for which $R(x,y)$ equal to set of $y$'s s.t. $R'(x,y)$.
+  \item $R \leftarrow R_p, R_p^{-1}, R_p^*, [R_p^{-1}]^*$: primitive, inverse, and transitive closure of both.
 %     \item $R \leftarrow R \circ R'$, role composition: $(x,y)$ in $R \circ R'$ if $(x,z)$ in $R$ and $(z,y)$ in $R'$.
@@
 
-The primitive predicates $p$ used in the grammar must be common to all instances $P$ of the generalized problem $\Q$.
-%They will usually be common domain predicates.
-Given that parameters $x$ are common as well,
-the compositional semantics of DL defines, for any concept $C$ and role $R$ resulting
-from these rules and any state $s$ of $P$, a denotation $C^s$ ($R^s$)
-\emph{that can be determined from the denotation of the primitive predicates in $s$ alone}.
-Below we see examples of derived concepts such as $\exists\, on^* . C_x$, whose denotation
-in any Blocksworld state $s$ is the set of blocks that are above $x$ in $s$.
-The numerical features $n_C$ are defined by the cardinality of such denotations $C^s$.
-
-\subsection{Pool}
-
-We define the \emph{complexity} of a concept or role as the minimum
-number of the above rules needed to generate it. We denote by $\G^k$
-the set of all concepts and roles with complexity no larger than $k$.
-We prune from $\G^k$ \emph{redundant} concepts whose denotation
-coincides with that of another concept over all sampled states
-in $\S$. 
-%
-The pool of features $\F=\F^k$  is defined from concepts $C \in \G^k$, and contains
-the following features $f$ that represent functions $\phi_f(s)$:
+The denotations   $C^s$  and $R^s$ of derived concepts and roles follows from the rules and the denotation of primitive
+concepts and roles. For example, for the concept  $C: \exists\, on^* . C_x$ in  the  Blocksworld, 
+$C^s$  denotes  the set of objects  that are above block $x$. The numerical feature
+$n_C$ associated with $C$  represents the number of such objects $|C^s|$.
+
+\subsection{Candidate Features}
+
+The \emph{complexity} of a concept or role is the minimum number of grammar  rules needed to generate it.
+The set of concepts and roles with complexity no greater than  $k$ is referred to as  $\G^k$.
+When generating  $\G^k$,  redundant  concepts are incrementally pruned. A concept is deemed
+redundant when its  denotation coincides with the denotation of a previously generated concept 
+over all  states in $\S$.  The set $\F=\F^k$, defined from $G^k$, contains the following candidate features:
 \begin{enumerate}[{\small$\bullet$}]
@@
   $n$ such that for $x_1, \ldots, x_n$, $C_1^s(x_1)$, $C_2^s(x_{n})$, and
-  $(R{:}C)(x_i,x_{i+1})$ for $i=1, \ldots, n$. 
+  $(R{:}C)(x_i,x_{i+1})$ for $i=1, \ldots, n$. The denotation $(R{:}C)^s$ 
+  contains all pairs $(x,y)$ in $R^s$ s.t. $y$ is in $C^s$.
 \end{enumerate}
 
-The denotation $(R{:}C)^s$ of role $R{:}C$ contains all pairs $(x,y)$ in $R^s$
-s.t. $y$ is in $C^s$.
-We also prune from this space of features all features whose denotation is either constant or coincident with that of another feature over all states in $\S$.
+% We also prune from this space of features all features whose denotation is either constant
+% or coincident with that of another feature over all states in $\S$.
 The measure $cost(f)$ is set to the complexity of $C$ for $n_C$ and $b_C$,
-to $0$ for $b_p$ and to the sum of complexities of $C_1$, $R$, $C$, and $C_2$, for $\textit{dist}(C_1,R{:}C,C_2)$.
-Only features with cost bounded by $k$ are allowed in $\F$.
+to $0$ for $b_p$,  and to the sum of the complexities of $C_1$, $R$, $C$, and $C_2$, for $\textit{dist}(C_1,R{:}C,C_2)$.
+Only features with cost bounded by $k$ are allowed in $\F^k$.
 
@@
 
-In summary, the steps for computing general plans are:
+The steps for computing general plans are then:
 \begin{enumerate}[1.]
@@
     pruning),
-  \item a satisfying assignment $\sigma$ of the theory $T(\S,\F)$ or $T_G(\S,\F)$
+  \item a satisfying assignment $\sigma$ of $T(\S,\F)$ or $T_G(\S,\F)$
     that minimizes $\sum_{f \in F_{\sigma}} cost(f)$ is obtained using a
-    Weighted-MAX SAT solver,
+    Max SAT solver,
   \item the features $F$ and abstract actions $A_F$ are extracted from $\sigma$
     as $F_{\sigma}$ and $A_{\sigma}$,
-  \item the problem $Q_F=\tup{V_F,I_F,G_F,A_F}$ is defined with initial and goal
+  \item abstraction $Q_F=\tup{V_F,I_F,G_F,A_F}$ is defined with initial and goal
     conditions $I_F$ and $G_F$ provided by hand to match $\Q$, 
-  \item the FOND problem $Q^+_F$, constructed automatically from $Q_F$, is solved by an off-the-shelf
+  \item the FOND problem $Q^+_F$, constructed automatically from $Q_F$ is solved by an off-the-shelf
     FOND planner.
@@
 
-Notice that the policy $\pi$ that results from the last step deals with
-propositional symbols that correspond to atoms $p$ and $n=0$ for boolean
-and numerical features $p$ and $n$ in $F$.
-When this policy is applied to an instance $P$ of $\Q$, however, such features denote state functions
-which are determined by the  semantics of the concepts and roles involved
-(e.g., $n_C$ stands for feature function $\phi_{n_c}(s)=|C^s|$).
+The policy $\pi$ that results from the last step deals with propositional symbols that correspond to atoms $p$ and $n=0$ for boolean
+and numerical features $p$ and $n$ in $F$. When this policy is applied to an instance $P$ of $\Q$, the  value of the features $f$
+is obtained from the functions $\phi_f(s)$ that they denote. For example, if $n=n_C$ for a concept $C$, then $\phi_{n}(s)=|C^s|$.
 
@@
 \begin{theorem}
-If the abstract actions $A_F$ that are sound relative to the sample set $\S$  are actually sound,
+If the abstract actions $A_F$ that are sound relative to the sample set $\S$  are  sound,
 then a   policy $\pi$ that solves  $Q^+_F$  is guaranteed to solve  all instances $P$ of $\Q$
@@
 
-This means that a policy $\pi$ obtained with a FOND planner from $Q^+_F$ 
-solves the generalized problem $\Q$, provided that the abstract actions
-learned are sound relative to $\Q$ and that $\Q$ is restricted to the instances $P$ whose
-initial and goal conditions comply with   $I_F$ and $G_F$.
-However,  for a policy to be correct, the abstract actions do
-not have to be sound relative to all reachable states $s$ in the instances $P$ of $\Q$, it  suffices
-if they are sound on the states $s$ that are reachable with the policy. We see an example of this below.
+% This means that a policy $\pi$ obtained with a FOND planner from $Q^+_F$ 
+% solves the generalized problem $\Q$, provided that the abstract actions
+% learned are sound relative to $\Q$ and that $\Q$ is restricted to the instances $P$ whose
+% initial and goal conditions comply with   $I_F$ and $G_F$.
+
+\Omit{
+For the policy $\pi$  to be correct, however, the abstract actions do not have to be sound relative
+to \emph{all} reachable states $s$ in the instances $P$ of $\Q$, it  suffices if they are sound on the
+states $s$ that are reachable with the policy. We see an example of this below.
+}
 
@@
 
-We evaluate the computational model on four generalized problems $\Q$. For each $\Q$, we have selected  a few and small  ``training''  instances $P$ in $\Q$ by hand, 
-from which we draw  the sample sets $\cal S$.  $\S$ is constructed by collecting  the first $m$ states
-generated by  a breadth-first search (where $m$ is typically never larger than a few hundreds)
-along with  the states generated in an \emph{optimal} plan. This is done for each
-training instance  $P$. The plans ensure that $\cal S$ contains some goal states, and they are used to mark states and transitions for
-constructing  the theory $T_G({\cal S,\cal F})$, which is the one used in the experiments. $\S$ is closed by fully expanding the states selected
-while the bound $k$ for generating  $\cal F$ is set to $8$. The  distance features are used only in the last problem.
-The Weighted-Max Solver is Open-WBO \cite{martins2014open} and the FOND planner is SAT-FOND \cite{tomas:fond-sat}.
-The translation from $Q'_F$ to $Q^+_F$ is very fast, in  the order of $0.01$ seconds in all cases.  
-The whole computational pipeline summarized by the steps 1--6 above is processed on Intel Xeon E5-2660 CPUs
-with time and memory cutoffs of 1h and 32GB. Table~\ref{tab:exp-result-data} summarizes the relevant data for 
-the problems.
+We evaluate the computational model on four generalized problems $\Q$.
+For each $\Q$, we select a few  ``training''  instances $P$ in $\Q$ by hand, 
+from which   the sample sets $\cal S$ are drawn.  $\S$ is constructed by collecting  the first $m$ states
+generated by  a breadth-first search, along with  the states generated in an \emph{optimal} plan. 
+The plans ensure that $\cal S$ contains some goal states and provide the state transitions that are marked
+as goal relevant when constructing  the theory $T_G({\cal S,\cal F})$, which is the one used in the experiments.
+$\S$ is closed by fully expanding the states selected.  The value of $m$ is chosen so that
+the resulting number of transitions in $\S$, which depends on the branching factor,
+is around 500. The  bound $k$ for $\F=\F^k$ is set to $8$.
+Distance features $dist$  are used only in the last problem. The Weighted-Max Solver is Open-WBO \cite{martins2014open}
+and the FOND planner is SAT-FOND \cite{tomas:fond-sat}. The translation from $Q'_F$ to $Q^+_F$ is very fast, in  the order
+of $0.01$ seconds in all cases.  The whole computational pipeline summarized by the steps 1--6 above is processed on Intel Xeon E5-2660 CPUs
+with time and memory cutoffs of 1h and 32GB. Table~\ref{tab:exp-result-data} summarizes the relevant data for the problems,
+including the size of the CNF encodings corresponding to the theories $T$ and $T_G$.
 
@@
   %}
-  \caption{Results on 4 different benchmarks: $n$ is number of training instances $P$,
+  \caption{Results: $n$ is number of training instances $P$,
     $\abs{\S}$ is number of transitions in $\S$,
@@
     $\abs{\F}$ is size of feature pool,
-    $np$ and $nc$ are the number of propositions and clauses in theories $T(\S,\F)$ and $T_G(\S,\F)$,
-    $t_{\text{SAT}}$ is time for SAT solver on theory $T_G$, 
-    $\abs{F}$ and  $\abs{A_F}$ are the resulting number of 
+    $np$ and $nc$ are  numbers of propositions and clauses in $T(\S,\F)$ and $T_G(\S,\F)$,
+    $t_{\text{SAT}}$ is time for SAT solver on $T_G$, 
+    $\abs{F}$ and  $\abs{A_F}$ are  number of 
     selected features and abstract actions, 
-    $t_{\text{FOND}}$ is time for the FOND planner, and
-    $\abs{\pi}$ is the size of the resulting policy.
+    $t_{\text{FOND}}$ is time for FOND planner, and
+    $\abs{\pi}$ is  size of the resulting policy.
     %$time_{\text{SAT}}$, $T_{\text{FOND}}$: computation time (sec.) of the SAT and FOND solvers, resp.;
@@
 $\Q_{clear}$ contains the Blocksworld instances  with goals of the form  $clear(x)$
-and stack/unstack actions. The primitive predicates, i.e., those appearing in all instances $P$ of $\Q_{clear}$, 
-are $on(\cdot,\cdot)$, $clear(\cdot)$,
-$ontable(\cdot)$, $holding(\cdot)$, and $handempty$. 
-For this problem, a  single training instance $P$  with  5 blocks
-sufficed to learn an abstract model from which a general plan was computed.
-The  set of  features $F$ obtained from the theory $T_G$ is:
+and stack/unstack actions. The primitive predicates, i.e., those appearing in the  instances $P$ of $\Q_{clear}$, 
+are $on(\cdot,\cdot)$, $clear(\cdot)$, $ontable(\cdot)$, $holding(\cdot)$, and $handempty$. 
+For this problem, a  single training instance $P$  with  5 blocks sufficed to learn an abstract model from
+which a general plan was computed. The  set of  features $F$ learned  from the theory $T_G$ is:
 
@@
 
-\noindent The set of abstract actions obtained from $T_G$ is:\footnote{
+\noindent The set $A_F$  of abstract actions learned  from $T_G$ is:\footnote{
 Feature and action names are provided   to make their intuitive meaning explicit.
-Their formal meaning is given by their definitions.}
+The formal meaning of features  follows from their syntactic form.}
 
@@
 
-The extra precondition $\neg X$ is not needed and would  be removed by drawing more samples,
-yet it does not prevent the computation of a solution to $\Q_{clear}$ using  $I_F = \{\neg H, \neg X, n(x) > 0\}$ and $G_F=\{n(x)=0\}$.
-The translator produces a FOND problem in 0.01 seconds that is solved
-by the FOND  planner in 0.46 seconds. The resulting plan implements  a loop that applies
-the pick-above-$x$ action followed by put-aside,  until $x$ becomes clear.
+The  abstraction $Q_F=\tup{V_F,I_F,G_F,A_F}$ for $\Q_{clear}$ is set up with 
+$I_F = \{\neg H, \neg X, n(x) > 0\}$ and $G_F=\{n(x)=0\}$.
+Since no actions increment the variable $n(x)$, 
+$Q^+_F=Q'_F$. The policy that solves the generalized problem $\Q_{clear}$
+is obtained from $Q^+_F$ by the FOND planner in 0.46 seconds. The plan implements
+a loop that applies the pick-above-$x$ action followed by put-aside,
+until $x$ becomes clear.
 
@@
 For this problem, three training instances $P$ were used.
-The set $F$ obtained from $T_G$ contains
+The set $F$ learned  from $T_G$ contains
 % 
@@
 for the number of blocks above $x$ and $y$ respectively.
-The DL  definition of the  features $X$, $n(x)$, and $n(y)$ are as above
-while $E$ is $handempty$ and $G$ stands for $(\exists\,on.C_y)  \sqcap C_x$.
+% The DL  definition of the  features $X$, $n(x)$, and $n(y)$ are as above
+% while $E$ is $handempty$ and $G$ stands for $(\exists\,on.C_y)  \sqcap C_x$.
 % Interestingly, this last feature is the concept-based representation
@@
 %% 4 bool[handempty] 1 holding(a) 0 n(b) 1 on(a,b)_2 1
-The learned set $A_F$ includes actions to pick up the clear
-block above $x$ or $y$, two actions to put aside the block being held,
-and actions to pick block $x$, to place $x$ on $y$, and to place
-$x$ on the table:
+The set of actions  $A_F$ that is learned is:
 \begin{enumerate}[--]
@@
 
-The resulting abstraction with $I_F\!=\!\{ E, \neg X, \neg G, n(x)\!>\!0, n(y)\!>\!0\}$
+The abstraction $Q_F$  with $I_F\!=\!\{ E, \neg X, \neg G, n(x)\!>\!0, n(y)\!>\!0\}$
 and $G_F = \{ \neg X, G, n(x)=0, n(y)\!>\!0\}$, is translated into the FOND problem
-$Q_F^+$  in 0.01 seconds,
-which is solved by the FOND planner  in 7.56 seconds.
-The policy implements a loop  to clear block $x$ as above, 
-followed by a loop to clear block $y$, and finishes by picking
-$x$ and placing it on $y$.
+$Q_F^+$  in 0.01 seconds, which is solved by the FOND planner  in 7.56 seconds.
+The resulting policy solves the generalized problem $\Q_{on}$.
+It implements a loop  to clear block $x$ as above,  followed by a second  loop to clear block $y$,
+picking then block $x$ and placing it on  $y$.
 
 \paragraph{Gripper.}
-$\Q_{gripper}$ contains classical Gripper instances  where a robot with
-some grippers needs to move a number of balls from one room into another
-\emph{target} room $x$. Each gripper may carry one ball
-at a time.
-The standard STRIPS predicates are $at\text{-}robby(l)$, $at\text{-}ball(b,l)$,
+$\Q_{gripper}$ involves  a robot with some grippers
+that needs to move a number of balls from one room into another,
+\emph{target} room $x$. Each gripper may carry one ball at a time.
+The STRIPS predicates are $at\text{-}robby(l)$, $at\text{-}ball(b,l)$,
 $free(g)$, $carry(b,g)$ that denote, respectively, whether the robot
@@
 gripper $g$ carries ball $b$, plus unary type predicates $room$, $ball$,
-and $gripper$. The type predicates are also included.
-
-Pipeline 1--6 is fed  2 training instances $P$ from $\Q_{gripper}$  with 2
+and $gripper$. % The type predicates are also included.
+
+Pipeline 1--6 is fed  two instances $P$ from $\Q_{gripper}$  with two
 rooms each, one with 4 balls and the other with 5. The set of features that
-is obtained from the theory $T_G$ is:
+is learned  from the theory $T_G$ is:
 \begin{enumerate}[--]
@@
 
-\noindent The resulting abstract actions are:
+\noindent The set of  abstract actions $A_F$  that is learned is:
 
@@
 
-Once again, although this set of abstract actions  is not complete,
-it is sufficient for generating a general plan. The abstraction with
-$I_F = \{ B>0, \neg X, G>0, C=0\}$ and $G_F = \{ B=0, G>0, C=0 \}$ is
+% Once again, although this set of abstract actions  is not complete,
+% it is sufficient for generating a general plan.
+The abstraction $Q_F$ with $I_F = \{ B>0, \neg X, G>0, C=0\}$ and $G_F = \{ B=0, G>0, C=0 \}$ is
 translated into the FOND problem $Q^+_F$, which is solved by the
-planner in 171.92 seconds. 
-The resulting policy picks the balls in the source room $y$, one by one, 
-until the grippers are full or there are no more balls in $y$.
-It then moves to room $x$, with one of the two move actions above
+planner in 171.92 seconds. The resulting policy solves $\Q_{gripper}$,
+meaning that it works for any  number of grippers and balls.
+The policy picks the balls in the source room, one by one, 
+until the grippers are full or there are no more balls.
+It then moves to the target room, with one of the two move actions above, 
 according to whether the grippers are fully loaded or not,
-and then drops the balls in room $x$ one by one, until the grippers
-become all empty. It then moves to room $y$ again if there are balls
-still to be moved, and the process repeats. Like before, the extra conditions
-that are not strictly required (in the move actions), do not preclude
-the solution of the problem, that results in a general plan that 
-works  for robots with any  number of grippers and  any number of balls.
+and then drops the balls one by one, until the grippers
+are all empty.  It then moves back to the other if there   are still  balls
+to be moved, and repeating the  process.
+
+% Like before, the extra conditions
+% that are not strictly required (in the move actions), do not preclude
+% the solution of the problem, that results in a general plan that 
+% works  for robots with any  number of grippers and  any number of balls.
 
@@
 $\Q_{reward}$ contains instances where  an agent needs to navigate a rectangular
-grid to pick up  rewards spread in a grid  while 
-avoiding \emph{blocked} cells. A variation of the example
-has been used recently in \cite{garnelo2016towards}.
-The STRIPS instances have the primitive predicates $reward(\cdot)$, $at(\cdot)$, $blocked(\cdot)$
+grid to pick up  rewards spread on a grid  while 
+avoiding \emph{blocked} cells. This is  variation of an
+example  in \cite{garnelo2016towards}. The STRIPS instances have the primitive predicates $reward(\cdot)$, $at(\cdot)$, $blocked(\cdot)$
 and $adjacent(\cdot,\cdot)$ that denote the position of the rewards, of
@@
 Pipeline 1--6  is fed with two instances of the problem of sizes $4 \times 4$
-and $5\times 5$, and two different  given distributions of blocked cells  and rewards.
+and $5\times 5$, and two different  distributions of blocked cells  and rewards.
 From the theory $T_G$, two numerical  features are learned:
@@
   \item $R: \abs{reward}$ (number of remaining rewards), 
-  \item $D: dist(at, adjacent{:}(\neg blocked), reward)$ (distance from current
+  \item $D: dist(at, adjacent{:}\neg blocked, reward)$ (dist. from current
     cell to closest cell with reward, traversing adjacent, unblocked cells only).
 \end{enumerate}
-Likewise, two abstract actions are obtained:
+The set of  abstract actions  learned is:
 \begin{enumerate}[--]
@@
 
-The abstraction with $I_F = \{ R>0, D>0 \}$ and $G_F=\{R=0,D>0\}$
+The resulting  abstraction $Q_F$ with $I_F = \{ R>0, D>0 \}$ and $G_F=\{R=0,D>0\}$
 is translated into the  FOND problem $Q^+_F$ which yields
-a policy  in 1.36 seconds. The policy moves the agent
-one step at a time, towards  the uncollected  reward
-that is closest, as measured by the numerical feature $D$.
-Once the reward is reached, the reward is consumed, and
+a policy  that solves $\Q_{reward}$ in 1.36 seconds. The policy moves the agent
+one step at a time towards  the uncollected  reward that is closest, as measured by the
+numerical feature $D$. Once the reward is reached, the reward is consumed, and
 the process repeats until there are no more rewards to be collected.
 
-\subsection{Observations}
-
-We conclude the section with a few observations.
-
-\paragraph{Soundness.} The abstract action \text{pick-ab-$y$} in $\Q_{on}$
-is sound relative to the sample set $\S$ used  but is not sound relative
-to $\Q_{on}$. In any state $s$ where $x$ is a few blocks above $y$ and clear,
-there is no concrete action that instantiates \text{pick-ab-$y$},
-as picking up the topmost block above $y$ would necessarily make feature $X$ (holding block $x$)
-true, which the abstract action does not. Still, the resulting policy is correct
-for $\Q_{on}$ because such states $s$ are not  visited by the policy
-from the given initial conditions.
-
-\paragraph{Expressivity.} The general DL grammar that we have described
-is adequate for many  domains but may fall short  on others. 
-For example,  in Towers of Hanoi, the first action to be taken
-(whether to move the smallest disk to target or buffer peg) depends
-on the number of disks being even or  odd, which is not a feature
-that  would result  from the primitive predicates used in the STRIPS
-instances and the general grammar. Yet, for some reason,
-Tower of Hanoi is regarded as a puzzle.  For other problems, the  use of \emph{goal predicates};
-namely, copies of some primitive predicates  evaluated in the goal
-(e.g., for solving arbitrary Blocksworld instances) would be  necessary  \cite{martin-geffner:generalized}.
-
-\paragraph{Scalability.} The current bottleneck of the approach is 
-in the size of the SAT  theories used to derive the features
-and the actions. This puts a limit on the complexity of the
-feature grammars, the bound $k$, and the number of state
-transitions in the sample, as they  result otherwise
-in SAT theories that are just too large (see Table~\ref{tab:exp-result-data}).
-This is indeed the reason why have chosen to use the theories $T_G$
+\section{Discussion}
+
+The computational  bottleneck of the approach
+is  in the size of the SAT  theories used to derive the features
+and the actions.
+% This puts a limit on the complexity of the
+% feature grammars, the bound, and the number of state transitions in the sample,
+% as they  result otherwise in SAT theories that are just too large (see Table~\ref{tab:exp-result-data}).
+This is indeed the reason why we have chosen to use the theories $T_G$
 with goal marking. Additional ideas, however, are required
 to improve scalability and to make the computational model
-captured by  steps 1--6 more robust.
-
-\section{Discussion}
+captured by  steps 1--6 more robust. One possibility that we haven't explored
+is the use of incomplete (suboptimal) Max SAT solvers.
+
+A second limitation is the assumption that 
+the features can be obtained from a large pool of candidate  features defined 
+from the primitive predicates and a  general grammar.
+It's not fully clear whether this assumption is reasonable in general.
+A relevant argument made in  \cite{bonet:ijcai2018} is that generalized problems
+over domains with low width \cite{nir:ecai2012}, have compact policies
+in terms of general features that can be computed in polynomial time.
+In addition, for dealing with arbitrary goals, the same grammar  appears
+to be suitable, once  \emph{goal predicates}, namely, copies $p_G$ of the  primitive domain predicates $p$
+that appear in the goal,  added to the set of primitive predicates and are evaluated in the goal
+\cite{martin-geffner:generalized}.
+
+\section{Summary and Future Work}
 
@@
 The number of samples required for obtaining correct general plans is
-very small as the learner does not have to learn how to plan, but
-just identify the features that are relevant for the  planner to track.
-Moreover, unlike purely learning approaches, the scope and correctness
+small as the learner doesn't  have to produce plans; it just
+has to learn   the relevant features for the  planner to track.
+Unlike purely learning approaches, the features and the policies
+are transparent, and the the scope and the correctness
 of the resulting general plans can be assessed.
-We have carried out a number of experiments over several domains to 
-illustrate the versatility of the approach. In the future, we want 
-to study the formal characterization  of the generalized problem $\Q$ and the possibilities of
-testing soundness over $\Q$ automatically, and to improve the scalability of the approach,
-for example, by collecting the samples intelligently and incrementally
-from the states where the computed policy fails, and by reusing the
-features and actions obtained for some goal for more complex goals.
-The learned features may also have other uses.
+
+There is however an interesting  relation to the ideas
+of dimensionality reduction and embeddings in machine learning
+\cite{hamilton:embeddings}. 
+The learned abstraction   maps the  states of the problem instances, 
+whose size is not bounded, into valuations over a small and bounded
+set of features. Moreover,  this is done preserving the   essential properties
+of states; namely,  how they are changed by actions and whether they  are goal states or not.
+The notion of \emph{sound} and \emph{complete} abstract action sets expresses this requirement
+that is captured in the SAT formulation.
+
+An   interesting challenge for the future  is to generalize the proposed methods  to other   contexts,   such as learning  from ``screen states''
+in video games,  where the sampled  states have no known structure and  there are no primitive predicates.
+One way for  achieving this, is  precisely by learning embeddings that yield sound and complete abstractions or
+suitable approximations.
+
+\medskip
+
+\noindent \textbf{Acknowledgements.}
+H. Geffner is partially funded by grant TIN-2015-67959-P, MINECO, Spain.
 

### 2018-11-06 doc/camera-ready/paper.tex
Correct missing bib ref

@@
 that can be decreased or increased by unspecified amounts, is the language of the \emph{qualitative numerical problems (QNPs)},
-that unlike standard numerical planning problems \cite{helmert:numerical-undecidable}
+that unlike standard numerical planning problems \cite{helmert:numerical}
 are decidable and can be solved effectively by means of  FOND planners \cite{srivastava:aaai2011,bonet:ijcai2017}.

### 2018-11-07 doc/camera-ready/paper.tex
Hector's camera ready second iteration

@@
 Our approach differs from first-order approaches in the use of propositional
-planners, and from purely learning approaches, in the \emph{formal guarantees} that
+planners, and from purely learning approaches, in the formal guarantees that
 are characteristic of planning: if the learned abstract actions are  sound,
@@
 and $\neg p$, that abbreviate the  atoms $p=true$ and $p=false$, while
-preconditions over  numerical state variables $n$ are of  the form $n=0$ and $n > 0$,
-and effects of the form  $n\mminus$ (decrements) and  $n\pplus$ (increments).
+preconditions and effects over  numerical state variables $n$ are of  the form $n=0$ and $n > 0$,
+and   $n\mminus$ (decrements) and  $n\pplus$ (increments), respectively.
 The language for the abstraction,  that combines boolean with numerical variables
@@
 of the generalized problem $\Q$, but  to \emph{state variables} in  the abstraction.
-Likewise, an \emph{abstract state} is  a \emph{truth valuation} over
+An \emph{abstract state} is  a \emph{truth valuation} over
 the atoms $p$ and $n=0$ defined over the state variables $p$ and $n$
@@
 
-An action $a$ and an abstract action $\bar{a}=\abst{Pre}{\Eff}$ are said to \emph{have  the same qualitative effects over the features $F$
+An action $a$ and an abstract action $\bar{a}=\abst{Pre}{\Eff}$  \emph{have  the same qualitative effects over the features $F$
 in a state $s$}, when they are both applicable in $s$,  and have the same effects on the boolean features $p$ and
-the same \emph{qualitative effects} on the numerical features $n$. Namely, if 1)~$p$ becomes true in the state $s'$
+the same \emph{qualitative effects} on the numerical features $n$. This means: 1)~$p$ becomes true in the state $s'$
 that results from $a$ in $s$, i.e., $\phi_p(s')$ is true and $\phi_p(s)$  is false,
@@
 Let $\Q_{clear}$ stand for  all  Blocksworld instances with stack and unstack actions and goal $clear(x)$,
-and let $F$ represent the two features above,  $F=\{H,n(x)\}$.  If $A$ is the clear block above $x$
- and the gripper is empty in a reachable state $s$ of some instance $P$ in $\Q_{clear}$, 
-the abstract action $\bar{a}=\abst{\neg H,n(x)>0}{H,n(x)\mminus}$ has the same qualitative
-effects over the features as the action   $a=unstack(A,B)$ in the state $s$:
-$\bar{a}$ and $a$  are both applicable in $s$, and while $\abs{a}$ makes the variable $H$ true and decreases
+let $F=\{H,n(x)\}$   represent the two features above, and let $s$ be a state in an instance $P$
+where the gripper is empty, the atoms $on(A,B)$ and $clear(A)$ are true, and block $A$ is above
+$x$. In the state $s$,
+the abstract action $\bar{a}=\abst{\neg H,n(x)>0}{H,n(x)\mminus}$
+and the concrete  action   $a=unstack(A,B)$ in $P$ have the same effect over the features.
+Indeed, $\bar{a}$ and $a$  are both applicable in $s$, and while $\bar{a}$ makes the variable $H$ true and decreases
 $n(x)$, the action $a$ results in a state $s'$ where $\phi_{H}(s')$ becomes true and $\phi_{n(x)}(s')$ is
@@
 
-The notions  of sound and complete sets of actions are key:
+Soundness and completeness are the two key  properties that  enable us to reason about the collection of
+instances $P$ in $\Q$ in terms of abstract actions that operate at the level of the  features:
 
@@
 
-\Omit{
-If we say that  action $a$ of an  instance $P$ in $\Q$   \emph{represents} or \emph{instantiates}
-an abstract action $\bar{a}$ in a state $s$ of $P$ when $a$ and $\bar{a}$ are both applicable in $s$
-and have  the \emph{same qualitative effects} over the features, then soundness means that an   applicable abstract action
-can always be instantiated by
-a  concrete action, while completeness  means that
-an applicable concrete actions  is always the instantiation of an abstract action.
-}
-
-Soundness and completeness are the two key  properties that  enable us to reason about the collection of
-instances $P$ in $\Q$ in terms of abstract actions that operate at the level of the  features directly.
+If we say that  action $a$  \emph{instantiates} an abstract action $\bar{a}$ in a state $s$
+and that $\bar{a}$ captures $a$ in $s$,  when $a$ and $\bar{a}$ are both applicable in $s$ and have  the same qualitative effects over the features,
+then soundness means that an   applicable abstract action can always be instantiated by
+a  concrete action, while completeness  means that an applicable concrete action is always captured 
+by an  applicable abstract action.
+
+It may well be possible  that a  feature set $F$ does not  support a set of abstract actions $A_F$
+that is both sound and complete relative to $\Q$. Soundness, however, is  crucial for deriving valid general plans.
+An important intuition is that the features $F$ support a set of sound abstraction action $A_F$ over $F$ when  the
+effects of the concrete actions over $F$  can be \emph{predicted} in the following sense:  there are no pairs of
+state transitions $(s,s')$ and $(t,t')$  over instances in $\Q$ where a feature $f$ in $F$ changes in  different
+ways (e.g., $f$ becomes true in one transition  but not in the other  or $f$ increases in one transition  but not in the other), 
+when all the  features start with the same values in $s$ and $t$ (i.e., $\bar{s}=\bar{t}$).
+This intuition will be made explicit  in the SAT encoding below  when enforcing 
+a  notion  of soundness and completeness over sampled sets of transitions.
 
 \begin{example}
-  For the features  $F=\{H,n(x)\}$ and  $\Q_{clear}$ above, 
-  the abstract actions    $\bar{a}=\abst{\neg H,n(x)>0}{H,n(x)\mminus}$
-  and $\bar{a'}=\abst{H}{\neg H}$ are  sound but not complete.
-  For this set to be complete,  it will have to contain abstract actions to represent
-  all concrete actions in the problems such as stacking a block above $x$, picking up a block that is not above $x$, and so on.
+  The set of  abstract actions  $\bar{a}=\abst{\neg H,n(x)>0}{H,n(x)\mminus}$
+  and $\bar{a'}=\abst{H}{\neg H}$ over the features  $F=\{H,n(x)\}$
+  is sound but not complete relative to $\Q_{clear}$. 
+  This captures the concrete actions that pick blocks from above $x$ and puts them away
+  but no  other actions like stacking a block above $x$ or picking up a block that is not above $x$. \qed
 \end{example}
-
-It may well be possible that a particular feature set $F$ does not  support a set of abstract actions $A_F$
-that is both sound and complete relative to $\Q$. Soundness, however, will be  crucial for deriving valid general plans.
 
@@
 \begin{example}
-  A solution for $\Q_{clear}$ with features  $F=\{H,n(x)\}$ is the policy $\pi$ given by the rules
+  A solution for $\Q_{clear}$ with   $F=\{H,n(x)\}$ is the policy $\pi$ given by the rules
   $\Rule{\neg H, n(x)>0}{\bar{a}}$ and $\Rule{H, n(x)>0}{\bar{a}'}$
@@
   $Q'_F$ is like $Q_F$ but with $n(x)=0$ regarded as a propositional symbol, $n(x) > 0$
-  as its negation, and the effect $n(x)\mminus$ of  $\bar{a}$ replaced by  $n(x) > 0 \, | \, n(x)=0$.
+  as its negation, and the effect $n(x)\mminus$ replaced by  $n(x) > 0 \, | \, n(x)=0$.
   Since  no action in $A_F$  increases $n(x)$,   $Q^+_F=Q'_F$, and the strong cyclic solution  of $Q'_F$,
@@
 states $s$ appearing \emph{first} in pairs $(s,s')$ must be fully expanded in $\S$;
-namely, all possible transitions $(s,s'')$ must be also in the sample. We call them
+namely, all possible transitions $(s,s'')$ must be  in the sample. We call them
 the \emph{expanded} states in $\S$.
@@
 For defining soundness and completeness \emph{over a sample set}, we say that  a transition $(s,s')$ and an abstract action $\bar{a}$
-have  the same qualitative effects over the features,  exactly when  the action $a$ that maps $s$ into $s'$ and the abstract action
-$\bar{a}$ have the same qualitative effects over the features in $s$:
+have  the same qualitative effects over the features,   when in the  state $s$,  the action $a$ that maps $s$ into $s'$ and the abstract action
+$\bar{a}$ have the same qualitative effects over the features:
 
@@
 \noindent where $f$ ranges over the features in $\F$ with  different qualitative
-values in $\bar{s}$ and $\bar{t}$; namely, the  boolean features $p$ with different
+values in $s$ and $t$; namely, the  boolean features $p$ with different
 truth values in $s$ and $t$, and the numerical features $n$ for which 
@@
 has a number of atoms given by $m^2 * (b^2 + 1) + |\F|$,  where $m$ is the number of expanded
-states in $\S$ and $b$ is their average branching factor (transitions per state),
-and a number of clauses that grows with $m^2 * b^2 * |\F|$.
+states in $\S$ and $b$ is their average branching factor (transitions per state).
+The number of clauses grows with $m^2 * b^2 * |\F|$.
 %
@@
 that \emph{capture} all transitions $(s,s')$ in $\S$, in the case of theory $T$,
-and only the goal-relevant transitions $(s,s')$ in $\S$, in the case of $T_G$.
+and the goal-relevant transitions $(s,s')$ in $\S$, in the case of $T_G$.
 % with duplicates and redundancies removed.
@@
 differ only in the sign of a precondition, the two abstract actions are    merged into one with
-such a  precondition dropped. The key properties of $T(\S,\F)$ and $T_G(\S,\F)$ are:
+the precondition dropped. The key properties of $T(\S,\F)$ and $T_G(\S,\F)$ are:
 
@@
 whose terms  correspond to the abstract states over the selected features in $F_{\sigma}$ 
-that are true in some  goal state of the sample  and false in all non-goal states.
-This DNF formula can also  be simplified in standard ways.}
+that are true in some  goal state of the sample  and false in all non-goal states.}
 
@@
 where clauses resulting from the theories $T$ or $T_G$ are taken as \emph{hard} clauses
-and \emph{soft} unit clauses $\neg selected(f)$ are added with  positive weights $cost(f)$
+and \emph{soft} unit clauses $\neg selected(f)$ are added with  weights $cost(f)$
 for all $f \in \F$.
@@
 The feature pool $\F$ used in $T(\S,\F)$ is obtained 
-from  predicates used to encode the instances in $\Q$.
+from the   predicates encoding  the instances in $\Q$.
 For this, these predicates are assumed to be common to
@@
 denoting the functions $\phi_{p_r}(s)$ and $\phi_{n_r}(s)$: the first is $0$
-if $|r^s|=0$ and else is $1$, the second represents the cardinality of $r$ in $s$,
-i.e., $|r^s|$.
-
-New unary predicates $r(x)$ can be created  defined by definitions of the form $r(x)$ iff $\exists y. q(x,y) \land t(y)$
-where $q$ and $t$ are primitive predicates of arity $2$ and $1$. we  use   instead the syntax and  grammar rules
+if $|r^s|=0$ and else is $1$, the second represents the cardinality $|r^s|$ of $r$ in $s$.
+
+New unary predicates $r(x)$ can be created by definitions of the form ``$r(x)$ iff $\exists y. q(x,y) \land t(y)$''
+where $q$ and $t$ are  primitive  predicates of arity $2$ and $1$. We  use   instead the syntax and  grammar
 of  description logics \cite{dl-handbook}.  In description logics, unary and binary predicates
-are referred to as \emph{concepts} $C$  and \emph{roles} $R$, and they have been used before
-for capturing general policies using purely learning methods  \cite{martin-geffner:generalized,fern:bias}.
+are referred to as \emph{concepts} $C$  and \emph{roles} $R$. Description logics  have been used 
+for capturing general policies using purely learning methods in  \cite{martin-geffner:generalized,fern:bias}.
 
@@
 \begin{enumerate}[{\small$\bullet$}]
-  \item $C_A \ \leftarrow\ \ C_p, C_u, C_x$,  primitive, universal, nominals: $C_u$ denotes whole universe,
+  \item $C_A \ \leftarrow\ \ C_p, C_u, C_x$,  primitive, universal, nominals: $C_u$ denotes  universe,
    $C_x$ denotes $\set{x}$ for parameter $x$ if any,
@@
   \item $C \ \leftarrow\   C \sqcap C'$,   conjunctions,
-  \item $C \ \leftarrow\  \exists R.C, \forall R.C$, existential and universal roles:  first denotes $x$'s  s.t.\  $R(x,y) \land C(y)$ for some  $y$, 
-  second denotes $x$'s  s.t. $R(x,y) \rightarrow C(y)$ for all $y$,
-  \item $C \ \leftarrow\  R=R'$, stands for $x$'s  s.t. $R(x,y)$ iff $R'(x,y)$ for all $y$,
+  \item $C \ \leftarrow\  \exists R.C, \forall R.C$, first denotes $x$'s  s.t.\  $R(x,y) \land C(y)$ for some  $y$, 
+  second,  $x$'s  s.t. $R(x,y) \rightarrow C(y)$ for all $y$,
+  \item $C \ \leftarrow\  R=R'$,  $x$'s  s.t. $R(x,y)$ iff $R'(x,y)$ for all $y$,
   %  set of $y$'s  for which $R(x,y)$ equal to set of $y$'s s.t. $R'(x,y)$.
@@
 The denotations   $C^s$  and $R^s$ of derived concepts and roles follows from the rules and the denotation of primitive
-concepts and roles. For example, for the concept  $C: \exists\, on^* . C_x$ in  the  Blocksworld, 
+concepts and roles. For example, for the concept  $C: \exists\, on^* . C_x$ in    Blocksworld, 
 $C^s$  denotes  the set of objects  that are above block $x$. The numerical feature
@@
     as $F_{\sigma}$ and $A_{\sigma}$,
-  \item abstraction $Q_F=\tup{V_F,I_F,G_F,A_F}$ is defined with initial and goal
+  \item the abstraction $Q_F=\tup{V_F,I_F,G_F,A_F}$ is defined with initial and goal
     conditions $I_F$ and $G_F$ provided by hand to match $\Q$, 
@@
     selected features and abstract actions, 
-    $t_{\text{FOND}}$ is time for FOND planner, and
+    $t_{\text{FOND}}$ is time for planner, and
     $\abs{\pi}$ is  size of the resulting policy.
@@
     %$\#Act$, $\#At$: number if actions and atoms in $Q^+_F$;
-    All times are in seconds. %figures that depend on the SAT theory are given for $T_G(\S,\F)$.
+    Times in seconds. %figures that depend on the SAT theory are given for $T_G(\S,\F)$.
   }
@@
 
-The  abstraction $Q_F=\tup{V_F,I_F,G_F,A_F}$ for $\Q_{clear}$ is set up with 
+The  abstraction $Q_F=\tup{V_F,I_F,G_F,A_F}$ for $\Q_{clear}$ is set with 
 $I_F = \{\neg H, \neg X, n(x) > 0\}$ and $G_F=\{n(x)=0\}$.
@@
 The resulting policy solves the generalized problem $\Q_{on}$.
-It implements a loop  to clear block $x$ as above,  followed by a second  loop to clear block $y$,
+It implements a loop  to clear block $x$,  followed by a second  loop to clear block $y$,
 picking then block $x$ and placing it on  $y$.
@@
 $\Q_{gripper}$ involves  a robot with some grippers
-that needs to move a number of balls from one room into another,
-\emph{target} room $x$. Each gripper may carry one ball at a time.
+that needs to move a number of balls from one room into a target
+room $x$. Each gripper may carry one ball at a time.
 The STRIPS predicates are $at\text{-}robby(l)$, $at\text{-}ball(b,l)$,
@@
 meaning that it works for any  number of grippers and balls.
+\Omit{ %% Add if space
 The policy picks the balls in the source room, one by one, 
@@
 and then drops the balls one by one, until the grippers
-are all empty.  It then moves back to the other if there   are still  balls
-to be moved, and repeating the  process.
+are all empty.  It then moves back to the other room if there   are still  balls
+to be moved,  repeating the  process.
+}
 
@@
 In addition, for dealing with arbitrary goals, the same grammar  appears
-to be suitable, once  \emph{goal predicates}, namely, copies $p_G$ of the  primitive domain predicates $p$
-that appear in the goal,  added to the set of primitive predicates and are evaluated in the goal
+to be suitable, once  \emph{goal predicates}, namely, copies $p_G$ of the  primitive predicates $p$
+that appear in the goal,  are added to the set of primitive predicates and are evaluated in the goal
 \cite{martin-geffner:generalized}.
@@
 
-There is however an interesting  relation to the ideas
+There is  an interesting  relation to the ideas
 of dimensionality reduction and embeddings in machine learning
-\cite{hamilton:embeddings}. 
-The learned abstraction   maps the  states of the problem instances, 
+\cite{hamilton:embeddings}. The learned abstraction
+maps the  states of the problem instances, 
 whose size is not bounded, into valuations over a small and bounded
-set of features. Moreover,  this is done preserving the   essential properties
-of states; namely,  how they are changed by actions and whether they  are goal states or not.
+set of features. Moreover,  this is done while preserving the   essential properties
+of states; namely,  how they are changed by actions and whether they denote  goal states or not.
 The notion of \emph{sound} and \emph{complete} abstract action sets expresses this requirement
-that is captured in the SAT formulation.
+that is enforced by  the SAT formulation.
 

### 2018-11-08 doc/camera-ready/paper.tex
spellcheck

@@
 or, more generally, $\sum_{f \in F_{\sigma}} cost(f)$. We achieve this by casting our problem
-into a constrained optimization problem, and more especifically, a Weighted Max-SAT problem,
+into a constrained optimization problem, and more specifically, a Weighted Max-SAT problem,
 where clauses resulting from the theories $T$ or $T_G$ are taken as \emph{hard} clauses
@@
 H. Geffner is partially funded by grant TIN-2015-67959-P, MINECO, Spain.
+G. Franc\`{e}s is funded by the European Research Council project ``State Space Exploration: Principles, Algorithms and Applications''.
 

### 2018-11-09 doc/camera-ready/paper.tex
Minor improvements to the camera ready

--- 2018-11-07 doc/camera-ready/paper.tex
@@
 \usepackage{graphicx}  %Required
-\usepackage{booktabs}  %Required
 
@@
 %\newcommand\bb[1]{\sethlcolor{cyan}\hl{BB: #1}}
-
 \usepackage{amsmath}
@@
 \usepackage{enumerate}
+\usepackage{booktabs}
 %\usepackage{upgreek}
 %\usepackage[normalem]{ulem}
-\usepackage{multirow}
+% \usepackage{multirow}
 % \usepackage[small]{caption}
-
-\usepackage{xspace}
+% \usepackage{xspace}
 % \usepackage{ifthen}
-\usepackage{tikz}
+% \usepackage{tikz}
 % \usetikzlibrary{shapes.geometric}
 % \pgfdeclareimage[width=1cm]{dirt}{dirt2}
-
 % \usetikzlibrary{shapes.geometric}
@@
 
-\def\withproofs{0} % to include proofs in the draft: 0=don't include, 1=include
-\newcommand{\nosuchproof}{\textcolor{red}{\bf No such proof yet...}}
-\newcommand{\problem}{\textcolor{red}{\bf **** PROBLEM ****}}
-\newcommand{\alert}[1]{\textcolor{red}{\bf #1}}
-\newcommand{\CHECK}[1]{\textcolor{red}{\bf Note: #1}}
+% \def\withproofs{0} % to include proofs in the draft: 0=don't include, 1=include
+% \newcommand{\nosuchproof}{\textcolor{red}{\bf No such proof yet...}}
+% \newcommand{\problem}{\textcolor{red}{\bf **** PROBLEM ****}}
+% \newcommand{\alert}[1]{\textcolor{red}{\bf #1}}
+% \newcommand{\CHECK}[1]{\textcolor{red}{\bf Note: #1}}
 
@@
 % Examples
-\newcommand{\Example}{\medskip\noindent\textbf{Example.}\xspace}
+% \newcommand{\Example}{\medskip\noindent\textbf{Example.}\xspace}
 
@@
 
-% \author{Blai  Bonet, G. Frances, H. Geffner} % Alphabetical
-
-\author{
-Blai Bonet$^1$,
-Guillem Franc\`es$^2$,
-Hector Geffner$^3$
-\\
-$^1$ Universidad Sim\'on Bol\'{\i}var, Caracas, Venezuela\\
-$^2$ University of Basel, Basel, Switzerland \\
-$^3$ ICREA \& Universitat Pompeu Fabra, Barcelona, Spain\\
-%
-bonet@usb.ve,
-guillem.frances@unibas.ch,
-hector.geffner@upf.edu
+\author{%
+Blai Bonet\\ Universidad Sim\'on Bol\'{\i}var \\ Caracas, Venezuela \\ bonet@usb.ve
+\And Guillem Franc\`{e}s \\ University of Basel \\ Basel, Switzerland \\ guillem.frances@unibas.ch
+\And Hector Geffner \\ ICREA \& Universitat Pompeu Fabra \\ Barcelona, Spain \\ hector.geffner@upf.edu
 }
 
-% 
+% Alternative style:
+% \author{
+% Blai Bonet, \textsuperscript{1}
+% Guillem Franc\`{e}s,\textsuperscript{2}
+% Hector Geffner,\textsuperscript{3} \\
+% \textsuperscript{1} Universidad Sim\'on Bol\'{\i}var, Caracas, Venezuela\\
+% \textsuperscript{2} University of Basel, Basel, Switzerland \\
+% \textsuperscript{3} ICREA \& Universitat Pompeu Fabra, Barcelona, Spain\\
+% bonet@usb.ve,
+% guillem.frances@unibas.ch,
+% hector.geffner@upf.edu
+% }
+
 % \author{\textbf{Submission \#5695}}
-
-\nocopyright
+% \nocopyright
 
@@
 among  all the instances. More recently, this formulation has been extended by
-\citeay{bonet:ijcai2018} to account for relational domains like Blocksworld  where the set of
+\citeay{bonet:ijcai2018} to account for relational domains like Blocksworld  where the sets of
 objects and  actions change from instance to instance.
@@
 and numerical features $F$ and a set of \emph{abstract actions} $A_F$.
-The  abstract actions are  \emph{sound} and \emph{complete} if they track the
+These abstract actions are  \emph{sound} and \emph{complete} if they track the
 effects  of the actions on  the features in a suitable way.
@@
 finite set of values (boolean or not), $I$ is a set of atoms over $V$ defining an
-initial state $s_0$, $G$ is a set of atoms or literals over $V$ describing the goal
+initial state $s_0$, $G$ is a set of literals over $V$ describing the goal
 states, and $A$ is a set of actions $a$ with their preconditions and effects that
@@
 A \emph{feature} $f$ for a class $\Q$ of problems represents a function $\phi_f$
-that takes an   instance $P$ from $\Q$ and an  state $s$ reachable in $P$, and results
+that takes an   instance $P$ from $\Q$ and a state $s$ reachable in $P$, and results
 in a value denoted as $\phi_f(s)$. 
@@
 While instances $P$ in a generalized problem $\Q$ would normally share some structure \cite{bonet:ijcai2015},
-like the same planning domain and the same predicate symbols, this is not strictly  necessary.
+like the same planning domain and predicate symbols, this is not strictly  necessary.
 On the other hand,  the features $f$  must be common to all the instances  in $\Q$ and 
@@
 and   $n\mminus$ (decrements) and  $n\pplus$ (increments), respectively.
-The language for the abstraction,  that combines boolean with numerical variables
-that can be decreased or increased by unspecified amounts, is the language of the \emph{qualitative numerical problems (QNPs)},
-that unlike standard numerical planning problems \cite{helmert:numerical}
+The language for the abstraction,  that combines boolean and numerical variables
+that can be decreased or increased by unspecified amounts, is the language of \emph{qualitative numerical problems (QNPs)}.
+Unlike standard numerical planning problems \cite{helmert:numerical}, QNPs
 are decidable and can be solved effectively by means of  FOND planners \cite{srivastava:aaai2011,bonet:ijcai2017}.
@@
 An action $a$ and an abstract action $\bar{a}=\abst{Pre}{\Eff}$  \emph{have  the same qualitative effects over the features $F$
-in a state $s$}, when they are both applicable in $s$,  and have the same effects on the boolean features $p$ and
-the same \emph{qualitative effects} on the numerical features $n$. This means: 1)~$p$ becomes true in the state $s'$
-that results from $a$ in $s$, i.e., $\phi_p(s')$ is true and $\phi_p(s)$  is false,
+in a state $s$}, when both are applicable in $s$ and have the same effects on the boolean features $p$ and
+the same \emph{qualitative effects} on the numerical features $n$. If $s'$ is the result of applying $a$ in $s$, this means: 
+1)~$p$ becomes true in $s'$,
+i.e., $\phi_p(s')$ is true and $\phi_p(s)$  is false,
 iff $p \in \Eff$ and $\neg p \in Pre$, 2)~$p$ becomes false in $s'$, i.e., $\phi_p(s')$ is false and $\phi_p(s)$  is true,
@@
 where the gripper is empty, the atoms $on(A,B)$ and $clear(A)$ are true, and block $A$ is above
-$x$. In the state $s$,
+$x$. In this state $s$,
 the abstract action $\bar{a}=\abst{\neg H,n(x)>0}{H,n(x)\mminus}$
@@
 Indeed, $\bar{a}$ and $a$  are both applicable in $s$, and while $\bar{a}$ makes the variable $H$ true and decreases
-$n(x)$, the action $a$ results in a state $s'$ where $\phi_{H}(s')$ becomes true and $\phi_{n(x)}(s')$ is
-smaller than $\phi_{n(x)}(s)$. \qed
+$n(x)$, the action $a$ results in a state $s'$ where $\phi_{H}(s')$ becomes true and $\phi_{n(x)}(s') < \phi_{n(x)}(s)$. \qed
 \end{example}
@@
 Soundness and completeness are the two key  properties that  enable us to reason about the collection of
-instances $P$ in $\Q$ in terms of abstract actions that operate at the level of the  features:
+instances $\Q$ in terms of abstract actions that are defined at the level of the features:
 
@@
 
-If we say that  action $a$  \emph{instantiates} an abstract action $\bar{a}$ in a state $s$
-and that $\bar{a}$ captures $a$ in $s$,  when $a$ and $\bar{a}$ are both applicable in $s$ and have  the same qualitative effects over the features,
-then soundness means that an   applicable abstract action can always be instantiated by
-a  concrete action, while completeness  means that an applicable concrete action is always captured 
-by an  applicable abstract action.
+Let us say that  action $a$ \emph{instantiates} abstract action $\bar{a}$ in a state $s$,
+and $\bar{a}$ \emph{captures} $a$ in $s$,  when $a$ and $\bar{a}$ are both applicable in $s$ and have the same qualitative effects over the features.
+Then, soundness means that any abstract action that is applicable in some reachable state can
+always be instantiated by a concrete action, while completeness  means that any concrete action that is applicable in some reachable state 
+is always captured by an abstract action.
 
 It may well be possible  that a  feature set $F$ does not  support a set of abstract actions $A_F$
-that is both sound and complete relative to $\Q$. Soundness, however, is  crucial for deriving valid general plans.
-An important intuition is that the features $F$ support a set of sound abstraction action $A_F$ over $F$ when  the
-effects of the concrete actions over $F$  can be \emph{predicted} in the following sense:  there are no pairs of
-state transitions $(s,s')$ and $(t,t')$  over instances in $\Q$ where a feature $f$ in $F$ changes in  different
-ways (e.g., $f$ becomes true in one transition  but not in the other  or $f$ increases in one transition  but not in the other), 
-when all the  features start with the same values in $s$ and $t$ (i.e., $\bar{s}=\bar{t}$).
-This intuition will be made explicit  in the SAT encoding below  when enforcing 
-a  notion  of soundness and completeness over sampled sets of transitions.
+that is both sound and complete relative to $\Q$; soundness, however, is  crucial for deriving valid general plans.
+An important intuition is that the features $F$ support a set of sound abstract actions $A_F$ when the
+effects of the concrete actions over $F$ can be \emph{predicted}
+in the sense that the qualitative values of the features in a state $s$ of an instance $P$, 
+as captured by $\bar{s}$, \emph{determine} the possible
+ways in which the actual values of the features may change in the 
+transitions from $s$ to a successor state $s'$.
+Hence, if $a$ and $b$ are two actions applicable in $s$ such that one makes 
+$p$ true and the other increases the value of $n$, then in any other state $t$ 
+such that $\bar{t}=\bar{s}$,
+there should be applicable actions $a'$ and $b'$ with the same effects 
+over $p$ and $n$.
 
@@
   The set of  abstract actions  $\bar{a}=\abst{\neg H,n(x)>0}{H,n(x)\mminus}$
-  and $\bar{a'}=\abst{H}{\neg H}$ over the features  $F=\{H,n(x)\}$
+  and $\bar{a}'=\abst{H}{\neg H}$ over the features  $F=\{H,n(x)\}$
   is sound but not complete relative to $\Q_{clear}$. 
-  This captures the concrete actions that pick blocks from above $x$ and puts them away
-  but no  other actions like stacking a block above $x$ or picking up a block that is not above $x$. \qed
+  They capture the concrete actions that pick blocks from above $x$ and put them away,
+  but do not capture others like stacking a block above $x$ or picking up a block that is not above $x$. \qed
 \end{example}
@@
   $\Rule{\neg H, n(x)>0}{\bar{a}}$ and $\Rule{H, n(x)>0}{\bar{a}'}$
-  for the abstract actions $\bar{a}$ and $\bar{a'}$ above.
+  for the abstract actions $\bar{a}$ and $\bar{a}'$ above.
   The policy picks blocks above $x$ and puts them aside (never  above $x$)
@@
 
-\begin{theorem}[\cite{bonet:ijcai2018}]
-If the abstract actions  $A_F$ are sound relative to the generalized problem $\Q$,   solutions  to $Q^+_F$
+\begin{theorem}[\citeay{bonet:ijcai2018}]
+If the abstract actions  $A_F$ are sound relative to the generalized problem $\Q$, the  solutions  to $Q^+_F$
 computed by FOND planners off-the-shelf  are solutions to $\Q$.
@@
   Let us restrict  $\Q_{clear}$  to those  instances where  the gripper is initially empty
-  and there are blocks on top of $x$. For $F=\{H,n(x)\}$  and $A_F=\{\bar{a},\bar{a'}\}$ as above, 
+  and there are blocks on top of $x$. For $F=\{H,n(x)\}$  and $A_F=\{\bar{a},\bar{a}'\}$ as above, 
   $Q_F=\tup{V_F,I_F,G_F,A_F}$ can be defined with $I_F=\{\neg H, n(x) > 0\}$ and $G_F=\{n(x)=0\}$.
@@
   such that 1)~$P$ is an instance of $\Q$, 2)~$s$ is a reachable state in $P$, 3)~$s'$
-  is a possible successor of $s$ in $P$; i.e., $s'=f(a,s)$ for some action $a$ applicable in $s$,
-  4)~$\S$ is closed in the sense  that if $(s,s',P)$ is in $\S$, then $(s,s'',P)$ is in $\S$  if $s''$ is another
-  successor of $s$ in $P$.
+  is a successor of $s$ in $P$; i.e., $s'=f(a,s)$ for some action $a \in A(s)$,
+  and
+  4)~$\S$ is closed in the sense  that if $(s,s',P) \in \S$, then $(s,s'',P) \in \S$  for any successor $s''$ 
+  of $s$ in $P$.
 \end{definition}
@@
   A set of abstract actions $A_F$ is sound \emph{relative to a  sample set} $\S$ for $\Q$
-  iff for  any  abstract action $\bar{a}$ in $A_F$ that is applicable in  a expanded state $s$ of $\S$,
-  there is a transition $(s,s')$ in $\S$ that has the same qualitative  effects over $F$ as $\bar{a}$.
+  iff for  any  abstract action $\bar{a}$ in $A_F$ applicable in an expanded state $s$ of $\S$,
+  there is a transition $(s,s') \in \S$ with the same qualitative  effects over $F$ as $\bar{a}$.
 \end{definition}
@@
 
-Clearly, for a sufficiently large sample set, the approximate and exact  soundness and completeness notions  converge.
+For a sufficiently large sample set, the approximate and exact notions of soundness and completeness   converge.
 
@@
 
-For learning the features and abstract actions from a sample set $\S$ for $\Q$, 
-we define a propositional formula $T(\S,\F)$ where  $\cal F$ represents a large
-pool of candidate features. The formula is  \emph{satisfiable} iff there is
+In order to learn the features and abstract actions from a sample set $\S$ for $\Q$, 
+we define a propositional formula $T(\S,\F)$, where  $\cal F$ represents a large
+pool of candidate features, that is \emph{satisfiable} iff there is
 a set of abstract actions $A_F$ over  $F \subseteq \F$ that is sound and complete
@@
   \item $D_2(s, s', t, t')$ for each $(s, s')$ and $(t, t')$ in $\S$,
-    true iff selected features $f$ distinguish the two transitions; i.e., if
+    true iff some selected feature $f$ distinguishes the two transitions; i.e., iff
     $\Delta_f(s, s')\not=\Delta_f(t,t')$. 
@@
 \noindent where $f$ ranges over the features in $\F$ with  different qualitative
-values in $s$ and $t$; namely, the  boolean features $p$ with different
-truth values in $s$ and $t$, and the numerical features $n$ for which 
+values in $s$ and $t$; namely, boolean features $p$ with different
+truth values in $s$ and $t$ and numerical features $n$ for which 
 the atom $n=0$ has different truth values in $s$ and $t$.
@@
 
-\noindent where $f$ ranges over the features $f \in \F$ that have the same
+\noindent where $f$ ranges over the features in $\F$ that have the same
 qualitative values in $s$ and $t$ but which change differently in the two
@@
 and $t'$ ranges over the successors of $t$ in $\S$. This formula is crucial. 
-It says that \emph{if the selected features don't distinguish the state $s$ from $t$,
+It says that \emph{if the selected features do not distinguish state $s$ from $t$,
 then for each action $a$ that maps  $s$ into $s'$, there must be an action $b$ that maps the state
@@
 
-The fourth and  last set of formulas force the selected features to distinguish goal from non-goal states as:
+The fourth and  last class of formulas force the selected features to distinguish goal from non-goal states as:
 \begin{equation}
@@
 but  smaller. It is obtained by marking  some of the transitions $(s,s')$ in $\S$  as  \emph{goal-relevant}.
-Then rather than creating abstract  actions to account for all the transitions, we only create
+Then, rather than creating abstract  actions to account for all the transitions, we only create
 abstract actions to account for the marked transitions. This is achieved by drawing the states
@@
 from the satisfying assignment $\sigma$ as well, for simplicity, in the examples below, they are
-provided by hand.\footnote{$G_F = G_{\sigma}$ can be defined as the DNF formula
-whose terms  correspond to the abstract states over the selected features in $F_{\sigma}$ 
-that are true in some  goal state of the sample  and false in all non-goal states.}
+provided by hand.\footnote{
+$G_F = G_{\sigma}$ can be defined as the DNF formula whose terms 
+correspond to the abstract states over 
+$F_{\sigma}$ that correspond to some goal state in the sample.}
+% \footnote{$G_F = G_{\sigma}$ can be defined as the DNF formula
+% whose terms  correspond to the abstract states over the selected features in $F_{\sigma}$ 
+% that are true in some  goal state of the sample  and false in all non-goal states.}
 
@@
 
-Different assignments $\sigma$ give rise to different features $F_{\sigma}$
-and abstract actions $A_{\sigma}$. More meaningful features are found by looking for
-satisfying assignments $\sigma$ that minimize a cost measure such as $|F_{\sigma}|$
-or, more generally, $\sum_{f \in F_{\sigma}} cost(f)$. We achieve this by casting our problem
-into a constrained optimization problem, and more especifically, a Weighted Max-SAT problem,
-where clauses resulting from the theories $T$ or $T_G$ are taken as \emph{hard} clauses
-and \emph{soft} unit clauses $\neg selected(f)$ are added with  weights $cost(f)$
-for all $f \in \F$.
+Different assignments $\sigma$ result in different feature sets $F_{\sigma}$
+and abstract action sets $A_{\sigma}$. Simpler and more meaningful features can be found 
+by minimizing a cost measure such as $|F_{\sigma}|$
+or, more generally, $\sum_{f \in F_{\sigma}} cost(f)$.
+We achieve this by casting our problem
+into a constrained optimization problem, more specifically, a Weighted Max-SAT problem,
+where clauses resulting from the theories $T$ or $T_G$ are taken as \emph{hard} clauses,
+and we add, for each feature $f \in \F$, a
+\emph{soft} unit clause $\neg selected(f)$ with weight $cost(f)$.
+The cost measure we use will be defined below.
 
@@
   \item $C \ \leftarrow\   C \sqcap C'$,   conjunctions,
-  \item $C \ \leftarrow\  \exists R.C, \forall R.C$, first denotes $x$'s  s.t.\  $R(x,y) \land C(y)$ for some  $y$, 
+  \item $C \ \leftarrow\  \exists R.C, \forall R.C$, first denotes $x$'s  s.t.\  $R(x,y) \land C(y)$ for some  $y$;
   second,  $x$'s  s.t. $R(x,y) \rightarrow C(y)$ for all $y$,
@@
     conditions $I_F$ and $G_F$ provided by hand to match $\Q$, 
-  \item the FOND problem $Q^+_F$, constructed automatically from $Q_F$ is solved by an off-the-shelf
+  \item the FOND problem $Q^+_F$, constructed automatically from $Q_F$, is solved by an off-the-shelf
     FOND planner.
@@
 The abstraction $Q_F$  with $I_F\!=\!\{ E, \neg X, \neg G, n(x)\!>\!0, n(y)\!>\!0\}$
-and $G_F = \{ \neg X, G, n(x)=0, n(y)\!>\!0\}$, is translated into the FOND problem
+% and $G_F = \{ \neg X, G, n(x)=0, n(y)\!>\!0\}$,
+and $G_F = \{G\}$, 
+is translated into the FOND problem
 $Q_F^+$  in 0.01 seconds, which is solved by the FOND planner  in 7.56 seconds.
@@
 % it is sufficient for generating a general plan.
-The abstraction $Q_F$ with $I_F = \{ B>0, \neg X, G>0, C=0\}$ and $G_F = \{ B=0, G>0, C=0 \}$ is
+The abstraction $Q_F$ with $I_F = \{ B>0, \neg X, G>0, C=0\}$ and $G_F = \{ B=0, C=0 \}$ is
 translated into the FOND problem $Q^+_F$, which is solved by the
-planner in 171.92 seconds. The resulting policy solves $\Q_{gripper}$,
+planner in 171.92 seconds. The resulting policy provably solves $\Q_{gripper}$,
 meaning that it works for any  number of grippers and balls.
@@
 
-The resulting  abstraction $Q_F$ with $I_F = \{ R>0, D>0 \}$ and $G_F=\{R=0,D>0\}$
+The resulting  abstraction $Q_F$ with $I_F = \{ R>0, D>0 \}$ and $G_F=\{R=0\}$
 is translated into the  FOND problem $Q^+_F$ which yields
@@
 to improve scalability and to make the computational model
-captured by  steps 1--6 more robust. One possibility that we haven't explored
+captured by  steps 1--6 more robust. One possibility that we have not explored
 is the use of incomplete (suboptimal) Max SAT solvers.
@@
 from the primitive predicates and a  general grammar.
-It's not fully clear whether this assumption is reasonable in general.
+It is not fully clear whether this assumption is reasonable in general.
 A relevant argument made in  \cite{bonet:ijcai2018} is that generalized problems
@@
 The number of samples required for obtaining correct general plans is
-small as the learner doesn't  have to produce plans; it just
+small, as the learner does not have to produce plans; it just
 has to learn the relevant features for the planner to track.
 Unlike purely learning approaches, the features and the policies
-are transparent, and the the scope and the correctness
+are transparent, and the scope and correctness
 of the resulting general plans can be assessed.
@@
 
-\medskip
-
-\noindent \textbf{Acknowledgements.}
+\subsubsection{Acknowledgments.}
 H. Geffner is partially funded by grant TIN-2015-67959-P, MINECO, Spain.
+G. Franc\`{e}s is funded by the SSX project, European Research Council.
 

### 2018-11-10 doc/camera-ready/paper.tex
Fix some bib changes

@@
 \newcommand{\Move}{\text{Move}}
+
+\providecommand{\noopsort}[1]{}
 
@@
 % it is sufficient for generating a general plan.
-The abstraction $Q_F$ with $I_F = \{ B>0, \neg X, G>0, C=0\}$ and $G_F = \{ B=0, C=0 \}$ is
+The abstraction $Q_F$ with $I_F\!=\!\{ B\!>\!0, \neg X, G\!>\!0, C\!=\!0\}$ and $G_F\!=\!\{ B\!=\!0, C\!=\!0 \}$ is
 translated into the FOND problem $Q^+_F$, which is solved by the
@@
 \subsubsection{Acknowledgments.}
-H. Geffner is partially funded by grant TIN-2015-67959-P, MINECO, Spain.
+H. Geffner is partially funded by grant TIN-2015-67959-P, MINECO, Spain;
 G. Franc\`{e}s is funded by the SSX project, European Research Council.
@@
 \bibliography{control}
-
 \end{document}

### 2018-11-11 doc/camera-ready/paper.tex
Blai edit

@@
 \usepackage{graphicx}  %Required
+\usepackage{xcolor}  %Required
 
@@
 \newtheorem{corollary}[definition]{Corollary}
-\newtheorem{example}[definition]{Example}
+%\newtheorem{example}[definition]{Example}
+\newenvironment{example}{\noindent\textbf{Example.}}{\qed}
 
@@
 \cite{srivastava:generalized}, and general plans over finite horizons have been
-derived using first-order regression \cite{boutilier2001symbolic,wang2008first,van2012solving}.
+derived using first-order regression \cite{boutilier2001symbolic,wang2008first,van2012solving,sanner:practicalMDPs}.
 Our approach differs from first-order approaches in the use of propositional
@@
 if it results in a numerical value, that we assume to be a non-negative integer.
-$H$ and $n(x)$ represent two features in the Blocksworld:
+For example, $H$ and $n(x)$ represent two features in the Blocksworld:
 the first is a boolean feature that tracks whether the gripper is empty, while
@@
 Symbols like $x$ denote \emph{parameters} whose value depends on the
-instance $P$ in $\Q$. For example, if $\Q_{clear}$ denotes the Blocksworld (BW)
+instance $P$ in $\Q$. For example, if $\Q_{clear}$ denotes the Blocksworld
 instances with goals of the form $clear(x)$, then a problem 
-$P$ with goal $clear(a)$ will belong to $\Q_{clear}$ and 
-the value of $x$ in $P$ will be $a$.
+$P$ with goal $clear(A)$ will belong to $\Q_{clear}$ and 
+the value of $x$ in $P$ will be (the block) $A$.
 
@@
 An action $a$ and an abstract action $\bar{a}=\abst{Pre}{\Eff}$  \emph{have  the same qualitative effects over the features $F$
-in a state $s$}, when both are applicable in $s$ and have the same effects on the boolean features $p$ and
-the same \emph{qualitative effects} on the numerical features $n$. If $s'$ is the result of applying $a$ in $s$, this means: 
+in a state $s$} when both are applicable in $s$ and have the same effects on the boolean features and
+the same \emph{qualitative effects} on the numerical features. If $s'$ is the result of applying $a$ in $s$,
+this means \cite{bonet:ijcai2018}: 
 1)~$p$ becomes true in $s'$,
@@
 
+\medskip
 \begin{example}
 Let $\Q_{clear}$ stand for  all  Blocksworld instances with stack and unstack actions and goal $clear(x)$,
-let $F=\{H,n(x)\}$   represent the two features above, and let $s$ be a state in an instance $P$
-where the gripper is empty, the atoms $on(A,B)$ and $clear(A)$ are true, and block $A$ is above
-$x$. In this state $s$,
+let $F=\{H,n(x)\}$ be the set of the two features above, and let $s$ be a reachable state in an instance $P$
+in $\Q_{clear}$
+where the gripper is empty, the atoms $on(A,B)$ and $clear(A)$ are true, and the block $A$ is above $x$.
+In this state $s$,
 the abstract action $\bar{a}=\abst{\neg H,n(x)>0}{H,n(x)\mminus}$
-and the concrete  action   $a=unstack(A,B)$ in $P$ have the same effect over the features.
+and the concrete  action  $a=\text{Unstack}(A,B)$ in $P$ have the same effect over the features.
 Indeed, $\bar{a}$ and $a$  are both applicable in $s$, and while $\bar{a}$ makes the variable $H$ true and decreases
-$n(x)$, the action $a$ results in a state $s'$ where $\phi_{H}(s')$ becomes true and $\phi_{n(x)}(s') < \phi_{n(x)}(s)$. \qed
+$n(x)$, the action $a$ results in a state $s'$ where $\phi_{H}(s')$ is true and $\phi_{n(x)}(s') < \phi_{n(x)}(s)$.
 \end{example}
@@
 It may well be possible  that a  feature set $F$ does not  support a set of abstract actions $A_F$
-that is both sound and complete relative to $\Q$; soundness, however, is  crucial for deriving valid general plans.
-An important intuition is that the features $F$ support a set of sound abstract actions $A_F$ when the
-effects of the concrete actions over $F$ can be \emph{predicted}
-in the sense that the qualitative values of the features in a state $s$ of an instance $P$, 
-as captured by $\bar{s}$, \emph{determine} the possible
-ways in which the actual values of the features may change in the 
-transitions from $s$ to a successor state $s'$.
+that is both sound and complete relative to $\Q$. Soundness, however, is  crucial for deriving general plans that are valid.
+An important intuition is that the features in $F$ support a set of sound abstract actions $A_F$ when the
+effects of the concrete actions over $F$ can be \emph{predicted};
+that is, the qualitative values of the features in a state $s$ of an instance $P$, 
+as captured by $\bar{s}$, \emph{determine} the possible ways in which the actual
+values of the features may change in the transitions from $s$ to a successor state $s'$.
 Hence, if $a$ and $b$ are two actions applicable in $s$ such that one makes 
 $p$ true and the other increases the value of $n$, then in any other state $t$ 
-such that $\bar{t}=\bar{s}$,
-there should be applicable actions $a'$ and $b'$ with the same effects 
-over $p$ and $n$.
-
+such that $\bar{t}=\bar{s}$, there should be applicable actions $a'$ and $b'$
+with the same qualitative effects on $p$ and $n$.
+
+\medskip
 \begin{example}
-  The set of  abstract actions  $\bar{a}=\abst{\neg H,n(x)>0}{H,n(x)\mminus}$
+  The two abstract actions  $\bar a = \langle \neg H,n(x)>0; H,$ $n(x)\mminus\rangle$ %$\bar{a}=\abst{\neg H,n(x)>0}{H,n(x)\mminus}$
   and $\bar{a}'=\abst{H}{\neg H}$ over the features  $F=\{H,n(x)\}$
@@
   They capture the concrete actions that pick blocks from above $x$ and put them away,
-  but do not capture others like stacking a block above $x$ or picking up a block that is not above $x$. \qed
+  but do not capture other actions like stacking a block above $x$ or picking up a block that is not above $x$.
 \end{example}
@@
 3)~$a_i$ is applicable in $s_i$ and $s_{i+1}=f(a_i,s_i)$, and
-4)~$s_n$ is a goal state of $P$  or  $\pi(\bar{s}_n)$ is 
-undefined,  not applicable in $\bar{s_n}$, or no applicable action $a_n$ in $P$ instantiates $\pi(\bar{s}_n)$.
+4)~$s_n$ is a goal state of $P$, or $\pi(\bar{s}_n)$ is undefined,
+not applicable in $\bar s_n$, or no applicable action $a_n$ in $P$ instantiates $\pi(\bar{s}_n)$.
 The policy $\pi$ solves $P$ iff all trajectories induced by $\pi$ reach a goal state of $P$.
 
+\medskip
 \begin{example}
@@
   The policy picks blocks above $x$ and puts them aside (never  above $x$)
-  until $n(x)$ becomes zero. \qed
+  until $n(x)$ becomes zero.
 \end{example}
@@
     all states that satisfy $G_F$ must be goal states of $P$.
+    \textcolor{red}{\bf ** What is the format for $I_F$ and $G_F$? Conjunctions of lits, DNF, ...? **}
 %
@@
 
+\medskip
 \begin{example}
@@
   Since  no action in $A_F$  increases $n(x)$,   $Q^+_F=Q'_F$, and the strong cyclic solution  of $Q'_F$,
-  that corresponds to the policy above, solves $\Q_{clear}$. \qed
+  that corresponds to the policy above, solves $\Q_{clear}$.
+  \textcolor{red}{\bf ** Why do we need to enter into details about $Q^+_F$?
+  We are using a tool and should not worry about such details.
+  Otherwise, we are conveying the idea that the user should worry about them and
+  then know how the QNP2FOND translation work. **}
 \end{example}
@@
 The formulation and computational model above, from \cite{bonet:ijcai2018}, 
-rests  on features $F$ and  abstract actions $A_F$ that are provided by hand.
+assume that the features $F$ and abstract actions $A_F$ are given.
 The contribution of this work is a \emph{method for  learning   them automatically
@@
   \item $selected(f)$ for each $f \in \F$, true iff $f$ selected (in $F$).
-  \item $D_1(s,t)$ for states $s$ and $t$ expanded in $\S$,  true iff
-    selected features distinguish $s$ from $t$; i.e.\ if $s$ and $t$ disagree
-    on  the truth value of atom  $p$ or  $n=0$ for   selected  feature $p$ or $n$.
+  \item $D_1(s,t)$ for states $s$ and $t$ expanded in $\S$,  true iff the
+    selected features distinguish $s$ from $t$; i.e., if $s$ and $t$ disagree
+    on the truth value of some selected feature, either a boolean or numeric feature,
+    %on  the truth value of atom  $p$ or  $n=0$ for   selected  feature $p$ or $n$.
     %Since $D_1$ is symmetric, only symbols for $s < t$ are created for some
@@
   \item $D_2(s, s', t, t')$ for each $(s, s')$ and $(t, t')$ in $\S$,
-    true iff some selected feature $f$ distinguishes the two transitions; i.e., iff
-    $\Delta_f(s, s')\not=\Delta_f(t,t')$. 
+    true iff some selected feature $f$ distinguishes the two transitions; i.e.,
+    $\Delta_f(s, s')\not=\Delta_f(t,t')$ for some selected feature $f$.
     %As for $D_1$, the number of $D_2$ variables is cut in half due to symmetries.
@@
 values in $s$ and $t$; namely, boolean features $p$ with different
-truth values in $s$ and $t$ and numerical features $n$ for which 
+truth values in $s$ and $t$, and numerical features $n$ for which 
 the atom $n=0$ has different truth values in $s$ and $t$.
@@
 The third class of formulas relate $D_1$ and $D_2$ by enforcing \emph{soundness}
-and \emph{completeness} over  the samples, expressed as
+and \emph{completeness} over  the sample: %, expressed as
 \Omit{
@@
 
-The theory $T(\S,\F)$ given by formulas \eqref{eq:d1}--\eqref{eq:goal}
-has a number of atoms given by $m^2 * (b^2 + 1) + |\F|$,  where $m$ is the number of expanded
-states in $\S$ and $b$ is their average branching factor (transitions per state).
-The number of clauses grows with $m^2 * b^2 * |\F|$.
+The SAT theory $T(\S,\F)$ given by formulas \eqref{eq:d1}--\eqref{eq:goal}
+has $|\F| + m^2 (b^2 + 1)$ propositional variables
+%has a number of propositional variables equal to $m^2 \times (b^2 + 1) + |\F|$
+where $m$ is the number of expanded states in $\S$, and $b$ is their average branching factor (transitions per state).
+The number of clauses is bounded by $m^2(3 + b^2 + |\F|(1+b^2))$. %$m^2 \times b^2 \times |\F|$.
 %
@@
 
-We also consider an alternative theory $T_G(\S,\F)$ that is similar to $T(\S,\F)$
+We also consider an alternative SAT theory $T_G(\S,\F)$ that is similar to $T(\S,\F)$
 but  smaller. It is obtained by marking  some of the transitions $(s,s')$ in $\S$  as  \emph{goal-relevant}.
@@
 \begin{theorem}
-  If $\sigma$ is a satisfying assignment of $T(\S,\F)$,   $A_{\sigma}$ is sound and complete relative to $\S$.
+  If $\sigma$ is a satisfying assignment of $T(\S,\F)$, the set $A_{\sigma}$ of abstract actions over $F_{\sigma}$ is sound and complete relative to $\S$.
 \end{theorem}
@@
 \begin{theorem}
-  If $\sigma$ satisfies  $T_G(\S,\F)$, $A_{\sigma}$  is sound relative to $\S$ and complete relative to the marked
-  transitions in $\S$. 
+  If $\sigma$ is a satisfying assignment of $T_G(\S,\F)$, the set $A_{\sigma}$ of abstract actions over $F_{\sigma}$ is sound relative to $\S$ and complete relative to the marked transitions in $\S$.
+  %If $\sigma$ satisfies  $T_G(\S,\F)$, $A_{\sigma}$  is sound relative to $\S$ and complete relative to the marked transitions in $\S$. 
 \end{theorem}
@@
 from the satisfying assignment $\sigma$ as well, for simplicity, in the examples below, they are
-provided by hand.\footnote{
-$G_F = G_{\sigma}$ can be defined as the DNF formula whose terms 
-correspond to the abstract states over 
-$F_{\sigma}$ that correspond to some goal state in the sample.}
+provided by hand.\footnote{$G_F = G_{\sigma}$ can be defined as the DNF formula whose terms 
+  correspond to the abstract states over $F_{\sigma}$ that correspond to some goal state in
+  the sample. \textcolor{red}{\bf ** Related to previous note **}
+  }
 % \footnote{$G_F = G_{\sigma}$ can be defined as the DNF formula
@@
 \emph{soft} unit clause $\neg selected(f)$ with weight $cost(f)$.
-The cost measure we use will be defined below.
+The cost measure that is used in the experiments is defined next.
 
@@
 The feature pool $\F$ used in $T(\S,\F)$ is obtained 
-from the   predicates encoding  the instances in $\Q$.
-For this, these predicates are assumed to be common to
-all such instances, and to have arity no greater than 2.
+from the   predicates encoding  the instances in $\Q$
+that are assumed to be common to all such instances.
+%For this, these predicates are assumed to be common to
+%all such instances, and to have arity no greater than 2.
 From these \emph{primitive predicates} and some composition rules,
@@
 
-New unary predicates $r(x)$ can be created by definitions of the form ``$r(x)$ iff $\exists y. q(x,y) \land t(y)$''
-where $q$ and $t$ are  primitive  predicates of arity $2$ and $1$. We  use   instead the syntax and  grammar
+Unary predicates $r(x)$ can be created by definitions of the form ``$r(x)$ iff $\exists y[q(x,y) \land t(y)]$''
+where $q$ and $t$ are  primitive  predicates of arity $2$ and $1$ respectively.
+We  use   instead the syntax and  grammar
 of  description logics \cite{dl-handbook}.  In description logics, unary and binary predicates
 are referred to as \emph{concepts} $C$  and \emph{roles} $R$. Description logics  have been used 
-for capturing general policies using purely learning methods in  \cite{martin-geffner:generalized,fern:bias}.
+for capturing general policies using purely learning methods \cite{martin-geffner:generalized,fern:bias}.
 
@@
 
-The concepts and roles denoted as $C_p$ and $R_p$ represent the primitive predicates of arity 1 and 2 respectively.
+The concepts and roles denoted as $C_p$ and $R_p$ represent the \emph{primitive predicates} of arity 1 and 2 respectively.
 The grammar for generating new concepts and roles from them is:
@@
 \begin{enumerate}[{\small$\bullet$}]
-  \item $C_A \ \leftarrow\ \ C_p, C_u, C_x$,  primitive, universal, nominals: $C_u$ denotes  universe,
+  \item $C_A \leftarrow C_p, C_u, C_x$,  primitive, universal, nominals: $C_u$ denotes  universe,
    $C_x$ denotes $\set{x}$ for parameter $x$ if any,
-  \item $C \ \leftarrow\  \neg C_A$,   negation on primitive, universal, nominals,
-  \item $C \ \leftarrow\   C \sqcap C'$,   conjunctions,
-  \item $C \ \leftarrow\  \exists R.C, \forall R.C$, first denotes $x$'s  s.t.\  $R(x,y) \land C(y)$ for some  $y$;
-  second,  $x$'s  s.t. $R(x,y) \rightarrow C(y)$ for all $y$,
-  \item $C \ \leftarrow\  R=R'$,  $x$'s  s.t. $R(x,y)$ iff $R'(x,y)$ for all $y$,
+  \item $C \leftarrow \neg C_A$, negation on primitive, universal, nominals,
+  \item $C \leftarrow C \sqcap C'$, conjunctions,
+  \item $C \leftarrow \exists R.C,\forall R.C$, first denotes $\{x:\!\exists y[R(x,y) \land C(y)]\}$, %  $x$'s  s.t.\  $R(x,y) \land C(y)$ for some  $y$;
+    second $\{x:\forall y[R(x,y)\land C(y)]\}$, %,  $x$'s  s.t. $R(x,y) \rightarrow C(y)$ for all $y$,
+  \item $C \leftarrow R=R'$, denotes $\{ x : \forall y[R(x,y)=R'(x,y)]\}$, %$x$'s  s.t. $R(x,y)$ iff $R'(x,y)$ for all $y$,
   %  set of $y$'s  for which $R(x,y)$ equal to set of $y$'s s.t. $R'(x,y)$.
@@
 $C^s$  denotes  the set of objects  that are above block $x$. The numerical feature
-$n_C$ associated with $C$  represents the number of such objects $|C^s|$.
+$n_C$ associated with concept $C$ represents the number of objects in $C^s$. %$|C^s|$.
+\textcolor{red}{\bf ** $C: \exists\, on^* . C_x$ also counts block $x$ since * is reflexive! **}
 
@@
 redundant when its  denotation coincides with the denotation of a previously generated concept 
-over all  states in $\S$.  The set $\F=\F^k$, defined from $G^k$, contains the following candidate features:
+over all  states in $\S$.  The set $\F=\F^k$, defined from $\G^k$, contains the following features:
 \begin{enumerate}[{\small$\bullet$}]
-\item For each nullary primitive predicate $p$, a \emph{boolean} feature $b_p$ that is true in $s$ iff $p^s$ is true.
-
-\item For each concept $C$, 
-a \emph{boolean} feature $b_C$, if $|C^s| \in \set{0,1}$ for all sampled states $s$, 
-and a \emph{numerical} feature $n_C$ otherwise. The value of $b_C$ in $s$ is true iff $|C^s| = 1$;
-the value of $n_C$ is $|C^s|$.
-
-\item Numerical features $\textit{dist}(C_1,R{:}C,C_2)$, that represent the smallest
-  $n$ such that for $x_1, \ldots, x_n$, $C_1^s(x_1)$, $C_2^s(x_{n})$, and
-  $(R{:}C)(x_i,x_{i+1})$ for $i=1, \ldots, n$. The denotation $(R{:}C)^s$ 
-  contains all pairs $(x,y)$ in $R^s$ s.t. $y$ is in $C^s$.
+\item For each nullary primitive predicate $p$, a \emph{boolean} feature $b_p$ that is true
+  in $s$ iff $p$ is true in $s$.
+%
+\item For each concept $C$, a \emph{boolean} feature $b_C$, if $|C^s| \in \set{0,1}$ for all
+  sampled states $s$, and a \emph{numerical} feature $n_C$ otherwise.
+  The value of $b_C$ in $s$ is true iff $|C^s| > 0$; the value of $n_C$ is $|C^s|$.
+%
+\item Numerical features $\textit{dist}(C_1,R{:}C,C_2)$ that represent the smallest
+  $n$ such that there are objects $x_1, \ldots, x_n$ satisfying $C_1^s(x_1)$,
+  $C_2^s(x_{n})$, and $(R{:}C)^s(x_i,x_{i+1})$ for $i=1,\ldots,n$.
+  The denotation $(R{:}C)^s$ contains all pairs $(x,y)$ in $R^s$ such that $y\in C^s$.
 \end{enumerate}
@@
 are $on(\cdot,\cdot)$, $clear(\cdot)$, $ontable(\cdot)$, $holding(\cdot)$, and $handempty$. 
-For this problem, a  single training instance $P$  with  5 blocks sufficed to learn an abstract model from
-which a general plan was computed. The  set of  features $F$ learned  from the theory $T_G$ is:
-
+For this problem, a  single training instance $P$  with  5 blocks suffices to learn an abstract
+model on which a general plan is computed. The  set of features $F$ learned from the theory $T_G$ is:
 \begin{enumerate}[--]
@@
   \item $X: holding \sqcap C_x$ (whether block $x$ is being held),
-  \item $n(x): \abs{\exists\,on^* . C_x}$, (number of blocks above block $x$).
+  \item $n(x): \abs{\exists\,on^* . C_x}$, (number of blocks above block $x$). \footnote{\textcolor{red}{** See note about reflexive closure **}}
 \end{enumerate}
 
-\noindent The set $A_F$  of abstract actions learned  from $T_G$ is:\footnote{
-Feature and action names are provided   to make their intuitive meaning explicit.
-The formal meaning of features  follows from their syntactic form.}
+\noindent The set $A_F$  of abstract actions learned  from $T_G$ is:\footnote{Feature and action
+  names are provided to make their meaning explicit; the meaning follows from their syntactic form.}
 
@@
 $I_F = \{\neg H, \neg X, n(x) > 0\}$ and $G_F=\{n(x)=0\}$.
-Since no actions increment the variable $n(x)$, 
-$Q^+_F=Q'_F$. The policy that solves the generalized problem $\Q_{clear}$
-is obtained from $Q^+_F$ by the FOND planner in 0.46 seconds. The plan implements
-a loop that applies the pick-above-$x$ action followed by put-aside,
-until $x$ becomes clear.
+Since no actions increment the variable $n(x)$, $Q^+_F=Q'_F$.
+\textcolor{red}{\bf ** Do we need this comment about $Q^+_F$ **}
+The policy that solves the generalized problem $\Q_{clear}$
+is obtained from $Q^+_F$ with the planner in 0.46 seconds.
+The plan implements a loop that applies the pick-above-$x$
+action followed by put-aside, until $x$ becomes clear.
 
@@
 $\Q_{on}$ consists  of Blocksworld instances with goals of the form $on(x,y)$,
-and initial situations  where the blocks $x$ and $y$ are not in the same tower.
+and initial situations  where the blocks $x$ and $y$ are in \emph{different towers}.
 The primitive predicates are the same as in $\Q_{clear}$.
-For this problem, three training instances $P$ were used.
-The set $F$ learned  from $T_G$ contains
+Three training instances $P$ are used to learn $F$ from $T_G$
+that contains
 % 
@@
 % features, 2 numerical features, and 7 actions is learnt.
-the boolean features  $E$, $X$ and $G$,  that stand for the gripper
-being empty, the gripper holding block $x$, and the block $x$ being
-on block $y$, and the numerical features  $n(x)$ and $n(y)$ that stand
-for the number of blocks above $x$ and $y$ respectively.
+the boolean features $E$, $X$ and $G$, for the gripper being empty, the
+block $x$ being held, and the block $x$ being on block $y$, and the
+numerical features $n(x)$ and $n(y)$ that count the number of blocks
+above $x$ and $y$ respectively.
 % The DL  definition of the  features $X$, $n(x)$, and $n(y)$ are as above
@@
 
-The abstraction $Q_F$  with $I_F\!=\!\{ E, \neg X, \neg G, n(x)\!>\!0, n(y)\!>\!0\}$
+The abstraction $Q_F$  with $I_F=\{E,\neg X,\neg G,n(x)>0,$ $n(y)\!>\!0\}$
 % and $G_F = \{ \neg X, G, n(x)=0, n(y)\!>\!0\}$,
-and $G_F = \{G\}$, 
-is translated into the FOND problem
-$Q_F^+$  in 0.01 seconds, which is solved by the FOND planner  in 7.56 seconds.
-The resulting policy solves the generalized problem $\Q_{on}$.
-It implements a loop  to clear block $x$,  followed by a second  loop to clear block $y$,
-picking then block $x$ and placing it on  $y$.
+and $G_F = \{G\}$ is translated into the FOND problem $Q_F^+$
+which is then solved by the planner in 7.56 seconds.
+The resulting policy solves the generalized problem $\Q_{on}$;
+it implements a loop to clear block $x$, followed by a second
+loop to clear block $y$, picking then $x$ and placing it on $y$.
 
 \paragraph{Gripper.}
-$\Q_{gripper}$ involves  a robot with some grippers
-that needs to move a number of balls from one room into a target
+$\Q_{gripper}$ involves  a robot with grippers whose goal
+is to move a number of balls from one room into a target
 room $x$. Each gripper may carry one ball at a time.
@@
 The abstraction $Q_F$ with $I_F\!=\!\{ B\!>\!0, \neg X, G\!>\!0, C\!=\!0\}$ and $G_F\!=\!\{ B\!=\!0, C\!=\!0 \}$ is
-translated into the FOND problem $Q^+_F$, which is solved by the
-planner in 171.92 seconds. The resulting policy provably solves $\Q_{gripper}$,
-meaning that it works for any  number of grippers and balls.
+translated into $Q^+_F$ which is solved by the FOND planner in 171.92 seconds.
+The resulting policy provably solves $\Q_{gripper}$, meaning that
+it works for any  number of grippers and balls.
 \Omit{ %% Add if space
@@
 and then drops the balls one by one, until the grippers
-are all empty.  It then moves back to the other room if there   are still  balls
-to be moved,  repeating the  process.
+are all empty.  It then moves back to the other room if there
+are still  balls to be moved, repeating the process.
 }
@@
 avoiding \emph{blocked} cells. This is  variation of an
-example  in \cite{garnelo2016towards}. The STRIPS instances have the primitive predicates $reward(\cdot)$, $at(\cdot)$, $blocked(\cdot)$
+example  by \citeay{garnelo2016towards}.
+The STRIPS instances have the primitive predicates $reward(\cdot)$, $at(\cdot)$, $blocked(\cdot)$
 and $adjacent(\cdot,\cdot)$ that denote the position of the rewards, of
@@
   \item $R: \abs{reward}$ (number of remaining rewards), 
-  \item $D: dist(at, adjacent{:}\neg blocked, reward)$ (dist. from current
+  \item $D:\! dist(at, adjacent{:}\neg blocked, reward)$ (distance from current
     cell to closest cell with reward, traversing adjacent, unblocked cells only).
 \end{enumerate}
-The set of  abstract actions  learned is:
+The learned set of abstract actions is:
 \begin{enumerate}[--]
@@
 The resulting  abstraction $Q_F$ with $I_F = \{ R>0, D>0 \}$ and $G_F=\{R=0\}$
-is translated into the  FOND problem $Q^+_F$ which yields
-a policy  that solves $\Q_{reward}$ in 1.36 seconds. The policy moves the agent
-one step at a time towards  the uncollected  reward that is closest, as measured by the
-numerical feature $D$. Once the reward is reached, the reward is consumed, and
-the process repeats until there are no more rewards to be collected.
+is translated into $Q^+_F$ which is solved by the planning in 1.36 seconds.
+The policy moves the agent one step at a time towards the uncollected reward
+that is closest, as measured by the numerical feature $D$. Once the reward is
+reached, the reward is consumed, and the process repeats until there are no
+more rewards to be collected.
 
@@
 It is not fully clear whether this assumption is reasonable in general.
-A relevant argument made in  \cite{bonet:ijcai2018} is that generalized problems
+A relevant argument made by \citeay{bonet:ijcai2018} is that generalized problems
 over domains with low width \cite{nir:ecai2012}, have compact policies
@@
 \subsubsection{Acknowledgments.}
-H. Geffner is partially funded by grant TIN-2015-67959-P, MINECO, Spain;
-G. Franc\`{e}s is funded by the SSX project, European Research Council.
+H.\ Geffner is partially funded by grant TIN-2015-67959-P, MINECO, Spain.
+G.\ Franc\`{e}s is funded by the SSX project, European Research Council.
+B.\ Bonet is partially funded by 2018 C\'atedra de Excelencia UC3M-Banco Santander award.
 

### 2018-11-11 driver.py
Accept different maxsat solvers as an experiment option

@@ -393,7 +393,7 @@ class MaxsatProblemSolutionStep(Step):
         super().__init__(**kwargs)
 
     def get_required_attributes(self):
-        return []
+        return ['maxsat_solver']
 
     def get_required_data(self):
         return []

### 2018-11-11 experiments/abstractions_defaults.py

@@ -86,6 +86,9 @@ def generate_experiment(domain_dir, domain, **kwargs):
         # Set a random seed for reproducibility (default: 1)
         random_seed=1,
 
+        # the max-sat solver to use. Accepted: openwbo, openwbo-inc, wpm3, maxino
+        maxsat_solver='openwbo-inc',
+
         # The number of features and actions for the SAT encoding
         encoding_k=10,
         encoding_m=10,

### 2018-11-11 learn_actions.py

@@ -85,9 +85,7 @@ def generate_maxsat_problem(config, data, rng):
 
 
 def run_solver(config, data, rng):
-    # solution = solve(config.experiment_dir, config.cnf_filename, 'wpm3')
-    # solution = solve(config.experiment_dir, config.cnf_filename, 'maxino')
-    solution = solve(config.experiment_dir, config.cnf_filename, 'openwbo')
+    solution = solve(config.experiment_dir, config.cnf_filename, config.maxsat_solver)
     if not solution.solved and solution.result == "UNSATISFIABLE":
         raise CriticalPipelineError("MAXSAT encoding is UNSATISFIABLE")
     else:

### 2018-11-11 solvers/library.py

@@ -34,6 +34,18 @@ class Openwbo(object):
         return ['open-wbo_static', input_filename]
 
 
+class OpenwboInc(object):
+    TAG = 'openwbo-inc'
+
+    def __init__(self, rundir=None):
+        self.run_dir = rundir
+
+    @staticmethod
+    def command(input_filename):
+        # From competition script: open-wbo-inc -ca=1 -c=100000 -algorithm=6
+        return ['open-wbo-inc', '-ca=1', '-c=100000', '-algorithm=6', input_filename]
+
+
 class Glucose(object):
     TAG = 'glucose'
 

### 2018-11-11 doc/camera-ready/paper.tex
Blai edit

@@
 the abstract action $\bar{a}=\abst{\neg H,n(x)>0}{H,n(x)\mminus}$
-and the concrete  action  $a=\text{Unstack}(A,B)$ in $P$ have the same effect over the features.
+and the concrete  action  $a=\text{Unstack}(A,B)$ have the same effect over the features.
 Indeed, $\bar{a}$ and $a$  are both applicable in $s$, and while $\bar{a}$ makes the variable $H$ true and decreases
@@
 
-Soundness and completeness are the two key  properties that  enable us to reason about the collection of
+Soundness and completeness are the key  properties that  enable us to reason about the collection of
 instances $\Q$ in terms of abstract actions that are defined at the level of the features:
@@
     all states that satisfy $G_F$ must be goal states of $P$.
-    \textcolor{red}{\bf ** What is the format for $I_F$ and $G_F$? Conjunctions of lits, DNF, ...? **}
+    %\textcolor{red}{\bf ** What is the format for $I_F$ and $G_F$? Conjunctions of lits, DNF, ...? **}
 %
@@
   and there are blocks on top of $x$. For $F=\{H,n(x)\}$  and $A_F=\{\bar{a},\bar{a}'\}$ as above, 
-  $Q_F=\tup{V_F,I_F,G_F,A_F}$ can be defined with $I_F=\{\neg H, n(x) > 0\}$ and $G_F=\{n(x)=0\}$.
+  $Q_F=\tup{V_F,I_F,G_F,A_F}$ may be defined with $I_F=\{\neg H, n(x) > 0\}$ and $G_F=\{n(x)=0\}$.
   $Q'_F$ is like $Q_F$ but with $n(x)=0$ regarded as a propositional symbol, $n(x) > 0$
   as its negation, and the effect $n(x)\mminus$ replaced by  $n(x) > 0 \, | \, n(x)=0$.
-  Since  no action in $A_F$  increases $n(x)$,   $Q^+_F=Q'_F$, and the strong cyclic solution  of $Q'_F$,
-  that corresponds to the policy above, solves $\Q_{clear}$.
-  \textcolor{red}{\bf ** Why do we need to enter into details about $Q^+_F$?
-  We are using a tool and should not worry about such details.
-  Otherwise, we are conveying the idea that the user should worry about them and
-  then know how the QNP2FOND translation work. **}
+  Since  no action in $A_F$ increases $n(x)$, the strong solutions of $Q'_F$
+  are solutions to $\Q_{clear}$ \cite{bonet:ijcai2017}.
+  %$Q^+_F=Q'_F$, and the strong cyclic solution
+  %of $Q'_F$, that corresponds to the policy above, solves $\Q_{clear}$.
+  %\textcolor{red}{\bf ** Why do we need to enter into details about $Q^+_F$?
+  %We are using a tool and should not worry about such details.
+  %Otherwise, we are conveying the idea that the user should worry about them and
+  %then know how the QNP2FOND translation work. **}
 \end{example}
@@
 % For this, it is assumed that expanded states $s$ in $\S$ are labeled as goal or non-goal states.
-For this, it is assumed that  states $s$ sampled from an instance $P$, are labeled as goal or non-goal states
-of $P$.
+%For this, it is assumed that states $s$ sampled from an instance $P$, are labeled as goal or non-goal states of $P$.
+For this, it is assumed that the states in the sample are labeled as goal or non-goal states.
 
@@
 $T_G(\S,\F)$ guarantees soundness relative to $\S$
-but completeness only relative to the the subset of goal-relevant transitions in $\S$.
+but completeness only relative to the subset of goal-relevant transitions in $\S$.
 Yet this does not affect the resulting formal guarantees, which depend only on soundness. 
@@
 differ only in the sign of a precondition, the two abstract actions are    merged into one with
-the precondition dropped. The key properties of $T(\S,\F)$ and $T_G(\S,\F)$ are:
+the precondition dropped.
+%The key properties of $T(\S,\F)$ and $T_G(\S,\F)$ are:
 
@@
 from the satisfying assignment $\sigma$ as well, for simplicity, in the examples below, they are
-provided by hand.\footnote{$G_F = G_{\sigma}$ can be defined as the DNF formula whose terms 
-  correspond to the abstract states over $F_{\sigma}$ that correspond to some goal state in
-  the sample. \textcolor{red}{\bf ** Related to previous note **}
+provided by hand.\footnote{$G_F = G_{\sigma}$ may be defined as the DNF formula whose terms 
+  correspond to the abstract states over $F_{\sigma}$ that correspond to goal states in
+  the sample. For each such term, a new abstract action to reach a new goal literal
+  from such a term may be required.
   }
@@
 \emph{soft} unit clause $\neg selected(f)$ with weight $cost(f)$.
-The cost measure that is used in the experiments is defined next.
+The cost measure that is used in the experiments is defined below.
 
@@
   %  set of $y$'s  for which $R(x,y)$ equal to set of $y$'s s.t. $R'(x,y)$.
-  \item $R \leftarrow R_p, R_p^{-1}, R_p^*, [R_p^{-1}]^*$: primitive, inverse, and transitive closure of both.
+  \item $R \leftarrow R_p, R_p^{-1}, R_p^+, [R_p^{-1}]^+$: primitive, inverse, and transitive closure of both.
 %     \item $R \leftarrow R \circ R'$, role composition: $(x,y)$ in $R \circ R'$ if $(x,z)$ in $R$ and $(z,y)$ in $R'$.
@@
 
-The denotations   $C^s$  and $R^s$ of derived concepts and roles follows from the rules and the denotation of primitive
-concepts and roles. For example, for the concept  $C: \exists\, on^* . C_x$ in    Blocksworld, 
-$C^s$  denotes  the set of objects  that are above block $x$. The numerical feature
-$n_C$ associated with concept $C$ represents the number of objects in $C^s$. %$|C^s|$.
-\textcolor{red}{\bf ** $C: \exists\, on^* . C_x$ also counts block $x$ since * is reflexive! **}
+The denotations $C^s$ and $R^s$ for concepts and roles follows
+from the rules and the denotation of primitive concepts and roles.
+For example, for the concept $C: \exists\, on^+ . C_x$ in Blocksworld,
+$C^s$ denotes the set of blocks that are above $x$.
+%The numerical feature $n_C$ associated with concept $C$ represents
+%the number of objects in $C^s$. %$|C^s|$.
 
@@
 \begin{enumerate}[{\small$\bullet$}]
-\item For each nullary primitive predicate $p$, a \emph{boolean} feature $b_p$ that is true
-  in $s$ iff $p$ is true in $s$.
-%
-\item For each concept $C$, a \emph{boolean} feature $b_C$, if $|C^s| \in \set{0,1}$ for all
-  sampled states $s$, and a \emph{numerical} feature $n_C$ otherwise.
-  The value of $b_C$ in $s$ is true iff $|C^s| > 0$; the value of $n_C$ is $|C^s|$.
-%
-\item Numerical features $\textit{dist}(C_1,R{:}C,C_2)$ that represent the smallest
-  $n$ such that there are objects $x_1, \ldots, x_n$ satisfying $C_1^s(x_1)$,
-  $C_2^s(x_{n})$, and $(R{:}C)^s(x_i,x_{i+1})$ for $i=1,\ldots,n$.
+  \item For each nullary primitive predicate $p$, a \emph{boolean}
+    feature $b_p$ that is true in $s$ iff $p$ is true in $s$.
+  %
+  \item For each concept $C$, a \emph{boolean} feature $b_C$, if
+    $|C^s| \in \set{0,1}$ for all sampled states $s$, and a \emph{numerical}
+    feature $n_C$ otherwise. The value of $b_C$ in $s$ is true iff $|C^s| > 0$;
+    the value of $n_C$ is $|C^s|$.
+  %
+  \item Numerical features $\textit{dist}(C_1,R{:}C,C_2)$ that represent the
+    smallest $n$ such that there are objects $x_1, \ldots, x_n$ satisfying
+    $C_1^s(x_1)$, $C_2^s(x_{n})$, and $(R{:}C)^s(x_i,x_{i+1})$ for $i=1,\ldots,n$.
     The denotation $(R{:}C)^s$ contains all pairs $(x,y)$ in $R^s$ such that $y\in C^s$.
@@
     generalized problem $\Q$,
-  \item a pool of features $\F$ is obtained from the predicates used in the
-    instances $P$, the grammar, a  bound $k$, and the sample $\S$ (used for
+  \item a pool of features $\F$ is obtained from the predicates in the instances, %used in the instances $P$,
+    the grammar, a  bound $k$, and the sample $\S$ (used for
     pruning),
-  \item a satisfying assignment $\sigma$ of $T(\S,\F)$ or $T_G(\S,\F)$
+  \item an assignment $\sigma$ of $T(\S,\F)$ or $T_G(\S,\F)$
     that minimizes $\sum_{f \in F_{\sigma}} cost(f)$ is obtained using a
     Max SAT solver,
-  \item the features $F$ and abstract actions $A_F$ are extracted from $\sigma$
-    as $F_{\sigma}$ and $A_{\sigma}$,
+  \item features $F$ and abstract actions $A_F$ are extracted from $\sigma$,
+    %as $F_{\sigma}$ and $A_{\sigma}$,
   \item the abstraction $Q_F=\tup{V_F,I_F,G_F,A_F}$ is defined with initial and goal
     conditions $I_F$ and $G_F$ provided by hand to match $\Q$, 
-  \item the FOND problem $Q^+_F$, constructed automatically from $Q_F$, is solved by an off-the-shelf
+  \item the FOND problem $Q^+_F$, constructed automatically from $Q_F$, is solved with an off-the-shelf
     FOND planner.
@@
 $\Q$. The theorem then says that if soundness over the samples, ensure
-formal soundness over $\Q$, the policy that solves the FONDP problem $Q^+_F$,
+formal soundness over $\Q$, the policy that solves the FOND problem $Q^+_F$,
 solves $\Q$. The result holds whether $T$ or $T_G$ (goal marking) is used in step 3.
@@
   \item $X: holding \sqcap C_x$ (whether block $x$ is being held),
-  \item $n(x): \abs{\exists\,on^* . C_x}$, (number of blocks above block $x$). \footnote{\textcolor{red}{** See note about reflexive closure **}}
+  \item $n(x): \abs{\exists\,on^+ . C_x}$, (number of blocks above block $x$).
 \end{enumerate}
@@
 $I_F = \{\neg H, \neg X, n(x) > 0\}$ and $G_F=\{n(x)=0\}$.
-Since no actions increment the variable $n(x)$, $Q^+_F=Q'_F$.
-\textcolor{red}{\bf ** Do we need this comment about $Q^+_F$ **}
-The policy that solves the generalized problem $\Q_{clear}$
-is obtained from $Q^+_F$ with the planner in 0.46 seconds.
+As mentioned above, since no action increments the variable $n(x)$,
+the strong-cyclic solutions of $Q'_F$ solve $\Q_{clear}$.
+One such policy is found with the FOND planner in 0.46 seconds.
 The plan implements a loop that applies the pick-above-$x$
@@
 The primitive predicates are the same as in $\Q_{clear}$.
-Three training instances $P$ are used to learn $F$ from $T_G$
-that contains
+Three training instances are used to learn $F$ from $T_G$;
+$F$ contains
 % 
@@
 the boolean features $E$, $X$ and $G$, for the gripper being empty, the
-block $x$ being held, and the block $x$ being on block $y$, and the
+block $x$ being held, and $x$ being on block $y$, and the
 numerical features $n(x)$ and $n(y)$ that count the number of blocks
@@
 %% 4 bool[handempty] 1 holding(a) 0 n(b) 1 on(a,b)_2 1
-The set of actions  $A_F$ that is learned is:
+The learned set of actions $A_F$ is:
 \begin{enumerate}[--]
@@
 which is then solved by the planner in 7.56 seconds.
-The resulting policy solves the generalized problem $\Q_{on}$;
-it implements a loop to clear block $x$, followed by a second
-loop to clear block $y$, picking then $x$ and placing it on $y$.
+The resulting policy solves the generalized problem $\Q_{on}$:
+it implements a loop to clear block $x$, followed by a loop
+to clear block $y$, picking then $x$ and placing it on $y$.
 
@@
 
-Pipeline 1--6 is fed  two instances $P$ from $\Q_{gripper}$  with two
+Pipeline 1--6 is fed with two instances $P$ from $\Q_{gripper}$  with two
 rooms each, one with 4 balls and the other with 5. The set of features that
@@
 The abstraction $Q_F$ with $I_F\!=\!\{ B\!>\!0, \neg X, G\!>\!0, C\!=\!0\}$ and $G_F\!=\!\{ B\!=\!0, C\!=\!0 \}$ is
-translated into $Q^+_F$ which is solved by the FOND planner in 171.92 seconds.
+translated into $Q^+_F$ which is solved by the planner in 171.92 seconds.
 The resulting policy provably solves $\Q_{gripper}$, meaning that
@@
 $\Q_{reward}$ contains instances where  an agent needs to navigate a rectangular
-grid to pick up  rewards spread on a grid  while 
+grid to pick up  rewards spread on the grid  while 
 avoiding \emph{blocked} cells. This is  variation of an
@@
 Pipeline 1--6  is fed with two instances of the problem of sizes $4 \times 4$
-and $5\times 5$, and two different  distributions of blocked cells  and rewards.
-From the theory $T_G$, two numerical features are learned:
+and $5\times 5$, and having different distributions of blocked cells and rewards.
+Two numerical features are learned:
 \begin{enumerate}[--]
@@
 The resulting  abstraction $Q_F$ with $I_F = \{ R>0, D>0 \}$ and $G_F=\{R=0\}$
-is translated into $Q^+_F$ which is solved by the planning in 1.36 seconds.
+is translated into $Q^+_F$ which is solved by the planner in 1.36 seconds.
 The policy moves the agent one step at a time towards the uncollected reward
@@
 
-An   interesting challenge for the future  is to generalize the proposed methods  to other   contexts,   such as learning  from ``screen states''
-in video games,  where the sampled  states have no known structure and  there are no primitive predicates.
-One way for  achieving this, is  precisely by learning embeddings that yield sound and complete abstractions or
-suitable approximations.
+An interesting challenge for the future  is to generalize the proposed methods
+to other contexts, such as learning from ``screen states'' in video games,
+where the sampled states have no known structure and there are no primitive
+predicates. One way for achieving this, is precisely by learning embeddings
+that yield sound and complete abstractions or suitable approximations.
+
+On the other hand, contexts involving non-deterministic actions may be
+accommodated with a simple modification of the SAT theory.
+Indeed, it would be enough to replace the formula (3) that links $D1$ and $D2$
+with a formula that says that when $s$ and $t$ are indistinguishable
+(i.e.\ $\neg D1(s,t)$), then for each action $a$ for $s$, there is an action $b$
+for $t$ whose sets of possible outcomes are indistinguishable.
 

### 2018-11-11 driver.py
Allow timing out maxsat solver

@@ -393,7 +393,7 @@ class MaxsatProblemSolutionStep(Step):
         super().__init__(**kwargs)
 
     def get_required_attributes(self):
-        return ['maxsat_solver']
+        return ['maxsat_solver', 'maxsat_timeout']
 
     def get_required_data(self):
         return []

### 2018-11-11 experiments/abstractions_defaults.py

@@ -87,7 +87,8 @@ def generate_experiment(domain_dir, domain, **kwargs):
         random_seed=1,
 
         # the max-sat solver to use. Accepted: openwbo, openwbo-inc, wpm3, maxino
-        maxsat_solver='openwbo-inc',
+        maxsat_solver='openwbo',
+        maxsat_timeout=None,
 
         # The number of features and actions for the SAT encoding
         encoding_k=10,

### 2018-11-11 learn_actions.py

@@ -85,7 +85,7 @@ def generate_maxsat_problem(config, data, rng):
 
 
 def run_solver(config, data, rng):
-    solution = solve(config.experiment_dir, config.cnf_filename, config.maxsat_solver)
+    solution = solve(config.experiment_dir, config.cnf_filename, config.maxsat_solver, config.maxsat_timeout)
     if not solution.solved and solution.result == "UNSATISFIABLE":
         raise CriticalPipelineError("MAXSAT encoding is UNSATISFIABLE")
     else:

### 2018-11-11 solvers/common.py

@@ -1,6 +1,9 @@
+import logging
 import os
+import signal
 import subprocess
 import sys
+from threading import Timer
 
 from solvers import library
 
@@ -14,7 +17,7 @@ class Solution(object):
 
     def parse_result(self, line):
         self.result = line
-        self.solved = {"UNSATISFIABLE": False, "UNKNOWN": False, "OPTIMUM FOUND": True}[line]
+        self.solved = {"UNSATISFIABLE": False, "UNKNOWN": False, "OPTIMUM FOUND": True, "SATISFIABLE": True}[line]
 
 
 def parse_maxsat_output(filename):
@@ -43,12 +46,14 @@ def choose_solver(solver):
         return library.Maxino
     elif solver == 'openwbo':
         return library.Openwbo
+    elif solver == 'openwbo-inc':
+        return library.OpenwboInc
     raise RuntimeError('Unknown solver "{}"'.format(solver))
 
 
-def solve(rundir, cnf_filename, solver='wpm3'):
+def solve(rundir, cnf_filename, solver='wpm3', timeout=None):
     solver = choose_solver(solver)(rundir)
-    error, output = run(solver, rundir, cnf_filename, 'maxsat_solver')
+    error, output = run_solver(solver, rundir, cnf_filename, 'maxsat_solver', timeout=timeout)
 
     if error:
         raise RuntimeError("There was an error running the MAXSAT solver. Check error logs")
@@ -56,7 +61,7 @@ def solve(rundir, cnf_filename, solver='wpm3'):
     return parse_maxsat_output(output)
 
 
-def run(solver, rundir, input_filename, tag=None, stdout=None):
+def run_solver(solver, rundir, input_filename, tag=None, stdout=None, timeout=None):
     assert tag or stdout
     error = False
     if stdout is None:
@@ -65,10 +70,38 @@ def run(solver, rundir, input_filename, tag=None, stdout=None):
     with open(stdout, 'w') as driver_log:
         with open(os.path.join(rundir, '{}_run.err'.format(tag)), 'w') as driver_err:
             cmd = solver.command(input_filename)
-            print('Executing "{}" on directory "{}"'.format(cmd, rundir))
-            retcode = subprocess.call(cmd, cwd=rundir, stdout=driver_log, stderr=driver_err)
+            logging.info('Executing (timeout: {}) "{}" on directory "{}"'.format(timeout, cmd, rundir))
+
+            proc = subprocess.Popen(cmd, cwd=rundir, stdout=driver_log, stderr=driver_err)
[…]

### 2018-11-12 doc/camera-ready/paper.tex
Blai edit

@@
 
-The paper is organized as follows. We provide the background  first and show then
-how features and abstract actions can be learned by enforcing soundness and
-completeness over a set of samples, from a pool of candidate features derived from the domain
-predicates. The full computational model is then summarized, followed by experimental results
-and a discussion.
+The paper is organized as follows. We first provide background and then show 
+how an abstract actions can be learned by enforcing soundness and
+completeness over a set of samples from a pool of candidate features derived
+from the domain predicates. The computational model is then summarized,
+followed by experimental results and a discussion.
 
@@
 
-A solution to a generalized problem $\Q$ given the features $F$ and abstract
+A solution to a problem $\Q$ given the features $F$ and abstract
 actions $A_F$ is a partial function $\pi$ that maps \emph{abstract states}
@@
 We also consider an alternative SAT theory $T_G(\S,\F)$ that is similar to $T(\S,\F)$
-but  smaller. It is obtained by marking  some of the transitions $(s,s')$ in $\S$  as  \emph{goal-relevant}.
+but  smaller. It is obtained by marking  some transitions $(s,s')$ in $\S$  as  \emph{goal relevant}.
 Then, rather than creating abstract  actions to account for all the transitions, we only create
 abstract actions to account for the marked transitions. This is achieved by drawing the states
-$s$ and the transitions $(s,s')$ in formulas (\ref{eq:d1}) and (\ref{eq:d2}) from the
+$s$ and the transitions $(s,s')$ in formulas \eqref{eq:d1} and \eqref{eq:d2} from the
 set of marked transitions in $\S$. The states $t$ and the transitions $(t,t')$, on the other hand,
-are drawn from the sample set $\S$ exactly as before. This simplification preserves soundness over $\S$
-but the completeness is not over $\S$ but over the marked transitions. 
-The   transitions marked as goal relevant in $\S$  are obtained by computing one plan
+are drawn from the whole sample set $\S$ as before. This simplification preserves soundness over $\S$
+but completeness is not over $\S$ but over the marked transitions in $\S$. 
+The goal-relevant transitions in $\S$ are obtained by computing one plan
 for each sampled instance $P$ from $\Q$.
+%The transitions marked as goal relevant in $\S$  are obtained by computing one plan
+%for each sampled instance $P$ from $\Q$.
 
@@
   \item $C \leftarrow \exists R.C,\forall R.C$, first denotes $\{x:\!\exists y[R(x,y) \land C(y)]\}$, %  $x$'s  s.t.\  $R(x,y) \land C(y)$ for some  $y$;
-    second $\{x:\forall y[R(x,y)\land C(y)]\}$, %,  $x$'s  s.t. $R(x,y) \rightarrow C(y)$ for all $y$,
+    the second denotes $\{x:\forall y[R(x,y)\land C(y)]\}$, %,  $x$'s  s.t. $R(x,y) \rightarrow C(y)$ for all $y$,
   \item $C \leftarrow R=R'$, denotes $\{ x : \forall y[R(x,y)=R'(x,y)]\}$, %$x$'s  s.t. $R(x,y)$ iff $R'(x,y)$ for all $y$,
@@
 We have introduced a scheme for computing general plans that mixes
-learning and planning: a \emph{learner} infers a general abstraction made
-up of features and abstract actions by enforcing a notion of soundness
-and completeness over a  set of samples, and a \emph{planner}
+learning and planning: a \emph{learner} infers an abstraction made
+up of features and abstract actions by enforcing %a notion of
+soundness and completeness over the samples, and a \emph{planner}
 uses the abstraction, suitably transformed, to compute general plans. 
@@
 
-On the other hand, contexts involving non-deterministic actions may be
-accommodated with a simple modification of the SAT theory.
-Indeed, it would be enough to replace the formula (3) that links $D1$ and $D2$
-with a formula that says that when $s$ and $t$ are indistinguishable
-(i.e.\ $\neg D1(s,t)$), then for each action $a$ for $s$, there is an action $b$
-for $t$ whose sets of possible outcomes are indistinguishable.
-
-\subsubsection{Acknowledgments.}
+Finally, problems involving non-deterministic transitions may be
+accommodated with a simple fix to the SAT theory.
+It would be enough to replace the formula \eqref{eq:bridge1} that
+links $D1$ and $D2$ with a formula that says that when $s$ and $t$ are
+indistinguishable (i.e.\ $\neg D1(s,t)$), for each action $a$ on $s$,
+there is an action $b$ on $t$ whose transitions are indistinguishable
+from those of $a$ on $s$. For this, transitions $(s,s')$ in the sample
+would need to be tagged with action labels as $(s,a,s')$.
+
+%\bigskip
+\subsection*{Acknowledgments}
+B.\ Bonet is partially funded by 2018 C\'atedra de Excelencia UC3M-Banco Santander award.
+G.\ Franc\`{e}s is funded by the SSX project, European Research Council.
 H.\ Geffner is partially funded by grant TIN-2015-67959-P, MINECO, Spain.
-G.\ Franc\`{e}s is funded by the SSX project, European Research Council.
-B.\ Bonet is partially funded by 2018 C\'atedra de Excelencia UC3M-Banco Santander award.
 

### 2018-11-12 doc/camera-ready/paper.tex
Blai edit

@@
 \usepackage{courier}  %Required
-\usepackage{url}  %Required
+%\usepackage{url}  %Required
 \usepackage{graphicx}  %Required
 \usepackage{xcolor}  %Required
+\usepackage{xspace}  %Required
 
@@
 %\newtheorem{example}[definition]{Example}
-\newenvironment{example}{\noindent\textbf{Example.}}{\qed}
+\newenvironment{example}{\noindent\textbf{Example.}\xspace}{\qed}
 
@@
 /Title (Learning Features and Abstract Actions for Computing Generalized Plans)
-/Author (Blai Bonet, Guillem Frances, Hector Geffner)}
+/Author (Blai Bonet, Guillem Frances, Hector Geffner)
+}
 
@@
 there is an action $b$ on $t$ whose transitions are indistinguishable
-from those of $a$ on $s$. For this, transitions $(s,s')$ in the sample
+from those of $a$ on $s$. To do this, transitions $(s,s')$ in the sample
 would need to be tagged with action labels as $(s,a,s')$.

### 2018-11-12 doc/camera-ready/paper.tex
Fix typo

@@
 The paper is organized as follows. We first provide background and then show 
-how an abstract actions can be learned by enforcing soundness and
+how abstract actions can be learned by enforcing soundness and
 completeness over a set of samples from a pool of candidate features derived
@@
 Let $\Q_{clear}$ stand for  all  Blocksworld instances with stack and unstack actions and goal $clear(x)$,
-let $F=\{H,n(x)\}$ be the set of the two features above, and let $s$ be a reachable state in an instance $P$
+let $F=\{H,n(x)\}$ be the set with the two features above, and let $s$ be a reachable state in an instance $P$
 in $\Q_{clear}$

### 2018-11-13 doc/camera-ready/paper.tex
typo
Edit of last paragraph

@@
 grid to pick up  rewards spread on the grid  while 
-avoiding \emph{blocked} cells. This is  variation of an
+avoiding \emph{blocked} cells. This is a variation of an
 example  by \citeay{garnelo2016towards}.
@@
 Finally, problems involving non-deterministic transitions may be
-accommodated with a simple fix to the SAT theory.
-It would be enough to replace the formula \eqref{eq:bridge1} that
-links $D1$ and $D2$ with a formula that says that when $s$ and $t$ are
-indistinguishable (i.e.\ $\neg D1(s,t)$), for each action $a$ on $s$,
-there is an action $b$ on $t$ whose transitions are indistinguishable
-from those of $a$ on $s$. To do this, transitions $(s,s')$ in the sample
-would need to be tagged with action labels as $(s,a,s')$.
+accounted for with a simple change to the SAT theory.
+It would be sufficient to replace the formula \eqref{eq:bridge1}
+that links $D1$ and $D2$ with a formula that establish that when
+$s$ and $t$ are indistinguishable (i.e.\ $\neg D1(s,t)$), then
+for each action $a$ on $s$, there is an action $b$ on $t$ whose
+set of transitions from $t$ are indistinguishable from the set
+of transitions of $a$ on $s$.
+To do this, transitions $(s,s')$ in the sample would need to be
+tagged with action labels as $(s,a,s')$.
 

### 2018-11-13 doc/camera-ready/paper.tex
Edit of last paragraph

@@
 It would be sufficient to replace the formula \eqref{eq:bridge1}
-that links $D1$ and $D2$ with a formula that establish that when
+that links $D1$ and $D2$ with a formula that establishes that when
 $s$ and $t$ are indistinguishable (i.e.\ $\neg D1(s,t)$), then
-for each action $a$ on $s$, there is an action $b$ on $t$ whose
-set of transitions from $t$ are indistinguishable from the set
-of transitions of $a$ on $s$.
+for each action $a$ for $s$, there is an action $b$ for $t$ whose
+set of transitions on $t$ is indistinguishable from the set
+of transitions of $a$ on $s$; namely, for each transition $(s,a,s')$
+there is a transition $(t,b,t')$ with $\Delta_f(s,s')=\Delta_f(t,t')$
+for each feature $f$, and vice versa.
 To do this, transitions $(s,s')$ in the sample would need to be

### 2018-11-13 doc/camera-ready/paper.tex
Changes in discussion + summary

--- 2018-11-13 doc/camera-ready/paper.tex
@@
   %
-  \item For each concept $C$, a \emph{boolean} feature $b_C$, if
-    $|C^s| \in \set{0,1}$ for all sampled states $s$, and a \emph{numerical}
-    feature $n_C$ otherwise. The value of $b_C$ in $s$ is true iff $|C^s| > 0$;
-    the value of $n_C$ is $|C^s|$.
+  %\item For each concept $C$, a \emph{boolean} feature $b_C$, if $|C^s|\leq 1$
+  \item For each concept $C$, a \emph{boolean} feature $b_C$, if $|C^s| \in \set{0,1}$
+    for all sampled states $s$, and a \emph{numerical} feature $n_C$ otherwise.
+    The value of $b_C$ in $s$ is true iff $|C^s| > 0$; the value of $n_C$ is $|C^s|$.
   %
@@
 
-A second limitation is the assumption that 
-the features can be obtained from a large pool of candidate  features defined 
-from the primitive predicates and a  general grammar.
-It is not fully clear whether this assumption is reasonable in general.
+It is not fully clear whether it is reasonable in general
+to assume that the features can be obtained from a large
+pool of candidate features that is defined from the primitive
+predicates and a general concept grammar.
+%A second limitation is the assumption that 
+%the features can be obtained from a large pool of candidate  features defined 
+%from the primitive predicates and a  general grammar.
+%It is not fully clear whether this assumption is reasonable in general.
 A relevant argument made by \citeay{bonet:ijcai2018} is that generalized problems
@@
 In addition, for dealing with arbitrary goals, the same grammar  appears
-to be suitable, once  \emph{goal predicates}, namely, copies $p_G$ of the  primitive predicates $p$
-that appear in the goal,  are added to the set of primitive predicates and are evaluated in the goal
+to be suitable once  \emph{goal predicates}, namely, copies $p_G$ of the primitive predicates $p$
+that appear in the goal, are added to the set of primitive predicates and evaluated in the goal
 \cite{martin-geffner:generalized}.
+
+Finally, while we have dealt with instances with deterministic actions
+only, the formulation can be extended to handle non-deterministic actions.
+Indeed, it is sufficient to replace the formula \eqref{eq:bridge1}
+that links $D1$ and $D2$ with a formula to establish that when states
+$s$ and $t$ are indistinguishable (i.e.\ $\neg D1(s,t)$), then
+for each action $a$ for $s$, there is an action $b$ for $t$ whose
+set of transitions on $t$ is indistinguishable from the set
+of transitions of $a$ on $s$; namely, for each transition $(s,a,s')$
+there is a transition $(t,b,t')$ with $\Delta_f(s,s')=\Delta_f(t,t')$
+for each feature $f$, and vice versa.
+To do this, transitions $(s,s')$ in the sample would need to be
+tagged with action labels as $(s,a,s')$.
 
@@
 
-Finally, problems involving non-deterministic transitions may be
-accounted for with a simple change to the SAT theory.
-It would be sufficient to replace the formula \eqref{eq:bridge1}
-that links $D1$ and $D2$ with a formula that establish that when
-$s$ and $t$ are indistinguishable (i.e.\ $\neg D1(s,t)$), then
-for each action $a$ on $s$, there is an action $b$ on $t$ whose
-set of transitions from $t$ are indistinguishable from the set
-of transitions of $a$ on $s$.
-To do this, transitions $(s,s')$ in the sample would need to be
-tagged with action labels as $(s,a,s')$.
+%Finally, problems involving non-deterministic transitions may be
+%accounted for with a simple change to the SAT theory.
+%It would be sufficient to replace the formula \eqref{eq:bridge1}
+%that links $D1$ and $D2$ with a formula that establishes that when
+%$s$ and $t$ are indistinguishable (i.e.\ $\neg D1(s,t)$), then
+%for each action $a$ for $s$, there is an action $b$ for $t$ whose
+%set of transitions on $t$ is indistinguishable from the set
+%of transitions of $a$ on $s$; namely, for each transition $(s,a,s')$
+%there is a transition $(t,b,t')$ with $\Delta_f(s,s')=\Delta_f(t,t')$
+%for each feature $f$, and vice versa.
+%To do this, transitions $(s,s')$ in the sample would need to be
+%tagged with action labels as $(s,a,s')$.
 

### 2018-11-13 doc/camera-ready/paper.tex
add/remove some commas

@@
 in \emph{any} instance of the Blocksworld where the gripper is initially empty.
-Once this  general plan or  policy is derived it can be  applied to solve an infinite collection
+Once this  general plan or  policy is derived, it can be  applied to solve an infinite collection
 of instances that involve different initial states, different objects, and  different (ground)  actions.
@@
 a concrete action with the same effect  over the features. Moreover, the form of the abstract actions
-ensures that generalized plans can be computed using fully observable non-deterministic (FOND) planners
+ensures that generalized plans can be computed using fully observable non-deterministic (FOND) planners,
 once the generalized planning problem is transformed into a FOND problem.
@@
 Bonet and Geffner's generalized planning formulation assumes that the
-features and abstract actions are given. In this work we address this limitation
+features and abstract actions are given. In this work, we address this limitation
 by  \emph{showing  how the features and abstract actions  can be learned from
-the primitive predicates used to define the instances and sampled state transitions.}
+the primitive predicates used to define the instances and from sampled state transitions.}
 For example,  the general policy for achieving $clear(x)$ % of \citeay{bonet:ijcai2018}
 is obtained using a FOND planner on an abstraction that consists of  a  boolean
-feature $H$  that tracks whether the gripper holds a block,  a numerical feature
+feature $H$,  that tracks whether the gripper holds a block,  a numerical feature
 $n(x)$ that counts the number of blocks above $x$, and two abstract actions:
@@
 (decrement $n(x)$), and the other with precondition $H$ and effect $\neg H$.
-In this work, we show how to obtain such policies from STRIPS instances alone,
+We here show how to obtain such policies from STRIPS instances alone,
 without having to provide the features and the abstract actions by hand.
 
-The work has relation to a number of research threads in planning, knowledge
+This work relates to a number of research threads in planning, knowledge
 representation, and machine learning. We make use of SAT solvers and description
@@
 Our approach differs from first-order approaches in the use of propositional
-planners, and from purely learning approaches, in the formal guarantees that
+planners, and from purely learning approaches in the formal guarantees that
 are characteristic of planning: if the learned abstract actions are  sound,
@@
 initial state $s_0$, $G$ is a set of literals over $V$ describing the goal
-states, and $A$ is a set of actions $a$ with their preconditions and effects that
+states, and $A$ is a set of actions $a$ with their preconditions and effects, which
 define the set $A(s)$ of actions applicable in any state $s$, and the successor
@@
 \begin{example}
-  The two abstract actions  $\bar a = \langle \neg H,n(x)>0; H,$ $n(x)\mminus\rangle$ %$\bar{a}=\abst{\neg H,n(x)>0}{H,n(x)\mminus}$
-  and $\bar{a}'=\abst{H}{\neg H}$ over the features  $F=\{H,n(x)\}$
+  The abstract action set $A_F = \{\bar{a}, \bar{a}'\}$, where $\bar{a} = \langle \neg H,n(x)>0; H,$ $n(x)\mminus\rangle$ %$\bar{a}=\abst{\neg H,n(x)>0}{H,n(x)\mminus}$
+  and $\bar{a}'=\abst{H}{\neg H}$ , defined over the above feature set $F=\{H,n(x)\}$,
   is sound but not complete relative to $\Q_{clear}$. 
-  They capture the concrete actions that pick blocks from above $x$ and put them away,
-  but do not capture other actions like stacking a block above $x$ or picking up a block that is not above $x$.
+  $A_F$ captures the concrete actions that pick blocks from above $x$ and put them away,
+  but not other actions like stacking a block above $x$ or picking up a block that is not above $x$.
 \end{example}
@@
 While the initial and goal expressions $I_F$ and $G_F$ of the abstract problem can be learned 
-from the satisfying assignment $\sigma$ as well, for simplicity, in the examples below, they are
+from the satisfying assignment $\sigma$ as well, for simplicity, in the examples below they are
 provided by hand.\footnote{$G_F = G_{\sigma}$ may be defined as the DNF formula whose terms 

### 2018-11-14 doc/camera-ready/paper.tex
Another iteration by Hector

@@
 The SAT theory $T(\S,\F)$ given by formulas \eqref{eq:d1}--\eqref{eq:goal}
-has $|\F| + m^2 (b^2 + 1)$ propositional variables
+has  $|\F| + m^2 (b^2 + 1)$ propositional variables, 
 %has a number of propositional variables equal to $m^2 \times (b^2 + 1) + |\F|$
 where $m$ is the number of expanded states in $\S$, and $b$ is their average branching factor (transitions per state).
-The number of clauses is bounded by $m^2(3 + b^2 + |\F|(1+b^2))$. %$m^2 \times b^2 \times |\F|$.
+The number of clauses is bounded by $m^2(2 + b^2 + |\F|(b^2+1))$. %$m^2 \times b^2 \times |\F|$.
 %
@@
   correspond to the abstract states over $F_{\sigma}$ that correspond to goal states in
-  the sample. For each such term, a new abstract action to reach a new goal literal
-  from such a term may be required.
+  the sample.
+  % For each such term, a new abstract action to reach a new goal literal
+  % from such a term may be required.
   }
@@
 For this problem, a  single training instance $P$  with  5 blocks suffices to learn an abstract
-model on which a general plan is computed. The  set of features $F$ learned from the theory $T_G$ is:
+model from  which a general plan is computed. The  set of features $F$ learned from the theory $T_G$ is:
 \begin{enumerate}[--]
@@
 
-\noindent The set $A_F$  of abstract actions learned  from $T_G$ is:\footnote{Feature and action
+\noindent The set $A_F$  of abstract actions learned is:\footnote{Feature and action
   names are provided to make their meaning explicit; the meaning follows from their syntactic form.}
@@
 % as they  result otherwise in SAT theories that are just too large (see Table~\ref{tab:exp-result-data}).
-This is indeed the reason why we have chosen to use the theories $T_G$
-with goal marking. Additional ideas, however, are required
+This is the   reason why we have chosen to use the more compact theories $T_G$
+in the experiments.  Additional ideas, however, are required
 to improve scalability and to make the computational model
-captured by  steps 1--6 more robust. One possibility that we have not explored
-is the use of incomplete (suboptimal) Max SAT solvers.
-
-It is not fully clear whether it is reasonable in general
-to assume that the features can be obtained from a large
-pool of candidate features that is defined from the primitive
-predicates and a general concept grammar.
+captured by  steps 1--6 more robust.
+% One possibility that we have not explored
+% is the use of incomplete (suboptimal) Max SAT solvers.
+
+From the point of view of expressivity, it is not clear how restrictive
+is the assumption  that the  features can be obtained from a pool of  features
+defined from the primitive predicates and a general grammar.
 %A second limitation is the assumption that 
@@
 over domains with low width \cite{nir:ecai2012}, have compact policies
-in terms of general features that can be computed in polynomial time.
-In addition, for dealing with arbitrary goals, the same grammar  appears
-to be suitable once  \emph{goal predicates}, namely, copies $p_G$ of the primitive predicates $p$
-that appear in the goal, are added to the set of primitive predicates and evaluated in the goal
-\cite{martin-geffner:generalized}.
-
-Finally, while we have dealt with instances with deterministic actions
-only, the formulation can be extended to handle non-deterministic actions.
-Indeed, it is sufficient to replace the formula \eqref{eq:bridge1}
-that links $D1$ and $D2$ with a formula to establish that when states
-$s$ and $t$ are indistinguishable (i.e.\ $\neg D1(s,t)$), then
-for each action $a$ for $s$, there is an action $b$ for $t$ whose
-set of transitions on $t$ is indistinguishable from the set
-of transitions of $a$ on $s$; namely, for each transition $(s,a,s')$
-there is a transition $(t,b,t')$ with $\Delta_f(s,s')=\Delta_f(t,t')$
-for each feature $f$, and vice versa.
-To do this, transitions $(s,s')$ in the sample would need to be
-tagged with action labels as $(s,a,s')$.
+in terms of features that can be computed in polynomial time.
+It is an open question, however,  whether such features are  captured
+by the proposed grammar or a  suitable variation. The feature language, however, seems 
+adequate for  dealing with arbitrary goals, once  \emph{goal predicates} are added
+to the set of primitive predicates. Goal predicates are copies $p_G$ of the primitive predicates $p$ that
+appear in the goal and which are evaluated in the goal  \cite{martin-geffner:generalized}.
+
+Finally, the formulation can be extended to handle non-deterministic actions.
+For this, it is sufficient to replace the formula \eqref{eq:bridge1}
+that links $D_1$ and $D_2$ atoms  with a formula asserting that, 
+if two states $s$ and $t$ are indistinguishable (i.e.\ $\neg D_1(s,t)$), then
+for each action $a$ in $s$, there must be  an action $b$ in  $t$,
+such that the  state transitions $(t,b,t')$ generated by $b$ in $t$, 
+can't be distinguished from the transitions $(s,a,s')$ generated by $a$ in $s$. 
+For this, transitions $(s,s')$ in the sample would need to be tagged with the actions
+that generated them as  $(s,a,s')$.
 
@@
 
-There is  an interesting  relation to the ideas
-of dimensionality reduction and embeddings in machine learning
-\cite{hamilton:embeddings}. The learned abstraction
-maps the  states of the problem instances, 
-whose size is not bounded, into valuations over a small and bounded
-set of features. Moreover,  this is done while preserving the   essential properties
-of states; namely,  how they are changed by actions and whether they denote  goal states or not.
-The notion of \emph{sound} and \emph{complete} abstract action sets expresses this requirement
-that is enforced by  the SAT formulation.
-
+There is  an interesting  relation between sound and complete abstractions, on the one hand, 
+and the ideas of dimensionality reduction and embeddings in machine learning, on the other 
+\cite{hamilton:embeddings}. Sound and complete abstractions  map  the  states of the problem
+instances,  whose size is not bounded, into valuations over a small and bounded set of features,
+while preserving the   essential properties of states; namely,  how they are changed by actions
+and whether they denote  goal states or not.
+% The notion of \emph{sound} and \emph{complete} abstract action sets expresses this requirement
+% that is enforced by  the SAT formulation.
 An interesting challenge for the future  is to generalize the proposed methods
@@
 where the sampled states have no known structure and there are no primitive
-predicates. One way for achieving this, is precisely by learning embeddings
+predicates. One way for achieving this would be precisely by learning embeddings
 that yield sound and complete abstractions or suitable approximations.

### 2018-11-14 doc/camera-ready/paper.tex
Blai

@@
 one with preconditions $\neg H$ and $n(x) > 0$, and effects $H$ and $n(x)\mminus$
-(decrement $n(x)$), and the other with precondition $H$ and effect $\neg H$.
+(decrement of $n(x)$), and the other with precondition $H$ and effect $\neg H$.
 We here show how to obtain such policies from STRIPS instances alone,
@@
 and $n=0$ true iff $\phi_n(s) = 0$. An abstract action $\bar{a}$ is applicable in  $s$
-if its preconditions
-are true in $\bar{s}$.
+if its preconditions are true in $\bar{s}$.
 
@@
 \begin{example}
-  The abstract action set $A_F = \{\bar{a}, \bar{a}'\}$, where $\bar{a} = \langle \neg H,n(x)>0; H,$ $n(x)\mminus\rangle$ %$\bar{a}=\abst{\neg H,n(x)>0}{H,n(x)\mminus}$
-  and $\bar{a}'=\abst{H}{\neg H}$ , defined over the above feature set $F=\{H,n(x)\}$,
+  The abstract action set $A_F = \{\bar{a}, \bar{a}'\}$ where $\bar{a} = \abst{\neg H,n(x)>0}{H,n(x)\mminus}$
+  and $\bar{a}'=\abst{H}{\neg H}$, defined over the above feature set $F=\{H,n(x)\}$,
   is sound but not complete relative to $\Q_{clear}$.
-  $A_F$ captures the concrete actions that pick blocks from above $x$ and put them away,
-  but not other actions like stacking a block above $x$ or picking up a block that is not above $x$.
+  The actions in $A_F$ captures the concrete actions that pick blocks from above $x$ and put them away,
+  but not other actions like stacking a block above $x$, or picking a block that is not above $x$.
 \end{example}
@@
 
-\noindent where  $s$ and $t$  are expanded states  $\S$ such that exactly one of them is a goal state. 
+\noindent where  $s$ and $t$  are expanded states  in $\S$ such that exactly one of them is a goal state.
 % For this, it is assumed that expanded states $s$ in $\S$ are labeled as goal or non-goal states.
@@
 where $m$ is the number of expanded states in $\S$, and $b$ is their average branching factor (transitions per state).
-The number of clauses is bounded by $m^2(2 + b^2 + |\F|(b^2+1))$. %$m^2 \times b^2 \times |\F|$.
+The number of clauses is bounded by $m^2(2 + b^2 + b + |\F|(b^2 + 1))$. %$m^2 \times b^2 \times |\F|$.
 %
@@
 %It is not fully clear whether this assumption is reasonable in general.
-A relevant argument made by \citeay{bonet:ijcai2018} is that generalized problems
-over domains with low width \cite{nir:ecai2012}, have compact policies
-in terms of features that can be computed in polynomial time.
-It is an open question, however,  whether such features are  captured
-by the proposed grammar or a  suitable variation. The feature language, however, seems 
-adequate for  dealing with arbitrary goals, once  \emph{goal predicates} are added
-to the set of primitive predicates. Goal predicates are copies $p_G$ of the primitive predicates $p$ that
-appear in the goal and which are evaluated in the goal  \cite{martin-geffner:generalized}.
+A relevant argument made by \citeay{bonet:ijcai2018} is that generalized
+problems over domains with bounded width \cite{nir:ecai2012} have compact
+policies in terms of features $f$ whose value $\phi_f(s)$ can be computed
+in polynomial time for any state $s$. It is an open question whether
+such features are  captured by the proposed grammar, or a  suitable variation.
+The feature language, however, seems adequate for  dealing with arbitrary goals,
+once  \emph{goal predicates} are added to the set of primitive predicates.
+%Goal predicates are copies $p_G$ of the primitive predicates $p$ that appear
+%in the goal and which are evaluated in the goal  \cite{martin-geffner:generalized}.
+Goal predicates are ``copies'' $p_G$ of the primitive predicates $p$ that
+appear in the goal, and have fixed interpretation\footnote{If $p$ is
+  a primitive predicate that appears in the goal, for any state $s$ and
+  tuple $\bar u$ of objects, $s\vDash p_G(\bar u)$ iff $p(\bar u)$
+  holds in the goal.}
+\cite{martin-geffner:generalized}.
 
@@
 that links $D_1$ and $D_2$ atoms  with a formula asserting that,
-if two states $s$ and $t$ are indistinguishable (i.e.\ $\neg D_1(s,t)$), then
-for each action $a$ in $s$, there must be  an action $b$ in  $t$,
-such that the  state transitions $(t,b,t')$ generated by $b$ in $t$, 
-can't be distinguished from the transitions $(s,a,s')$ generated by $a$ in $s$. 
+if two states $s$ and $t$ are indistinguishable (i.e., $\neg D_1(s,t)$), then
+for each action $a$ in $s$, there is an action $b$ in  $t$
+such that the set of transitions $(t,b,t')$ generated by $b$ in $t$
+cannot be distinguished from the set of transitions $(s,a,s')$ generated by $a$ in $s$.
 For this, transitions $(s,s')$ in the sample would need to be tagged with the actions

### 2018-11-14 doc/camera-ready/paper.tex
Fix typo

@@
   is sound but not complete relative to $\Q_{clear}$.
-  The actions in $A_F$ captures the concrete actions that pick blocks from above $x$ and put them away,
+  The actions in $A_F$ capture the concrete actions that pick blocks from above $x$ and put them away,
   but not other actions like stacking a block above $x$, or picking a block that is not above $x$.

### 2018-11-14 doc/camera-ready/paper.tex
Last minute changes

@@
 
-Bonet and Geffner's generalized planning formulation assumes that the
+Bonet and Geffner's formulation of generalized planning  assumes that the
 features and abstract actions are given. In this work, we address this limitation
@@
 
-The paper is organized as follows. We first provide background and then show
-how abstract actions can be learned by enforcing soundness and
-completeness over a set of samples from a pool of candidate features derived
+The paper is organized as follows. We first provide the relevant background, and then show
+how features and abstract actions can be learned by enforcing soundness and
+completeness over a set of samples, from a pool of candidate features derived
 from the domain predicates. The computational model is then summarized,
@@
 that takes an instance $P$ from $\Q$ and a state $s$ reachable in $P$, and results
-in a value denoted as $\phi_f(s)$.
-A feature is \emph{boolean} if it results in a boolean value, and \emph{numeric}
-if it results in a numerical value, that we assume to be a non-negative integer.
+in a value $\phi_f(s)$.
+A feature is \emph{boolean} if it results in boolean values, and \emph{numeric}
+if it results in numerical values. %, that we assume to be a non-negative integer.
 For example, $H$ and $n(x)$ represent two features in the Blocksworld:
@@
 An action $a$ and an abstract action $\bar{a}=\abst{Pre}{\Eff}$  \emph{have  the same qualitative effects over the features $F$
-in a state $s$} when both are applicable in $s$ and have the same effects on the boolean features and
+in a state $s$} when both are applicable in $s$ with the same effects on the boolean features and
 the same \emph{qualitative effects} on the numerical features. If $s'$ is the result of applying $a$ in $s$,
 this means \cite{bonet:ijcai2018}:
-1)~$p$ becomes true in $s'$,
-i.e., $\phi_p(s')$ is true and $\phi_p(s)$  is false,
-iff $p \in \Eff$ and $\neg p \in Pre$, 2)~$p$ becomes false in $s'$, i.e., $\phi_p(s')$ is false and $\phi_p(s)$  is true,
-iff $\neg p \in \Eff$ and $p \in Pre$, 3)~$n$ increases in $s'$, i.e., $\phi_n(s') > \phi_n(s)$, iff $n\pplus \in \Eff$,
-and 4)~$n$ decreases  in $s'$, i.e., $\phi_n(s') < \phi_n(s)$, iff $n\mminus \in \Eff$.
+1)~$p \in \Eff$ and $\neg p \in Pre$ iff
+$\phi_p(s)$ is false and $\phi_p(s')$ is true (i.e., $p$ becomes true),
+%
+2)~$\neg p \in \Eff$ and $p \in Pre$ iff
+$\phi_p(s)$ is true and $\phi_p(s')$ is false ($p$ becomes false),
+%
+3)~$n\pplus \in \Eff$ iff $\phi_n(s') > \phi_n(s)$ ($n$ increases),
+and 
+%
+4)~$n\mminus \in \Eff$ iff $\phi_n(s') < \phi_n(s)$ ($n$ decreases).
 
@@
 Soundness and completeness are the key  properties that  enable us to reason about the collection of
-instances $\Q$ in terms of abstract actions that are defined at the level of the features:
+instances $\Q$ in terms of abstract actions that operate at the level of the features:
 
@@
   The abstract action set $A_F = \{\bar{a}, \bar{a}'\}$ where $\bar{a} = \abst{\neg H,n(x)>0}{H,n(x)\mminus}$
-  and $\bar{a}'=\abst{H}{\neg H}$, defined over the above feature set $F=\{H,n(x)\}$,
+  and $\bar{a}'=\abst{H}{\neg H}$, defined over the feature set $F=\{H,n(x)\}$,
   is sound but not complete relative to $\Q_{clear}$.
@@
 and we add, for each feature $f \in \F$, a
-\emph{soft} unit clause $\neg selected(f)$ with weight $cost(f)$.
-The cost measure that is used in the experiments is defined below.
+\emph{soft} unit clause $\neg selected(f)$ with weight $cost(f)$, defined below.
 
@@
 
-Finally, the formulation can be extended to handle non-deterministic actions.
+Finally, our formulation can be extended to handle non-deterministic actions.
 For this, it is sufficient to replace the formula \eqref{eq:bridge1}

### 2018-11-14 doc/camera-ready/paper.tex
Last minute changes

@@
 that takes an instance $P$ from $\Q$ and a state $s$ reachable in $P$, and results
-in a value $\phi_f(s)$.
+in a value denoted as $\phi_f(s)$.
 A feature is \emph{boolean} if it results in boolean values, and \emph{numeric}
 if it results in numerical values. %, that we assume to be a non-negative integer.
-For example, $H$ and $n(x)$ represent two features in the Blocksworld:
-the first is a boolean feature that tracks whether the gripper is empty, while
-the second is numerical and tracks the number of blocks above $x$.
+For example, $H$ and $n(x)$ represent two features in Blocksworld:
+$H$ is a boolean feature that tracks whether the gripper is empty, while
+$n(x)$ is numerical and tracks the number of blocks above $x$.
 

### 2018-11-18 doc/arxiv-version/paper.tex
Add different arxiv version sources with hyperref package tuned

--- 2018-11-14 doc/camera-ready/paper.tex
@@
 % \usepackage{soul}
-% \usepackage[hidelinks]{hyperref}
+\usepackage[hidelinks,breaklinks=true]{hyperref}
 % \usepackage[utf8]{inputenc}
@@
 A feature is \emph{boolean} if it results in boolean values, and \emph{numeric}
-if it results in numerical values. %, that we assume to be a non-negative integer.
-For example, $H$ and $n(x)$ represent two features in the Blocksworld:
-the first is a boolean feature that tracks whether the gripper is empty, while
-the second is numerical and tracks the number of blocks above $x$.
+if it results in numerical values, here assumed to be non-negative.
+For example, $H$ and $n(x)$ are two features in Blocksworld:
+$H$, boolean, tracks whether the gripper is empty, while
+$n(x)$, numerical, tracks the number of blocks above $x$.
 
@@
 the same \emph{qualitative effects} on the numerical features. If $s'$ is the result of applying $a$ in $s$,
-this means \cite{bonet:ijcai2018}:
+this means that \cite{bonet:ijcai2018}:
 1)~$p \in \Eff$ and $\neg p \in Pre$ iff
@@
 Different assignments $\sigma$ result in different feature sets $F_{\sigma}$
-and abstract action sets $A_{\sigma}$. Simpler and more meaningful features can be found
+and abstract action sets $A_{\sigma}$. Simpler and more meaningful features are found
 by minimizing a cost measure such as $|F_{\sigma}|$
 or, more generally, $\sum_{f \in F_{\sigma}} cost(f)$.
-We achieve this by casting our problem
+To achieve this, we cast our problem
 into a constrained optimization problem, more specifically, a Weighted Max-SAT problem,

### 2018-12-01 compact_encoding/encoder/build_alt_theory.cc
implement naive encoding

@@ -473,6 +473,7 @@ class Theory {
 
         AbstractAction() : selected_precondition_(0), precondition_(0), selected_effect_(0), effect_(0) { }
         AbstractAction(const string &name) : name_(name), selected_precondition_(0), precondition_(0), selected_effect_(0), effect_(0) { }
+
         void add_precondition(const string &feature, int prec) {
             if( (feature_map_ == nullptr) || (feature_map_->find(feature) == feature_map_->end()) ) {
                 cout << "error: add_precondition: feature '" << feature << "' not found in feature map; did you forget --add-features to add features in forced actions?" << endl;
@@ -559,13 +560,15 @@ class Theory {
         bool add_actions_;
         bool add_features_;
         bool only_soundness_;
+        bool log_feature_variables_;
         Options()
-          : ordering_(true),
+          : ordering_(false),
             decode_(false),
             explicit_goals_(false),
             add_actions_(false),
             add_features_(false),
-            only_soundness_(false) {
+            only_soundness_(false),
+            log_feature_variables_(true) {
         }
     };
 
@@ -574,11 +577,12 @@ class Theory {
     const Transitions &T_;
     const int K_;           // number of features in abstraction
     const int N_;           // number of actions in abstraction
-
     const Options options_;
-    Items<AbstractAction> actions_;
-    Items<string> features_;
-    Items<int> goals_;
+
+    // forced elements (added by "hand")
+    Items<AbstractAction> forced_actions_;
+    Items<string> forced_features_;
+    Items<int> forced_goals_;
 
     const int num_states_;
     const int num_features_;
@@ -598,7 +602,8 @@ class Theory {
 
   public:
     Theory(const Matrix &M, const Transitions &T, int K, int N, const Options &options)
-      : M_(M), T_(T), K_(K), N_(N), options_(options),
+      : M_(M), T_(T), K_(K), N_(N),
+        options_(options),
         num_states_(M_.num_states()),
         num_features_(M_.num_features()),
         log_num_features_(0),
@@ -657,15 +662,9 @@ class Theory {
         assert((-int(literals_.size()) <= index) && (index <= int(literals_.size())));
         return index > 0 ? *literals_[index - 1] : *literals_[variables_.size() + -index - 1];
     }
+
     int num_variables() const {
         return variables_.size();
-    }
-
-    void add_implication(const Implication *IP) {
[…]

### 2018-12-01 compact_encoding/doc/alt-encoding.tex
Fix bug in doc for alt encoding

--- 2018-07-22 compact-encoding/doc/alt-encoding.tex
@@
   &df(k,t) \land F^k_i \rightarrow F^{k+1}_i \qquad \text{for $i\in\{t+1,\ldots,L-1\}$} \,, \\
-  &df(k,t) \land \neg F^{k+1}_i \rightarrow F^k_i \qquad \text{for $i\in\{t+1,\ldots,L-1\}$} \,.
+  &df(k,t) \land \neg F^k_i \rightarrow \neg F^{k+1}_i \qquad \text{for $i\in\{t+1,\ldots,L-1\}$} \,.
 \end{alignat}

### 2020-07-28 experiments/2020/blocks.py
Blai's "local" separation encoding

@@ -29,7 +29,8 @@ def experiments():
         maxsat_encoding="separation",
         complete_only_wrt_optimal=True,
         prune_redundant_states=False,
-        optimal_selection_strategy="complete"
+        optimal_selection_strategy="complete",
+        # distinguish_transitions_locally=False,
     )
 
     exps["on"] = update_dict(

### 2020-07-28 src/features/src/blai/transitions.h

@@ -33,6 +33,7 @@ protected:
     transition_set_t marked_transitions_;
     transition_list_t unmarked_transitions_;
     transition_list_t unmarked_and_alive_transitions_;
+    std::vector<transition_list_t> unmarked_transitions_from_;
 
 public:
     TransitionSample(std::size_t num_states, std::size_t num_transitions, std::size_t num_marked_transitions)
@@ -43,7 +44,8 @@ public:
               alive_states_(num_states, false),
               marked_transitions_(),
               unmarked_transitions_(),
-              unmarked_and_alive_transitions_()
+              unmarked_and_alive_transitions_(),
+              unmarked_transitions_from_(num_states)
      {}
 
     ~TransitionSample() = default;
@@ -68,6 +70,10 @@ public:
 
     const transition_list_t& unmarked_and_alive_transitions() const {
         return unmarked_and_alive_transitions_;
+    }
+
+    const transition_list_t& unmarked_transitions_starting_at(unsigned s) const {
+        return unmarked_transitions_from_[s];
     }
 
     bool marked(const transition_t& p) const {
@@ -127,6 +133,7 @@ public:
 
                     if (!marked(src, dst)) {
                         unmarked_transitions_.push_back(std::make_pair(src, dst));
+                        unmarked_transitions_from_[src].push_back(std::make_pair(src, dst));
                     }
                     n_total_transitions++;
                 }

### 2020-07-28 src/features/src/cnf/generator.h

@@ -145,11 +145,16 @@ public:
         return sample_.transitions().unmarked_and_alive_transitions();
     }
 
-    const transition_list_t& get_relevant_unmarked_transitions(bool use_only_alive_states) const {
-        if (use_only_alive_states) {
-            return sample_.transitions().unmarked_and_alive_transitions();
+    const transition_list_t& get_relevant_unmarked_transitions(unsigned s) const {
+        if (options.distinguish_transitions_locally) {
+            assert(is_alive(s));
+            return sample_.transitions().unmarked_transitions_starting_at(s);
         } else {
-            return sample_.transitions().unmarked_transitions();
+            if (options.use_only_alive_unmarked_states) {
+                return sample_.transitions().unmarked_and_alive_transitions();
+            } else {
+                return sample_.transitions().unmarked_transitions();
+            }
         }
     }
 
@@ -202,7 +207,7 @@ public:
     std::pair<bool, CNFWriter> write_basic_maxsat(std::ostream &os);
 
     //! Generate and write the CNF instance for the transition-separation encoding as we go
-    std::pair<bool, CNFWriter> write_separation_maxsat(std::ostream &os);
+    std::pair<bool, CNFWriter> write_transition_classification_maxsat(std::ostream &os);
 };
 
 using isomorphism_t = std::unordered_map<unsigned, unsigned>;

### 2020-07-28 src/sltp/cnfgen.py

@@ -24,6 +24,7 @@ def run(config, data, rng):
     args += ["--encoding", config.maxsat_encoding]
 
     args += ["--use-only-unmarked-alive-transitions"] if config.use_only_alive_unmarked_transitions else []
+    args += ["--distinguish-transitions-locally"] if config.distinguish_transitions_locally else []
 
     retcode = execute([cmd] + args)
 

### 2020-07-28 src/sltp/tester.py

@@ -150,7 +150,7 @@ def create_pyperplan_abstract_policy_based_search(pyperplan, search_policy):
         closed = set()
 
         while not task.goal_reached(node.state):
-            # logging.info(f'Expanding: {sorted(node.state)}')
+            logging.info(f'Expanding: {sorted(node.state)}')
 
             closed.add(node.state)
             expanded += 1

### 2020-07-28 src/sltp/util/defaults.py

@@ -187,6 +187,10 @@ def generate_experiment(domain_dir, domain, **kwargs):
         # In the transition-separation CNF encoding, whether to distinguish good transitions *only from*
         # unmarked transitions that start in an alive state
         use_only_alive_unmarked_transitions=True,
+
+        # In the transition-separation CNF encoding, whether to distinguish good transitions *only from*
+        # unmarked transitions that start in the same state as the good transition
+        distinguish_transitions_locally=True,
     )
 
     parameters = {**defaults, **kwargs}  # Copy defaults, overwrite with user-specified parameters

### 2020-07-28 doc/bonet-et-al-aaai2019/paper.tex
Move paper within own folder

--- 2018-09-06 doc/paper.tex
@@
 \begin{example}
-  Let restrict  $\Q_{clear}$  to those  instances where  the gripper is initially empty
+  Let us restrict  $\Q_{clear}$  to those  instances where  the gripper is initially empty
   and there are blocks on top of $x$. For $F=\{H,n(x)\}$  and $A_F=\{\bar{a},\bar{a'}\}$ as above, 
@@
 over the samples. Like before,  we take a transition $(s,s')$ and an abstract action $\bar{a}$
-to have  the same qualitative effects over a boolean  feature  iff they both make it true , they both make it false,  or
+to have  the same qualitative effects over a boolean  feature  iff they both make it true, they both make it false,  or
 they both leave it unchanged,  and over a numerical feature  iff they both decrease it,  they both increase it,  or they both leave it unchanged.
@@
 decreases (resp.\ increases) in the transition from $s$ to $s'$. 
-If two abstract actions $\bar{a}$ and $\bar{a}$ differ only in the sign of a
+If two abstract actions $\bar{a}$ and $\bar{a}'$ differ only in the sign of a
 precondition, the precondition can be dropped, removing the duplicates and
-iterating until no more simplifications can be done. Their key properties are:
+iterating until no more simplifications can be done. The key properties of $T(\S,\F)$ and $T_G(\S,\F)$ are:
 

### 2020-07-29 doc/encodings/paper.tex
New encoding starting to work in BW

\documentclass[a4paper]{article}
%\def\year{2020}\relax
%\usepackage{aaai20}  % DO NOT CHANGE THIS
\usepackage{times}  % DO NOT CHANGE THIS
\usepackage{helvet} % DO NOT CHANGE THIS
\usepackage{courier}  % DO NOT CHANGE THIS
\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
\urlstyle{rm} % DO NOT CHANGE THIS
\def\UrlFont{\rm}  % DO NOT CHANGE THIS
\frenchspacing  % DO NOT CHANGE THIS
\setlength{\pdfpagewidth}{8.5in}  % DO NOT CHANGE THIS
\setlength{\pdfpageheight}{11in}  % DO NOT CHANGE THIS

% our packages
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{bm}
\usepackage{booktabs}
%\usepackage[inline]{enumitem}
%\usepackage{mathtools}
%\usepackage{multirow}
%\usepackage[mode=buildnew,subpreambles=true]{standalone}
\usepackage{subcaption}
%\usepackage{todonotes}
\usepackage[table]{xcolor}  % TODO Comment out for final submission
%\usepackage{pgfplots}
%\usepackage{tikz}
\usepackage{natbib}

% Some useful macros
%\newcommand{\inlinecite}[1]{\citeauthor{#1}~(\citeyear{#1})}
%\newcommand{\citealp}[1]{\citeauthor{#1}~\citeyear{#1}}

\newcommand{\free}[1]{\ensuremath{\mathrm{free}(#1)}}
\newcommand{\vars}{\ensuremath{\mathrm{vars}}}
\newcommand{\pre}{\ensuremath{\mathrm{pre}}}
\newcommand{\add}{\ensuremath{\mathrm{add}}}
\newcommand{\del}{\ensuremath{\mathrm{del}}}
\newcommand{\effs}{\ensuremath{\mathrm{effs}}}

\newcommand{\tup}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\tuple}[1]{\tup{#1}}  % Just an alias
\newcommand{\set}[1]{\ensuremath{\left\{#1 \right\}}}
\newcommand{\setst}[2]{\ensuremath{\left\{#1 \mid #2 \right\}}}

\newtheorem{definition}{Definition}
\newtheorem{definitionandtheorem}[definition]{Definition and Theorem}
\newtheorem{theorem}[definition]{Theorem}
\newtheorem{lemma}[definition]{Lemma}
\newtheorem{proposition}[definition]{Proposition}
\newtheorem{corollary}[definition]{Corollary}
\newtheorem{example}[definition]{Example}

\newcommand{\wip}[1]{{\color{red} #1}}  % From "work in progress" :-)
\newcommand{\gfm}[1]{\footnote{\color{red}{[Guillem] #1}}}

\newcommand{\numtasks}[1]{\small{(#1)}}

\setcounter{secnumdepth}{0} %May be changed to 1 or 2 if section numbers are desired.

%PDF Info Is REQUIRED.
% For /Author, add all authors within the parentheses, separated by commas. No accents or commands.
% For /Title, add Title in Mixed Case. No accents or commands. Retain the parentheses.
 \pdfinfo{
/Title ()
/Author ()
} %Leave this

\title{Representation Learning for Generalized Planning - SAT Encodings}
%Your title must be in mixed case, not sentence case.
% That means all verbs (including short verbs like be, is, using,and go),
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
\author{GFM}

\begin{document}

\maketitle

\section{One-Shot Learning of Generalized Features \& Policy}

Encoding is parametrized by
\begin{itemize}
 \item $D$: maximum allowed value for $V(s)$ for any $s$ in the training set.
 \item max. concept complexity.
\end{itemize}

\subsection{Terminology}
A state is called \emph{reachable} if there is a path to it from $s_0$, and
it is called \emph{solvable} if there is a path from it to a goal
state.
A state is \emph{alive} if it is solvable, reachable and not a
goal state~\cite{frances-et-al-ijcai2019}.

\subsection{Variables}

\subsubsection{Main Variables}
\begin{itemize}
 \item $Good(s, s')$ for $s$ alive, $s'$ solvable.
 Note that if $s$ is alive but $s'$ is not solvable,
 it cannot make sense for a policy to recommend transition $(s, s')$.

 \item $Vleq(s, d)$, for $s$ alive, and $0 \leq d \leq D$.
 $Vleq(s, d)$ stands for $V(s) \leq d$.
 Note that for states $s$ that are a goal, we want $V(s)=0$, hance $V(s)\leq d$ for all $d$.
 Thus, we can restrict SAT variables $Vleq(s, d)$ to those states $s$ that are alive.

 \item $Select(f)$, for each feature $f$ in the feature pool.
\end{itemize}

\subsubsection{Auxiliary Variables}
\begin{itemize}
 \item $GV(s, s', d) \rightarrow Good(s, s') \land Vleq(s', d)$ for $s, s'$ alive, $d \in [0, D)$.
\end{itemize}

\subsection{Constraints}

\noindent The policy has to be complete with respect to all alive states:
\begin{align}
\bigvee_{s' \text{ solvable child of } s} Good(s, s'),&\;\; \text{for $s$ alive.}
\end{align}

\noindent Labeling a transition $(s, s')$ as ``good'' necessarily upper-bounds $V(s)$:
\begin{align}
 Good(s, s') \land V(s') \leq d \rightarrow V(s) \leq d+1,&\;\; \text{for $s, s'$ alive, $d \in [0, D)$.} \\
 Good(s, s') \rightarrow V(s) \leq d+1,&\;\; \text{for $s$ alive, $s'$ goal, $d \in [0, D)$.}
\end{align}

\noindent Any upper bound on $V(s)$ (for $s$ not a goal) needs to be justified:
\begin{align}
 V(s) \leq d+1 \rightarrow \bigvee_{s' \text{ goal child of } s} Good(s, s') \lor
                           \bigvee_{s' \text{ alive child of } s} GV(s, s', d),&
                           \;\; \text{for $s$ alive, $d \in [0, D)$.} \\
 \neg V(s) \leq 0,&\;\; \text{for $s$ not a goal.}
\end{align}

\noindent Atoms $V(s) \leq d$ are consistent among them:
\begin{align}
 V(s) \leq d \rightarrow V(s) \leq d+1,&\;\; \text{for $s$ alive, $d \in [0, D)$.}
\end{align}

\noindent Good transitions can be distinguished from bad transitions:
\begin{align}
 Good(s, s') \rightarrow Good(t, t') \lor
 Distinguish(s, s', t, t'),&\;\; \text{for $s, t$ alive, $s', t'$ solvable.} \\
 Good(s, s') \rightarrow
 Distinguish(s, s', t, t'),&\;\; \text{for $s, t$ alive, $s'$ solvable, $t'$ unsolvable,}
\end{align}

\noindent where $Distinguish(s, s', t, t')$ is merely a presentational shorthand for $\bigvee_{f \in D2(s, s', t, t')} Select(f)$.

\bibliographystyle{plain}
% Cross-referenced entries need to go after the entries that cross-reference them
\bibliography{abbrv-short,literatur,references,crossref-short}
\end{document}

### 2020-07-29 setup.py

@@ -60,6 +60,7 @@ def main():
             'setuptools',
             'psutil',
             'bitarray',
+            'natsorted',
             'numpy',
             "tarski @ git+ssh://git@github.com/aig-upf/tarski.git@1c107d5#egg=tarski-devel"
         ],

### 2020-07-29 src/features/src/blai/transitions.h

@@ -28,7 +28,9 @@ protected:
     const std::size_t num_marked_transitions_;
     // trdata_[s] contains the IDs of all neighbors of s in the state space
     std::vector<std::vector<unsigned>> trdata_;
-    std::vector<bool> alive_states_;
+    std::vector<bool> is_state_alive_;
+
+    std::vector<unsigned> alive_states_;
 
     transition_set_t marked_transitions_;
     transition_list_t unmarked_transitions_;
@@ -41,7 +43,8 @@ public:
               num_transitions_(num_transitions),
               num_marked_transitions_(num_marked_transitions),
               trdata_(num_states),
-              alive_states_(num_states, false),
+              is_state_alive_(num_states, false),
+              alive_states_(),
               marked_transitions_(),
               unmarked_transitions_(),
               unmarked_and_alive_transitions_(),
@@ -84,8 +87,10 @@ public:
     }
 
     bool is_alive(unsigned state) const {
-        return alive_states_.at(state);
+        return is_state_alive_.at(state);
     }
+
+    const std::vector<unsigned>&  all_alive() const { return alive_states_; }
 
     //! Print a representation of the object to the given stream.
     friend std::ostream& operator<<(std::ostream &os, const TransitionSample& o) { return o.print(os); }
@@ -169,12 +174,13 @@ public:
             for(unsigned j = 0; j < count; ++j) {
                 is >> s;
                 assert(s < num_states_);
-                alive_states_[s] = true;
+                is_state_alive_[s] = true;
+                alive_states_.push_back(s);
             }
         }
 
         for (const auto tx:unmarked_transitions_) {
-            if (alive_states_[tx.first]) unmarked_and_alive_transitions_.push_back(tx);
+            if (is_state_alive_[tx.first]) unmarked_and_alive_transitions_.push_back(tx);
         }
     }
 
@@ -208,7 +214,7 @@ public:
             unsigned mapped_s = mapping.at(s);
             assert(mapped_s < nstates);
 
-            if (alive_states_.at(s)) alive_states.at(mapped_s) = true;
+            if (is_state_alive_.at(s)) alive_states.at(mapped_s) = true;
 
             for (unsigned sprime:successors(s)) {
                 // mapping must contain all successors of the states in selected

### 2020-07-29 src/features/src/cnf/cnfwriter.h

@@ -20,13 +20,22 @@ protected:
 
     std::ostream& os_;
 
+    std::ostream* varnamestream_;
+
 public:
     // Variable IDs must start with 1
-    explicit CNFWriter(std::ostream &os) : next_var_id_(1), accumulated_weight_(0), nclauses_(0), os_(os) {}
+    explicit CNFWriter(std::ostream &os, std::ostream* varnamestream = nullptr)
+        : next_var_id_(1), accumulated_weight_(0), nclauses_(0), os_(os), varnamestream_(varnamestream)
+    {
+    }
 
-    cnfvar_t variable() {
+    cnfvar_t variable(const std::string& name = "") {
+        const auto id = next_var_id_;
+        if (varnamestream_ != nullptr) {
+            *varnamestream_ << id << "\t" << name << std::endl;
+        }
         next_var_id_ += 1;
-        return next_var_id_-1;
+        return id;
     }
 
     uint32_t nvars() const { return next_var_id_ - 1; }
@@ -41,16 +50,17 @@ public:
     ulong top() const { return accumulated_weight_ + 1; }
 
     //! Print the given clause - if weight is negative, it is assumed to be TOP
-    void print_clause(const cnfclause_t& clause, int weight = -1) {
+    void print_clause(const cnfclause_t& clause, unsigned weight = std::numeric_limits<unsigned int>::max()) {
         //  w <literals> 0
         assert(!clause.empty());
         if (weight == 0) throw std::runtime_error("Cannot use weight-0 clauses");
-        auto size = clause.size();
-        if (weight < 0) os_ << "TOP ";
+        if (weight == std::numeric_limits<unsigned int>::max()) os_ << "TOP ";
         else {
             accumulated_weight_ += weight;
             os_ << weight << " ";
         }
+
+        auto size = clause.size();
         for (unsigned i = 0; i < size; ++i) {
             os_ << clause[i] << " ";
         }
@@ -58,4 +68,11 @@ public:
 
         nclauses_ += 1;
     }
+
+    //! A couple shorthands
+    cnfvar_t var(const std::string& name) { return variable(name); }
+    static inline cnflit_t lit(cnfvar_t var, bool polarity) { return literal(var, polarity); }
+    inline void cl(const cnfclause_t& clause, unsigned weight = std::numeric_limits<unsigned int>::max()) { print_clause(clause, weight); }
+
+
 };

### 2020-07-29 src/features/src/cnf/generator.h

@@ -119,6 +119,8 @@ public:
         }
     }
 
+    const std::vector<unsigned>& all_alive() const { return sample_.transitions().all_alive(); }
+
     bool is_goal(unsigned s) const { return sample_.matrix().goal(s); }
 
     bool is_alive(unsigned s) const { return sample_.transitions().is_alive(s); }

### 2020-07-29 src/features/src/cnfgen.cpp

@@ -160,17 +160,17 @@ int main(int argc, const char **argv) {
     std::cout << "Preprocessing time: " << preprocess_time << std::endl;
     std::cout << "Total-time: " << total_time << std::endl;
     std::cout << "CNF Theory: "  << writer.nvars() << " vars + " << writer.nclauses() << " clauses" << std::endl;
-    std::cout << "Clause breakdown: " << std::endl;
-    std::cout << "\tSelected(f): " <<  gen.n_selected_clauses << std::endl;
-    std::cout << "\tD2-definition: " <<  gen.n_d2_clauses << std::endl;
-    std::cout << "\tBridge: " <<  gen.n_bridge_clauses << std::endl;
-    std::cout << "\tGoal-distinguishing: " <<  gen.n_goal_clauses << std::endl;
-    std::cout << "\tDead-end-distinguishing: " <<  gen.n_deadend_clauses << std::endl;
-    std::cout << "\tTransition-Separation: " <<  gen.n_separation_clauses << std::endl;
-    std::cout << "\tGood-transition: " <<  gen.n_good_tx_clauses << std::endl;
-    std::cout << "\tTOTAL: " <<
-              gen.n_selected_clauses + gen.n_d2_clauses + gen.n_separation_clauses +
-              gen.n_bridge_clauses + gen.n_goal_clauses + gen.n_deadend_clauses + gen.n_good_tx_clauses << std::endl;
+//    std::cout << "Clause breakdown: " << std::endl;
+//    std::cout << "\tSelected(f): " <<  gen.n_selected_clauses << std::endl;
+//    std::cout << "\tD2-definition: " <<  gen.n_d2_clauses << std::endl;
+//    std::cout << "\tBridge: " <<  gen.n_bridge_clauses << std::endl;
+//    std::cout << "\tGoal-distinguishing: " <<  gen.n_goal_clauses << std::endl;
+//    std::cout << "\tDead-end-distinguishing: " <<  gen.n_deadend_clauses << std::endl;
+//    std::cout << "\tTransition-Separation: " <<  gen.n_separation_clauses << std::endl;
+//    std::cout << "\tGood-transition: " <<  gen.n_good_tx_clauses << std::endl;
+//    std::cout << "\tTOTAL: " <<
+//              gen.n_selected_clauses + gen.n_d2_clauses + gen.n_separation_clauses +
[…]

### 2020-07-29 src/sltp/driver.py

@@ -401,6 +401,7 @@ class TransitionClassificationPolicyStep(Step):
 
     def process_config(self, config):
         config["wsat_varmap_filename"] = compute_info_filename(config, "varmap.wsat")
+        config["wsat_allvars_filename"] = compute_info_filename(config, "allvars.wsat")
         return config
 
     def get_required_data(self):

### 2020-07-29 src/sltp/separation.py

@@ -1,4 +1,7 @@
 import itertools
+import sys
+
+from natsort import natsorted
 
 from tarski.dl import FeatureValueChange
 
@@ -21,6 +24,24 @@ def compute_good_transitions(assignment, wsat_varmap_filename):
     return list(sorted(good))
 
 
+def print_maxsat_solution(assignment, filename):
+    """ """
+    solution = []
+    vleqs = dict()
+    with open(filename, 'r') as f:
+        for line in f:
+            id_, name = line.rstrip().split("\t")
+            if assignment[int(id_)] is True:
+                # e.g. Vleq(4, 1)
+                if name.startswith("Vleq("):
+                    state, value = map(int, name[5:-1].split(", "))
+                    vleqs[state] = min(value, vleqs.get(state, sys.maxsize))
+                else:
+                    solution.append(name)
+    print('\n'.join(natsorted(solution, key=lambda y: y.lower())))
+    print('\n'.join(f"V({s}) = {vleqs[s]}" for s in sorted(vleqs.keys())))
+
+
 def compute_transition_classification_policy(config, data, rng):
     if config.transition_classification_policy is not None:
         rules = config.transition_classification_policy()
@@ -31,12 +52,14 @@ def compute_transition_classification_policy(config, data, rng):
     solution = data.cnf_solution
     assert solution.solved
 
+
     # CNF variables "selected(f)" take range from 1 to num_features+1
     selected_feature_ids = [i - 1 for i in range(1, data.num_features + 1) if solution.assignment[i] is True]
 
     features = load_selected_features(selected_feature_ids, config.domain, config.serialized_feature_filename)
     features = [IdentifiedFeature(f, i, config.feature_namer(str(f))) for i, f in zip(selected_feature_ids, features)]
 
+    print_maxsat_solution(solution.assignment, config.wsat_allvars_filename)
     good_transitions = compute_good_transitions(solution.assignment, config.wsat_varmap_filename)
 
     policy = TransitionClassificationPolicy(features)

### 2020-07-29 src/sltp/util/names.py

@@ -86,6 +86,7 @@ def blocksworld_names(feature):
         "And(And(And(Not(Exists(Star(on),Nominal(a))),Not(Exists(Star(Inverse(on)),Nominal(a)))),Not(Nominal(a))),Not(holding))": "m(a)",
         "And(And(Forall(Star(on),Not(Nominal(a))),Forall(Star(Inverse(on)),Not(Nominal(a)))),And(Not(holding),Not(Nominal(a))))": "m(a)",
         "Exists(Star(on),Exists(on,Nominal(b)))": "n-at-least-2-above-b",
+        "Not(clear)": "num-unclear",
     }
     return extend_namer_to_all_features(base).get(s, s)
 

### 2020-07-31 src/sltp/separation.py
Compute more compact policy exploiting some obvious invariants

@@ -3,7 +3,7 @@ import sys
 
 from natsort import natsorted
 
-from tarski.dl import FeatureValueChange
+from tarski.dl import FeatureValueChange, NullaryAtomFeature, EmpiricalBinaryConcept, ConceptCardinalityFeature
 
 from .language import parse_pddl
 from .util.serialization import unserialize_feature
@@ -52,32 +52,49 @@ def compute_transition_classification_policy(config, data, rng):
     solution = data.cnf_solution
     assert solution.solved
 
-
     # CNF variables "selected(f)" take range from 1 to num_features+1
     selected_feature_ids = [i - 1 for i in range(1, data.num_features + 1) if solution.assignment[i] is True]
 
     features = load_selected_features(selected_feature_ids, config.domain, config.serialized_feature_filename)
     features = [IdentifiedFeature(f, i, config.feature_namer(str(f))) for i, f in zip(selected_feature_ids, features)]
 
-    print_maxsat_solution(solution.assignment, config.wsat_allvars_filename)
+    # print_maxsat_solution(solution.assignment, config.wsat_allvars_filename)
     good_transitions = compute_good_transitions(solution.assignment, config.wsat_varmap_filename)
 
     policy = TransitionClassificationPolicy(features)
 
     for (s, t) in good_transitions:
-        m1 = data.model_cache.get_feature_model(s)
-        m2 = data.model_cache.get_feature_model(t)
+        m_s = data.model_cache.get_feature_model(s)
+        m_t = data.model_cache.get_feature_model(t)
 
         clause = []
         for f in features:
-            clause += [DNFAtom(f, f.denotation(m1) != 0),
-                       DNFAtom(f, f.feature.diff(f.denotation(m1), f.denotation(m2)))]
+            den_s = f.denotation(m_s)
+            den_t = f.denotation(m_t)
+            clause += compute_policy_clauses(f, den_s, den_t)
 
         policy.add_clause(frozenset(clause))
 
     policy.minimize()
     policy.print()
     return ExitCode.Success, dict(transition_classification_policy=policy)
+
+
+def compute_policy_clauses(f, den_s, den_t):
+    diff = f.feature.diff(den_s, den_t)
+
+    if isinstance(f.feature, NullaryAtomFeature) or isinstance(f.feature, EmpiricalBinaryConcept):
+        if diff in (FeatureValueChange.DEL, FeatureValueChange.ADD):
[…]

### 2020-07-31 src/sltp/tester.py
Remove verbose print

@@ -150,7 +150,7 @@ def create_pyperplan_abstract_policy_based_search(pyperplan, search_policy):
         closed = set()
 
         while not task.goal_reached(node.state):
-            logging.info(f'Expanding: {sorted(node.state)}')
+            # logging.info(f'Expanding: {sorted(node.state)}')
 
             closed.add(node.state)
             expanded += 1

### 2020-07-31 experiments/2020/blocks.py
Reorganize blocksworld experiments

@@ -15,12 +15,15 @@ def experiments():
     exps["clear"] = update_dict(
         base,
         pipeline="transition_classifier",
-        # instances=["sample-2balls.pddl", "sample-small.pddl"],
-        instances=["instance_5_clear_x_1.pddl"],
-        test_instances=["instance_5_clear_x_2.pddl"],
-        test_policy_instances=["instance_5_clear_x_2.pddl"],
-        # num_states=2000,
-        # num_sampled_states=300,
+        instances=["training_clear_5.pddl"],
+        test_instances=[
+            "instance_5_clear_x_1.pddl",
+            "instance_5_clear_x_2.pddl",
+        ],
+        test_policy_instances=[
+            "instance_5_clear_x_1.pddl",
+            "instance_5_clear_x_2.pddl",
+        ],
         num_states="all",
         max_concept_size=8,
         # concept_generation_timeout=120,  # in seconds
@@ -30,7 +33,15 @@ def experiments():
         complete_only_wrt_optimal=True,
         prune_redundant_states=False,
         optimal_selection_strategy="complete",
-        # distinguish_transitions_locally=False,
+    )
+
+    exps["clear_fn"] = update_dict(
+        exps["clear"],
+        domain="domain_fstrips.pddl",
+        test_domain="domain_fstrips.pddl",
+        instances=["training_clear_5_fs.pddl"],
+        test_instances=[],
+        test_policy_instances=[],
     )
 
     exps["on"] = update_dict(
@@ -40,7 +51,11 @@ def experiments():
         #            "inst_on_x_y_14.pddl",
         #            "holding_a_b_unclear.pddl",
         #            ],
-        instances=["inst_on_x_y_5.pddl"],
+        instances=[
+            "training_on_2.pddl",
+            "training_on_3.pddl",
+            "training_on_5.pddl",
+        ],
         test_instances=["inst_on_x_y_7.pddl"],
         test_policy_instances=[
             "inst_on_x_y_7.pddl",
@@ -63,6 +78,15 @@ def experiments():
         complete_only_wrt_optimal=True,
         prune_redundant_states=False,
         optimal_selection_strategy="complete"
+    )
+
+    exps["on_fn"] = update_dict(
+        exps["on"],
+        domain="domain_fstrips.pddl",
+        test_domain="domain_fstrips.pddl",
+        instances=["training_on_5_fs.pddl"],
+        test_instances=[],
+        test_policy_instances=[],
     )
 
     exps["all"] = update_dict(

### 2020-07-31 src/sltp/util/names.py

@@ -75,18 +75,23 @@ def blocksworld_names(feature):
         "And(holding,Nominal(a))": "holding(a)",
         "And(holding,Nominal(b))": "holding(b)",
         "And(Exists(on,Nominal(b)),Nominal(a))": "on(a,b)",
+        "And(Exists(loc,Nominal(b)),Nominal(a))": "on(a,b)",  # FSTRIPS
         "And(Exists(Inverse(on),Nominal(a)),Nominal(b))": "on(a,b)",
         "And(Exists(Star(on),Nominal(b)),Nominal(a))": "above(a,b)",
         "And(Not(Nominal(a)),holding)": "H",
         "Exists(Inverse(on),Nominal(a))": "Z",
         "Exists(Star(on),Nominal(a))": "n(a)",
         "Exists(Star(on),Nominal(b))": "n(b)",
+        "Exists(Star(loc),Nominal(a))": "n(a)",  # FSTRIPS
+        "Exists(Star(loc),Nominal(b))": "n(b)",  # FSTRIPS
         "And(ontable,Nominal(a))": "ontable(a)",
         "And(Forall(on,Nominal(b)),Nominal(a))": "a_on_b_ontable_or_held",
         "And(And(And(Not(Exists(Star(on),Nominal(a))),Not(Exists(Star(Inverse(on)),Nominal(a)))),Not(Nominal(a))),Not(holding))": "m(a)",
         "And(And(Forall(Star(on),Not(Nominal(a))),Forall(Star(Inverse(on)),Not(Nominal(a)))),And(Not(holding),Not(Nominal(a))))": "m(a)",
         "Exists(Star(on),Exists(on,Nominal(b)))": "n-at-least-2-above-b",
         "Not(clear)": "num-unclear",
+        "And(clear,ontable)": "n-single-blocks",
+        "Atom[handempty]": "handempty",
     }
     return extend_namer_to_all_features(base).get(s, s)
 

### 2020-08-25 doc/encodings/paper.tex
Tweaking the encoding
Update PDF with encoding

@@
 
+\newcommand{\smallpar}[1]{{\vspace{10pt}\noindent \bf #1.}}
+
 \newcommand{\free}[1]{\ensuremath{\mathrm{free}(#1)}}
@@
 \newcommand{\numtasks}[1]{\small{(#1)}}
+
+\newcommand{\badtx}{\ensuremath{\mathrm{BAD}}}
 
@@
 \subsection{Terminology}
+We use $T$ to denote the set of all transitions in the training sample.
 A state is called \emph{reachable} if there is a path to it from $s_0$, and
@@
 
+\paragraph{``Bad'' transitions.}
+We use \badtx{} to denote the set of transitions that have been determined at preprocessing as necessarily
+\emph{not} good.
+At the moment, this set contains all transitions that go from an alive to a dead state and, if using
+equivalence relations, all those other transitions in whose equivalence class there is some other ``bad'' transition.
+
 \subsection{Variables}
@@
 \begin{itemize}
- \item $Good(s, s')$ for $s$ alive, $s'$ solvable.
- Note that if $s$ is alive but $s'$ is not solvable,
- it cannot make sense for a policy to recommend transition $(s, s')$.
+ \item $Good(s, s')$ for $s$ alive, $s'$ solvable and $(s, s') \not\in \badtx$.
 
- \item $Vleq(s, d)$, for $s$ alive, and $0 \leq d \leq D$.
- $Vleq(s, d)$ stands for $V(s) \leq d$.
- Note that for states $s$ that are a goal, we want $V(s)=0$, hance $V(s)\leq d$ for all $d$.
+ \item $Vleq(s, d)$, for $s$ alive, and $d \in [0, D]$, which stands for $V(s) \leq d$.
+ Note that for states $s$ that are a goal, we want $V(s)=0$, hance $V(s) \leq d$ for all $d$.
  Thus, we can restrict SAT variables $Vleq(s, d)$ to those states $s$ that are alive.
+ We'll use the notation $V(s) \leq d$ in what follows.
 
@@
 \begin{itemize}
- \item $GV(s, s', d) \rightarrow Good(s, s') \land Vleq(s', d)$ for $s, s'$ alive, $d \in [0, D)$.
+ \item $GV(s, s', d) \rightarrow Good(s, s') \land V(s') \leq d$ for $s, s'$ alive, $(s, s') \not\in \badtx$, $d \in [0, D)$.
 \end{itemize}
 
-\subsection{Constraints}
+\newpage
 
-\noindent The policy has to be complete with respect to all alive states:
+\subsection{Hard Constraints}
+
+\smallpar{C1}
+The policy has to be complete with respect to all alive states:
 \begin{align}
-\bigvee_{s' \text{ solvable child of } s} Good(s, s'),&\;\; \text{for $s$ alive.}
+\bigvee_{s' \text{ s.t. } (s, s') \in T \setminus \badtx} Good(s, s'),&\;\; \text{for $s$ alive.}
 \end{align}
 
-\noindent Labeling a transition $(s, s')$ as ``good'' necessarily upper-bounds $V(s)$:
+\smallpar{C2-3} Labeling a transition $(s, s')$ as ``good'' necessarily upper-bounds $V(s)$:
 \begin{align}
- Good(s, s') \land V(s') \leq d \rightarrow V(s) \leq d+1,&\;\; \text{for $s, s'$ alive, $d \in [0, D)$.} \\
- Good(s, s') \rightarrow V(s) \leq d+1,&\;\; \text{for $s$ alive, $s'$ goal, $d \in [0, D)$.}
+ Good(s, s') \land V(s') \leq d \rightarrow V(s) \leq d+1,&\;\; \text{for $s, s'$ alive, $(s, s') \not\in \badtx$, $d \in [0, D)$.} \\
+ Good(s, s') \rightarrow V(s) \leq d+1,&\;\; \text{for $s$ alive, $s'$ goal, $(s, s') \not\in \badtx$, $d \in [0, D)$.}
 \end{align}
 
-\noindent Any upper bound on $V(s)$ (for $s$ not a goal) needs to be justified:
+\smallpar{C3'} All transitions chosen as \emph{Good} need to have the same q-value:
 \begin{align}
- V(s) \leq d+1 \rightarrow \bigvee_{s' \text{ goal child of } s} Good(s, s') \lor
-                           \bigvee_{s' \text{ alive child of } s} GV(s, s', d),&
+ V(s) \leq d+1 \land \neg V(s') \leq d \rightarrow -Good(s, s'),&\;\; \text{for $s, s'$ alive, $(s, s') \not\in \badtx$, $d \in [0, D)$.} \tag{\theequation${}^\prime$}
+\end{align}
+
+\smallpar{C4-5} Any upper bound on $V(s)$ (for $s$ not a goal) needs to be justified:
+\begin{align}
+ V(s) \leq d+1 \rightarrow \bigvee_{\substack{
+ s' \text{ goal child of } s\\
+ (s, s') \not\in \badtx}} Good(s, s') \lor
+ \bigvee_{\substack{
+ s' \text{ alive child of } s\\
+ (s, s') \not\in \badtx}} GV(s, s', d),&
  \;\; \text{for $s$ alive, $d \in [0, D)$.} \\
@@
 
-\noindent Atoms $V(s) \leq d$ are consistent among them:
+\smallpar{C6}
+At least one atom $V(s) \leq d$ is true for each alive state $s$:
+\begin{align}
+ \bigvee_{d \in [0, D]} V(s) \leq d,&\;\; \text{for $s$ alive.}
+\end{align}
+
+\smallpar{C7}
+Atoms $V(s) \leq d$ are consistent among them:
 \begin{align}
@@
 
-\noindent Good transitions can be distinguished from bad transitions:
+\smallpar{C8-9}
+Good transitions can be distinguished from bad transitions.
+Let $(s, s')$ and $(t, t')$ be \emph{representative} transitions
+of two different equivalence classes such that $(s, s') \not\in \badtx$
+(which implies that $s'$ is solvable).
+
 \begin{align}
  Good(s, s') \rightarrow Good(t, t') \lor
- Distinguish(s, s', t, t'),&\;\; \text{for $s, t$ alive, $s', t'$ solvable.} \\
+ Dist(s, s', t, t'),&\;\; \text{for $s, t$ alive, $(t, t') \not\in \badtx$.} \\
  Good(s, s') \rightarrow
- Distinguish(s, s', t, t'),&\;\; \text{for $s, t$ alive, $s'$ solvable, $t'$ unsolvable,}
+ Dist(s, s', t, t'),&\;\; \text{for $s, t$ alive, $(t, t') \in \badtx$.}
 \end{align}
 
-\noindent where $Distinguish(s, s', t, t')$ is merely a presentational shorthand for $\bigvee_{f \in D2(s, s', t, t')} Select(f)$.
+%\begin{align}
+% Good(s, s') \rightarrow Good(t, t') \lor
+% Distinguish(s, s', t, t'),&\;\; \text{for $s, t$ alive, $s', t'$ solvable.} \\
+% Good(s, s') \rightarrow
+% Distinguish(s, s', t, t'),&\;\; \text{for $s, t$ alive, $s'$ solvable, $t'$ unsolvable,}
+%\end{align}
+
+\noindent where $Dist(s, s', t, t')$ is merely a presentational shorthand for $\bigvee_{f \in D2(s, s', t, t')} Select(f)$.
+
+% Note that the clauses below are not necessary, as we don't even declare SAT variables Good(s, s') when (s, s') \in BAD
+%\smallpar{C10} Necessarily-bad transitions can be declared as such:
+%\begin{align}
+% \neg Good(s, s'),&\;\; \text{for $(s, s') \in \badtx$.}
+%\end{align}
+
+\subsection{Soft Constraints}
+We simply post a constraint $\neg Select(f)$ for each feature $f$ in the pool, with weight equal to the complexity of $f$.
 

### 2020-08-25 experiments/2020/blocks.py
Add "use_equivalence_classes" option

@@ -33,6 +33,7 @@ def experiments():
         complete_only_wrt_optimal=True,
         prune_redundant_states=False,
         optimal_selection_strategy="complete",
+        use_equivalence_classes=True,
     )
 
     exps["clear_fn"] = update_dict(
@@ -77,7 +78,8 @@ def experiments():
         maxsat_encoding="separation",
         complete_only_wrt_optimal=True,
         prune_redundant_states=False,
-        optimal_selection_strategy="complete"
+        optimal_selection_strategy="complete",
+        use_equivalence_classes=True,
     )
 
     exps["on_fn"] = update_dict(
@@ -113,7 +115,8 @@ def experiments():
         maxsat_encoding="separation",
         complete_only_wrt_optimal=True,
         prune_redundant_states=False,
-        optimal_selection_strategy="complete"
+        optimal_selection_strategy="complete",
+        use_equivalence_classes=True,
     )
 
     return exps

### 2020-08-25 experiments/2020/gripper.py

@@ -30,7 +30,8 @@ def experiments():
         maxsat_encoding="separation",
         complete_only_wrt_optimal=True,
         prune_redundant_states=False,
-        optimal_selection_strategy="complete"
+        optimal_selection_strategy="complete",
+        use_equivalence_classes=True,
     )
 
     exps["quick"] = update_dict(

### 2020-08-25 experiments/2020/reward.py

@@ -30,7 +30,8 @@ def experiments():
         maxsat_encoding="separation",
         complete_only_wrt_optimal=True,
         prune_redundant_states=False,
-        optimal_selection_strategy="complete"
+        optimal_selection_strategy="complete",
+        use_equivalence_classes=True,
     )
 
 

### 2020-08-25 experiments/2020/spanner.py

@@ -49,6 +49,7 @@ def experiments():
         prune_redundant_states=False,
         optimal_selection_strategy="complete",
         use_only_alive_unmarked_transitions=True,
+        use_equivalence_classes=True,
         # transition_classification_policy=debug_policy
     )
 

### 2020-08-25 src/features/src/cnf/generator.h

@@ -18,12 +18,31 @@ struct Options {
         TransitionSeparation
     };
 
+    //! The path of the workspace where output files will be left
     std::string workspace;
-    bool prune_redundant_states;
-    bool verbose;
+
+    //! The type of encoding we want to use
     Encoding encoding;
+
+    //! In the transition-separation encoding, whether we want to exploit the equivalence relation
+    //! among transitions given by the feature pool
+    bool use_equivalence_classes;
+
+    //! Whether to prune states that appear redundant for the given feature pool
+    bool prune_redundant_states;
+
+    //! Whether to be more verbose in the generation of the encoding
+    bool verbose;
+
+    //! In the transition-separation CNF encoding, whether to distinguish good transitions *only from*
+    //! unmarked transitions that start in an alive state
     bool use_only_alive_unmarked_states;
+
+    //! In the transition-separation CNF encoding, whether to distinguish good transitions *only from*
+    //! unmarked transitions that start in the same state as the good transition
     bool distinguish_transitions_locally;
+
+    //! A list of user-provided feature IDs for which we want to enforce selection
     std::vector<unsigned> enforced_features;
 
 

### 2020-08-25 src/features/src/cnfgen.cpp

@@ -27,30 +27,34 @@ sltp::cnf::Options parse_options(int argc, const char **argv) {
                                         "matrix and transition samples.\n\nOptions");
 
     description.add_options()
-            ("help,h", "Display this help message and exit.")
-            ("verbose,v", "Show extra debugging messages.")
+        ("help,h", "Display this help message and exit.")
+        ("verbose,v", "Show extra debugging messages.")
 
-            ("workspace,w", po::value<std::string>()->required(),
-                    "Directory where the input files (feature matrix, transition sample) reside, "
-                    "and where the output .wsat file will be left.")
+        ("workspace,w", po::value<std::string>()->required(),
+                "Directory where the input files (feature matrix, transition sample) reside, "
+                "and where the output .wsat file will be left.")
 
-            ("prune-redundant-states,p",
-                    "Whether to prune those states that appear redundant for the given feature pool.")
+        ("prune-redundant-states,p",
+                "Whether to prune those states that appear redundant for the given feature pool.")
 
-            ("enforce-features,e", po::value<std::string>()->default_value(""),
-             "Comma-separated (no spaces!) list of IDs of feature we want to enforce in the abstraction.")
+        ("enforce-features,e", po::value<std::string>()->default_value(""),
+         "Comma-separated (no spaces!) list of IDs of feature we want to enforce in the abstraction.")
 
-            ("encoding", po::value<std::string>()->default_value("basic"),
-                 "The encoding to be used (options: {basic, d2tree, separation}).")
+        ("encoding", po::value<std::string>()->default_value("basic"),
+             "The encoding to be used (options: {basic, d2tree, separation}).")
 
-            ("use-only-unmarked-alive-transitions",
-             "In the transition-separation CNF encoding, whether to distinguish good transitions *only from*"
-             " unmarked transitions that start in an alive state")
+        ("use-only-unmarked-alive-transitions",
+         "In the transition-separation CNF encoding, whether to distinguish good transitions *only from*"
+         " unmarked transitions that start in an alive state")
 
-            ("distinguish-transitions-locally",
-             "In the transition-separation CNF encoding, whether to distinguish good transitions *only from*"
-             " unmarked transitions starting in the same state")
-            ;
+        ("use-equivalence-classes",
+         "In the transition-separation encoding, whether we want to exploit the equivalence relation "
[…]

### 2020-08-25 src/sltp/cnfgen.py

@@ -25,6 +25,7 @@ def run(config, data, rng):
 
     args += ["--use-only-unmarked-alive-transitions"] if config.use_only_alive_unmarked_transitions else []
     args += ["--distinguish-transitions-locally"] if config.distinguish_transitions_locally else []
+    args += ["--use-equivalence-classes"] if config.use_equivalence_classes else []
 
     retcode = execute([cmd] + args)
 

### 2020-08-25 src/sltp/util/defaults.py

@@ -191,6 +191,10 @@ def generate_experiment(expid, domain_dir, domain, **kwargs):
         # In the transition-separation CNF encoding, whether to distinguish good transitions *only from*
         # unmarked transitions that start in the same state as the good transition
         distinguish_transitions_locally=True,
+
+        # In the transition-separation encoding, whether we want to exploit the equivalence relation
+        # among transitions given by the feature pool
+        use_equivalence_classes=False
     )
 
     parameters = {**defaults, **kwargs}  # Copy defaults, overwrite with user-specified parameters

### 2020-08-25 src/sltp/util/names.py

@@ -60,6 +60,7 @@ def spanner_names(feature):
         "Exists(at,Forall(Inverse(at),man))": "bob-in-empty-loc",
         "Exists(Star(link),Exists(Inverse(at),spanner))": "num-locs-from-which-some-spanner-is-reachable",
         "Exists(Inverse(Star(link)),Exists(Inverse(at),<universe>))": "num-locs-reachable-from-a-loc-with-things",
+        "And(Forall(at,<empty>),useable)": "picked-up-spanners",
     }
     d = extend_namer_to_all_features(base)
     return d.get(s, s)

### 2020-08-25 src/features/src/cnf/transition_classification.h
Exploit "necessarily bad transitions" + some fixes

@@ -21,7 +21,8 @@ public:
             transition_ids_inv_(),
             class_representatives_(),
             from_transition_to_eq_class_(),
-            types_()
+            types_(),
+            necessarily_bad_transitions_()
     {
         compute_equivalence_relations();
     }
@@ -30,12 +31,16 @@ public:
     std::pair<bool, CNFWriter> write(std::ostream &os) override;
 
     inline unsigned get_transition_id(uint16_t s, uint16_t t) const { return transition_ids_.at(state_pair(s, t)); }
+
     inline unsigned get_representative_id(unsigned tx) const { return from_transition_to_eq_class_.at(tx); }
 
     inline unsigned get_class_representative(uint16_t s, uint16_t t) const {
         return get_representative_id(get_transition_id(s, t));
     }
 
+    inline bool is_necessarily_bad(unsigned tx) const {
+        return necessarily_bad_transitions_.find(tx) != necessarily_bad_transitions_.end();
+    }
 
 protected:
     // A mapping from pairs of state to the assigned transition id
@@ -51,6 +56,8 @@ protected:
 
     // A mapping from the ID of the transition to its transition type
     std::vector<transition_type> types_;
+
+    std::unordered_set<unsigned> necessarily_bad_transitions_;
 
     //!
     void compute_equivalence_relations();

### 2020-08-25 src/features/src/cnf/generator.h
Small refactoring

@@ -71,6 +71,10 @@ std::vector<feature_t> compute_d1_distinguishing_features(const Sample::Sample& 
 std::vector<feature_t> compute_d2_distinguishing_features(const Sample::Sample& sample,
         unsigned s, unsigned sprime, unsigned t, unsigned tprime);
 
+//! Return a sorted vector with those features that d2-distinguish transition (s, s') from (t, t')
+std::vector<feature_t> compute_d1d2_distinguishing_features(const Sample::Sample& sample,
+                                                          unsigned s, unsigned sprime, unsigned t, unsigned tprime);
+
 struct transition_pair {
     transition_pair(uint16_t s, uint16_t sprime, uint16_t t, uint16_t tprime) :
             s(s), sprime(sprime), t(t), tprime(tprime)

### 2020-08-25 doc/encodings/paper.tex
Fix typo in PDF
Improve encoding PDF a little bit

@@
 % directly follow a colon or long dash
-\author{GFM}
+%\author{GFM}
 
@@
 
-\section{One-Shot Learning of Generalized Features \& Policy}
+\section{Learning of Generalized Features \& Policy}
 
@@
 \begin{itemize}
- \item $D$: maximum allowed value for $V(s)$ for any $s$ in the training set.
- \item max. concept complexity.
+ \item Pool $F$ of description logic features $f$, each with given feature complexity $\mathcal{K}(f)$.
+ \item Training set consisting of a sample of transitions from a number of instances of the same domain.
+       At the moment we're assuming the sample is complete, and we have full information on whether each state
+       in the sample is a goal state, an unsolvable state, or otherwise (see below for definitions). We also have access
+       to the minimum distance to a goal $V^*(s)$ for each state $s$.
+ \item $D$: maximum allowed value for $V(s)$ for any $s$ in the training set.\footnote{This is to be improved based on $V^*(s)$.}
 \end{itemize}
@@
 \subsection{Terminology}
-We use $T$ to denote the set of all transitions in the training sample.
 A state is called \emph{reachable} if there is a path to it from $s_0$, and
@@
 
+We use $T$ to denote the set of all transitions $(s, s')$ in the training sample such that $s$ is alive.
+
+\subsection{Improvements}
+
+\paragraph{Non-distinguishability of transitions as an equivalence relation.}
+Any fixed, given pool of features $F$ implicitly defines an equivalence relation where to transitions are
+equivalent iff they cannot be distinguished \emph{by any feature in $F$}.
+If two transitions cannot be distinguished by any feature, then clearly either the policy computed by the SAT solver
+considers all of them as ``good'', or as ``bad''.
+We'll exploit this by using one single SAT variable to denote whether \emph{all transitions in a given equivalence
+class} are good or bad. When exploiting this notion of equivalence (which is implemented as an optional feature of
+the CNF generator), then every mention below to SAT variable $Good(s, s')$ needs to be read as $Good(s_{\star}, s_{\star}')$,
+where $(s_{\star}, s_{\star}')$ is the \emph{representative} transition of the equivalence class to which $(s, s')$ belongs.
+
 \paragraph{``Bad'' transitions.}
@@
 \emph{not} good.
-At the moment, this set contains all transitions that go from an alive to a dead state and, if using
+At the moment, this set contains all transitions that go from an alive to an unsolvable state and, if using
 equivalence relations, all those other transitions in whose equivalence class there is some other ``bad'' transition.
+
+\paragraph{Feature dominance.}
+Work In Progress.\footnote{This is implemented, but needs to be adapted to the equivalence classes improvement.}
 
@@
 
- \item $Vleq(s, d)$, for $s$ alive, and $d \in [0, D]$, which stands for $V(s) \leq d$.
- Note that for states $s$ that are a goal, we want $V(s)=0$, hance $V(s) \leq d$ for all $d$.
- Thus, we can restrict SAT variables $Vleq(s, d)$ to those states $s$ that are alive.
- We'll use the notation $V(s) \leq d$ in what follows.
+ \item $V(s) \leq d$ for $s$ alive, and $d \in [0, D]$.
+ Note that for states $s$ that are a goal, we know $V(s)=0$, hence $V(s) \leq d$ for all $d$,
+ and for states $s$ that are unsolvable, we know that $V(s) \not\leq d$ for all $d$.
+ Thus, we can restrict SAT variables $V(s) \leq d$ to those states $s$ that are alive.
 
@@
 \begin{itemize}
- \item $GV(s, s', d) \rightarrow Good(s, s') \land V(s') \leq d$ for $s, s'$ alive, $(s, s') \not\in \badtx$, $d \in [0, D)$.
+ \item We'll use $GV(s, s', d)$ to denote that $Good(s, s')$ and $V(s') \leq d$. This is enforced with the constraint
+ $GV(s, s', d) \rightarrow Good(s, s') \land V(s') \leq d$ for $s, s'$ alive, $(s, s') \not\in \badtx$, $d \in [0, D)$.
+ The other direction of the implication is not necessary.
 \end{itemize}
@@
 \begin{align}
- V(s) \leq d+1 \land \neg V(s') \leq d \rightarrow -Good(s, s'),&\;\; \text{for $s, s'$ alive, $(s, s') \not\in \badtx$, $d \in [0, D)$.} \tag{\theequation${}^\prime$}
+ V(s) \leq d+1 \land \neg V(s') \leq d \rightarrow \neg Good(s, s'),&\;\; \text{for $s, s'$ alive, $(s, s') \not\in \badtx$, $d \in [0, D)$.} \tag{\theequation${}^\prime$}
 \end{align}
@@
 of two different equivalence classes such that $(s, s') \not\in \badtx$
-(which implies that $s'$ is solvable).
+(which implies that $s'$ is solvable). Then,
 
@@
 
-%\begin{align}
-% Good(s, s') \rightarrow Good(t, t') \lor
-% Distinguish(s, s', t, t'),&\;\; \text{for $s, t$ alive, $s', t'$ solvable.} \\
-% Good(s, s') \rightarrow
-% Distinguish(s, s', t, t'),&\;\; \text{for $s, t$ alive, $s'$ solvable, $t'$ unsolvable,}
-%\end{align}
-
-\noindent where $Dist(s, s', t, t')$ is merely a presentational shorthand for $\bigvee_{f \in D2(s, s', t, t')} Select(f)$.
-
-% Note that the clauses below are not necessary, as we don't even declare SAT variables Good(s, s') when (s, s') \in BAD
-%\smallpar{C10} Necessarily-bad transitions can be declared as such:
-%\begin{align}
-% \neg Good(s, s'),&\;\; \text{for $(s, s') \in \badtx$.}
-%\end{align}
+\noindent where $Dist(s, s', t, t')$ is shorthand for $\bigvee_{f \in D2(s, s', t, t')} Select(f)$.
 
 \subsection{Soft Constraints}
-We simply post a constraint $\neg Select(f)$ for each feature $f$ in the pool, with weight equal to the complexity of $f$.
+We simply post a constraint $\neg Select(f)$ for each feature $f$ in the pool, with weight equal to its complexity $\mathcal{K}(f)$.
 

### 2020-08-25 experiments/2020/spanner.py
Remove unused option "use_only_alive_unmarked_transitions"

@@ -48,7 +48,6 @@ def experiments():
         complete_only_wrt_optimal=True,
         prune_redundant_states=False,
         optimal_selection_strategy="complete",
-        use_only_alive_unmarked_transitions=True,
         use_equivalence_classes=True,
         # transition_classification_policy=debug_policy
     )

### 2020-08-25 src/features/src/cnf/generator.h

@@ -35,10 +35,6 @@ struct Options {
     bool verbose;
 
     //! In the transition-separation CNF encoding, whether to distinguish good transitions *only from*
-    //! unmarked transitions that start in an alive state
-    bool use_only_alive_unmarked_states;
-
-    //! In the transition-separation CNF encoding, whether to distinguish good transitions *only from*
     //! unmarked transitions that start in the same state as the good transition
     bool distinguish_transitions_locally;
 
@@ -210,11 +206,7 @@ public:
             assert(is_alive(s));
             return sample_.transitions().unmarked_transitions_starting_at(s);
         } else {
-            if (options.use_only_alive_unmarked_states) {
-                return sample_.transitions().unmarked_and_alive_transitions();
-            } else {
-                return sample_.transitions().unmarked_transitions();
-            }
+            return sample_.transitions().unmarked_transitions();
         }
     }
 

### 2020-08-25 src/features/src/cnfgen.cpp

@@ -43,10 +43,6 @@ sltp::cnf::Options parse_options(int argc, const char **argv) {
         ("encoding", po::value<std::string>()->default_value("basic"),
              "The encoding to be used (options: {basic, d2tree, separation}).")
 
-        ("use-only-unmarked-alive-transitions",
-         "In the transition-separation CNF encoding, whether to distinguish good transitions *only from*"
-         " unmarked transitions that start in an alive state")
-
         ("use-equivalence-classes",
          "In the transition-separation encoding, whether we want to exploit the equivalence relation "
          "among transitions given by the feature pool")
@@ -75,7 +71,6 @@ sltp::cnf::Options parse_options(int argc, const char **argv) {
     options.workspace = vm["workspace"].as<std::string>();
     options.prune_redundant_states = vm.count("prune-redundant-states") > 0;
     options.verbose = vm.count("verbose") > 0;
-    options.use_only_alive_unmarked_states = vm.count("use-only-unmarked-alive-transitions") > 0;
     options.distinguish_transitions_locally = vm.count("distinguish-transitions-locally") > 0;
     options.use_equivalence_classes = vm.count("use-equivalence-classes") > 0;
 

### 2020-08-25 src/sltp/cnfgen.py

@@ -23,7 +23,6 @@ def run(config, data, rng):
 
     args += ["--encoding", config.maxsat_encoding]
 
-    args += ["--use-only-unmarked-alive-transitions"] if config.use_only_alive_unmarked_transitions else []
     args += ["--distinguish-transitions-locally"] if config.distinguish_transitions_locally else []
     args += ["--use-equivalence-classes"] if config.use_equivalence_classes else []
 

### 2020-08-25 src/sltp/util/defaults.py

@@ -185,10 +185,6 @@ def generate_experiment(expid, domain_dir, domain, **kwargs):
         transition_classification_policy=None,
 
         # In the transition-separation CNF encoding, whether to distinguish good transitions *only from*
-        # unmarked transitions that start in an alive state
-        use_only_alive_unmarked_transitions=True,
-
-        # In the transition-separation CNF encoding, whether to distinguish good transitions *only from*
         # unmarked transitions that start in the same state as the good transition
         distinguish_transitions_locally=True,
 

### 2020-08-26 doc/encodings/paper.tex
Implement an adaptive value for the upper bound on V(s)

@@
        to the minimum distance to a goal $V^*(s)$ for each state $s$.
- \item $D$: maximum allowed value for $V(s)$ for any $s$ in the training set.\footnote{This is to be improved based on $V^*(s)$.}
+ \item A parameter $\delta$ which is a ``slack'' value to determine the maximum deviation from the optimal $V^*(s)$
+ what we will allow in our policy. This will be made clearer in the encoding below.
 \end{itemize}
@@
 
- \item $V(s) \leq d$ for $s$ alive, and $d \in [0, D]$.
+ \item $V(s) \leq d$ for $s$ alive, and $d \in [0, D]$, where $D = \max_{s} \delta \cdot V^*(s)$.
  Note that for states $s$ that are a goal, we know $V(s)=0$, hence $V(s) \leq d$ for all $d$,
@@
 
+\smallpar{C10}
+$V(s)$ cannot be much larger than $V^+(s)$:
+\begin{align}
+ V(s) \leq \delta \cdot V^*(s),&\;\; \text{for $s$ alive.}
+\end{align}
+
 \subsection{Soft Constraints}

### 2020-08-26 experiments/2020/spanner.py

@@ -50,6 +50,7 @@ def experiments():
         optimal_selection_strategy="complete",
         use_equivalence_classes=True,
         # transition_classification_policy=debug_policy
+        v_slack=2,
     )
 
     return exps

### 2020-08-26 src/features/src/blai/transitions.h

@@ -32,6 +32,8 @@ protected:
 
     std::vector<unsigned> alive_states_;
 
+    std::vector<int> vstar_;
+
     transition_set_t marked_transitions_;
     transition_list_t unmarked_transitions_;
     transition_list_t unmarked_and_alive_transitions_;
@@ -58,6 +60,8 @@ public:
     std::size_t num_states() const { return num_states_; }
     std::size_t num_transitions() const { return num_transitions_; }
     std::size_t num_marked_transitions() const { return num_marked_transitions_; }
+
+    int vstar(unsigned sid) const { return vstar_.at(sid); }
 
     const std::vector<unsigned>& successors(unsigned s) const {
         return trdata_.at(s);
@@ -181,6 +185,14 @@ public:
 
         for (const auto tx:unmarked_transitions_) {
             if (is_state_alive_[tx.first]) unmarked_and_alive_transitions_.push_back(tx);
+        }
+
+        // Store the value of V^*(s) for each state s
+        int vstar;
+        vstar_.reserve(num_states_);
+        for (s=0; s < num_states_; ++s) {
+            is >> vstar;
+            vstar_.push_back(vstar);
         }
     }
 

### 2020-08-26 src/features/src/cnf/generator.h

@@ -34,6 +34,9 @@ struct Options {
     //! Whether to be more verbose in the generation of the encoding
     bool verbose;
 
+    //! The slack value for the maximum allowed value for V_\pi(s) = slack * V^*(s)
+    double v_slack;
+
     //! In the transition-separation CNF encoding, whether to distinguish good transitions *only from*
     //! unmarked transitions that start in the same state as the good transition
     bool distinguish_transitions_locally;

### 2020-08-26 src/features/src/cnf/transition_classification.h

@@ -42,6 +42,23 @@ public:
         return necessarily_bad_transitions_.find(tx) != necessarily_bad_transitions_.end();
     }
 
+    inline int get_max_v(unsigned s) const {
+        int vstar = sample_.transitions().vstar(s);
+        if (vstar < 0) return -1;
+        return std::ceil(options.v_slack * vstar);
+    }
+
+    inline unsigned compute_D() const {
+        // return 20;
+        // D will be the maximum over the set of alive states of the upper bounds on V_pi
+        unsigned D = 0;
+        for (const auto s:all_alive()) {
+            auto max_v_s = get_max_v(s);
+            if (max_v_s > D) D = max_v_s;
+        }
+        return D;
+    }
+
 protected:
     // A mapping from pairs of state to the assigned transition id
     std::unordered_map<state_pair, unsigned, boost::hash<state_pair>> transition_ids_;

### 2020-08-26 src/features/src/cnfgen.cpp

@@ -40,6 +40,10 @@ sltp::cnf::Options parse_options(int argc, const char **argv) {
         ("enforce-features,e", po::value<std::string>()->default_value(""),
          "Comma-separated (no spaces!) list of IDs of feature we want to enforce in the abstraction.")
 
+
+        ("v_slack", po::value<double>()->default_value(2),
+         "The slack value for the maximum allowed value for V_pi(s) = slack * V^*(s)")
+
         ("encoding", po::value<std::string>()->default_value("basic"),
              "The encoding to be used (options: {basic, d2tree, separation}).")
 
@@ -73,6 +77,7 @@ sltp::cnf::Options parse_options(int argc, const char **argv) {
     options.verbose = vm.count("verbose") > 0;
     options.distinguish_transitions_locally = vm.count("distinguish-transitions-locally") > 0;
     options.use_equivalence_classes = vm.count("use-equivalence-classes") > 0;
+    options.v_slack = vm["v_slack"].as<double>();
 
     auto enc = vm["encoding"].as<std::string>();
     if (enc == "basic") options.encoding = sltp::cnf::Options::Encoding::Basic;

### 2020-08-26 src/sltp/cnfgen.py

@@ -25,6 +25,7 @@ def run(config, data, rng):
 
     args += ["--distinguish-transitions-locally"] if config.distinguish_transitions_locally else []
     args += ["--use-equivalence-classes"] if config.use_equivalence_classes else []
+    args += ["--v_slack", str(config.v_slack)]
 
     retcode = execute([cmd] + args)
 

### 2020-08-26 src/sltp/outputs.py

@@ -16,12 +16,15 @@ def print_transition_matrix(state_ids, transitions, transitions_filename):
             print("{} {}".format(s, " ".join("{}".format(sprime) for sprime in succ)), file=f)
 
 
-def print_sat_transition_matrix(state_ids, transitions, alive, optimal_transitions, transitions_filename):
+def print_sat_transition_matrix(state_ids, transitions, alive, vstar, optimal_transitions, transitions_filename):
     num_transitions = sum(len(targets) for targets in transitions.values())
     num_states = len(transitions.keys())
     num_optimal_transitions = len(optimal_transitions)
     logging.info(f"Printing SAT transition matrix with {len(state_ids)} states,"
                  f" {num_states} expanded states and {num_transitions} transitions to '{transitions_filename}'")
+
+    # State Ids should start at 0 and be contiguous
+    assert state_ids == list(range(0, len(state_ids)))
 
     with open(transitions_filename, 'w') as f:
         # first line: <#states> <#transitions> <#marked-transitions>
@@ -40,3 +43,6 @@ def print_sat_transition_matrix(state_ids, transitions, alive, optimal_transitio
 
         # Next: A list of all alive states
         print(f'{len(alive)} ' + " ".join(map(str, alive)), file=f)
+
+        # Next: A space-separated list of V^*(s) values, one per each state s, where -1 denotes infinity
+        print(' '.join(str(vstar.get(s, -1)) for s in state_ids), file=f)

### 2020-08-26 src/sltp/sampling.py

@@ -29,6 +29,7 @@ class TransitionSample:
         self.instance = dict()  # A mapping between states and the problem instances they came from
         # self.transition_schemas = dict()  # A mapping between transitions and the action schema that induced them
         self.remapping = dict()
+        self.vstar = {}
 
     def add_transitions(self, states, transitions, schemas, instance_id, deadends):
         assert not any(s in self.states for s in states.keys())  # Don't allow repeated states
@@ -106,6 +107,8 @@ class TransitionSample:
         optimal = {(remapping[x], remapping[y]) for x, y in self.optimal_transitions
                    if x in remapping and y in remapping}
 
+        vstar = {remapping[x]: self.vstar[x] for x in ordered_sample if x in self.vstar}
+
         instance = dict()
         states = OrderedDict()
         for i, s in self.states.items():  # Iterate following the order in self.states
@@ -130,6 +133,7 @@ class TransitionSample:
         resampled.mark_as_optimal(optimal)
         resampled.mark_as_unsolvable(unsolvable)
         resampled.mark_as_alive(alive_states)
+        resampled.vstar = vstar
         resampled.remapping = remapping
         _ = [resampled.mark_as_root(r) for r in roots]
         return resampled
@@ -190,9 +194,9 @@ def run_backwards_brfs(g, parents, mincosts, minactions):
 def mark_all_optimal(goals, parents):
     """ Collect all transitions that lie on at least one optimal plan starting from some alive state (i.e. solvable,
      reachable, and not a goal). """
-    mincosts, minactions = {}, {}
+    vstar, minactions = {}, {}
     for g in goals:
-        run_backwards_brfs(g, parents, mincosts, minactions)
+        run_backwards_brfs(g, parents, vstar, minactions)
 
     # minactions contains a map between state IDs and a list with those successors that represent an optimal transition
     # from that state
@@ -204,9 +208,9 @@ def mark_all_optimal(goals, parents):
     # the backwards brfs (i.e. for which mincost[s] is actually defined). Note that the "reachable" part of a state
     # being alive is already guaranteed by the fact that the state is on the sample, as we sample states with a simple
     # breadth first search.
-    alive = {s for s, cost in mincosts.items() if cost > 0}
+    alive = {s for s, cost in vstar.items() if cost > 0}
 
-    return optimal_txs, alive
+    return optimal_txs, alive, vstar
 
 
 def print_atom(atom):
@@ -288,7 +292,7 @@ def print_transition_matrices(sample, config):
[…]

### 2020-08-26 src/sltp/util/defaults.py

@@ -190,7 +190,10 @@ def generate_experiment(expid, domain_dir, domain, **kwargs):
 
         # In the transition-separation encoding, whether we want to exploit the equivalence relation
         # among transitions given by the feature pool
-        use_equivalence_classes=False
+        use_equivalence_classes=False,
+
+        # The slack value for the maximum allowed value for V_pi(s) = slack * V^*(s)
+        v_slack=2,
     )
 
     parameters = {**defaults, **kwargs}  # Copy defaults, overwrite with user-specified parameters

### 2020-09-03 src/sltp/cnfgen.py
Make policy minimization optional

@@ -21,7 +21,7 @@ def run(config, data, rng):
     maxiterations = 9999 if config.use_incremental_refinement else 1
     approach_name = "incremental refinement" if config.use_incremental_refinement else "standard"
 
-    policy = None
+    solution = None
     for it in range(1, maxiterations+1):
         logging.info(f"Starting iteration {it} of the {approach_name} approach")
 
@@ -32,6 +32,10 @@ def run(config, data, rng):
         if exitcode == ExitCode.IterativeMaxsatApproachSuccessful:
             # Here we subtract 1 because last iteration was only the check
             print(f"Iterative approach finished successfully after {it-1} iterations")
+
+            # Recover the policy one more time, but doing policy minimization
+            _, _, policy = generate_policy_from_sat_solution(config, solution, data.model_cache, minimize_policy=True)
+
             return ExitCode.Success, dict(transition_classification_policy=policy)
 
         solution = solve(config.experiment_dir, config.cnf_filename, config.maxsat_solver, config.maxsat_timeout)
@@ -45,7 +49,7 @@ def run(config, data, rng):
 
         # print_maxsat_solution(solution.assignment, config.wsat_allvars_filename)
         features, good_transitions, policy = generate_policy_from_sat_solution(
-            config, solution, data.model_cache)
+            config, solution, data.model_cache, minimize_policy=False)
 
     return ExitCode.Success, dict(transition_classification_policy=None)
 

### 2020-09-03 src/sltp/separation.py

@@ -61,7 +61,7 @@ def compute_transition_classification_policy(config, data, rng):
     return ExitCode.Success, dict(transition_classification_policy=policy)
 
 
-def generate_policy_from_sat_solution(config, solution, model_cache, print_policy=True):
+def generate_policy_from_sat_solution(config, solution, model_cache, minimize_policy=True, print_policy=True):
     assert solution.solved
     features = extract_features_from_sat_solution(config, solution)
     # print_maxsat_solution(solution.assignment, config.wsat_allvars_filename)
@@ -78,7 +78,10 @@ def generate_policy_from_sat_solution(config, solution, model_cache, print_polic
             clause += compute_policy_clauses(f, den_s, den_t)
 
         policy.add_clause(frozenset(clause))
-    policy.minimize()
+
+    if minimize_policy:
+        policy.minimize()
+
     if print_policy:
         policy.print()
     return features, good_transitions, policy

### 2020-09-03 doc/encodings/paper.tex
Update encoding description with goal-distinguish constraints
Reorganize blocks experiments

@@
 \frenchspacing  % DO NOT CHANGE THIS
-\setlength{\pdfpagewidth}{8.5in}  % DO NOT CHANGE THIS
-\setlength{\pdfpageheight}{11in}  % DO NOT CHANGE THIS
+%\setlength{\pdfpagewidth}{8.5in}  % DO NOT CHANGE THIS
+%\setlength{\pdfpageheight}{11in}  % DO NOT CHANGE THIS
+
+\usepackage[margin=1in]{geometry}
 
@@
 
-\noindent where $Dist(s, s', t, t')$ is shorthand for $\bigvee_{f \in D2(s, s', t, t')} Select(f)$.
+\noindent where $Dist(s, s', t, t')$ is shorthand for $\bigvee_{f \in D1\&2(s, s', t, t')} Select(f)$.
 
@@
 
+\smallpar{C11}
+Goals can be distinguished from non-goals.
+\begin{align}
+\bigvee_{f \in D1(s, s')} Select(f),&\;\; \text{for $s$ goal, $s$ not a goal}
+\end{align}
+
 \subsection{Soft Constraints}
 We simply post a constraint $\neg Select(f)$ for each feature $f$ in the pool, with weight equal to its complexity $\mathcal{K}(f)$.
+
+\section{Empirical Results}
+
+\subsection{Blocks:clear}
+Two instances of 5 blocks.
+1 iteration, 19759 vars + 69255 clauses
+Number of transitions/equivalence classes: 1161/55
+Optimal MAXSAT solution with cost 5 found
+
+\begin{verbatim}
+0. n(a) DECs AND n-ontable NILs AND n-ontable>0
+1. n(a) NILs AND n-ontable INCs AND n-ontable>0
+\end{verbatim}
+
+\subsection{Blocks:on}
+\begin{verbatim}
+0. clear(a) ADDs AND holding(·) ADDs AND n-single-blocks INCs AND not-clear(b) NILs
+1. clear(a) DELs AND holding(·) ADDs AND n-single-blocks DECs AND not-clear(b) NILs AND not-clear(b)=0
+2. clear(a) NILs AND clear(a)=0 AND holding(·) ADDs AND n-single-blocks NILs AND not-clear(b) NILs
+3. clear(a) ADDs AND holding(·) DELs AND n-single-blocks NILs AND not-clear(b) ADDs
+4. clear(a) ADDs AND holding(·) DELs AND n-single-blocks INCs AND not-clear(b) NILs AND not-clear(b)>0
+5. clear(a) NILs AND clear(a)>0 AND holding(·) ADDs AND n-single-blocks NILs AND not-clear(b) NILs AND not-clear(b)>0
+6. clear(a) DELs AND holding(·) ADDs AND n-single-blocks NILs AND not-clear(b) NILs
+7. clear(a) NILs AND clear(a)=0 AND holding(·) DELs AND n-single-blocks INCs AND not-clear(b) NILs AND not-clear(b)>0
+8. clear(a) NILs AND holding(·) DELs AND n-single-blocks INCs AND n-single-blocks=0 AND not-clear(b) NILs AND not-clear(b)=0
+9. clear(a) NILs AND clear(a)=0 AND holding(·) DELs AND n-single-blocks DECs AND not-clear(b) NILs
+10. clear(a) ADDs AND holding(·) DELs AND n-single-blocks DECs AND not-clear(b) NILs AND not-clear(b)>0
+11. clear(a) NILs AND clear(a)>0 AND holding(·) DELs AND n-single-blocks INCs AND n-single-blocks>0 AND not-clear(b) NILs
+12. clear(a) ADDs AND holding(·) DELs AND n-single-blocks DECs AND not-clear(b) ADDs
+13. clear(a) DELs AND holding(·) ADDs AND n-single-blocks INCs AND not-clear(b) NILs AND not-clear(b)=0
+14. clear(a) NILs AND clear(a)=0 AND holding(·) ADDs AND n-single-blocks INCs AND not-clear(b) DELs
+15. clear(a) NILs AND clear(a)=0 AND holding(·) DELs AND n-single-blocks INCs AND n-single-blocks>0 AND not-clear(b) NILs AND not-clear(b)=0
+16. clear(a) NILs AND holding(·) ADDs AND n-single-blocks NILs AND not-clear(b) DELs
+17. clear(a) NILs AND clear(a)>0 AND holding(·) ADDs AND n-single-blocks INCs AND n-single-blocks>0 AND not-clear(b) DELs
+18. clear(a) ADDs AND holding(·) ADDs AND n-single-blocks NILs AND not-clear(b) NILs
+19. clear(a) NILs AND clear(a)>0 AND holding(·) DELs AND n-single-blocks INCs AND n-single-blocks=0 AND not-clear(b) NILs AND not-clear(b)>0
+20. clear(a) NILs AND clear(a)>0 AND holding(·) DELs AND n-single-blocks DECs AND not-clear(b) NILs AND not-clear(b)>0
+\end{verbatim}
 

### 2020-09-03 experiments/2020/blocks.py

@@ -5,54 +5,73 @@ from sltp.util.names import blocksworld_names, blocksworld_parameters_for_clear,
 def experiments():
     base = dict(
         domain_dir="blocks",
-        domain="domain.pddl",
-        test_domain="domain.pddl",
         feature_namer=blocksworld_names,
         pipeline="transition_classifier",
-    )
-
-    exps = dict()
-
-    exps["clear"] = update_dict(
-        base,
-        instances=["training_clear_5.pddl"],
-        test_instances=[
-            "instance_5_clear_x_1.pddl",
-            "instance_5_clear_x_2.pddl",
-        ],
-        test_policy_instances=[
-            "instance_5_clear_x_1.pddl",
-            "instance_5_clear_x_2.pddl",
-        ],
-        num_states="all",
-        max_concept_size=8,
-        # concept_generation_timeout=120,  # in seconds
-        concept_generator=None,
-        parameter_generator=blocksworld_parameters_for_clear,
         maxsat_encoding="separation",
         complete_only_wrt_optimal=True,
         prune_redundant_states=False,
         optimal_selection_strategy="complete",
+        num_states="all",
+        concept_generator=None,
+        parameter_generator=None,
+
+        # concept_generation_timeout=120,  # in seconds
+        maxsat_timeout=None,
+    )
+
+    exps = dict()
+
+    strips_base = update_dict(
+        base,
+        domain="domain.pddl",
+        test_domain="domain.pddl",
+    )
+
+    fn_base = update_dict(
+        base,
+        domain="domain_fstrips.pddl",
+        test_domain="domain_fstrips.pddl",
+    )
+
+    strips_atomic_base = update_dict(
+        base,
+        domain="domain_atomic_move.pddl",
+        test_domain="domain_atomic_move.pddl",
+    )
+
+    exps["clear"] = update_dict(
+        strips_base,
+        instances=["training_clear_5.pddl"],
+        test_instances=[
+            "instance_5_clear_x_1.pddl",
+            # "instance_5_clear_x_2.pddl",
+        ],
+        test_policy_instances=[
+            "testing_clear_10_0.pddl",
+            "testing_clear_10_1.pddl",
+        ],
+
+        max_concept_size=8,
+        parameter_generator=blocksworld_parameters_for_clear,
         use_equivalence_classes=True,
-        use_feature_dominance=True,
+        # use_feature_dominance=True,
         use_incremental_refinement=True,
     )
 
     exps["clear_fn"] = update_dict(
[…]

### 2020-09-03 src/sltp/cnfgen.py

@@ -51,7 +51,9 @@ def run(config, data, rng):
         features, good_transitions, policy = generate_policy_from_sat_solution(
             config, solution, data.model_cache, minimize_policy=False)
 
-    return ExitCode.Success, dict(transition_classification_policy=None)
+    # Recover the policy one more time, but doing policy minimization
+    _, _, policy = generate_policy_from_sat_solution(config, solution, data.model_cache, minimize_policy=True)
+    return ExitCode.Success, dict(transition_classification_policy=policy)
 
 
 def print_good_transition_list(good_txs, filename):

### 2020-09-03 src/sltp/util/names.py

@@ -112,6 +112,8 @@ def blocksworld_names(feature):
         "Equal(Inverse(loc_g),Inverse(loc))": "n-right-under-target",  # FSTRIPS
         "Forall(Star(loc),Equal(loc_g,loc))": "n-all-below-well-placed",  # FSTRIPS
         "Forall(Star(on),Equal(on_g,on))": "n-all-below-well-placed",
+        "Exists(on,Nominal(b))": "not-clear(b)",
+        "Exists(on,Nominal(a))": "not-clear(a)",
     }
 
     return extend_namer_to_all_features(base).get(s, s)

### 2020-09-03 doc/encodings/paper.tex
Experiments adjustments

@@
 \usepackage{natbib}
+\usepackage{fancyvrb}
 
@@
 
-\begin{verbatim}
+\begin{Verbatim}[fontsize=\footnotesize]
 0. n(a) DECs AND n-ontable NILs AND n-ontable>0
 1. n(a) NILs AND n-ontable INCs AND n-ontable>0
-\end{verbatim}
+\end{Verbatim}
 
 \subsection{Blocks:on}
-\begin{verbatim}
+\begin{Verbatim}[fontsize=\footnotesize]
 0. clear(a) ADDs AND holding(·) ADDs AND n-single-blocks INCs AND not-clear(b) NILs
@@
 20. clear(a) NILs AND clear(a)>0 AND holding(·) DELs AND n-single-blocks DECs AND not-clear(b) NILs AND not-clear(b)>0
-\end{verbatim}
+\end{Verbatim}
+
+\subsection{Gripper}
+One simple (smallest) standard instance, 4 blocks, two rooms.
+1 iteration
+CNF Theory: 34418 vars + 124935 clauses
+835 features
+Number of transitions/equivalence classes: 1140/65
+concept size bound=10,
+Optimal MAXSAT solution with cost 5 found
+
+\begin{Verbatim}[fontsize=\footnotesize]
+0. ncarried NILs AND ncarried=0 AND robby-at-B DECs
+1. ncarried DECs AND robby-at-B NILs AND robby-at-B>0
+2. ncarried INCs AND ncarried=0 AND robby-at-B NILs AND robby-at-B=0
+3. ncarried NILs AND ncarried>0 AND robby-at-B INCs AND robby-at-B=0
+\end{Verbatim}
+
+\subsection{Spanner}
+8 instances with different spanner configurations, goal is always the same.
+1 iteration
+CNF Theory: 466575 vars + 1732026 clauses
+724 features
+Number of transitions/equivalence classes: 11053/133
+concept size bound=8,
+Optimal MAXSAT solution with cost 9 found
+
+NOTE: Features of the form $f1>f2$ are not necessary for this solution.
+
+\begin{Verbatim}[fontsize=\footnotesize]
+0. bobs-loc-empty NILs AND bobs-loc-empty=0 AND n-items-somewhere DECs AND n-tightened NILs AND n-tightened=0
+1. bobs-loc-empty NILs AND bobs-loc-empty>0 AND n-items-somewhere NILs AND n-items-somewhere>0 AND n-tightened NILs AND n-tightened=0
+2. bobs-loc-empty DELs AND n-items-somewhere NILs AND n-items-somewhere>0 AND n-tightened NILs AND n-tightened=0
+3. bobs-loc-empty NILs AND bobs-loc-empty=0 AND n-items-somewhere NILs AND n-items-somewhere>0 AND n-tightened INCs
+4. bobs-loc-empty ADDs AND n-items-somewhere DECs AND n-tightened NILs AND n-tightened=0
+\end{Verbatim}
+
+\subsection{Pick Reward}
+1 5x5 training instance
+3 iteration
+CNF Theory: 18519 vars + 76267 clauses
+514 features
+Number of transitions/equivalence classes: 432/361
+concept size bound=8,
+Optimal MAXSAT solution with cost 7 found
+
+\begin{Verbatim}[fontsize=\footnotesize]
+ 1. n-rewards NILs AND n-rewards>0 AND unblocked-dist-to-closest-reward DECs AND unblocked-dist-to-closest-reward>0
+ 2. n-rewards DECs AND unblocked-dist-to-closest-reward INCs AND unblocked-dist-to-closest-reward=0
+\end{Verbatim}
 

### 2020-09-03 experiments/2020/blocks.py

@@ -14,6 +14,7 @@ def experiments():
         num_states="all",
         concept_generator=None,
         parameter_generator=None,
+        v_slack=2,
 
         # concept_generation_timeout=120,  # in seconds
         maxsat_timeout=None,

### 2020-09-03 experiments/2020/gripper.py

@@ -7,37 +7,42 @@ def experiments():
         # domain_dir="gripper-m",
         domain_dir="gripper",
         domain="domain.pddl",
+        test_domain="domain.pddl",
         feature_namer=gripper_names,
+        pipeline="transition_classifier",
+        maxsat_encoding="separation",
+        complete_only_wrt_optimal=True,
+        prune_redundant_states=False,
+        optimal_selection_strategy="complete",
+        num_states="all",
+        concept_generator=None,
+        parameter_generator=None,
+        v_slack=2,
+
+        # concept_generation_timeout=120,  # in seconds
+        maxsat_timeout=None,
     )
 
     exps = dict()
 
     exps["small"] = update_dict(
         base,
-        pipeline="transition_classifier",
         # instances=["sample-2balls.pddl", "sample-small.pddl"],
         instances=["prob01.pddl"],
-        test_domain="domain.pddl",
-        # test_instances=[],
-        test_instances=[f"prob{i:02d}.pddl" for i in range(3, 11)],
+        # test_instances=[f"prob{i:02d}.pddl" for i in range(3, 11)],
+        test_instances=[],
         test_policy_instances=[f"prob{i:02d}.pddl" for i in range(3, 21)],
-        num_states="all",
+
         max_concept_size=10,
-        # concept_generation_timeout=120,  # in seconds
-        concept_generator=None,
         parameter_generator=gripper_parameters,
-        # parameter_generator=None
-        maxsat_encoding="separation",
-        complete_only_wrt_optimal=True,
-        prune_redundant_states=False,
-        optimal_selection_strategy="complete",
         use_equivalence_classes=True,
+        # use_feature_dominance=True,
+        use_incremental_refinement=True,
     )
 
-    exps["quick"] = update_dict(
+    exps["small_goal_preds"] = update_dict(
         exps["small"],
-        test_instances=[],
-        test_policy_instances=["prob04.pddl"],
+        parameter_generator=None,
     )
 
     return exps

### 2020-09-03 experiments/2020/reward.py

@@ -1,37 +1,43 @@
 from sltp.util.misc import update_dict
-from sltp.util.names import gripper_names, gripper_parameters, reward_names, no_parameter
+from sltp.util.names import reward_names, no_parameter
 
 
 def experiments():
     base = dict(
         domain_dir="pick-reward",
         domain="domain.pddl",
+        test_domain="domain.pddl",
         feature_namer=reward_names,
+        pipeline="transition_classifier",
+        maxsat_encoding="separation",
+        complete_only_wrt_optimal=True,
+        prune_redundant_states=False,
+        optimal_selection_strategy="complete",
+        num_states="all",
+        concept_generator=None,
+        parameter_generator=None,
+        v_slack=2,
+
+        # concept_generation_timeout=120,  # in seconds
+        maxsat_timeout=None,
     )
 
     exps = dict()
 
     exps["small"] = update_dict(
         base,
-        pipeline="transition_classifier",
         instances=["training_5x5.pddl"],
         # instances=["instance_5.pddl", "instance_4_blocked.pddl"],
-        test_domain="domain.pddl",
-        # test_instances=[],
         test_instances=[],
         test_policy_instances=["instance_7.pddl", "instance_10.pddl", "instance_20.pddl"],
-        num_states="all",
+
         max_concept_size=8,
-        distance_feature_max_complexity=8,
-        # concept_generation_timeout=120,  # in seconds
-        concept_generator=None,
+        # distance_feature_max_complexity=8,
         parameter_generator=no_parameter,
         # parameter_generator=None
-        maxsat_encoding="separation",
-        complete_only_wrt_optimal=True,
-        prune_redundant_states=False,
-        optimal_selection_strategy="complete",
         use_equivalence_classes=True,
+        # use_feature_dominance=True,
+        use_incremental_refinement=False,
     )
 
 

### 2020-09-03 experiments/2020/spanner.py

@@ -4,11 +4,23 @@ from sltp.util.names import spanner_names
 
 def experiments():
     base = dict(
+        # domain_dir="gripper-m",
         domain_dir="spanner-small",
         domain="domain.pddl",
         test_domain="domain.pddl",
-        complete_only_wrt_optimal=True,
         feature_namer=spanner_names,
+        pipeline="transition_classifier",
+        maxsat_encoding="separation",
+        complete_only_wrt_optimal=True,
+        prune_redundant_states=False,
+        optimal_selection_strategy="complete",
+        num_states="all",
+        concept_generator=None,
+        parameter_generator=None,
+        v_slack=2,
+
+        # concept_generation_timeout=120,  # in seconds
+        maxsat_timeout=None,
     )
 
     exps = dict()
@@ -38,19 +50,13 @@ def experiments():
             "prob-10-10-10-1540903568.pddl",
             "prob-15-10-8-1540913795.pddl"
         ],
-        num_states="all",
+
         max_concept_size=8,
-        comparison_features=True,
-        # concept_generation_timeout=120,  # in seconds
-        concept_generator=None,
-        parameter_generator=None,
-        maxsat_encoding="separation",
-        complete_only_wrt_optimal=True,
-        prune_redundant_states=False,
-        optimal_selection_strategy="complete",
-        use_equivalence_classes=True,
         # transition_classification_policy=debug_policy
-        v_slack=2,
+        # comparison_features=True,
+        use_equivalence_classes=True,
+        # use_feature_dominance=True,
+        use_incremental_refinement=True,
     )
 
     return exps

### 2020-09-03 src/sltp/separation.py

@@ -210,7 +210,7 @@ class TransitionClassificationPolicy:
 
     def print(self):
         print("Transition-classification policy with the following transitions labeled as good:")
-        for i, clause in enumerate(self.dnf, start=0):
+        for i, clause in enumerate(self.dnf, start=1):
             print(f"  {i}. " + self.print_clause(clause))
 
     @staticmethod

### 2020-09-03 src/sltp/util/names.py

@@ -31,6 +31,7 @@ def gripper_names(feature):
         # "Not(And(Forall(at-robby,And(Not(Nominal(roomb)),Exists(Inverse(at),<universe>))),Forall(carry,<empty>)))":
         #     ""
         "And(Exists(carry,<universe>),Exists(at_g,at-robby))": "if-robot-at-B-then-num-carried-balls-else-emptyset",
+        "Exists(at_g,at-robby)": "robby-at-B",  # This one only works for a single target room
     }
     return extend_namer_to_all_features(base).get(s, s)
 
@@ -53,14 +54,18 @@ def spanner_names(feature):
         "Exists(at,Exists(link,Exists(Inverse(at),man)))": "n-spanners-on-cell-left-to-man",
         "Exists(Inverse(at),spanner)": "locations-with-a-spanner",
         "Exists(carrying,useable)": "bob-is-carrying-a-usable-spanner",
-        "tightened": "num-tightened-nuts",
+        "tightened": "n-tightened",
         "Exists(Star(link),Exists(Inverse(at),man))": "n-unreachable-locations",
         "Exists(at,Exists(Star(link),Exists(Inverse(at),man)))": "n-unreachable-spanners",
         "LessThan{Num[Exists(Inverse(carrying),<universe>)]}{Num[loose]}": "not-carrying-enough-spanners",
-        "Exists(at,Forall(Inverse(at),man))": "bob-in-empty-loc",
-        "Exists(Star(link),Exists(Inverse(at),spanner))": "num-locs-from-which-some-spanner-is-reachable",
-        "Exists(Inverse(Star(link)),Exists(Inverse(at),<universe>))": "num-locs-reachable-from-a-loc-with-things",
+        "Exists(at,Forall(Inverse(at),man))": "bobs-loc-empty",
+        "Exists(Star(link),Exists(Inverse(at),spanner))": "n-locs-from-which-some-spanner-is-reachable",
+        "Exists(Inverse(Star(link)),Exists(Inverse(at),<universe>))": "n-locs-reachable-from-a-loc-with-things",
         "And(Forall(at,<empty>),useable)": "picked-up-spanners",
+
+        "useable": "n-useable-spanners",
+        "Exists(at,<universe>)": "n-items-somewhere",
+        "Forall(at,<empty>)": "n-picked",
     }
     d = extend_namer_to_all_features(base)
[…]

### 2020-09-05 doc/encodings/paper.tex
Update PDF encoding

--- 2020-08-26 doc/encodings/paper.tex
@@
 
+\newpage
+
 \subsection{Variables}
@@
 
- \item $V(s) \leq d$ for $s$ alive, and $d \in [0, D]$, where $D = \max_{s} \delta \cdot V^*(s)$.
- Note that for states $s$ that are a goal, we know $V(s)=0$, hence $V(s) \leq d$ for all $d$,
- and for states $s$ that are unsolvable, we know that $V(s) \not\leq d$ for all $d$.
- Thus, we can restrict SAT variables $V(s) \leq d$ to those states $s$ that are alive.
+ \item $V(s, d)$ for $s$ alive, and $d \in [0, D]$, where $D = \max_{s} \delta \cdot V^*(s)$,
+ with intended denotation $V(s)=d$.
+ Note that for states $s$ that are a goal, we know $V(s)=0$,
+ and for states $s$ that are unsolvable, we know that $V(s) \neq d$ for all $d$.
+ Thus, we can restrict SAT variables $V(s, d)$ to those states $s$ that are alive.
 
@@
 
-\subsubsection{Auxiliary Variables}
-\begin{itemize}
- \item We'll use $GV(s, s', d)$ to denote that $Good(s, s')$ and $V(s') \leq d$. This is enforced with the constraint
- $GV(s, s', d) \rightarrow Good(s, s') \land V(s') \leq d$ for $s, s'$ alive, $(s, s') \not\in \badtx$, $d \in [0, D)$.
- The other direction of the implication is not necessary.
-\end{itemize}
-
-\newpage
+%\subsubsection{Auxiliary Variables}
+%\begin{itemize}
+% \item We'll use $GV(s, s', d)$ to denote that $Good(s, s')$ and $V(s') \leq d$. This is enforced with the constraint
+% $GV(s, s', d) \rightarrow Good(s, s') \land V(s') \leq d$ for $s, s'$ alive, $(s, s') \not\in \badtx$, $d \in [0, D)$.
+% The other direction of the implication is not necessary.
+%\end{itemize}
 
@@
 
-\smallpar{C2-3} Labeling a transition $(s, s')$ as ``good'' necessarily upper-bounds $V(s)$:
+\smallpar{C2} $V$ is always descending along Good transitions:
 \begin{align}
- Good(s, s') \land V(s') \leq d \rightarrow V(s) \leq d+1,&\;\; \text{for $s, s'$ alive, $(s, s') \not\in \badtx$, $d \in [0, D)$.} \\
- Good(s, s') \rightarrow V(s) \leq d+1,&\;\; \text{for $s$ alive, $s'$ goal, $(s, s') \not\in \badtx$, $d \in [0, D)$.}
+ Good(s, s') \land V(s', d) \rightarrow \bigvee_{d < k \leq D} V(s, k),&\;\; \text{for $s, s'$ alive, $(s, s') \not\in \badtx$, $d \in [1, D)$.} \\
+ V(s', D) \rightarrow \neg Good(s, s'),&\;\; \text{for $s, s'$ alive, $(s, s') \not\in \badtx$.} \tag{\theequation${}^\prime$}
 \end{align}
 
-\smallpar{C3'} All transitions chosen as \emph{Good} need to have the same q-value:
+%\smallpar{C4-5} Any upper bound on $V(s)$ (for $s$ not a goal) needs to be justified:
+%\begin{align}
+% V(s) \leq d+1 \rightarrow \bigvee_{\substack{
+% s' \text{ goal child of } s\\
+% (s, s') \not\in \badtx}} Good(s, s') \lor
+% \bigvee_{\substack{
+% s' \text{ alive child of } s\\
+% (s, s') \not\in \badtx}} GV(s, s', d),&
+% \;\; \text{for $s$ alive, $d \in [0, D)$.} \\
+% \neg V(s) \leq 0,&\;\; \text{for $s$ not a goal.}
+%\end{align}
+
+\smallpar{C3-4}
+Variables $V(s, d)$ define a function that is total over the set of alive states,
+and such that $V(s)$ is within lower bound $V^*(s)$ and upper bound $\delta \cdot V^*(s)$:
 \begin{align}
- V(s) \leq d+1 \land \neg V(s') \leq d \rightarrow \neg Good(s, s'),&\;\; \text{for $s, s'$ alive, $(s, s') \not\in \badtx$, $d \in [0, D)$.} \tag{\theequation${}^\prime$}
+ \bigvee_{V^*(s) \leq d \leq \delta \cdot V^*(s)} V(s,d),&\;\; \text{for $s$ alive.} \\
+ \neg V(s, d) \lor \neg V(s, d')&\;\; \text{for $s$ alive, $1 \leq d < d' \leq D$.}
 \end{align}
 
-\smallpar{C4-5} Any upper bound on $V(s)$ (for $s$ not a goal) needs to be justified:
-\begin{align}
- V(s) \leq d+1 \rightarrow \bigvee_{\substack{
- s' \text{ goal child of } s\\
- (s, s') \not\in \badtx}} Good(s, s') \lor
- \bigvee_{\substack{
- s' \text{ alive child of } s\\
- (s, s') \not\in \badtx}} GV(s, s', d),&
- \;\; \text{for $s$ alive, $d \in [0, D)$.} \\
- \neg V(s) \leq 0,&\;\; \text{for $s$ not a goal.}
-\end{align}
-
-\smallpar{C6}
-At least one atom $V(s) \leq d$ is true for each alive state $s$:
-\begin{align}
- \bigvee_{d \in [0, D]} V(s) \leq d,&\;\; \text{for $s$ alive.}
-\end{align}
-
-\smallpar{C7}
-Atoms $V(s) \leq d$ are consistent among them:
-\begin{align}
- V(s) \leq d \rightarrow V(s) \leq d+1,&\;\; \text{for $s$ alive, $d \in [0, D)$.}
-\end{align}
-
-\smallpar{C8-9}
+\smallpar{C5-6}
 Good transitions can be distinguished from bad transitions.
@@
 
-\noindent where $Dist(s, s', t, t')$ is shorthand for $\bigvee_{f \in D2(s, s', t, t')} Select(f)$.
+\noindent where $Dist(s, s', t, t')$ is shorthand for $\bigvee_{f \in D1\&2(s, s', t, t')} Select(f)$.
 
-\smallpar{C10}
-$V(s)$ cannot be much larger than $V^+(s)$:
+\smallpar{C7 (Optional)}
+Goals are distinguishable from non-goals.
 \begin{align}
- V(s) \leq \delta \cdot V^*(s),&\;\; \text{for $s$ alive.}
+\bigvee_{f \in D1(s, s')} Select(f),&\;\; \text{for $s$ goal, $s$ not a goal}
 \end{align}
+
+\smallpar{C8 (Optional)}
+All selected features need to have some Good transition that takes them to $0$:
+\begin{align}
+ Selected(f) \rightarrow \bigvee_{(s, s') \in Z(f)} Good(s, s'),&\;\; \text{for $f$ in pool}
+\end{align}
+
+\noindent where $Z(f)$ is the set transitions starting in an alive state that change the denotation of $f$ from something
+larger than 0 to 0.
 
@@
 We simply post a constraint $\neg Select(f)$ for each feature $f$ in the pool, with weight equal to its complexity $\mathcal{K}(f)$.
+
+\section{Empirical Results}
+
+\subsection{Blocks:clear}
+\begin{itemize}
+ \item Two instances of 5 blocks.
+ \item 1 iteration, 19759 vars + 69255 clauses
+ \item Number of transitions/equivalence classes: 1161/55
+ \item Optimal MAXSAT solution with cost 5 found
+\end{itemize}
+
+\begin{Verbatim}[fontsize=\footnotesize]
+0. n(a) DECs AND n-ontable NILs AND n-ontable>0
+1. n(a) NILs AND n-ontable INCs AND n-ontable>0
+\end{Verbatim}
+
+\subsection{Blocks:on}
+\begin{Verbatim}[fontsize=\footnotesize]
+ 1. clear(a) NILs AND clear(b) ADDs AND holding(·) ADDs AND n-clear NILs AND n-clear>0
+ 2. clear(a) ADDs AND clear(b) NILs AND clear(b)=0 AND holding(·) DELs AND n-clear INCs AND n-clear>0
+ 3. clear(a) ADDs AND clear(b) DELs AND holding(·) DELs AND n-clear NILs AND n-clear>0
+ 4. clear(a) DELs AND clear(b) NILs AND clear(b)>0 AND holding(·) ADDs AND n-clear DECs
+ 5. clear(a) NILs AND clear(a)=0 AND clear(b) DELs AND holding(·) ADDs AND n-clear NILs AND n-clear>0
+ 6. clear(a) ADDs AND clear(b) NILs AND holding(·) ADDs AND n-clear NILs AND n-clear>0
+ 7. clear(a) NILs AND clear(b) NILs AND holding(·) DELs AND n-clear INCs AND n-clear>0
+ 8. clear(a) ADDs AND clear(b) DELs AND holding(·) ADDs AND n-clear NILs AND n-clear>0
+ 9. clear(a) DELs AND clear(b) NILs AND holding(·) ADDs AND n-clear NILs AND n-clear>0
+ 10. clear(a) NILs AND clear(b) ADDs AND holding(·) DELs AND n-clear INCs AND n-clear>0
+ 11. clear(a) NILs AND clear(a)=0 AND clear(b) NILs AND holding(·) ADDs AND n-clear NILs AND n-clear>0
+ 12. clear(a) NILs AND clear(a)>0 AND clear(b) NILs AND clear(b)=0 AND holding(·) ADDs AND n-clear NILs AND n-clear>0
+\end{Verbatim}
+
+\subsection{Blocks:arbitrary}
+(Atomic move encoding, experiment name: all$\_$at$\_$5)
+
+\begin{itemize}
+ \item 1 instance 5 blocks
+ \item 11 iterations
+ \item CNF Theory: 21795 vars + 189561 clauses
+ \item 12659 features
+ \item concept size bound=10,
+ \item Number of transitions/equivalence classes: 11053/133
+ \item Optimal MAXSAT solution with cost 9 found
+\end{itemize}
+
+\begin{Verbatim}[fontsize=\footnotesize]
+1. n-clear INCs AND n-clear>0 AND n-ontarget DECs AND n-superficially-well-placed NILs AND n-superficially-well-placed>0
+2. n-clear NILs AND n-clear>0 AND n-ontarget INCs AND n-ontarget>0 AND n-superficially-well-placed INCs AND n-superficially-well-placed>0
+3. n-clear INCs AND n-clear>0 AND n-ontarget NILs AND n-ontarget>0 AND n-superficially-well-placed NILs AND n-superficially-well-placed>0
+4. n-clear NILs AND n-clear>0 AND n-ontarget NILs AND n-ontarget>0 AND n-superficially-well-placed INCs AND n-superficially-well-placed>0
+5. n-clear INCs AND n-clear>0 AND n-ontarget INCs AND n-ontarget>0 AND n-superficially-well-placed INCs AND n-superficially-well-placed>0
+6. n-clear DECs AND n-ontarget INCs AND n-ontarget>0 AND n-superficially-well-placed INCs AND n-superficially-well-placed>0
+7. n-clear INCs AND n-clear>0 AND n-ontarget NILs AND n-ontarget>0 AND n-superficially-well-placed DECs
+\end{Verbatim}
+
+\subsection{Gripper}
+One simple (smallest) standard instance, 4 blocks, two rooms.
+1 iteration
+CNF Theory: 34418 vars + 124935 clauses
+835 features
+Number of transitions/equivalence classes: 1140/65
+concept size bound=10,
+Optimal MAXSAT solution with cost 5 found
+
+\begin{Verbatim}[fontsize=\footnotesize]
+0. ncarried NILs AND ncarried=0 AND robby-at-B DECs
+1. ncarried DECs AND robby-at-B NILs AND robby-at-B>0
+2. ncarried INCs AND ncarried=0 AND robby-at-B NILs AND robby-at-B=0
+3. ncarried NILs AND ncarried>0 AND robby-at-B INCs AND robby-at-B=0
+\end{Verbatim}
+
+\subsection{Spanner}
+
+\begin{itemize}
+ \item 8 instances with different spanner configurations, goal is always the same.
+ \item 1 iteration
+ \item CNF Theory: 466575 vars + 1732026 clauses
+ \item 724 features
+ \item concept size bound=8,
+ \item Number of transitions/equivalence classes: 11053/133
+ \item Optimal MAXSAT solution with cost 9 found
+\end{itemize}
+
+NOTE: Features of the form $f1>f2$ are not necessary for this solution.
+
+\begin{Verbatim}[fontsize=\footnotesize]
+0. bobs-loc-empty NILs AND bobs-loc-empty=0 AND n-items-somewhere DECs AND n-tightened NILs AND n-tightened=0
+1. bobs-loc-empty NILs AND bobs-loc-empty>0 AND n-items-somewhere NILs AND n-items-somewhere>0 AND n-tightened NILs AND n-tightened=0
+2. bobs-loc-empty DELs AND n-items-somewhere NILs AND n-items-somewhere>0 AND n-tightened NILs AND n-tightened=0
+3. bobs-loc-empty NILs AND bobs-loc-empty=0 AND n-items-somewhere NILs AND n-items-somewhere>0 AND n-tightened INCs
+4. bobs-loc-empty ADDs AND n-items-somewhere DECs AND n-tightened NILs AND n-tightened=0
+\end{Verbatim}
+
+\subsection{Pick Reward}
+1 5x5 training instance
+3 iteration
+CNF Theory: 18519 vars + 76267 clauses
+514 features
+Number of transitions/equivalence classes: 432/361
+concept size bound=8,
+Optimal MAXSAT solution with cost 7 found
+
+\begin{Verbatim}[fontsize=\footnotesize]
+ 1. n-rewards NILs AND n-rewards>0 AND unblocked-dist-to-closest-reward DECs AND unblocked-dist-to-closest-reward>0
+ 2. n-rewards DECs AND unblocked-dist-to-closest-reward INCs AND unblocked-dist-to-closest-reward=0
+\end{Verbatim}
 

### 2020-09-05 src/features/src/cnf/transition_classification.h
Update CNF generator to new encoding

@@ -44,10 +44,14 @@ public:
         return necessarily_bad_transitions_.find(tx) != necessarily_bad_transitions_.end();
     }
 
+    inline int get_vstar(unsigned s) const {
+        int vstar = tr_set_.transitions().vstar(s);
+        return vstar < 0 ? -1 : vstar;
+    }
+
     inline int get_max_v(unsigned s) const {
         int vstar = tr_set_.transitions().vstar(s);
-        if (vstar < 0) return -1;
-        return std::ceil(options.v_slack * vstar);
+        return vstar < 0 ? -1 : std::ceil(options.v_slack * vstar);
     }
 
     inline unsigned compute_D() const {

### 2020-09-05 src/sltp/separation.py
Update CNF solution printer to new encoding

@@ -38,19 +38,21 @@ def read_selected_variable_ids(assignment, filename):
 def print_maxsat_solution(assignment, filename):
     """ """
     solution = []
-    vleqs = dict()
+    vs = dict()
     with open(filename, 'r') as f:
         for line in f:
             id_, name = line.rstrip().split("\t")
             if assignment[int(id_)] is True:
-                # e.g. Vleq(4, 1)
-                if name.startswith("Vleq("):
-                    state, value = map(int, name[5:-1].split(", "))
-                    vleqs[state] = min(value, vleqs.get(state, sys.maxsize))
+                # e.g. V(4, 1)
+                if name.startswith("V("):
+                    state, value = map(int, name[2:-1].split(", "))
+                    if state in vs:
+                        raise RuntimeError(f"Maxsat solution has more than one value V(s) for state {state}")
+                    vs[state] = value
                 else:
                     solution.append(name)
     print('\n'.join(natsorted(solution, key=lambda y: y.lower())))
-    print('\n'.join(f"V({s}) = {vleqs[s]}" for s in sorted(vleqs.keys())))
+    print('\n'.join(f"V({s}) = {vs[s]}" for s in sorted(vs.keys())))
 
 
 def compute_transition_classification_policy(config, data, rng):

### 2020-09-05 experiments/2020/blocks.py
Update experiment configs

@@ -186,7 +186,7 @@ def experiments():
         use_equivalence_classes=True,
         use_feature_dominance=False,
 
-        force_zeros=True,
+        # force_zeros=True,
     )
 
     exps["all_at_testing"] = update_dict(

### 2020-09-05 experiments/2020/reward.py

@@ -32,7 +32,7 @@ def experiments():
         test_policy_instances=["instance_7.pddl", "instance_10.pddl", "instance_20.pddl"],
 
         max_concept_size=8,
-        # distance_feature_max_complexity=8,
+        distance_feature_max_complexity=8,
         parameter_generator=no_parameter,
         # parameter_generator=None
         use_equivalence_classes=True,
@@ -40,6 +40,7 @@ def experiments():
         use_incremental_refinement=False,
     )
 
+    exps["debug"] = update_dict(exps["small"], feature_generator=debug_features)
 
     # One reason for overfitting: in a 3x3 grid, with 2 booleans per dimension you can perfectly represent any position
     # exps['sample_2x2_1reward'] = update_dict(exps['sample_1x3'], instances=["sample_2x2_1reward.pddl"])
@@ -51,5 +52,13 @@ def experiments():
     # Same but using goal-concepts instead of goal parameters:
     # exps["instance_5_gc"] = update_dict(exps["instance_5"], parameter_generator=None)
 
-
     return exps
+
+
+def debug_features(lang):
+    unblocked_dist = "Dist[at;Restrict(adjacent,unblocked);reward]"
+    nrewards = "Num[reward]"
+    return [
+        unblocked_dist,
+        nrewards,
+    ]

### 2020-09-05 experiments/2020/spanner.py

@@ -56,7 +56,7 @@ def experiments():
         # comparison_features=True,
         use_equivalence_classes=True,
         # use_feature_dominance=True,
-        use_incremental_refinement=True,
+        use_incremental_refinement=False,
     )
 
     return exps

### 2020-09-05 experiments/2020/visitall.py

@@ -34,15 +34,14 @@ def experiments():
         test_instances=[],
         test_policy_instances=all_test_instances(),
 
-        max_concept_size=8,
-        distance_feature_max_complexity=8,
-        cond_feature_max_complexity=8,
+        max_concept_size=6,
+        distance_feature_max_complexity=6,
         use_equivalence_classes=True,
         # use_feature_dominance=True,
         # use_incremental_refinement=True,
     )
 
-    exps["test"] = update_dict(
+    exps["debug"] = update_dict(
         base,
         instances=[
             # 'problem02-full.pddl',

### 2020-09-05 src/sltp/util/names.py

@@ -139,8 +139,8 @@ def reward_names(feature):
     s = str(feature)
     base = {
         "reward": "n-rewards",
-        "Dist[at, adjacent, reward]": "dist-to-closest-reward",
-        "Dist[at, Restrict(adjacent,unblocked), reward]": "unblocked-dist-to-closest-reward",
+        "Dist[at;adjacent;reward]": "dist-to-closest-reward",
+        "Dist[at;Restrict(adjacent,unblocked);reward]": "unblocked-dist-to-closest-reward",
     }
 
     return extend_namer_to_all_features(base).get(s, s)

### 2020-09-07 src/sltp/separation.py
Small fixes to testing of manually-provided policies

@@ -1,6 +1,5 @@
 import itertools
 import os
-import sys
 
 from natsort import natsorted
 
@@ -180,6 +179,14 @@ class DNFAtom:
         return self.feature == other.feature and self.value == other.value
 
 
+def changes_are_analogous(change1, change2):
+    normalization = {
+        FeatureValueChange.ADD: FeatureValueChange.INC,
+        FeatureValueChange.DEL: FeatureValueChange.DEC,
+    }
+    return normalization.get(change1, change1) == normalization.get(change2, change2)
+
+
 class TransitionClassificationPolicy:
     def __init__(self, features):
         self.features = features
@@ -206,7 +213,7 @@ class TransitionClassificationPolicy:
                     return False
             else:
                 tx_val = feat.feature.diff(feat.denotation(m0), feat.denotation(m1))
-                if tx_val != atom.value:
+                if not changes_are_analogous(tx_val, atom.value):
                     return False
         return True
 

### 2020-09-07 src/sltp/tester.py

@@ -2,7 +2,7 @@ import logging
 import sys
 
 from .models import FeatureModel
-from .separation import TransitionClassificationPolicy
+from .separation import TransitionClassificationPolicy, generate_user_provided_policy
 from .util.tools import Abstraction, AbstractAction, generate_qualitative_effects_from_transition
 from .features import generate_model_cache, create_model_factory, compute_static_atoms
 from .returncodes import ExitCode
@@ -202,8 +202,12 @@ def test_transition_classification_policy(config, data, rng):
         return ExitCode.Success, dict()
 
     def create_policy(model_factory, static_atoms):
-        return create_action_selection_function_from_transition_policy(model_factory, static_atoms,
-                                                                       data.transition_classification_policy)
+        if config.transition_classification_policy is not None:
+            policy = generate_user_provided_policy(config)
+        else:
+            policy = data.transition_classification_policy
+
+        return create_action_selection_function_from_transition_policy(model_factory, static_atoms, policy)
 
     # Test that the policy reaches a goal when applied on all test instances
     res = apply_policy_on_test_instances(config, create_policy)

### 2020-09-07 doc/encodings/paper.tex
Optional constraints in CNF to force descending transitions to be selected as Good

@@
 
+\smallpar{C3 (Optional)} All descending transitions must be considered Good:
+\begin{align}
+ V(s, d) \land V(s', d') \rightarrow Good(s, s'),&\;\; \text{for $s, s'$ alive, $(s, s') \not\in \badtx$, $1 \leq d' < d \leq D$,} \\
+ Good(s, s'),&\;\; \text{for $s$ alive, $s'$ goal.} \tag{\theequation${}^\prime$}
+\end{align}
+
 %\smallpar{C4-5} Any upper bound on $V(s)$ (for $s$ not a goal) needs to be justified:

### 2020-09-07 src/features/src/cnf/generator.h

@@ -53,6 +53,9 @@ struct Options {
     //! decrease to zero in some transition
     bool force_zeros;
 
+    //! In the transition-separation encoding, whether to force any V-descending transition to be labeled as Good
+    bool decreasing_transitions_must_be_good;
+
     //! A list of user-provided feature IDs for which we want to enforce selection
     std::vector<unsigned> enforced_features;
 

### 2020-09-07 src/features/src/cnfgen.cpp

@@ -66,6 +66,9 @@ sltp::cnf::Options parse_options(int argc, const char **argv) {
          "In the transition-separation encoding, whether to post constraints to ensure that all selected features "
          "decrease to zero in some transition")
 
+        ("decreasing_transitions_must_be_good",
+         "In the transition-separation encoding, whether to force any V-descending transition to be labeled as Good")
+
         ("use-feature-dominance",
          "In the transition-separation encoding, whether we want to exploit the dominance among features to ignore "
          "dominated features and reduce the size of the encoding.")
@@ -94,6 +97,7 @@ sltp::cnf::Options parse_options(int argc, const char **argv) {
     options.use_feature_dominance = vm.count("use-feature-dominance") > 0;
     options.use_incremental_refinement = vm.count("use-incremental-refinement") > 0;
     options.distinguish_goals = vm.count("distinguish-goals") > 0;
+    options.decreasing_transitions_must_be_good = vm.count("decreasing_transitions_must_be_good") > 0;
     options.force_zeros = vm.count("force_zeros") > 0;
     options.cross_instance_constraints = vm.count("cross_instance_constraints") > 0;
     options.v_slack = vm["v_slack"].as<double>();

### 2020-09-07 src/sltp/cnfgen.py

@@ -41,7 +41,7 @@ def run(config, data, rng):
         solution = solve(config.experiment_dir, config.cnf_filename, config.maxsat_solver, config.maxsat_timeout)
 
         if solution.result == "UNSATISFIABLE":
-            return ExitCode.MaxsatModelUnsat, dict()
+            return ExitCode.MaxsatModelUnsat, dict(transition_classification_policy=None)
         elif solution.result == "SATISFIABLE":
             logging.info(f"Possibly suboptimal MAXSAT solution with cost {solution.cost} found")
         else:
@@ -83,6 +83,7 @@ def generate_cnf(config, data):
     args += ["--distinguish-goals"] if config.distinguish_goals else []
     args += ["--cross_instance_constraints"] if config.cross_instance_constraints else []
     args += ["--force_zeros"] if config.force_zeros else []
+    args += ["--decreasing_transitions_must_be_good"] if config.decreasing_transitions_must_be_good else []
     retcode = execute([cmd] + args)
 
     if retcode == 0:

### 2020-09-07 src/sltp/util/defaults.py
Optional constraints in CNF to force descending transitions to be selected as Good

@@ -211,6 +211,9 @@ def generate_experiment(expid, domain_dir, domain, **kwargs):
         # In the transition-separation encoding, whether to post constraints to ensure that all selected features
         # decrease to zero in some transition
         force_zeros=False,
+
+        # In the transition-separation encoding, whether to force any V-descending transition to be labeled as Good
+        decreasing_transitions_must_be_good=False,
     )
 
     parameters = {**defaults, **kwargs}  # Copy defaults, overwrite with user-specified parameters

### 2020-09-07 experiments/2020/childsnack.py
Miconic (working) and childnsack (untested) domains

@@ -1,0 +1,66 @@
+from sltp.util.misc import update_dict, extend_namer_to_all_features
+from sltp.util.names import childsnack_names
+
+
+def experiments():
+    base = dict(
+        domain_dir="childsnack-opt14-strips",
+        domain="domain.pddl",
+        test_domain="domain.pddl",
+        feature_namer=childsnack_names,
+        pipeline="transition_classifier",
+        maxsat_encoding="separation",
+        complete_only_wrt_optimal=True,
+        prune_redundant_states=False,
+        optimal_selection_strategy="complete",
+        num_states="all",
+        concept_generator=None,
+        parameter_generator=None,
+        v_slack=2,
+
+        # concept_generation_timeout=120,  # in seconds
+        maxsat_timeout=None,
+    )
+
+    exps = dict()
+
+    exps["small"] = update_dict(
+        base,
+        # instances=['sample{:02d}.pddl'.format(i) for i in range(1, 5)],
+        instances=[
+            # 'sample_mini.pddl',
+            'sample01.pddl',
+            'sample02.pddl',
+            # 'sample03.pddl',
+
+            # 'child-snack_pfile01-2.pddl'  # STATE SPACE TOO LARGE
+        ],
+        test_instances=[
+            # 'child-snack_pfile01-2.pddl',
+        ],
+        test_policy_instances=all_test_instances(),
+
+        # clean_workspace=False,
+        # goal_selector=goal_selector,
+        # create_goal_features_automatically=True,
+        max_concept_size=10,
+        use_equivalence_classes=True,
+        # use_feature_dominance=True,
+        use_incremental_refinement=True,
+    )
+
+    return exps
+
+
+def all_test_instances():
+    return ['child-snack_pfile01.pddl', 'child-snack_pfile01-2.pddl', 'child-snack_pfile02.pddl',
+            'child-snack_pfile04-2.pddl', 'child-snack_pfile05.pddl', 'child-snack_pfile07-2.pddl',
+            'child-snack_pfile08.pddl', 'child-snack_pfile10-2.pddl', 'child-snack_pfile01.pddl',
+            'child-snack_pfile03-2.pddl', 'child-snack_pfile04.pddl', 'child-snack_pfile06-2.pddl',
+            'child-snack_pfile07.pddl', 'child-snack_pfile09-2.pddl', 'child-snack_pfile10.pddl',
[…]

### 2020-09-07 experiments/2020/miconic.py

@@ -5,8 +5,13 @@ from sltp.util.names import miconic_names
 def experiments():
     base = dict(
         domain_dir="miconic",
-        domain="domain.pddl",
-        test_domain="domain.pddl",
+        # domain="domain.pddl",
+        # test_domain="domain.pddl",
+
+        # Without the fix, the "board" action allows to board passengers that are not on the floor anymore!
+        test_domain="domain-with-fix.pddl",
+        domain="domain-with-fix.pddl",
+
         feature_namer=miconic_names,
         pipeline="transition_classifier",
         maxsat_encoding="separation",
@@ -27,11 +32,13 @@ def experiments():
     exps["small"] = update_dict(
         base,
         instances=[
-            's2-0.pddl',
+            # 's2-0.pddl',
             # 's2-1.pddl',
             # 's2-2.pddl',
             # 's2-3.pddl',
-            's3-0.pddl',
+            # 's3-0.pddl',
+            's4-0.pddl',
+            'training2.pddl',
         ],
         test_instances=[
         ],
@@ -42,6 +49,35 @@ def experiments():
         use_equivalence_classes=True,
         # use_feature_dominance=True,
         use_incremental_refinement=True,
+        decreasing_transitions_must_be_good=True,
+    )
+
+    exps["small2"] = update_dict(
+        exps["small"],
+        instances=[
+            's3-0.pddl',
+            'training2.pddl',
+        ],
+        use_incremental_refinement=False,
+    )
+
+    exps["debug"] = update_dict(
+        exps["small"],
+        instances=[
+            # 'debug.pddl',
+            's4-0.pddl',
+            'training2.pddl',
+        ],
+
+        feature_generator=debug_features,
+        # transition_classification_policy=debug_policy,
+        print_denotations=True,
+        use_incremental_refinement=False,
+
+        test_policy_instances=all_test_instances(),
+        # test_policy_instances=[
+        #     's4-0.pddl',
+        # ]
     )
 
     return exps
@@ -53,3 +89,40 @@ def all_test_instances():
         for j in range(0, 5):  # Each x has 5 subproblems
             instances.append("s{}-{}.pddl".format(i, j))
     return instances
+
+
+nserved = "Num[served]"  # k = 1
+nboarded = "Num[boarded]"  # k = 1
+lift_at_dest_some_boarded_pass = "Bool[And(lift-at,Exists(Inverse(destin),boarded))]"  # k = 6
+
[…]

### 2020-09-07 experiments/childsnack.py

@@ -1,4 +1,5 @@
 from sltp.util.misc import update_dict, extend_namer_to_all_features
+from sltp.util.names import childsnack_names
 
 
 def experiments():
@@ -8,7 +9,7 @@ def experiments():
         domain="domain.pddl",
         test_domain="domain.pddl",
         complete_only_wrt_optimal=True,
-        feature_namer=feature_namer,
+        feature_namer=childsnack_names,
     )
 
     exps = dict()
@@ -60,32 +61,6 @@ def experiments():
     )
 
     return exps
-
-
-def feature_namer(feature):
-    s = str(feature)
-    base = {
-        # "": "",
-        "served": "num-served-children",
-        "And(Not(served),child)": "num-unserved-children",
-        "notexist": "num-unprepared-sandwiches",
-        "no_gluten_sandwich": "num-sandwiches-wo-gluten",
-        "at_kitchen_bread": "num-breads-at-kitchen",
-        "at_kitchen_content": "num-fillings-at-kitchen",
-        "at_kitchen_sandwich": "num-sandwiches-at-kitchen",
-        "Exists(ontray,<universe>)": "num-sandwiches-on-some-tray",
-        "And(allergic_gluten,served)": "num-allergic-served",
-        "Exists(at,Nominal(kitchen))": "num-trays-on-kitchen",
-        "And(Not(served),not_allergic_gluten)": "num-unallergic-unserved",
-        "And(Not(served),allergic_gluten)": "num-allergic-unserved",
-        "And(Not(no_gluten_content),content-portion)": "num-gluten-free-fillings",
-        "And(Not(no_gluten_bread),bread-portion)": "num-gluten-free-breads",
-        "And(Not(no_gluten_sandwich),sandwich)": "num-sandwiches-with-gluten",
-        "Exists(at,Exists(Inverse(waiting),<universe>))": "num-trays-on-place-with-some-child",
-        "Exists(ontray,Exists(at,Nominal(kitchen)))": "num-sandwiches-on-some-tray-in-kitchen",
-    }
-
-    return extend_namer_to_all_features(base).get(s, s)
 
 
 def all_test_instances():

### 2020-09-07 src/sltp/util/names.py

@@ -190,10 +190,18 @@ def gridworld_names(feature):
 def miconic_names(feature):
     s = str(feature)
     base = {
-        "": "",
-        "": "",
-        "": "",
-        "": "",
+        "And(lift-at,Exists(Inverse(destin),boarded))": "lift_at_dest_some_boarded_pass",
+        "served": "n-served",
+        "boarded": "n-boarded",
+
+        # num passengers unboarded s.t. lift is on their origin floor
+        "And(And(Not(boarded),Not(served)),Exists(origin,lift-at))": "n_pass_ready_to_board",
+        "And(lift-at,Exists(Inverse(origin),And(Not(boarded),Not(served))))": "lift_at_origin_some_awaiting_pass",
+
+        "Exists(origin,lift-at)": "n-pass-waiting-on-lifts-floor",
+        "And(Exists(destin,lift-at),boarded)": "n-boarded-and-at-destiny",
+        "Forall(origin,lift-at)": "n-ready-to-board",
+
     }
 
     return extend_namer_to_all_features(base).get(s, s)
@@ -204,8 +212,6 @@ def satellite_names(feature):
     base = {
         "calibrated": "n-calibrated-instruments",
         "power_on": "n-on-instruments",
-        "": "",
-        "": "",
         "": "",
         "": "",
     }
@@ -235,4 +241,30 @@ def visitall_names(feature):
         "visited": "n-visited",
         "Not(visited)": "n-unvisited",
     }
+    return extend_namer_to_all_features(base).get(s, s)
+
+
+def childsnack_names(feature):
+    s = str(feature)
+    base = {
+        # "": "",
+        "served": "num-served-children",
+        "And(Not(served),child)": "num-unserved-children",
+        "notexist": "num-unprepared-sandwiches",
+        "no_gluten_sandwich": "num-sandwiches-wo-gluten",
+        "at_kitchen_bread": "num-breads-at-kitchen",
+        "at_kitchen_content": "num-fillings-at-kitchen",
+        "at_kitchen_sandwich": "num-sandwiches-at-kitchen",
+        "Exists(ontray,<universe>)": "num-sandwiches-on-some-tray",
+        "And(allergic_gluten,served)": "num-allergic-served",
+        "Exists(at,Nominal(kitchen))": "num-trays-on-kitchen",
+        "And(Not(served),not_allergic_gluten)": "num-unallergic-unserved",
[…]
